### 2023-02-27 Texfiles/Notes.tex
Add all code files
Adding files that I was previously working on in OneDrive.

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{enumerate}
\setlength{\parindent}{0pt}
\usepackage[parfill]{parskip}
\date{Lent 2023}
\title{Dissertation notes}
\author{Tobias Leigh-Wood}
\usepackage[letterpaper, margin=1.1in]{geometry}
\linespread{1.1}

\begin{document}
\maketitle

\section*{To Do List}
\begin{itemize}
    \item What actually happened?
          Annuity provision peaked in 2012. Prior to the March 2011 budget individuals had to annuitise their pension pot.
          After 2011 they only had to annuitise up to a minimum income requirement (MIR) of £20,000. This minimum income kept most people using annuities.

          In March 2014 there was a surprise announcement that the MIR would be scrapped and individuals would be able to 'draw down' their pension pot instead.
          This meant the compulsory market for pension annuities basically stopped.

    \item When did the changes take place?
    \item When were the changes announced?
\end{itemize}

\end{document}

### 2023-02-27 R/1__explore.R

@@ -1,0 +1,109 @@
+setwd(dirname(rstudioapi::getSourceEditorContext()$path))
+
+library(data.table)
+library(dplyr)
+library(readxl)
+
+# Item expenditure data set
+item_expend <- fread("Data/2011_UKDA-7272-tab/tab/2011_dv_set89_ukanon.tab")
+
+# Household info
+household_info <- fread("Data/2011_UKDA-7272-tab/tab/2011_dvhh_ukanon.tab")
+
+# Individual info
+household_members_info <- fread("Data/2011_UKDA-7272-tab/tab/2011_dvper_ukanon_v2.tab")
+
+# Expenditure look up
+expend_lookup <- read_excel("Data/2011_UKDA-7272-tab/mrdoc/excel/7272volume_d_expenditure_codes_2011.xls", sheet = "Part 1")
+library(httpgd)
+install.packages("httpgd")
+regions <- fread("Gor_name, Gor
+North East, 1
+North West,	2
+Merseyside,	3
+Yorkshire and the Humber,	4
+East Midlands,	5
+West Midlands,	6
+Eastern,     	7
+London,	8
+South East (except Greater London),	9
+South West,	10
+Wales,	11
+Scotland,	12
+Northern Ireland,	13", header = T)
+
+
+dim(household_info[, 1:2])
+
+household_info[, 1:2]
+household_info$Gor
+household_info$OAC3D
+household_info$incanon
+
+household_info[, .(income = mean(incanon)), by = .(Gor)][regions, on = "Gor"] %>%
+  arrange(income)
+
+x <- c(1:1000 / 100)
+y <- gamma(x)
+
+ggplot2::ggplot(data = data.frame(x = x, y = y)) +
+  geom_
+
+
+
+head(expend_lookup)
+
+# Ok so government region is included in 2011
+# I can assume it is also included in every other year
+# Let's try and match it though for individuals and then see their total bread expenditure
+
+
+
+# Cool so we can get some basic summary stats
+
+# How to calculate inflation indices from the data though
+# Need to look at changes in cost of the same basket
+# And then update the basket each year?
+
+# I could check some of these against housing costs from census
+
+dim(household_info)
+household_info$
+
+  dim(item_expend)
+names(item_expend)
+
+hist(item_expend$Person)
+
+length(unique(item_expend$case))
+
+
+# Ok, so we match case from item_expend with case from household
+
+household_info$case
+
+
+nrow(data)
+
+head(data)
+
+length(unique(item_expend$COI_PLUS))
+
[…]

### 2023-02-27 R/3__aggregate_consumption.R

@@ -1,0 +1,78 @@
+library(data.table)
+library(dplyr)
+library(stringr)
+library(ggplot2)
+
+
+# Join consumption data from the various waves of data
+# Hopefully this isnt too difficult
+
+dir("Data/household_files")
+
+test_hh <- fread("Data/derived_files/2014_dvhh_ukanon.tab")
+test_person <- fread("Data/derived_files/2014_dvper_ukanon.tab")
+
+
+# So we need the age variable from the person files
+# And the consumption variable from the household files
+
+# Is case still the joining variable
+
+
+consumption_data <- test_hh[, .(case, total_expend = P600)]
+age_data <- test_person[Person == 1, .(case, age = a005p)] # Select head of household
+
+joined_data <- full_join(consumption_data, age_data,
+    by = "case"
+)
+
+## Ok so we want to do the aboce for all years
+
+derived_files <- dir("Data/derived_files")
+
+str_extract_all(derived_files, "\\d{4}")
+# Do 2019-20 data refer to 2019 or 2020 - always the lower year.
+
+JoinYear <- function(year_of_data = 2004) {
+    files_for_year <- grep(year_of_data, derived_files, value = T)
+
+    full_name_for_files <- paste0("Data/derived_files/", files_for_year)
+
+    # household file
+    hh_file <- grep("dvhh", full_name_for_files, value = T)
+    hh_data <- fread(hh_file)
+    setnames(hh_data, str_to_lower)
+    hh_consump <- hh_data[, list(case, total_expend = p600)] # nolint: object_usage_linter.
+
+    hh_consump[, data_wave := year_of_data] # nolint: object_usage_linter.
+
+    # person files
+    per_file <- grep("dvper", full_name_for_files, value = TRUE)
+    per_data <- fread(per_file)
+    setnames(per_data, str_to_lower)
+    per_age <- per_data[person == 1, list(
+        case, # nolint: object_usage_linter.
+        age = a005p, # nolint: object_usage_linter.
+        annuity_income = b346 # nolint: object_usage_linter.
+    )]
+
+    consump_age <- full_join(hh_consump, per_age, by = "case")
+    return(consump_age)
+}
+
+JoinYear(year_of_data = 2004)
+
+all_data <- lapply(
+    2004:2020,
+    JoinYear
+) %>%
+    rbindlist() %>%
+    mutate(birth_year = data_wave - age) %>%
[…]

### 2023-02-27 R/4__simple_diff_in_diff.R

@@ -1,0 +1,6 @@
+# Not going to run anything yet
+
+# I actually want to see whether the above graphs change with different consumption levels?
+
+# There is also an income from annuitys variable I could look for
+b346

### 2023-02-27 R/5__annuity_data.r

@@ -1,0 +1,33 @@
+library(data.table)
+library(dplyr)
+library(ggplot2)
+
+all_data <- fread("Data/clean_data/joined_household_individual.csv")
+names(all_data)
+
+# Plot mean annuity income
+# Plot proportion who have an annuity income
+income <- all_data %>%
+    mutate(any_annuity_inc = fifelse(annuity_income > 0, 1, 0)) %>%
+    group_by(age, birth_year) %>%
+    summarise(
+        prop_annuity = mean(any_annuity_inc),
+        mean_annuity = mean(annuity_income[any_annuity_inc == 1])
+    )
+
+income %>%
+    filter(birth_year == 1945)
+
+# Let's plot any annuity income for over 65s
+
+plot_data <- income %>%
+    filter(age == 65)
+
+ggplot(
+    plot_data,
+    aes(
+        x = birth_year,
+        y = prop_annuity
+    )
+) +
+    geom_point()

### 2023-02-27 R/6__ELSA_data.R

@@ -1,0 +1,38 @@
+# This script will sort out the ELSA data.
+
+library(data.table)
+library(dplyr)
+
+# Set the correct working directory.
+# Note you need to run the whole script to get this to work.
+# I.e crtl+shift+enter.
+cur_dir <- getSrcDirectory(function(x) {
+    x
+})
+setwd(cur_dir)
+stop() # Remove if running the whole thing through
+
+
+# I don't think there is consumption data but there is income data
+# And there is pension data and hopefully annuity data as well
+if (!dir.exists("Data/ELSA/elsa_unziped/UKDA-5050-tab/tab/")) {
+    readLines("Do you want to unzip the ELSA files")
+    unzip(
+        zipfile = "Data/ELSA/5050tab_BDDEA8F799D95FB6A9F6FB5ADBE43D55F7B50718D8615C88A14F94298AADF043_V1.zip",
+        exdir = "Data/ELSA/elsa_unziped"
+    )
+}
+
+financial_files <- dir("Data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "wave",
+    full.names = T
+) %>%
+    grep("financial_", ., value = T)
+
+fread(financial_files[[1]]) %>%
+    names() %>%
+    grep("exp", ., value = T)
+# Join across waves I think. Is there a unique joiner in each household
+
+
+lapply(financial_files, \(x) "idauniq" %in% names(fread(x)))

### 2023-02-28 Texfiles/Notes.tex
updating notes
Create and practice with bibliography

@@
     \item What actually happened?
-          Annuity provision peaked in 2012. Prior to the March 2011 budget individuals had to annuitise their pension pot.
+          Annuity provision peaked in 2012 (\cite{cannon_effect_2016}) Prior to the March 2011 budget individuals had to annuitise their pension pot.
           After 2011 they only had to annuitise up to a minimum income requirement (MIR) of £20,000. This minimum income kept most people using annuities.
@@
           This meant the compulsory market for pension annuities basically stopped.
+Moreover, the amount needed to be classed as a 'small pot' increased meaning there were more pots that could be taken out all at once.
+          In particular this reform meant the number of small annuities in the compulsory market was drastically reduced.
 
@@
 
+\bibliographystyle{chicago}
+\bibliography{references}
 \end{document}

### 2023-03-13 Texfiles/Notes.tex
Timeline of changes
new notes

@@
 
-    \item When did the changes take place?
-    \item When were the changes announced?
+      \item \textbf{When did the changes take place?}
+
+            First round in 2012 and a second round of changes in March 2014. Which age groups were particularly impacted?
+
+      \item \textbf{When were the changes announced?}
+
+            There were expected changes announced in March 2011. Unexpected changes made in March 2014.
+            The changes made in 2011 would still have been quite unexpected so we could probably still expect changes in early retirement consumption.
+
+            So perhaps we could look at people just above and just below the cut offs.
+      \item \textbf{What were the pension rules before?}
+            Were there cutoffs that changed peoples behaviour?
+
 \end{itemize}

### 2023-03-13 R/6__ELSA_data.R

@@ -2,6 +2,7 @@
 
 library(data.table)
 library(dplyr)
+library(stringr)
 
 # Set the correct working directory.
 # Note you need to run the whole script to get this to work.
@@ -12,8 +13,10 @@ cur_dir <- getSrcDirectory(function(x) {
 setwd(cur_dir)
 stop() # Remove if running the whole thing through
 
+setwd(dirname(rstudioapi::getSourceEditorContext()$path))
 
 # I don't think there is consumption data but there is income data
+# there must be consumption data
 # And there is pension data and hopefully annuity data as well
 if (!dir.exists("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/")) {
     readLines("Do you want to unzip the ELSA files")
@@ -23,16 +26,35 @@ if (!dir.exists("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/")) {
     )
 }
 
+
 financial_files <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
     pattern = "wave",
     full.names = T
 ) %>%
     grep("financial_", ., value = T)
 
+
+
 fread(financial_files[[1]]) %>%
     names() %>%
     grep("exp", ., value = T)
+
 # Join across waves I think. Is there a unique joiner in each household
-
-
 lapply(financial_files, \(x) "idauniq" %in% names(fread(x)))
+
+ReadAndSetWave <- function(filename) {
+    wave <- str_extract("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_5_financial_derived_variables.tab", "wave_\\d{1}")
+    data <- fread(filename)
+    data[, wave := wave]
+    return(data)
+}
+
+all_finance <- lapply(financial_files, ReadAndSetWave) %>%
+    rbindlist(fill = T, use.names = T)
+
+
+
+all_finance
+
+## aim is to figure out what change there was in the spending habit of retirees.
+# where is the documentation?

### 2023-05-22 R/6__ELSA_data.R
recent changes to code

@@ -1,4 +1,4 @@
-# This script will sort out the ELSA data.
+# This script will unzip the elsa zip file
 
 library(data.table)
 library(dplyr)
@@ -25,36 +25,3 @@ if (!dir.exists("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/")) {
         exdir = "../../data/ELSA/elsa_unziped"
     )
 }
-
-
-financial_files <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-    pattern = "wave",
-    full.names = T
-) %>%
-    grep("financial_", ., value = T)
-
-
-
-fread(financial_files[[1]]) %>%
-    names() %>%
-    grep("exp", ., value = T)
-
-# Join across waves I think. Is there a unique joiner in each household
-lapply(financial_files, \(x) "idauniq" %in% names(fread(x)))
-
-ReadAndSetWave <- function(filename) {
-    wave <- str_extract("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_5_financial_derived_variables.tab", "wave_\\d{1}")
-    data <- fread(filename)
-    data[, wave := wave]
-    return(data)
-}
-
-all_finance <- lapply(financial_files, ReadAndSetWave) %>%
-    rbindlist(fill = T, use.names = T)
-
-
-
-all_finance
-
-## aim is to figure out what change there was in the spending habit of retirees.
-# where is the documentation?

### 2023-05-22 R/7__ELSA_pension.r

@@ -26,9 +26,9 @@ pension_files <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
 head(pension_files)
 
 
-fread(financial_files[[1]]) %>%
+fread(pension_files[[1]]) %>%
     names() %>%
-    grep("exp", ., value = T)
+    grep("", ., value = T)
 
 # Join across waves I think. Is there a unique joiner in each household
 lapply(financial_files, \(x) "idauniq" %in% names(fread(x)))
@@ -40,12 +40,14 @@ ReadAndSetWave <- function(filename) {
     return(data)
 }
 
-all_finance <- lapply(financial_files, ReadAndSetWave) %>%
+all_pension <- lapply(pension_files, ReadAndSetWave) %>%
     rbindlist(fill = T, use.names = T)
 
-# threr is partner income
+# there is partner income
 
-all_finance
+head(all_pension)
 
-## aim is to figure out what change there was in the spending habit of retirees.
-# where is the documentation?
+# do annuity rates change in the data
+
+# was eric french's value function iteration paper based on elsa data
+# could i do that?

### 2023-05-22 R/8__ELSA_financial.r

@@ -1,0 +1,88 @@
+library(data.table)
+library(dplyr)
+library(stringr)
+
+# Set the correct working directory.
+# Note you need to run the whole script to get this to work.
+# I.e crtl+shift+enter.
+cur_dir <- getSrcDirectory(function(x) {
+    x
+})
+setwd(cur_dir)
+stop() # Remove if running the whole thing through
+
+setwd(dirname(rstudioapi::getSourceEditorContext()$path))
+
+ReadAndSetWave <- function(filename) {
+    wave <- str_extract(filename, "wave_\\d{1}")
+    data <- fread(filename)
+    data[, wave := wave]
+    return(data)
+}
+
+
+# Compile finance files
+
+financial_files <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "wave",
+    full.names = TRUE
+) %>%
+    grep("financial_", ., value = TRUE)
+
+# Join across waves I think. Is there a unique joiner in each household
+lapply(financial_files, \(x) "idauniq" %in% names(fread(x)))
+
+
+all_finance <- lapply(financial_files, ReadAndSetWave) %>%
+    rbindlist(fill = TRUE, use.names = TRUE)
+
+all_ifs <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "wave",
+    full.names = TRUE
+) %>%
+    grep("ifs", ., value = TRUE) %>%
+    lapply(ReadAndSetWave) %>%
+    rbindlist(fill = TRUE, use.names = TRUE)
+
+
+all_pension <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "wave",
+    full.names = TRUE
+) %>%
+    grep("pen", ., value = TRUE) %>%
+    lapply(ReadAndSetWave) %>%
+    rbindlist(fill = TRUE, use.names = TRUE)
+
+
+fwrite(all_pension, "../../data/ELSA/elsa_to_use/elsa_pensions.csv")
+fwrite(all_ifs, "../../data/ELSA/elsa_to_use/elsa_pensions.csv")
+fwrite(all_finance, "../../data/ELSA/elsa_to_use/elsa_pensions.csv")
+
+head(all_finance$idauniq)
+
+full_join(all_pension, all_ifs)
+
+sum(all_pension[wave == "wave_1"]$idauniq %in%
[…]

### 2023-07-07 Texfiles/Draft1.tex
Create first draft

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{enumerate}
\setlength{\parindent}{0pt}
\usepackage[parfill]{parskip}
\date{Lent 2023}
\title{Dissertation draft}
\author{Tobias Leigh-Wood}
\usepackage[letterpaper, margin=1.1in]{geometry}
\linespread{1.1}

\begin{document}
\maketitle

Standard economic theory suggests that retirement annuities should be highly prized by individuals as a way to
insure against the risk of late death. However, in developed countries rates of annuitization are far below the
levels that theory predicts. In this paper I test two competing hypothesises for the annuity problem: bequests and
pessimistic life expectancy. Under the coalition government in the UK the law regarding the use of private defined
contribution pensions changed. Individuals were no longer forced to annuitise their pension pots and could access
them in a variety of ways such as lump sum withdrawrels or income drawdown and subsequently the number of annuities
sold in the UK dropped precipitously.

Depending on the reason for the lack of annuitization in the UK, the consumption response of retirees to the pension
reform will differ. If individuals do not annuitise because of pessimistic life expectancy I will show that their
consumption should increase. If, on the other hand, individuals do not annuitise because of a bequest motive, consumption
should not change much as a result of the reform. I will solve lifecycle models for both of these cases and simulate
consumption decisions with, and without, forced annuitisation. I will then use a variety of empirical models to measure
the consumption change in early retirement that resulted from the policy reform. The size and magnitude of this change
will be indicative of the mechanism causing the annuitization problem.

The number of individuals over the age of 65 is expected to increase dramatically in the 21st century.

\subsection{Rough plan}
\begin{itemize}
      \item Intro
      \item Lit review
      \item Models
      \item Empirical
            \begin{enumerate}
                  \item Diff in diff
                  \item RDD
                  \item Matching?
                  \item can I use anything I learnt in panel?
                        In some sense the decision to annuitise is a discrete choice problem so I could use something from there.
            \end{enumerate}
      \item Conclusion
\end{itemize}

\bibliographystyle{chicago}
\bibliography{references}
\end{document}

### 2023-07-07 Texfiles/Notes.tex
recent changes to code
Create first draft

--- 2023-02-27 Texfiles/Notes.tex
@@
 \usepackage{enumerate}
+\usepackage{hyperref}
+
+\hypersetup{
+    colorlinks,
+    citecolor=black,
+    filecolor=black,
+    linkcolor=black,
+    urlcolor=black
+}
 \setlength{\parindent}{0pt}
 \usepackage[parfill]{parskip}
-\date{Lent 2023}
-\title{Dissertation notes}
+\date{June 2023}
+\title{Dissertation reading notes}
 \author{Tobias Leigh-Wood}
@@
 \linespread{1.1}
-
 \begin{document}
 \maketitle
+\tableofcontents
+\newpage
 
-\section*{To Do List}
+\section{Comparing Retirement Wealth Trajectories on Both Sides of the Pond}
+\cite{Blundell_et_al_2016_fs}
+
+\subsection{Introduction}
 \begin{itemize}
-    \item What actually happened?
-          Annuity provision peaked in 2012. Prior to the March 2011 budget individuals had to annuitise their pension pot.
-          After 2011 they only had to annuitise up to a minimum income requirement (MIR) of £20,000. This minimum income kept most people using annuities.
+      \item Compare wealth decumulation in retirement in the US and the UK using HRS and ELSA data.
+            ELSA was modelled on HRS so they are comparable. Both follow elderly people and have similar questions
+            and methodologies. Data from 2002 to 2010.
 
-          In March 2014 there was a surprise announcement that the MIR would be scrapped and individuals would be able to 'draw down' their pension pot instead.
-          This meant the compulsory market for pension annuities basically stopped.
+      \item Median wealth at ages 70-74 is similar. But UK individuals tend to have slightly higher income.
+            They also have lower medical expenditures because of the NHS. But UK households unexpectedly decumulate assets more slowly.
+            This is partly due to the rise in house prices in the UK during that period.
 
-    \item When did the changes take place?
-    \item When were the changes announced?
+\end{itemize}
+\subsection{Empirical facts on asset decumulation}
+\begin{itemize}
+      \item Wealth in the US has more of a right-skew at the start of retirement. Medians are similar.
+      \item Primary housing wealth accounts for 2/3rds of UK retirement wealth but only 1/3 of US retirement wealth.
+      \item Life expectancies are pretty similar. US household income is higher than the UKs but it drops with age
+            whereas UK income rises with age -- this could be down to different indexation practices between the UK and US. UK pensions
+            are often indexed above prices.
+
+      \item Median net wealth grew in the UK but fell in the US. Even once we strip out housing wealth assets are decumulated faster
+            in America. Non-housing assets are much larger in the US so this could be because in the UK people want to keep enough assets to
+            finance funeral costs. But this is still surprising.
+
+\end{itemize}
+\subsection{Potential explanations of cross country differences}
+\begin{itemize}
+
+      \item Check whether the life-cycle model explains these differences.
+      \item \textbf{Bequest motives}
+            \begin{itemize}
+                  \item Empirical literature in the US points to this being a channel. Less of a literature in the UK but
+                        surveys have shown that people want to pass on wealth.
             \end{itemize}
 
+      \item \textbf{Precautionary savings}
+            \begin{itemize}
+                  \item \cite{de_nardi_et_al_2010} have shown that pre-cautionary saving for medical expenses explain alot of the slow decumulation
+                        of assets in America.
+
+                  \item There is not comparable data for late life medical expenses in the UK. And although the NHS covers some medical
+                        expenses they do not cover care home costs which can be a large proportion of medical expenses.
+                  \item \textit{Note to self: } \cite{french_jones_jae_2004} estimate the stochastic process for medical expenses in retirement.
+
+                  \item But because there is not enough data we cannot compare the US and the UK.
+            \end{itemize}
+
+      \item \textbf{Housing}
+            \begin{itemize}
+                  \item \cite{Nakajima_Telyukova_ier_2020} make a life-cycle model that takes into account the extra utility from housing
+                        and the fact that it is difficult to borrow against. Without housing assets they find that assets would up to 55\% lower.
+                        Combination of utility from housing and housing being difficult to get rid of.
+
+                  \item Later work has shown that housing is not a pre-cautionary saving tool.
+                  \item In the US there is a government administered equity release scheme, the UK has several private operators. In 2011 2\%
+                        of homeowners had some reverse mortgage.
+
+                  \item Nature of housing as an asset impacts decumulation. Empirical work has not been able to distinguish mechanisms from each
+                        other -- consumption value, transaction costs or risk-return matrix.
+
+            \end{itemize}
+
+\end{itemize}
+
+\subsection{Conclusions}
+\begin{itemize}
+      \item Housing can explain difference between the UK and US. But so far there is not a satisfactory life-cycle model
+            that can account for the difference in asset decumulation.
+\end{itemize}
+
+\section{Savings After Retirement: A Survey}
+\cite{de_nardi_et_al_are_2016}
+\subsection{Introduction}
+\begin{itemize}
+      \item 1/3rd of US wealth is held by those over 65. Standard life-cycle models with known time of death cannot account for the
+            slow decumulation of these assets, particularly for those who are wealthy.
+
+      \item Two groups of mechanisms that could explain this. 1. \textbf{Risks} -- uncertainty over time of death and medical expenses.
+            2. \textbf{Bequests}.
+
+      \item Both of these have similar implications for saving in old age.
+\end{itemize}
+\subsection{Facts about retired households}
+\begin{itemize}
+      \item Three main sources of retirement wealth -- pension, social security and private savings.
+      \item To understand how assets are decumulated across retirement \citeauthor*{de_nardi_et_al_are_2016} use
+            data from AHEAD and HRS. They split up the data by postretirement permanent income (PI). This is the average nonasset,
+            non means tested social security income over retirement.
+
+      \item Mortality bias -- assets in the unbalanced panel are 50\% of assets in the balanced panel at the start of the
+            period.
+
+      \item Show an interesting graph of decomposed medical expenses by age and by source of spending. Social spending
+            programs make up the difference in out-of-pocked medical spending as opposed to differences in the use of medical services.
+
+      \item Medical spending has a massive right skew. Top 5\% of the distribution account for over a third of the total spend.
+
+      \item Wealthy and the rich have significantly lower mortality and live longer.
+      \item The literature is very mixed on bequests -- some papers find that lots of people leave bequests whereas others find no bequests
+            for a majority of the population and then large bequests for the wealthiest.
+
+\end{itemize}
+
+\section{HOME EQUITY IN RETIREMENT}
+\cite{Nakajima_Telyukova_ier_2020}
+\subsection{Overview}
+\begin{itemize}
+      \item Look at the role of housing equity in the retirement saving problem.
+      \item Find, contrary to \cite{de_nardi_et_al_2010}, that medical expenses do not explain the decumulation problem.
+      \item This is the closest in content to the paper that I want to write.
+\end{itemize}
+
+\section{THE RETIREMENT-SAVINGS PUZZLE REVIEWED: THE ROLE OF HOUSING AND BEQUESTS}
+\cite{Suari_Andreu_et_al_jes_2019}
+
+\subsection{Introduction}
+\begin{itemize}
+      \item Three categories of explanation for the retirement saving problem (RSP) -- uncertain mortality, bequests and precautionary saving.
+      \item Parallel literature looking at housing equity in retirement (HER), find that home-owners are unlikely to draw down
+            housing equity because they like living in their home, and partial drawdowns are costly.
+
+      \item \cite{Nakajima_Telyukova_ier_2020} augment the model from \cite{de_nardi_et_al_2010}. The NT model treats housing
+            as a different asset. They find that home ownership interacts with the bequest motive.
+\end{itemize}
+\subsection{The retirement-savings problem}
+\begin{itemize}
+      \item Three strands to the RSP literature. The first looks at the impact of lifetime uncertainty on retirement savings.
+            Initiated by \cite{yaari_65} with recent contributions by \cite{de_nardi_et_al_2010}. The risk of outliving their wealth leads
+            individuals to decumulate more slowly than they would if time of death was certain.
+
+      \item The second set of literature studies bequest motives. Altruistic, egoistic, and strategic motives.
+            There is also a small literature on inter-vivo.
+
+      \item More recently, the impact that uncertain medical expenses can have on retirement savings has been studied.
+            \cite{de_nardi_et_al_2010} show that this can explain a large part of the savings of singles.
+\end{itemize}
+
+\subsection{Home equity in retirement}
+
+\begin{itemize}
+      \item Housing is special because of its role as a consumption and investment good, and because it is infrequently traded due to
+            high transaction costs.
+
+      \item There is at most a modest decline in housing wealth among older people. \cite{venti_wise_nber_2004} have shown that this is
+            caused by the subset of households that experience widowhood or a nursing home event. They conclude that
+            housing wealth is generally not used for consumption; this explains the low demand for reverse mortgages and implies we should
+            not count housing wealth as retirement wealth.
+
+      \item Less moving in the UK than the US. Higher taxes and lower variation in climate and institutional choice.
+            Confirmed by \cite{Blundell_et_al_2016_fs}.
+
+      \item Few international studies. A couple of dutch papers have shown also that older people tend not to move or use reverse
+            mortgages.
+
+\end{itemize}
+
+\subsection{Models of retirement saving with housing}
+\begin{itemize}
+      \item First key contribution comes from \cite{yang_red_2009}. She builds a model that explicity models housing and its features.
+            Owner occupied housing counts contributes to utility and can be used as collateral.
+
+      \item Households can either be owners or renters. They can become renters exogenously or endogenously. She matches the model
+            to parameter estimates from the literature and uses it to explain features of the life-cycle pattern of housing and non-housing
+            consumption.
+
+      \item She finds that the transaction cost associated with moving makes older households much less likely to sell their house.
+            If transaction costs are 0 then households would sell houses at a faster rate.
+
+      \item But the model leaves out key things features that are in the RSP literature; death uncertainty, medical expense uncertainty and
+            bequests which could also explain the pattern.
+
+      \item \cite{Nakajima_Telyukova_ier_2020} adds housing consumption to the model from \cite{de_nardi_et_al_2010}.
+            Quite a complicated model but I don't see why I could not copy it. \citeauthor*{Nakajima_Telyukova_ier_2020} have a single health
+            shock as opposed to a medical expense shock.
+
+      \item NT find that bequest motives and housing are the big drivers of retirement saving. But \cite{de_nardi_et_al_2010} only use
+            singles and they do not match the housing distribution in their method of simulated moments. (Is this process, in general, robust
+            to choosing different moments?)
+
+\end{itemize}
+
+\subsubsection{Bequest motives}
+\begin{itemize}
+      \item Authors consider three different motivations for bequests.
+      \item \textbf{Altruistic} NT model bequest motvie as an egoistic motive. Households just want to have positive
+            wealth at death no matter who they are passing on the bequests to. An altruistic bequester models the utility of the heirs and
+            seeks to maximise this. With concave utility, bequests are decreasing in the life time income of the heirs because their marginal
+            utility from money is lower.
+
+      \item \textbf{Strategic}. Bequests should be seen in the context of inter-generational transfers where parents care about
+            what their children do for them in old age. Re-write utility with a parameter which increases parents utility but decreases the heirs.
+            Some evidence for this channel but I think it is a bit shaky. There is also some evidence that inter-vivo transfers increase if someone
+            is a caregiver.
+
+\end{itemize}
+
+\section{Personal notes}
+\subsection{ELSA notes}
+
+\subsection{\textbf{06-07}}
+\begin{itemize}
+      \item Do I have death data? check UKDS.
+            Harmonized data does have year of death but only for very few
+            people. It doesn't have it for the majority of exits.
+
+      \item I can change pension for individuals and asset levels and subjective life expectancy.
+      \item
+
+\end{itemize}
+
+\subsection{\textbf{04-07}}
+\begin{itemize}
+      \item How to make a consumption floor? Is it a question of giving income?
+      \item Also check
+      \item should think about the impact of housing on the effects of the change
+            -- will their be heterogeneity
+
+      \item To do
+            \begin{enumerate}
+                  \item Make julia module
+                  \item add gender into life cycle
+                  \item get subjective life expectancies -- have got the code from
+                        david and cormac but I think they have birth dates and death dates which
+                        arent included in the normal elsa data.
+                  \item generate correct rates of annutisation
+            \end{enumerate}
+\end{itemize}
+\subsection{\textbf{30-06}}
+\begin{itemize}
+
+      \item Continue working on retirement lifecycle code.
+            Once we have a working version of a simple model we can then attempt
+            to add the extensions.
+
+      \item I think I should also check the \cite{odea_sturrock_rest_2023} paper,
+            I think they just use a simple model without bequests.
+
+      \item Simple model is working. Now we want to add medical risk?
+      \item First clean up code, put it into a function where we can change
+            everything easily.
+
+      \item Plot consumption - done. Also sorted out some problems with the code.
+      \item Then I need to augment with annuity products -
+            I should be able to do this. At age 65 they can make a decision to buy an annuity or not.
+            Check expected utility from consumption path.
+            Then do the same with bequests.
+            Shall I asked ODea and Sturrock for codes?
+
+      \item Lockwood makes a discrete space that is exponential.
+            Seems a bit strange tbh. I'd like to look at Odea's code.
+            How to draft the email?
+
+      \item So the grid of assets is created from the consumption choices?
+            -- Yes I think that is how it works and then consumption should always be in the grid.
+
+      \item Also should compare to income drawdown. Most pensioners have income
+            below the taxable amount, and therefore the optimal amount of drawdown
+            should not be impacted by taxes.
+      \item And state pension?
+      \item I guess in it's current
+
+\end{itemize}
+
+\subsection{\textbf{28-06}}
+\begin{itemize}
+      \item Not sure what I should work ok. I could do more data work or work on the simulation part.
+            I think simulation part. Data is probably not far from being done tbh.
+
+      \item Aim is to get a small working version just with risky retirement life expectancy.
+      \item
+\end{itemize}
+
+\subsection{\textbf{27-06}}
+\begin{itemize}
+      \item Pretty bad week of work. Understand the data a bit better.
+            But have made little progress elsewhere. I want to work on my Julia code
+            to replicate the analysis in \cite{odea_sturrock_rest_2023} and \cite{lockwood_red_2012}.
+
+      \item Lets have a look at the matlab code from Lockwood. Is actually kinda hard but I think doable.
+            That would be the bit I enjoy the most.
+            Why do people keep on coming up with new life-cycle models though.
+      \item
+      \item What are the implications for the reason no annuities being the returns in the stockmarket.
+            - could someone look at actual returns that retirees get. Presumably many switch to portfolios with bonds.
+
+      \item I think the implication would be increased
+      \item \cite{inkman_et_al_rfs_2011} -- reading this now. Say that there is an annuity question in the
+            Income and Assets section of the ELSA data set. "IAAIM" this variable. Makes sense now how they got
+            such a small amount. No way to tell whether people are using income drawdown products or not.
+            Guess we could see if non-annuity income increases after.  "IAPPEN" then this variable is the amount of
+            pension income which includes annuity income. Not sure whether we should see this go up or down around 2013.
+
+\end{itemize}
+
+\subsection{\textbf{19-06}}
+\begin{itemize}
+      \item Get familiar with the pension data and the different pension variables. \cite{odea_sturrock_rest_2023} -- in this paper they calculate money's
+            worth using subjective mortality risk as opposed to normal life tables. No simulations, just
+      \item
+\end{itemize}
+
+\subsection{\textbf{18-06}}
+\begin{itemize}
+      \item Dear Eric, thanks for your advice -- I'll stay clear of trying to estimate a savings model. That paper looks
+            very comprehensive so I am not sure what I could add to the topic in four weeks.
+
+            I've come back to the ideas around the reform to annuities and how it fits with the implications that different models
+            of retirement saving would predict. Models that have been estimated, such as the Lockwood paper, give quite sharp
+            predictions about what should happen to savings and consumption in retirement when the annuity constraint is lifted.
+
+            My analysis could take several of the most recent models in the lifecycle literature that explain retirement saving
+            and estimate the expected changes in consumption of retirees in the first few years of retirement conditional on not
+            being forced to annuitise pension wealth
+
+      \item If Lockwood paper is correct and bequests are the driver of retirement savings then we should see a decrease in savings
+            from income. If individuals dont buy annuities because they aren't priced fairly for them then we should see a rise in consumption
+            when annuities are no longer compulsory. One section of the paper could fit these two simple models of non-annuitisation using
+            the drop in demand for annuities.
+
+      \item ok found quite an interesting new paper \cite{inkman_et_al_rfs_2011}. Look
+            at annuity demand and characteristics that are associated with it. Also,
+            estimate a life-cycle model that can explain the portfolio choices
+            of individuals in the data.
+
+      \item Mmm could I also compare those with a DB pension and those with a DC pension.
+            Nothing changed for DB pension pots. Could I do a diff in diff using the DB pension group
+            as the control? I think that would work.
+
+      \item Have a look at the pension data today. Can we distinguish the point at which individuals retire?
+            When they convert pension to annuity income? Or when they keep on running it down as normal?
+
+      \item I also need to let Eric know and just run this past him as well.
+
+      \item Does consumption increase with bequest motive as well? Only if people under forced bequests were
+            saving more at the start of retirement to regain what they were forced to annuitise.
+
+\end{itemize}
+
+\subsection{\textbf{16-06}}
+\begin{itemize}
+      \item Have a better understanding of the data.
+
+      \item But I still need a better idea. Read Lee Lockwood RED paper.
+      \item Also figure out pre-retirement pension DC pots. i.e. what they can spend
+            on annuities.
+
+      \item I think a basic easier question would be what impact has the change had on consumption profiles.
+      \item Could I estimate a small model somehow?
+
+      \item New idea is to estimate reduced form estimates of the impact compulsory annuitisation
+            had on wealth.
+
+\end{itemize}
+
+\subsection{\textbf{13-06}}
+
+\begin{itemize}
+      \item There is data on consumption and spending in ELSA. The technical reports have good information on what is available
+            in the data sets.
+
+      \item Questions: Is health expenditure data available? There is some social care cost data but there is not any medical expense data.
+            Or I cannot find it at least.
+
+      \item \textit{Questions/to do:}
+            \begin{enumerate}
+                  \item Minimum distance estimation and method of simulated moments in Julia? Solving the model in Julia?
+                        \begin{enumerate}
+                              \item Minimum distance theory vs method of simulated moments -- check Hansen.
+                              \item Perhaps get a small working example going and then play around with it. There are some on quant econ that I can use \hyperref{text}
+                                    https://github.com/fediskhakov/CompEcon -- this has some nice python notebooks that I could try and replicate.
+                                    I do not know how to start this though? Should I just start trying to bang away at it?
+                                    Do I even understand how the method works? Do I solve for optimal policy functions for a given
+                              \item Good textbook treatments of SMM are found in Adda and Cooper (2003, pp. 87-100) and Davidson and MacKinnon (2004, pp. 383-394).
+                        \end{enumerate}
+                  \item
+                  \item Medical expense data?
+                  \item Read \cite{banks_et_al_aej_2019} -- this looks at differences in consumption across retirement in the US and the UK.
+                  \item Read \cite{gourinchas_parker_ec_2003} -- big contribution to the field it seems.
+                  \item \textbf{Scrap this idea. Eric advised against it.}
+            \end{enumerate}
+
+      \item I am back to the annuity idea. I could document changes in annuities in ELSA and FES and also
+            see if consumption changed because of this. I like the idea of fitting a bequests model to the change in annuitisation
+            rates like Lee Lockwood did but I do not know if that is possible. What moments would I match it to?
+
+      \item I have just broken my version of R annoyingly.
+            I want to have a separate set of packages downloaded for my thesis.
+            Done and it is working again nicely now.
+
+\end{itemize}
+
+\bibliographystyle{chicago}
+\bibliography{references}
 \end{document}

### 2023-07-12 Texfiles/Draft1.tex
spelling
Draft and more notes

@@
 Standard economic theory suggests that retirement annuities should be highly prized by individuals as a way to
-insure against the risk of late death. However, in developed countries rates of annuitization are far below the
-levels that theory predicts. In this paper I test two competing hypothesises for the annuity problem: bequests and
+insure against the risk of late death \cite{yaari_65}. However, in developed countries rates of annuitization are far below the
+levels that theory predicts. In this paper I test two competing hypotheses for the annuity problem: bequests and
 pessimistic life expectancy. Under the coalition government in the UK the law regarding the use of private defined
 contribution pensions changed. Individuals were no longer forced to annuitise their pension pots and could access
-them in a variety of ways such as lump sum withdrawrels or income drawdown and subsequently the number of annuities
+them in a variety of ways such as lump sum withdrawals or income drawdown and subsequently the number of annuities
 sold in the UK dropped precipitously.
@@
 should not change much as a result of the reform. I will solve lifecycle models for both of these cases and simulate
-consumption decisions with, and without, forced annuitisation. I will then use a variety of empirical models to measure
+consumption decisions with, and without, forced annuitization. I will then use a variety of empirical models to measure
 the consumption change in early retirement that resulted from the policy reform. The size and magnitude of this change
@@
 
-The number of individuals over the age of 65 is expected to increase dramatically in the 21st century.
+The importance of retirement policy to individuals in the UK is growing. The number of individuals of
+pensionable age is expected to grow from 11.9 million in 2020 to 15.2 million in
+2045 according to the latest ONS statistics and for every 1000 people of working age there will be 341 of pensionable
+age in 20145 compared to 280 in 2020 \cite{ons_population_predictions_2020}. The increase in absolute and relative
+numbers of elderly makes retirement policy more important. Moreover, private, defined contribution (DC), pensions are
+becoming increasingly common and are predicted to grow as current cohorts age \cite{cribb_karjalainen_ifs_2023}.
+Therefore, policies regarding how private pensions can be accessed will have a larger impact on overall welfare for
+retirees.
+
+The so called "pensions freedom act" received Royal Asset in December 2014 marked the end of a series
+of pension reforms carried out by the coalition government between 2010 and 2015. The reform was announced
+in the Spring budget and made it possible to withdraw money from a private pension pot subject to the
+marginal rate of income tax that an individual faced. In the June 2010 budget the government made a first
+reform to the annuitization rules, creating an minimum income requirement above which individuals would not need to
+annuitise more \cite{finance_act_hmt_2011}. However, this was set at £20,000 and therefore few individuals
+were eligible. The minimum income requirement was scrapped in the 2014 bill finally eliminating the
+compulsory annuities market. The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
+Using data from the Association of British Insurers they show that annuity demand dropped by 75\% from
+its maximum.
+
+\subsection{Literature review}
+My paper draws on three main strands of literature. The annuity problem, the retirement saving
+problem and lifecycle models. \textbf{this intro is rubbish}
+
+\cite{yaari_65} was the first to show that under standard assumptions we would expect individuals to
+annuitise all of their wealth at retirement to insure against the risk of long life. Since then there
+has been much literature discussing possible reasons that people do not annuitise. \cite{finkelstein_porteba_2002}
+and \cite{finkelstein_porteba_2004} find evidence of adverse selection, thereby making the 'money's worth'
+of annuities lower for the general population as opposed to the population of annuitants. However, they
+also find that theory would still predict annuitization.
+
+\cite{friedman_warshawsky_qje_1990} show that annuitization decisions can be fully explained by a mixture of
+bequest motives and actuarily unfair annuities. They solve an augmented life-cycle model with a range of
+parameters on how severe the rate of return is on the annuity versus market rates. For plausible values
+they find that individuals would optimally not annuitise much wealth. Similarly to \cite{finkelstein_porteba_2004},
+\cite{friedman_warshawsky_chicago_1988} show that there is a significant difference between the life expectancy of
+annuitants and the general population in the American annuity market but this cannot fully explain the annuitization
+problem. Only when bequest motives are added to the model can annuitization rates be rationalized.
+
+\cite{lockwood_red_2012} builds on this and shows that a realistic bequest motive in lifecycle simulations achieves
+realistic annuitization rates. He solves a simple lifecycle model with bequest motives taken from several recent papers
+in the literature. The bequest motives he picks therefore match other important aspects of the lifecycle model such
+as how much individuals actually bequest and how rich individuals are when they bequest.
+
+\cite{mitchell_et_al_1999} - this is empirical
+\cite{friedman_warshawsky_qje_1990} - this is theory
+
+\cite{friedman_warshawsky_chicago_1988} - this is more theory. Probably good to have read.
+
+\cite{lockwood_aer_2018}
+
+. \textit{Continue
+      chat about how elderly population is growing and private pension pots are growing in size, could be worth referencing
+      latest census data so we seem super up to date, look at IFS data about the growth of DC pensions as well. }
+\textbf{I will do this this afternoon}
 
@@
       \item Intro
+            \begin{enumerate}
+                  \item I think Eric French wrote something about population that I could use in then intro to say why it is important.
+                  \item "Latest data from HM Revenue Customs (published in April) showed more than £45bn has been taken from pots since 2015."
+                        https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
+
+                  \item Need annuity theory references -- Yarri?
+            \end{enumerate}
       \item Lit review

### 2023-07-12 Texfiles/Notes.tex
Draft and more notes

@@
 
+\subsection{Government policy timeline}
+\begin{itemize}
+      \item Freedom and choice in pensions announced in March \cite{pen_freedoms_hmt_2014}. Gains royal
+            assent in December becoming the Taxation of Pensions Act. \cite{tax_pensions_hmg_2014}
+
+      \item What other reforms were there? 2010 budget annuitisation
+\end{itemize}
+
+\section{The Cost of Annuities: Implications for Saving Behavior and Bequests}
+\cite{friedman_warshawsky_qje_1990}
+
+\begin{itemize}
+      \item Interesting point re annuities being low because of state pension.
+            I could test consumption response to forced annuitization if that is the
+            case as well. I think it would just mean increased spending?
+
+      \item Make a good point that other explanations for the AP are to do with unfairly priced
+            annuities. This could be subjectively or not and it would have the same impact (or
+            a mixture of the two). So I don't think there is massive need in estimating subjective
+            survival curves.
+
+      \item I think there has been some work by cannon and tokes on estimating this in the
+            early 2010s so I am not sure whether there is point me trying to do this again using the
+            CMI annuitants tables.
+
+      \item Also test model with both bequests and unfair annuities. This seems most obvious tbh.
+            Maybe whole idea is rubbish because of this.
+
+      \item Could I compare how a mixture of load factors vs a mixture of subjective death probs
+            vs a mixture of adverse selection interacts. Disentangling the extent to which each causes
+            people not to annuitise would be difficult though. But can use evidence from ELSA and CMI for
+            subjective and adverse selection amounts respectively.
+
+      \item \begin{equation*}
+                  \mathbb{E}(U) = \sum_{0}^{w - x - 1}[p_{t}U_{t}(c_{t}) + p_{t}q_{t}V_{t+1}(G_{t+1})]
+            \end{equation*}
+            Where $V_{t+1}(G_{t+1})$ is the utility from bequests next period and $q_{t}$ is the probability of death.
+
+      \item Does some simulations showing how bequests and acturarial unfairness interact with each other.
+            This type of table could be a good thing to include in my paper. Then a section showing implied consumption
+            changes and then the data work. I think I am kinda close to being able to make that anyway.
+
+      \item What is expected yield differential? D
+      \item \textit{" but investigating it seriously would require data on the
+                  degree of actuarial unfairness in the pricing of annuities not just for
+                  individuals but for double or even multiple lives."} This is interesting.
+            Could I compare annuitization decisions of couples and singles? Singles
+            should be more likely to buy annuities if this is true.
+
+      \item Couples should also have a larger consumption response. Can we solve
+            two person models? Probably? But I think joint annuities would face
+            more than double the problem with actuarially unfair annuities worse --
+            impact would be multiplied and not summed.
+
+      \item Have not seen this before? Could be interesting. I think Eric solved a life-cycle model
+            with a couple in so I could try and look at that paper. (No I think it was another paper from S130)
+
+\end{itemize}
+
+\section{Annuity Prices and Saving Behavior in the United States}
+\begin{itemize}
+      \item Solve life cycle models again. The variable of interest they use is the
+            loading factor.
+
+      \item Need bequest motives to get annuity rates that are observed.
+      \item "See \cite{friedman_warshawsky_nber_1985} for a brief discussion of several
+            of these other possible explanations." that is for other explanations for
+            lack of annuitization. Had a look at this and it wasn't super useful. Just
+            seemed like a first draft of the paper.
+
+\end{itemize}
+
+\section{Bequest motives and the annuity puzzle}
+\begin{itemize}
+      \item Literature (Abel 2004?? Find when landed) claims that bequest motives
+            can only explain part of the AP. But Lockwood shows that it can explain why people
+            annuitise no wealth given loading factors (prices etc) of annuities.
+
+      \item  (Davidoff et al., 2005??) is another paper on annuities I should read.
+
+\end{itemize}
+
 \section{Personal notes}
+
 \subsection{ELSA notes}
+
+\subsection{\textbf{12-07}}
+\begin{itemize}
+
+      \item Got CMI access.
+\end{itemize}
+
+\subsection{\textbf{11-07}}
+\begin{itemize}
+      \item Good idea about couples. Good idea about table showing \% of annuititants
+            with param values and then also consumption changes.
+
+      \item How to split consumption? Guess have more grids? -- Don't need to since only one
+            continuation value and there is no heterogeneity between partners.
+      \item How do behaviourel models change the DP problem. Could present bias
+            be part of the explanation.
+
+      \item Also introduction of lifetime cap in care costs could be interesting
+            hypothetical policy issue to mention.
+
+      \item
+\end{itemize}
+
+\subsection{\textbf{08-07}}
+\begin{itemize}
+      \item Trying to understand the weibull code from \cite{odea_sturrock_rest_2023}. Non-linear
+            least squares
+      \item Big docs have questions in that we can use.
+\end{itemize}
 

### 2023-07-12 Julia/4_retirement_lifecycle.jl

@@ -22,6 +22,51 @@ cum_death_probs = CSV.File("Julia/death_probs.csv") |>
                   vec
 # Increase every prob of death by 0.01
 
+# Mortality data from the ONS
+const base_year = 2005
+const base_age = 50
+
+male_data = CSV.File("../data/ONS/male_transition_probs.csv") |>
+            Tables.matrix
+
+female_data = CSV.File("../data/ONS/female_transition_probs.csv") |>
+              Tables.matrix
+# Each column is the individuals age in 2010. 
+# each row is a year after that, transition probs from 
+# age to age + 1
+
+# these people are aged 50 in 2005
+plot([male_data[:, 21] male_data[:, 1]])
+
+# So then if we want to pdf of probs in a given age 
+# ONS life tables
+function death_probabilities(data, year=2012, age=65)
+    # return probs from this age and year onwards
+    col_index = year - (base_year - 1) # selects the correct column
+    start_row_index = age - base_age # first row of interest
+
+    death_probs = data[start_row_index:51, col_index]
+
+    return death_probs
+end
+
+# If I dont update 
+death_probabilities(male_data, 2010, 63)
+death_probabilities(female_data, 2010, 63)
+# these finish at age 100. and start at the age specified
+
+
+# ok so now I can set gender and use realistic lifetables
+# I could have a look at annuitant tables as well and then use a 
+# scaling factor on the implied fair price to calucalte the price
+# that normal people face. 
+
+# can set wealth from elsa data as well 
+
+
+
+
+
 function constrained_add(a, b)
     # This is just so I can play around with subjective 
     # probabilities and get strictly increasing probs. 

### 2023-07-12 R/12__ELSA_harmonised.R

@@ -77,11 +77,12 @@ harm_long <- melt(
         "age_at_interview", "in_wave",
         "date_month", "date_year",
         "total_monthly_consumption"
-    )
+    ),
+    variable.name = "wave"
 )
 
 fwrite(harm_long, "../../data/ELSA/elsa_to_use/harmonised_data_long.csv")
-
+stop()
 harm_long[, unique(variable)] # what is this variable?
 
 harm_long[, .(number_of_waves = sum(in_wave)), by = .(idauniq)][order(number_of_waves)]

### 2023-07-12 R/13__ELSA_eol.R

@@ -4,106 +4,190 @@ library(stringr)
 
 source("__ELSA_functions.R")
 
-dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-    pattern = "core",
-    full.names = TRUE
+
+check_files_for_variable_name <- function(var_name) {
+    # returns the files which contain the variable name we are looking for
+    # Just helpful for seeing where variables originally came from
+    files_with_var_name <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+        pattern = "",
+        full.names = TRUE
+    ) %>%
+        lapply(\(filename) names(fread(filename, nrows = 1))) %>%
+        lapply(function(col_names) {
+            grep(var_name,
+                col_names,
+                value = T,
+                ignore.case = T
+            )
+        }) %>%
+        sapply(length) %>%
+        sapply(\(x) x != 0)
+
+    dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+        pattern = "",
+        full.names = TRUE
+    )[files_with_var_name]
+}
+
+check_files_for_variable_name("eida")
+# year of death
+check_files_for_variable_name("exlo")
+# life expec
+
+search_names(
+    fread(
+        "../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_6_elsa_data_v2.tab",
+        nrows = 1
+    ),
+    "exlo",
+    ignore.case = TRUE
+)
+# "ExLo80" "ExLo90"
+
+expectation_cols <- c(
+    "idauniq",
+    "ExLo80",
+    "ExLo90"
+)
+
+## get expectation data
+files_with_exp <- check_files_for_variable_name("exlo")
+
+data_with_exp <- lapply(files_with_exp, ReadAndSetWave) %>%
+    lapply(\(data) setnames(data, str_to_lower)) %>%
+    rbindlist(use.names = T, fill = T) %>%
+    select(idauniq, file_wave, exlo80, exlo90) %>%
+    mutate(wave = as.numeric(str_extract(file_wave, "\\d{1}"))) # make wave numeric so that we can join later
+
+
+# ok this is the data.
+# need to do some cleaning on it
+# get rid of negative values and set values of 0 and 1
+# to be just above and below respectively
+
+# use scaled life expectancy as the third point of data.
+# I think we need rough estimates to plug into the life cycle model.
+
[…]

### 2023-07-12 R/14__ONS_lifeexpec.R

@@ -1,0 +1,243 @@
+library(data.table)
+library(dplyr)
+library(stringr)
+library(readxl)
+
+source("__ELSA_functions.R")
+
+# I want to make survival curves from ONS
+
+# Also could make survival curves from annutitant tables that
+# the institute and faculty of actuaries makes produces annuity tables.
+# one in 2008 vs the latest could make for an interesting comparison
+
+# How do we make cumulative death probabilities?
+### qx is the mortality rate between age x and (x +1),
+# that is the probability that a person aged x exact
+# will die before reaching age (x +1).
+
+
+# sheets_to_get <- excel_sheets("../../data/ONS/nationallifetables3yruk.xlsx") %>%
+#     grep("\\d{4}", ., value = TRUE)
+
+# what form shall we put it in?
+# json?
+# or long csv with year, age and gender?
+
+# ons_lifeexp <- lapply(sheets_to_get, function(sheet_name) {
+#     suppressMessages({
+#         data <- read_excel("../../data/ONS/nationallifetables3yruk.xlsx",
+#             sheet = sheet_name,
+#             skip = 5
+#         ) %>%
+#             select(-7) %>% # drop gap in the sheet between m & f lifexp
+#             setnames(col_names) %>%
+#             mutate(years = sheet_name)
+#     })
+#     return(data)
+# })
+
+
+# all_data <- rbindlist(ons_lifeexp)
+
+# qx is the prob of reaching the next age, conditional on being that age.
+# so if we have a constant birth year
+# we can trance an age through and get all of their conditional life expectancies.
+
+
+# we really want predictions though? not real life expectancy
+# these are mortality predictions
+
+# predictions_12 <- read_excel("../../data/ONS/wukprincipal12qxr.xls")
+# excel_sheets("../../data/ONS/wukprincipal12qxr.xls")
+
+# I think i want period lx and then we trace up the diagonal
+
+
+# return_prob_for_age_year <- function(pred, year, age_to_get) {
+#     # This function takes the data and selects the transition
+#     # prob from an age in a year to the next age
+#     col_name <- paste0("period_", year)
+
+#     pred[age == age_to_get][, COL, env = list(COL = col_name)]
+# }
+
+# return_all_probs <- function(pred, start_year, start_age) {
+#     # This finds all the transition probs
[…]

### 2023-07-12 R/__ELSA_functions.R

@@ -1,12 +1,23 @@
-
-ReadAndSetWave <- function(filename) {
+ReadAndSetWave <- function(
+    filename,
+    select_cols = "all") {
     # Read a tab file and set a wave number.
 
     # Extract wave number from filename
     WAVE_num <- str_extract(filename, "wave_\\d{1}")
-    data <- fread(filename)
+
+    # get data, if we know the cols we want then we just select those
+
+    if (select_cols[1] == "all") { # bit sloppy
+        data <- fread(filename)
+    } else {
+        data <- fread(filename,
+            select = select_cols
+        )
+    }
     # Set wave in data
     data[, file_wave := WAVE_num]
+
     return(data)
 }
 
@@ -30,4 +41,11 @@ ReadRTF <- function(filename) {
     return(clean)
 }
 
-search_names <- function(data, string) grep(string, names(data), value = T)
+search_names <- function(data, string, ...) {
+    grep(
+        string,
+        names(data),
+        value = TRUE,
+        ...
+    )
+}

### 2023-07-13 Julia/5__solve_retirement_for_elsa.jl
Mainly creating subjective life tables

@@ -1,0 +1,36 @@
+using LinearAlgebra
+using DelimitedFiles
+using CSV
+using DataFrames
+using Plots
+using Pipe: @pipe
+
+
+# so first get functions from 4
+# then read in elsa data
+# then run solve using real wealth amounts and
+
+elsa_dt = CSV.File("../data/ELSA/elsa_to_use/for_julia.csv",) |>
+          DataFrame
+
+# that's slow
+#  
+Base.print_matrix(stdout, names(elsa_dt))
+
+# so for each individual get gender and financial wealth 
+# then run for all combos and save as jsons?
+# does dc pension count in that?
+small_elsa = @pipe filter(row -> row[:in_wave] == 1, elsa_dt) |>
+    filter(row -> row[:retired] == 1, _)
+ |> 
+select(_, [:ragender, :age_at_interview,
+    :public_pension, :fin_wealth, :retired_age]) 
+    
+elsa_dt.retired
+
+# now round financial wealth to nearest 250
+# pension to nearest 100?
+
+
+small_elsa[small_elsa., :public_pension]
+

### 2023-07-13 Texfiles/Draft1.tex
Mainly subjective life tables

@@
 
-\cite{mitchell_et_al_1999} - this is empirical
-\cite{friedman_warshawsky_qje_1990} - this is theory
-
-\cite{friedman_warshawsky_chicago_1988} - this is more theory. Probably good to have read.
-
 \cite{lockwood_aer_2018}
 
-. \textit{Continue
-      chat about how elderly population is growing and private pension pots are growing in size, could be worth referencing
-      latest census data so we seem super up to date, look at IFS data about the growth of DC pensions as well. }
-\textbf{I will do this this afternoon}
+\cite{vidalmelia_lejarragagarcia_munich_2004} have some interesting results. Need to talk about that.
+
+\subsection{Methodology}
+
+I first solve a modified retirement lifecycle model. Retirees face
+\begin{equation*}
+      E(U) = \sum_{t = x}^{110 - w}[u(c_{t})]
+\end{equation*}
 

### 2023-07-13 Julia/4_retirement_lifecycle.jl

@@ -15,8 +15,6 @@ using Plots
 
 # I can also make my life cycle code faster I think although not sure how
 
-# to
-
 cum_death_probs = CSV.File("Julia/death_probs.csv") |>
                   Tables.matrix |>
                   vec
@@ -135,7 +133,8 @@ end
 # can we just solve with annuitisation 
 # and then compare with starting on the grid higher or lower. 
 
-# Ok so run with annuity incomes of 
+# Ok so run with annuity incomes
+
 
 function LifeCycleSolve(annuity_cost; moneysworth=0.75, bequests=false, death_probs=cum_death_probs, pension=0)
 
@@ -155,7 +154,6 @@ function LifeCycleSolve(annuity_cost; moneysworth=0.75, bequests=false, death_pr
             consumption_matrix[x, y] = bc(assets, next_assets, annual_annuity_payment, pension)
         end
     end
-
 
     # # Check that low assets this period and high the next implies negative consumption
     # # and visa versa

### 2023-07-13 R/12__ELSA_harmonised.R

@@ -58,6 +58,10 @@ harm[, table(radyear)]
 harm[, unique(radyear)]
 harm[, sum(!is.na(radyear))]
 
+# H1ATOTF net value of non-housing financial wealth
+# H1ATOTH value of primary wealth
+
+
 
 harm_long <- melt(
     harm,
@@ -68,7 +72,8 @@ harm_long <- melt(
         "r\\d{1}retage", "r\\d{1}wretage",
         "r\\d{1}agey", "inw\\d{1}$",
         "r\\d{1}iwindm", "r\\d{1}iwindy",
-        "hh\\d{1}ctot1m"
+        "hh\\d{1}ctot1m", "h\\d{1}atotf",
+        "h\\d{1}atoth", "r\\d{1}ipubpen"
     ),
     value.name = c(
         "retired", "pension_type_job1",
@@ -76,7 +81,9 @@ harm_long <- melt(
         "retired_age", "expected_retired_age",
         "age_at_interview", "in_wave",
         "date_month", "date_year",
-        "total_monthly_consumption"
+        "total_monthly_consumption",
+        "fin_wealth", "house_value",
+        "public_pension"
     ),
     variable.name = "wave"
 )

### 2023-07-13 R/13__ELSA_eol.R

@@ -1,7 +1,9 @@
 library(data.table)
 library(dplyr)
 library(stringr)
-
+library(ggplot2)
+library(nls2)
+library(jsonlite)
 source("__ELSA_functions.R")
 
 
@@ -134,7 +136,7 @@ joined_harm <- left_join(
 
 library(tidyr)
 
-joined_harm %>%
+ready_for_nls <- joined_harm %>%
     select(idauniq, exlo80, exlo90, prob110, wave, age = age_at_interview) %>%
     filter(!is.na(exlo80) & !is.na(exlo90) & !is.na(prob110)) %>%
     filter(exlo80 >= 0, exlo80 <= 100) %>%
@@ -157,66 +159,248 @@ joined_harm %>%
     # generate some intial parameter values
     mutate(k = log(log(ex1) / log(ex2)) / (log(t1) - log(t2))) %>%
     mutate(lambda = (1 / t1) * ((-log(ex1))^(1 / k))) %>%
-    mutate(invlambda = lambda^(-1)) %>% 
-    filter(!is.infinite(lambda))
+    mutate(invlambda = lambda^(-1)) %>%
+    filter(!is.infinite(lambda)) %>%
+    filter(ex1 > ex2)
 
-nl ( exlo =  exp(-( {lambda}*t)^{k})) if id==`id', initial(lambda `initlambda' k `initk')    
-nls
 
-# gen     t1 = 75-age if age<66
-# replace t1 = 80-age if age>65
-# gen     t2 = 85-age
-# gen t3 = 110 - age
+hist(ready_for_nls$k)
 
-# gen k = ln(ln(exlo80)/ln(exlo90))/(ln(t1)-ln(t2))
-# gen lambda = (1/t1)*((-ln(exlo80))^(1/k))
-# gen invlambda = 1/lambda
-# *replace invlambda=. if invlambda>5000
-
-# /**********************************************************************************************/
-# /* Now, impose third point and then fit Weibull according to least squares					  */
-# /**********************************************************************************************/
-
-# * We are going to merge in the individual's probability of reaching age 110, according to life tables *
-# * We treat this as a third "answer" that the individual gives. *
+ready_for_nls[, mean(ex1 < ex2)]
 
 
 
+# convert to long, answer level
+long_ready <- ready_for_nls %>%
+    melt(
+        measure.vars = patterns(
+            t = "t",
+            exlo = "ex"
+        ),
+        variable.name = "names"
+    )
+
+
+long_ready[, id_wave := paste0(idauniq, "-", wave)]
[…]

### 2023-07-13 R/testing.R

@@ -1,0 +1,40 @@
+library(dplyr)
+library(tidyr)
+
+who %>% pivot_longer(
+    cols = new_sp_m014:newrel_f65,
+    names_to = c("diagnosis", "gender", "age"),
+    names_pattern = "new_?(.*)_(.)(.*)",
+    values_to = "count"
+)
+
+anscombe
+anscombe_2 <- anscombe %>%
+    pivot_longer(
+        everything(),
+        cols_vary = "slowest",
+        names_to = c(".value", "set"),
+        names_pattern = "(.)(.)"
+    )
+
+anscombe_2$set %>% table()
+
+
+
+library(dplyr)
+library(tidyr)
+
+longer <- pivot_longer(
+    dat,
+    cols = -1,
+    names_pattern = "(.*)(..)$",
+    names_to = c("limit", "name")
+) %>%
+    mutate(limit = ifelse(limit == "", "value", limit))
+
+answer <- pivot_wider(longer,
+    id_cols = c(group, name),
+    names_from = limit,
+    values_from = value,
+    names_repair = "check_unique"
+)

### 2023-07-14 Texfiles/Draft1.tex
added a small data section

@@
 
+\subsection{Data}
+The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA repeatedly interviews
+individuals over the age of 50 and asks them a range of questions relating to their income and wealth as well as expectations about
+the future. Importantly it also includes detailed information pensions including the type of pension that an individual holds when
+working, thus I am able to distinguish between individuals who have defined benefit and defined contribution pensions. For ease of access
+I use Harmonized ELSA \footnote{"This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
+      July 2021 developed by the Gateway to Global Aging Data. The development of the Harmonized ELSA was funded by the National
+      Institute on Aging (R01 AG030153, RC2 AG036619, R03 AG043052). For more information,
+      please refer to https://g2aging.org/.”} which ensures that variables are comparable across waves. Since this only includes a
+subset of the questions in ELSA I also supplement it with variables taken directly from the data.
+
+I also use ONS life tables data.
+
+I also use annuitant life tables data.
+
+I also use FCA product data.
+
 \subsection{Methodology}
 
-I first solve a modified retirement lifecycle model. Retirees face
+I first solve a modified retirement lifecycle model. The problem that retirees face is as follows:
+\textbf{This is not correct -- need to include probs}
 \begin{equation*}
@@
 \end{equation*}
+subject to their budget constraint
+\begin{equation*}
+      c_{t} =a_{t}(1 +r) + p_{t} + ann_{t} -  a_{t+1}
+\end{equation*}
+where $a_{t}$ are asset holdings in time t, $p_{t}$ is state pension income in time t, and $ann_{t}$ is annuity
+income in time t. I use the standard CRRA utility function.
+\begin{equation*}
+      u(c_{t}) = \frac{c_{t}^{1 - \sigma}}{1 - \sigma}
+\end{equation*} In some specifications retirees can leave bequests, I use the bequest function from
+\cite{lockwood_red_2012}.
+\begin{equation*}
+      u(b_{t}) = (\frac{m}{1 - m})^{\sigma}  \frac{((\frac{m c_{0}}{1 - m}) + b_{t})^(1 - \sigma)}{(1 - \sigma)}
+\end{equation*}
+Where $b_{t}$ is the amount left at death, $m$ is a measure of bequest motive strength and $c_{0}$ is
+a minimum amount of consumption that individuals want. \textbf{Check this}
+
+I solve the retirees problem using backward induction. At age 110 there is certainty of death so any leftover assets
+are carried over into the next period and bequested. This means that the value at the end of the final period is
+either 0 (if we do not allow a bequest motive) or the value of bequests. I then take this value function and solve
+an individuals final period problem, choosing assets next period (i.e. those to bequest) and how much to consume.
+
+Using the optimal policy function in the last period, I calculate the value of the last period, which the utility
+function and the value function evaluated at the maximum. This is then used in the problem the year before that.
+I repeat this process back to the age of retirement to obtain optimal consumption amounts for each year of retirement
+and associated value functions.
+
+To simulate the ELSA data I solve this retirement problems for each new retiree in the data set.
 

### 2023-07-14 Texfiles/Notes.tex
Mainly creating subjective life tables
added a small data section

@@
     \item Got CMI access.
+
+    \item To do: \begin{enumerate}
+              \item Start regression work - done
+              \item Finish making subjective survival curves - on their way
+              \item Finish lit review - done
+              \item Start methodology section - need to do
+              \item Write couple code - need to do
+              \item Write behaviourel model code need to do
+              \item Run lifecycle models with different starting points - this is to do.
+                    Do we have size of pension though? No. Only size of income.
+
+                    \item Do I want to add defined pension income to the model? Yes I think so. 
+          \end{enumerate}
 \end{itemize}

### 2023-07-14 Julia/4_retirement_lifecycle.jl

@@ -108,11 +108,14 @@ function utility(c)
     end
 end
 
+utility(6)
 
 # bequest function
 # from lockwood RED 2012
 bequest_lock(b; c0=5000, m=0.96) = (m / (1 - m))^σ * (((m / (1 - m)) * c0) + b)^(1 - σ) / (1 - σ)
 # but won't including bequests make utility lower
+
+bequest_lock(10)
 
 # Pricing of annuity
 # sum of payments divided by prob of death
@@ -135,62 +138,67 @@ end
 
 # Ok so run with annuity incomes
 
-
 function LifeCycleSolve(annuity_cost; moneysworth=0.75, bequests=false, death_probs=cum_death_probs, pension=0)
+    # Function definition with parameters:
+    # - annuity_cost: the cost of the annuity
+    # - moneysworth: a factor used to calculate the annual annuity payment
+    # - bequests: a boolean indicating whether bequests are considered
+    # - death_probs: an array of cumulative death probabilities
+    # - pension: a fixed pension amount
 
-    # Take the annuity cost and get the rough annual payment 
     annual_annuity_payment = annuity_payment_function(annuity_cost, moneysworth=moneysworth)
+    # Calculate the rough annual annuity payment based on the annuity cost and moneysworth factor
 
-    # Intialise consumption grid
+
     consumption_matrix = zeros(grid_points, grid_points)
+    # Initialize a matrix to store consumption values
 
-    # Now for each level of assets return how much consumption individuals get
+
+    # Nested loop to calculate consumption values for each combination of assets
     for x in eachindex(asset_grid)
         assets = asset_grid[x]
         for y in eachindex(asset_grid)
             next_assets = asset_grid[y]
-            # assets this period are 1st dimension
-            # assets next period are 2nd dimension
             consumption_matrix[x, y] = bc(assets, next_assets, annual_annuity_payment, pension)
         end
     end
 
-    # # Check that low assets this period and high the next implies negative consumption
-    # # and visa versa
-    # consumption_matrix[1, 500]
-    # consumption_matrix[1000, 1]
 
     utility_matrix = zeros(grid_points, grid_points)
+    # Initialize a matrix to store utility values
 
+    # Loop to calculate utility values for each consumption value in the consumption matrix
     for i in eachindex(utility_matrix)
         utility_matrix[i] = utility(consumption_matrix[i])
     end
 
+    # If bequests then we get utility at end of life from certain death
+    if bequests == true
[…]

### 2023-07-14 R/12__ELSA_harmonised.R

@@ -89,7 +89,10 @@ harm_long <- melt(
 )
 
 fwrite(harm_long, "../../data/ELSA/elsa_to_use/harmonised_data_long.csv")
+
 stop()
+
+## below this stop I explore the data
 harm_long[, unique(variable)] # what is this variable?
 
 harm_long[, .(number_of_waves = sum(in_wave)), by = .(idauniq)][order(number_of_waves)]
@@ -389,3 +392,17 @@ table(harm_long[, pension_type_job1], useNA = "ifany")
 harm_long[, .(mean(total_monthly_consumption, na.rm = T), sd(total_monthly_consumption, na.rm = T)),
     by = .(date_month)
 ][order(date_month)]
+
+
+
+## plot state pension
+names(harm_long)
+
+hist(harm_long[, public_pension])
+hist(harm_long[, fin_wealth])
+
+harm_long[, mean(fin_wealth < 0, na.rm = TRUE)]
+
+
+# will I actually be able to solve
+# enough life cycle models

### 2023-07-16 Texfiles/Draft1.tex
Lots of lifecycle work

@@
 levels that theory predicts. In this paper I test two competing hypotheses for the annuity problem: bequests and
-pessimistic life expectancy. Under the coalition government in the UK the law regarding the use of private defined
+pessimistic life expectancy. In most developed countries, individuals are forced to contribute to either a defined 
+benefit or defined contribution pension. A defined benefit pension accumulates throughout your job tenure as  
+ Under the coalition government in the UK the law regarding the use of private defined
 contribution pensions changed. Individuals were no longer forced to annuitise their pension pots and could access
@@
 
-                  \item Need annuity theory references -- Yarri?
+                  \item Add bit about DC/DB pensions in the intro. Also talk about heterogeneity across countries.
+                        Some countries want to move towards more annuitisation. This annual review is a good source of info
+                        \cite{banks_crawford_ar_2022}
             \end{enumerate}

### 2023-07-16 Texfiles/Notes.tex
Lots of lifecycle work

@@
 
+\subsection{\textbf{14-07}}
+\begin{itemize}
+      \item Pension wealth variables from IFS -- I could illustrate
+            the method with just financial wealth and then use pension wealth calculations
+            as well. Pension wealth gives us retirement income. This leaves financial wealth to be
+            invested in annuities.
+
+      \item Apparently there are pension wealth variables for DC pensions?
+            Cant find in harmonized but perhaps they are in the pension files?
+
+      \item Could subjective survival probs also explain slow asset decumulation. And then by the time they are old
+            they have medical risks.
+
+      \item "Allowing individuals to solve optimisation problems using hueristics"
+\end{itemize}
+
 \subsection{\textbf{12-07}}
@@
                   \item Do I want to add defined pension income to the model? Yes I think so.
+                  \item Can also check the consumption responses of individuals who think they are going to leave a bequest.
             \end{enumerate}

### 2023-07-16 Julia/0__functions.jl

@@ -1,0 +1,85 @@
+using CSV
+using LinearAlgebra
+using DataFrames
+using Plots
+using JSON3
+
+
+function get_death_prob_function()
+    # this function returns a function that gives death probs
+
+    # First get some meta info
+    first_two_lines = CSV.File("../data/ONS/males_transition_probs.csv", limit=2, header=false) |>
+                      Tables.matrix
+
+    header_info = first_two_lines[1, :]
+    # first column is age 
+    # next columns are year of birth
+
+    # The second element of first column is the min age
+    min_age = parse(Int, first_two_lines[2, 1])
+
+    # Max year
+    max_year = Int(header_info[2])
+
+
+    # Objective transition probs from ons
+    male_lftb = CSV.File("../data/ONS/males_transition_probs.csv") |>
+                Tables.matrix
+
+    female_lftb = CSV.File("../data/ONS/females_transition_probs.csv") |>
+                  Tables.matrix
+    # Each colum is a year of birth
+    # each row is an age 
+    return (male=male_lftb, female=female_lftb,
+        top_year=max_year, min_age=min_age)
+end
+
+death_prob_matrices = get_death_prob_function()
+
+# ONS life tables
+function death_probabilities(; gender, year, age)
+    if gender == "male"
+        data = death_prob_matrices.male
+    elseif gender == "female"
+        data = death_prob_matrices.female
+    end
+
+    birth_year = year - age
+
+    if birth_year < death_prob_matrices.top_year | age < death_prob_matrices.min_age
+        println("Person is too young, we dont have death prob")
+    end
+
+    # plus 2 because first column is ages, and plus one because of indexing
+    col_index = 2 + death_prob_matrices.top_year - birth_year # selects the correct column
+    # second column of data is birth year 1975, so if if birth_year = 1975 we will select it
+
+    start_row_index = 1 + age - death_prob_matrices.min_age # first row of interest
+
+    death_probs = data[start_row_index:end, col_index]
+
+    return death_probs
+end
+
+# utility function
+function utility(c)
+    if c > 0.0001
+        (c^(1 - σ)) / (1 - σ)
+    else
+        -Inf
+    end
+end
+
+
+
+# bequest function
+# from lockwood RED 2012
+bequest_lock(b; c0=5000, m=0.96) = (m / (1 - m))^σ * (((m / (1 - m)) * c0) + b)^(1 - σ) / (1 - σ)
[…]

### 2023-07-16 Julia/1__lifecycle_functions.jl

@@ -1,0 +1,228 @@
+include("0__functions.jl")
+
+# Ok I think I should make this into a module and export the functions 
+
+# Set parameters
+const β = 0.97
+const r = 1 / β - 1
+const σ = 2.5
+const grid_points = 100
+const grid_gap = 1000
+const terminal_age = 110
+
+# define consumption grid
+const asset_grid = [i * grid_gap for i in 1:grid_points]
+
+
+death_probabilities(gender="male", year=2010, age=63)
+death_probabilities(gender="female", year=2010, age=63)
+death_probabilities(gender="female", year=2015, age=63)
+
+
+
+
+# Pricing of annuity
+# sum of payments divided by prob of death
+# divided by loading factor
+# times interest rate
+# then we calc utility
+
+
+function LifeCycleSolve(
+    annuity_cost;
+    bequests=false,
+    pension=0,
+    age,
+    gender,
+    life_prob_types="objective",
+    year)
+    # Function definition with parameters:
+    # - annuity_cost: the cost of the annuity
+    # - moneysworth: a factor used to calculate the annual annuity payment
+    # - bequests: a boolean indicating whether bequests are considered
+    # - pension: a fixed pension amount
+    # - age: age to run the lifecycle model from, this impacts years of life left and annuity rates
+
+
+    # set file name to save json to. 
+    json_filename = string(age, "_beq", bequests, "_", gender, "_", life_prob_types, "_", year, "_pen", pension)
+    filename_with_dir = "../data/ELSA/lifecycle_outputs/" * json_filename
+
+    # If we have a json with those combos already then we skip
+    # otherwise run the life cycle model as usual
+    if isfile(filename_with_dir)
+        return "json_run_already"
+    end
+
+
+    # set death probabilities.
+    if life_prob_types == "objective"
+        death_probs = death_probabilities(gender=gender, age=55, year=year)
+    elseif life_prob_types == "subjective"
+        death_probs = "" # come and define later
+    end
+    # - death_probs: an array of cumulative death probabilities that are used in the optimisation stage
+
+
+
+    # Retirement cohort
+    if year > 2013
+        ret_cohort = 2016
+    else
+        ret_cohort = 2008
+    end
+    # - ret_cohort: either pre 2014 or post 2014 cohort of data. this impacts annuity rates
+
+    # using annuitant life tables estimate annuity payments 
+    # that are realistic
+    annual_annuity_payment = annuity_payment_function(
+        annuity_cost, age=age,
[…]

### 2023-07-16 Julia/4__lifecycle_with_income.jl

@@ -1,0 +1,274 @@
+using LinearAlgebra
+using DelimitedFiles
+using CSV
+using DataFrames
+using Plots
+using JSON3
+
+# Ok I think I should make this into a module and export the functions 
+
+# Set parameters
+const β = 0.97
+const r = 1 / β - 1
+const σ = 2.5
+const terminal_age = 110
+
+const asset_grid_points = 100
+const asset_grid_gap = 1000
+# define asset grid
+const asset_grid = [i * asset_grid_gap for i in 1:asset_grid_points]
+
+# define income grid
+const inc_grid_points = 100
+const inc_grid_gap = 500
+const income_grid = [i * inc_grid_gap for i in 1:inc_grid_points]
+
+include("0__functions.jl")
+
+consumption_array = ones(Float64, asset_grid_points, asset_grid_points, inc_grid_points)
+# Initialize a matrix to store consumption values
+
+# Nested loop to calculate consumption values for each combination of assets
+for z in eachindex(income_grid)
+    income = income_grid[z]
+    for x in eachindex(asset_grid)
+        cur_assets = asset_grid[x]
+        for y in eachindex(asset_grid)
+            next_assets = asset_grid[y]
+            consumption_array[x, y, z] = bc_inc(cur_assets, next_assets, income)
+        end
+    end
+end
+
+utility_array = ones(Float64, asset_grid_points, asset_grid_points, inc_grid_points)
+# Initialize a matrix to store utility values
+
+# utility function
+function utility(c)
+    if c > 0.0001
+        (c^(1 - σ)) / (1 - σ)
+    else
+        -Inf
+    end
+end
+
+# Loop to calculate utility values for each consumption value in the consumption matrix
+for i in eachindex(utility_array)
+    utility_array[i] = utility(consumption_array[i])
+end
+
+
+retirement_lifecycle_with_income = function (;
+    bequests=false,
+    age=60,
+    gender="male",
+    life_prob_types="objective",
+    year=2015)
+
+
+    # set file name to save json to. 
+    json_filename = string(age, "_beq", bequests, "_", gender, "_", life_prob_types, "_", year, ".json")
+    filename_with_dir = "../data/ELSA/lifecycle_outputs/" * json_filename
+
+    ## So now for each level of income and current assets we find best consumption next period
+    terminal_age = 110
+    start_age = age
+
+    death_probs = death_probabilities(gender=gender, age=age, year=year)
+
+    # Dimensions are the state variables
[…]

### 2023-07-16 Julia/4_retirement_lifecycle.jl

@@ -1,456 +1,0 @@
-using LinearAlgebra
-using DelimitedFiles
-using CSV
-using DataFrames
-using Plots
-
-# need life probabilities. 
-# copied for now from lockwood
-# Ok I think I should make this into a module and export the functions 
-# then I can use revise for a better workflow. 
-# Things to do: 
-# add gender and maybe health status into the model 
-# then find the expectancies and bequest motives which 
-# generate similar sized proportions of annuitisation
-
-# I can also make my life cycle code faster I think although not sure how
-
-cum_death_probs = CSV.File("Julia/death_probs.csv") |>
-                  Tables.matrix |>
-                  vec
-# Increase every prob of death by 0.01
-
-# Mortality data from the ONS
-const base_year = 2005
-const base_age = 50
-
-male_data = CSV.File("../data/ONS/male_transition_probs.csv") |>
-            Tables.matrix
-
-female_data = CSV.File("../data/ONS/female_transition_probs.csv") |>
-              Tables.matrix
-# Each column is the individuals age in 2010. 
-# each row is a year after that, transition probs from 
-# age to age + 1
-
-# these people are aged 50 in 2005
-plot([male_data[:, 21] male_data[:, 1]])
-
-# So then if we want to pdf of probs in a given age 
-# ONS life tables
-function death_probabilities(data, year=2012, age=65)
-    # return probs from this age and year onwards
-    col_index = year - (base_year - 1) # selects the correct column
-    start_row_index = age - base_age # first row of interest
-
-    death_probs = data[start_row_index:51, col_index]
-
-    return death_probs
-end
-
-# If I dont update 
-death_probabilities(male_data, 2010, 63)
-death_probabilities(female_data, 2010, 63)
-# these finish at age 100. and start at the age specified
-
-
-# ok so now I can set gender and use realistic lifetables
-# I could have a look at annuitant tables as well and then use a 
-# scaling factor on the implied fair price to calucalte the price
-# that normal people face. 
-
-# can set wealth from elsa data as well 
-
-
-
-
-
-function constrained_add(a, b)
-    # This is just so I can play around with subjective 
-    # probabilities and get strictly increasing probs. 
-    if (a + b) > (a + (1 - a) / 2)
-        return a + (1 - a) / 2
-    else
-        return a + b
-    end
-end
-
[…]

### 2023-07-16 Julia/5__annuity_code.jl

@@ -1,0 +1,50 @@
+
+include("0__functions.jl")
+
+ann_lf_df = CSV.File("../data/CMI/raw/cmi_annuitant_lifetables.csv") |>
+            DataFrame
+
+# get vectors for all diff years and genders
+ann_lf_male_08 = ann_lf_df[(ann_lf_df.gender.=="male").&(ann_lf_df.year.==2008), :][!, :prob]
+ann_lf_female_08 = ann_lf_df[(ann_lf_df.gender.=="female").&(ann_lf_df.year.==2008), :][!, :prob]
+ann_lf_male_16 = ann_lf_df[(ann_lf_df.gender.=="male").&(ann_lf_df.year.==2016), :][!, :prob]
+ann_lf_female_16 = ann_lf_df[(ann_lf_df.gender.=="female").&(ann_lf_df.year.==2016), :][!, :prob]
+
+
+function annuity_payment_function(
+    annuity_cost;
+    loading_factor=0.95,
+    age=60,
+    gender,
+    ann_life_table_year)
+
+    # if older than 60 the annuity should be priced according to mortality from that point
+    start_index = age - 60 + 1
+
+    if ann_life_table_year == 2008 && gender == "male"
+        death_probs = ann_lf_male_08
+    elseif ann_life_table_year == 2008 && gender == "female"
+        death_probs = ann_lf_female_08
+    elseif ann_life_table_year == 2016 && gender == "male"
+        death_probs = ann_lf_male_16
+    elseif ann_life_table_year == 2016 && gender == "female"
+        death_probs = ann_lf_female_16
+    end
+
+    # now want p we reach ages given we are age
+    alive_probs = 1 .- death_probs[start_index:end]
+
+    # Prob of reaching 61 is just top
+    # prob of reaching 62 is top two multiplied together
+    # So we take cumulative product of alive probability
+    cum_alive = cumprod(alive_probs)
+
+    loading_factor * (annuity_cost / (sum([cum_alive[age] * 1 / (((1 + r)^age)) for age in eachindex(cum_alive)])))
+end
+# cool now we can use this function.. 
+
+annuity_payment_function(100000, gender="male", age=61, ann_life_table_year=2016)
+## males get more which is right # is this roughly inline for a nominal non esclating pension
+
+annuity_payment_function(100000, gender="male", age=65, ann_life_table_year=2016)
+# also annuities increase with age which is good

### 2023-07-16 R/14__ONS_lifeexpec.R

@@ -195,7 +195,7 @@ source("__ELSA_functions.R")
 #### USE LIFE TABLES WITH AGE UNTIL 110 IN
 
 
-create_death_prob_matrix <- function(gender) {
+create_death_prob_matrix <- function(gender = "Males") {
     # Function that creates the transition matrix from ons data.
 
     sheet_to_get <- paste0(gender, " cohort lx")
@@ -222,10 +222,17 @@ create_death_prob_matrix <- function(gender) {
         arrange(age) %>%
         mutate(deaths = lead(lag(alive) - alive)) %>% # alive yesterday minus alive today, shifted back to today
         mutate(death_rate = deaths / alive) %>%
-        filter(birth_year >= 1935, birth_year <= 1955) %>%
+        ungroup() %>%
+        filter(birth_year >= 1935, birth_year <= 1975) %>%
         mutate(year = age + birth_year) %>%
         filter(year >= 2005) %>%
-        filter(age >= 50, age <= 110)
+        filter(age >= 50, age <= 110) %>%
+        arrange(-birth_year)
+
+    ons_surv %>%
+        filter(age == 59 & year == 2016)
+
+    min(ons_surv$birth_year)
 
     ons_surv_matrix <- ons_surv %>%
         tidyr::pivot_wider(
@@ -236,7 +243,7 @@ create_death_prob_matrix <- function(gender) {
 
     fwrite(
         ons_surv_matrix,
-        paste0("../../data/ONS/", gender, "_transition_probs.csv")
+        str_to_lower(paste0("../../data/ONS/", gender, "_transition_probs.csv"))
     )
 }
 

### 2023-07-16 R/15__regressions.R

@@ -66,20 +66,32 @@ harm_long[, ever_dc_pen := fifelse(
 by = .(idauniq)
 ]
 
-harm_long[, years_since_retirement := age_at_interview - retired_age]
 
-# table(harm_long$years_since_retirement)
+harm_long[, has_db_pension := case_when(
+    pension_type_job1 == 2 ~ "yes",
+    pension_type_job2 == 2 ~ "yes",
+    pension_type_job3 == 2 ~ "yes",
+    T ~ "no"
+)]
 
-# hist(harm_long[total_monthly_consumption < 0, total_monthly_consumption])
+harm_long[, ever_db_pen := fifelse(
+    any(has_db_pension == "yes"),
+    "yes", "no"
+),
+by = .(idauniq)
+]
+
+
+# Don't use either because it doesnt give us much info.
 
 # ever have a dc pension
 # 1 dc, 2 db, 3 either
 # want to find people who have had dc pension at some point
 
 
-fwrite(harm_long, "../../data/ELSA/elsa_to_use/for_julia.csv")
+harm_long[, years_since_retirement := age_at_interview - retired_age]
 
-stop()
+
 # now also save financial wealth
 # pension wealth
 # state pension
@@ -106,7 +118,6 @@ AddYearsAround <- function(year, add = 2) {
 harm_long$fuzzy_retirement <- lapply(harm_long$retired_age, AddYearsAround, 2)
 
 
-
 harm_long[, ":="(ret_in = mapply("%in%", age_at_interview, fuzzy_retirement),
     ret_equals = retired_age == age_at_interview), by = .(idauniq)]
 
@@ -114,6 +125,22 @@ harm_long[ret_in & ret_equals, .(retired_age, age_at_interview, ret_in)]
 # ok this works
 
 
+data_for_regressions <- harm_long[ret_in == TRUE & !is.na(pre_post_ref)]
+
+data_for_regressions[, id_wave := paste0(idauniq, "-", wave)]
+
+# nice good number of people in both.
+data_for_regressions[, table(ever_db_pen, ever_dc_pen)]
+
+fwrite(
+    data_for_regressions,
+    "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
+)
+
+nrow(data_for_regressions)
+
+
+#### Plot consumption
 consump_time_in <- harm_long[ret_in == TRUE,
     .(mean_consump = mean(total_monthly_consumption, na.rm = T)),
[…]

### 2023-07-16 R/16__ELSA_prep_julia.r

@@ -1,0 +1,98 @@
+library(data.table)
+library(dplyr)
+library(stringr)
+library(ggplot2)
+library(readxl)
+library(tidyr)
+library(lubridate)
+library(haven)
+library(patchwork)
+
+source("__ELSA_functions.R")
+
+# this is the distance that we use on the asset grid in Julia
+grid_distance <- 250
+
+
+reg_data <- fread(
+    "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
+)
+
+## now for julia we need to round financial assets
+# what to do with negative assets
+# reg_data$fin_wealth %>% hist()
+
+
+reg_data[, round_fin_wealth :=
+    round(fin_wealth / grid_distance, 0) * grid_distance] # round to nearest grid point
+
+# do we need some other things?
+# I am not sure...
+dt_for_julia <- reg_data[, .(id_wave,
+    gender = ragender, age = age_at_interview,
+    round_fin_wealth, public_pension,
+    ever_dc_pen, ever_db_pen,
+    year = date_year
+)]
+
+dt_for_julia[, gender := fcase(
+    gender == 1, "male",
+    gender == 2, "female"
+)]
+
+dt_for_julia[, public_pension := round(public_pension / 100, 0) * 100] # round pension to the nearest 100
+
+summary(dt_for_julia$round_fin_wealth)
+
+# , ever_dc_pen, ever_db_pen
+gender_age_year_unique <- dt_for_julia[!is.na(public_pension), .(gender, age, year)] %>%
+    unique()
+
+# check no NAs in any columns
+stopifnot(!any(sapply(gender_pension_age_unique, \(x) any(is.na(x)))))
+
+fwrite(
+    gender_age_year_unique,
+    "../../data/ELSA/elsa_to_use/gender_pension_age_unique.csv"
+)
+# We want to run the life cycle model for each of these gender-pension-age pairs.
+
+
+
+fwrite(
+    dt_for_julia,
+    "../../data/ELSA/elsa_to_use/for_julia.csv"
+)
+
+
+dir("../../data/ELSA/lifecycle_outputs/", full.names = T) %>%
+    length()
+# its not actually that slow!
+
+time_made <- dir("../../data/ELSA/lifecycle_outputs/", full.names = T) %>%
+    lapply(file.info) %>%
+    sapply(\(x) x$ctime) %>%
+    sort()
+
+# takes about a second
+summary(time_made - lag(time_made), na.rm = T)
[…]

### 2023-07-16 R/17__cmi_annuity.r

@@ -1,0 +1,84 @@
+# calculate annuity prices using the IofAs continuous mortality series data
+# this is the actual death probs of annuitants,
+# so we can add a loading factor to the implied price and use
+# that.
+
+
+library(readxl)
+library(dplyr)
+library(data.table)
+library(stringr)
+library(ggplot2)
+
+# annuitant rates
+
+# this is the 2008 annuitants mortality rates
+# Table	        Description
+# LML08	        Life annuities Male Lives
+# LFL08	        Life annuities Female Lives
+
+# annuity life tables in 2008
+ann_lt_08 <- read_excel(
+    "../../data/CMI/raw/92190-cmiwp81-final08series-annuities-rates.xlsx",
+    sheet = "08 annuities qx",
+    skip = 4
+) %>%
+    setnames(str_to_lower) %>%
+    select(age = "age x", "lml08", "lfl08") %>% # get annuitant life tables
+    filter(age >= 60) %>% # only have data for those people who are over 60
+    mutate(year = 2008) %>%
+    rename(
+        male_prob = lml08,
+        female_prob = lfl08
+    )
+
+
+
+###### 2016 data ######
+# Table name	Product type	Gender	Weighting
+# PML16	        All products	Male	Lives
+# PFL16	        All products	Female	Lives
+
+ann_lt_16 <- read_excel(
+    "../../data/CMI/raw/Final 16 Series pension annuitant mortality rates v01 2020-07-10.xlsx",
+    sheet = "q",
+    skip = 6
+) %>%
+    setnames(str_to_lower) %>%
+    select(age, pml16, pfl16) %>%
+    filter(age >= 60) %>%
+    mutate(year = 2016) %>%
+    rename(
+        male_prob = pml16,
+        female_prob = pfl16
+    )
+
+# these are just probs of reaching next year
+#   conditional on being a certain age.
+# I need to adjust if I am just considering from age 60.
+#
+
+
+both_years <- rbindlist(list(ann_lt_08, ann_lt_16)) %>%
+    tidyr::pivot_longer(
+        c(male_prob, female_prob),
+        names_to = "gender",
+        values_to = "prob"
+    ) %>%
+    mutate(gender = str_remove(gender, "_prob"))
+
+both_years %>%
+    mutate(gender_year = paste0(gender, "-", year)) %>%
+    ggplot(aes(
+        x = age, y = prob,
+        colour = gender_year
+    )) +
+    geom_point() +
[…]

### 2023-07-16 R/7__ELSA_pension.r

@@ -17,18 +17,54 @@ pension_grid <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
 #  %>%
 # rbindlist(fill = T, use.names = T)
 
-mapply(search_names, pension_grid, "wpapu")
+mapply(search_names, pension_grid, "wpapu", ignore.case = TRUE)
+
+pension_grid
+
+head(pension_grid[[1]])
+
+pension_grid1 <- pension_grid[[1]]
+
+head(pension_grid1)
 
 # oh maybe this is when the pension mapping grid becomes useful.
 # get pension grid lookups
-grid_lookups <- read_excel("../../data/ELSA/elsa_unziped/UKDA-5050-tab/mrdoc/excel/5050_wave_2_pension_grid_corresponding_variables.xls")
+grid_lookups <- read_excel(
+    "../../data/ELSA/elsa_unziped/UKDA-5050-tab/mrdoc/excel/5050_wave_2_pension_grid_corresponding_variables.xls"
+)
+
 names(grid_lookups)
 head(grid_lookups)
 # lets have a look at the excel files.
 # pension variables seem to be different in each
 
+search_files <- function(string, ...) {
+    dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+        pattern = "wave",
+        full.names = TRUE
+    ) %>%
+        grep(string, ., value = T, ...)
+}
 
+dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "wave",
+    full.names = TRUE
+) %>%
+    grep("pension(_wealth|wealth)", ., value = T)
+# Only up to wave 5
 
+search_files("fin")
+
+grid_5 <- fread("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_5_pension_wealth.tab")
+
+fin_7 <- fread("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_7_financial_derived_variables.tab")
+
+names(grid_5)[names(grid_5) %in% names(fin_7)]
+names(grid_5)
+names(fin_7)
+head(grid_5)
+
+fin_6 <- fread
 
 fread(pension_files[[1]]) %>%
     names() %>%
@@ -62,3 +98,6 @@ head(all_pension)
 
 # Why is this research interesting -- no satisfactory life-cycle model that explains the
 # dec
+
+
[…]

### 2023-07-16 Texfiles/Draft1.tex
end of day

@@
 \begin{equation*}
-      c_{t} =a_{t}(1 +r) + p_{t} + ann_{t} -  a_{t+1}
+      c_{t} =a_{t}(1 +r) -  a_{t+1} + y_{t}
 \end{equation*}
-where $a_{t}$ are asset holdings in time t, $p_{t}$ is state pension income in time t, and $ann_{t}$ is annuity
-income in time t. I use the standard CRRA utility function.
+where $a_{t}$ are asset holdings in time t and $y_{t}$ is income in period 2. Income can come either
+from state pensions, defined benefit pension plans or purchased annuities.
 \begin{equation*}
@@
 
-To simulate the ELSA data I solve this retirement problems for each new retiree in the data set.
+To simulate the ELSA data I solve this retirement problems for each new retiree in the data set dependent on their
+objective probability of death each period. I estimate with subjective life expectancies and objective life expectancies,
+I also estimate the model both with and without a bequest motive which was picked to fit the unforced real annuity
+rates seen in the data. I then estimate several empirical models with the simulated data.
+
+\subsection{Empirical models}
+
+In this section I outline the key empirical models I run with both the simulated consumption data from the
+lifecycle models and with the real data from ELSA. I then see which lifecycle model better fits the consumption
+response that happened as a result of the pension reform.
+
+I use a regression discontinuity design where I compare the consumption of recent retirees after the policy reform
+to recent retirees before the policy reform. The individuals who retired in 2014 and later were not forced to
+annuitise their defined contribution pension. The key assumption implicit in regression discontinuities is
+that nothing else changes at the time of the jump apart from the policy of interest. And that the policy occurs
+without individuals predicting it. The policy change was widely seen as a surprise by the media and financial planners,
+some of whom in fact complained that they had not been consulted enough. \textbf{How can I justify no other jumps at the jump}
+
+Retirement year is the running variable and individuals are treated if retirement year is greater than 2013 and less than 2016.
+I consumption of individuals up to 2 years into retirement so that the sample size is larger. So if someone retired in 2014 and
+had consumption data in 2014 and 2016 I include both values. An individual is considered not treated if they retire before 2013 and after 2011.
+
+The estimating equation is therefore
 

### 2023-07-16 Julia/2__run_models.jl

@@ -3,6 +3,7 @@ include("4__lifecycle_with_income.jl")
 
 # Could run this for each 
 
+
 # annuity_cost;
 # moneysworth=0.95,
 # bequests=false,
@@ -21,17 +22,8 @@ gender_age_year_combos = CSV.File("../data/ELSA/elsa_to_use/gender_pension_age_u
                          Tables.matrix
 # This is the combo of demographic info we need to run it for
 
-# drop those under 60 for now. Need to change annuity function to just be life probs with some loading factor
-sum([row[2] < 60 for row in eachrow(gender_age_year_combos)])
-gender_age_year_combos = gender_age_year_combos[gender_age_year_combos[:, 2].>=60, :]
-
 
 gender_age_year_combos
-
-gender = "male"
-age = 64
-year = 2011
-
 
 # This will take a while I think
 # I wonder how long
@@ -52,11 +44,10 @@ end
 
 
 
+
 # Nice this is working now!!
-# Just need to run for different level of forced annuities
-# different life expectancies 
-# different amount of state pension income
-# different years and genders 
+# different life expectancies & different years and genders (these are kinda the same)
+
 # get a unique df with all the types we will need
 
 # (opt_policy_function_ann, temp_value_ann, ann_consump) = LifeCycleSolve(100000, bequests=false)

### 2023-07-16 Julia/4__lifecycle_with_income.jl

@@ -13,7 +13,7 @@ const r = 1 / β - 1
 const σ = 2.5
 const terminal_age = 110
 
-const asset_grid_points = 100
+const asset_grid_points = 500
 const asset_grid_gap = 1000
 # define asset grid
 const asset_grid = [i * asset_grid_gap for i in 1:asset_grid_points]
@@ -24,6 +24,7 @@ const inc_grid_gap = 500
 const income_grid = [i * inc_grid_gap for i in 1:inc_grid_points]
 
 include("0__functions.jl")
+
 
 consumption_array = ones(Float64, asset_grid_points, asset_grid_points, inc_grid_points)
 # Initialize a matrix to store consumption values
@@ -42,15 +43,6 @@ end
 
 utility_array = ones(Float64, asset_grid_points, asset_grid_points, inc_grid_points)
 # Initialize a matrix to store utility values
-
-# utility function
-function utility(c)
-    if c > 0.0001
-        (c^(1 - σ)) / (1 - σ)
-    else
-        -Inf
-    end
-end
 
 # Loop to calculate utility values for each consumption value in the consumption matrix
 for i in eachindex(utility_array)
@@ -205,9 +197,7 @@ retrieve_policy_function = function (
     return pol_func_readin
 end
 
-
-pol_func = retrieve_policy_function()
-pol_func[income_grid_point, :, :]
+# pol_func[income_grid_point, :, :]
 # now we trace through this
 
 
@@ -242,29 +232,29 @@ function asset_path_function(opt_policy_function, asset_start_point, income_grid
 
 end
 
-for_plots_60 = asset_path_function(
-    retrieve_policy_function(false, "objective", 60, 2010, "male")[50, :, :],
-    70,
-    60)
+# for_plots_60 = asset_path_function(
+#     retrieve_policy_function(false, "objective", 60, 2010, "male")[50, :, :],
+#     70,
+#     60)
 
-for_plots.assets[6]
+# for_plots.assets[6]
 
-for_plots_65 = asset_path_function(
-    retrieve_policy_function(false, "objective", 65, 2015, "male")[50, :, :],
-    55,
-    60)
+# for_plots_65 = asset_path_function(
+#     retrieve_policy_function(false, "objective", 65, 2015, "male")[50, :, :],
+#     55,
+#     60)
 
 
-death_probabilities(gender="male", year=2010, age=60)[6:end]
+# death_probabilities(gender="male", year=2010, age=60)[6:end]
 
[…]

### 2023-07-19 Texfiles/Draft1.tex
Lifecycle work

@@
 
-\subsection{Data}
+\textit{There are some more papers to include here. Time to go and have dinner. }
+
+\section{Data}
 The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA repeatedly interviews
@@
 
-I also use ONS life tables data.
+ELSA also includes questions on expenditures. In particular individuals are asked how much they consume on a range of broad categories.
 
-I also use annuitant life tables data.
+% Include a summary stats table here of the individuals selected for the regression
 
-I also use FCA product data.
+I also use life tables from the UK's Office for National Statistics. These give us risk of death for each age group. I adjust them 
+to make death certain at age 110 as is common in the literature. I transform these so that I have risk of death conditional on being
+a given age since this is what is used in the life cycle simulations.  
+%  include a table showing death probs
 
-\subsection{Methodology}
+To illustrate the effect the reform had on sales of annuities in the UK I obtained product data from the Financial Conduct Authority. 
+These track the sale of different financial products overtime including data on annuity sales.
+% Plot showing the drop of in annuity sales
 
-I first solve a modified retirement lifecycle model. The problem that retirees face is as follows:
-\textbf{This is not correct -- need to include probs}
+To calculate subjective life expectancies I follow \cite{odea_sturrock_rest_2023}. Individuals are asked “What are the chances that you will 
+live to be age X or more?” where X changes depending on the age of the interviewee. If individuals were under 65 then X was 75, if individuals
+were 66 and older they were asked the age that was 11 to 15 years older than them and is a multiple of 5. From wave three respondents were 
+also asked “What are the chances that you will live to be age 85 or more?” if they were under 70. As most recent retirees are under 70 we 
+therefore have two data points. I carry out the following procedures: drop any individuals who think it is more likely they reach an older
+age than a younger age since; add, as a third data point their objective chance of reaching 110 according to the ONS life tables; fit these
+three points to a Weibull distribution using non-linear least squares. Then I create subjective survival tables.
+
+\textbf{I could go into more detail here? Maybe I should. Add equation etc}
+
+\section{Methodology}
+
+I first solve a modified retirement lifecycle model. The problem that retirees face is as follows. 
+Every period retirees solve:
 \begin{equation*}
-      E(U) = \sum_{t = x}^{110 - w}[u(c_{t})]
+  V_{t}(a_{t}, y) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V_{t+1}(a_{t+1}, y) \}
 \end{equation*}
@@
 \begin{equation*}
-      c_{t} =a_{t}(1 +r) -  a_{t+1} + y_{t}
+  c_{t} =a_{t}(1 +r) -  a_{t+1} + y
 \end{equation*}
-where $a_{t}$ are asset holdings in time t and $y_{t}$ is income in period 2. Income can come either
+where $a_{t}$ are asset holdings in time t, $y$ is constant income for all periods and . Income can come either
 from state pensions, defined benefit pension plans or purchased annuities.
@@
 \begin{equation*}
-      u(b_{t}) = (\frac{m}{1 - m})^{\sigma}  \frac{((\frac{m c_{0}}{1 - m}) + b_{t})^(1 - \sigma)}{(1 - \sigma)}
+  B(a_{t}) = \bigl( \frac{m}{1 - m} \bigr)^{\sigma}  \frac{(\frac{m}{1 - m}c_{0} + a_{t})^{(1 - \sigma)}}{1 - \sigma}
 \end{equation*}
@@
 
-I solve the retirees problem using backward induction. At age 110 there is certainty of death so any leftover assets
+First, I discretise the state space. I create a grid from 500 to 50,000 incrementing by 500 for income and 1000 to 
+500,000 incrementing by 1000 for financial assets. I solve the retirees problem using backward induction. At age 110 there is certainty of death so any leftover assets
 are carried over into the next period and bequested. This means that the value at the end of the final period is
@@
 
-\subsection{Empirical models}
+In retirees first year of retirement I allow them to choose to annuitise some of their wealth. In practical terms 
+this is moving down the asset grid but up the income grid and seeing if the value of being in that position is better 
+than where the individual is currently. To calculate this trade-off I calculate the annual annuity payment that follows 
+from a given annuity cost. I calculate this using objective life tables from the ONS using the following equation:
+
+\begin{equation*}
+ Ann = \delta * C * \biggl[\sum_{t = Retage}^{110}\frac{1 - p_{t|Retage}}{(1 + r)^{t - Retage}}\biggr]^{-1}
+\end{equation*}
+
+Where $C$ is the one-off payment, $\delta$ is a factor that controls the 'money's worth' of annuity and $p_{t|Retage}$ 
+is the probability of death at age $t$ conditional on being age $Retage$. So individuals can move C on the asset grid 
+for gaining Ann on the income grid for the rest of their lives.
+
+\section{Empirical models}
 
@@
 had consumption data in 2014 and 2016 I include both values. An individual is considered not treated if they retire before 2013 and after 2011.
+One benefit of using ELSA is that it includes data on pension type for individuals who are working. Therefore, I can differentiate 
+between individuals who have a defined contribution pension and those who have a defined benefit contribution. To account for the fact
+that the reform only impacted individuals who had accumulated defined contribution pension pots I interact the treatment dummy with 
+an indicator variable signalling whether the individual had ever held a DC pension pot.
 
 The estimating equation is therefore
+
+\begin{equation*}
+  Cons_{it} =  \gamma X_{it} + \beta PostReformDC_{it} + \epsilon
+\end{equation*}
+
+And $X_{it}$ is a set of controls including financial wealth etc
+
+The ta
 

### 2023-07-19 Julia/0__functions.jl

@@ -5,7 +5,15 @@ using Plots
 using JSON3
 
 
-function get_death_prob_function()
+# I will not change these so they can be in functions
+# Set parameters
+const β = 0.97
+const r = 1 / β - 1
+const σ = 2.5
+const terminal_age = 110
+
+
+function GetObjDeathProbMatrices()
     # this function returns a function that gives death probs
 
     # First get some meta info
@@ -35,10 +43,10 @@ function get_death_prob_function()
         top_year=max_year, min_age=min_age)
 end
 
-death_prob_matrices = get_death_prob_function()
+death_prob_matrices = GetObjDeathProbMatrices()
 
 # ONS life tables
-function death_probabilities(; gender, year, age)
+function GetObjectiveDeathProbs(; gender, year, age)
     if gender == "male"
         data = death_prob_matrices.male
     elseif gender == "female"
@@ -62,6 +70,22 @@ function death_probabilities(; gender, year, age)
     return death_probs
 end
 
+
+
+# --------- Read in subjective life expectancies ---------------
+# Have to set column types
+death_column_types = Dict(:id_wave => String, :age => Int128, :death_prob => Float64)
+
+subjective_death_df = CSV.File("../data/ELSA/subjective_tables/subjective_death_probs_for_julia.csv",
+    types=death_column_types) |>
+                      DataFrame
+
+# Function takes an id-wave and returns the subjective death probs
+GetSubjectiveDeathProbs = function (df = subjective_death_df; id_wave_to_get = id_wave_to_get)
+    df[df.id_wave.==id_wave_to_get, :][!, :death_prob]
+end
+
+
 # utility function
 function utility(c)
     if c > 0.0001
@@ -83,3 +107,196 @@ bequest_lock(b; c0=5000, m=0.96) = (m / (1 - m))^σ * (((m / (1 - m)) * c0) + b)
 # asset accumulation
 # budget constraint
 bc_inc(a, a′, income) = a * (1 + r) - a′ + income
+
+
+CreateConsumptionArray = function (asset_grid=asset_grid, income_grid=income_grid)
+    consumption_array = ones(Float64, length(asset_grid), length(asset_grid), length(income_grid))
+    # Initialize a matrix to store consumption values
+
+    # Nested loop to calculate consumption values for each combination of assets
+    for z in eachindex(income_grid)
+        income = income_grid[z]
+        for x in eachindex(asset_grid)
+            cur_assets = asset_grid[x]
[…]

### 2023-07-19 Julia/1__lifecycle_with_income_function.jl

@@ -7,12 +7,7 @@ using JSON3
 
 # Ok I think I should make this into a module and export the functions 
 
-# Set parameters
-const β = 0.97
-const r = 1 / β - 1
-const σ = 2.5
-const terminal_age = 110
-
+## Set grid points
 const asset_grid_points = 500
 const asset_grid_gap = 1000
 # define asset grid
@@ -23,50 +18,51 @@ const inc_grid_points = 100
 const inc_grid_gap = 500
 const income_grid = [i * inc_grid_gap for i in 1:inc_grid_points]
 
-include("0__functions.jl")
 
+consumption_array = CreateConsumptionArray()
+utility_array = CreateUtilityArray()
 
-consumption_array = ones(Float64, asset_grid_points, asset_grid_points, inc_grid_points)
-# Initialize a matrix to store consumption values
-
-# Nested loop to calculate consumption values for each combination of assets
-for z in eachindex(income_grid)
-    income = income_grid[z]
-    for x in eachindex(asset_grid)
-        cur_assets = asset_grid[x]
-        for y in eachindex(asset_grid)
-            next_assets = asset_grid[y]
-            consumption_array[x, y, z] = bc_inc(cur_assets, next_assets, income)
-        end
-    end
-end
-
-utility_array = ones(Float64, asset_grid_points, asset_grid_points, inc_grid_points)
-# Initialize a matrix to store utility values
-
-# Loop to calculate utility values for each consumption value in the consumption matrix
-for i in eachindex(utility_array)
-    utility_array[i] = utility(consumption_array[i])
-end
-
-
-retirement_lifecycle_with_income = function (;
+RetirementLifecycleWithIncome = function (;
     bequests=false,
     age=60,
     gender="male",
     life_prob_types="objective",
-    year=2015)
+    year=2015,
+    id_wave="")
+
+    # two ways to run this function
+    # either with objective probs given by age, gender and year
+    # or subjective probs given by id_wave
+
+    if life_prob_types == "objective"
+        # set file name to save json to. 
+        json_filename = string(age, "_beq", bequests, "_", gender, "_", life_prob_types, "_", year, ".json")
+        filename_with_dir = "../data/ELSA/lifecycle_outputs/" * json_filename
+
+        death_probs = GetObjectiveDeathProbs(gender=gender, age=age, year=year)
+
+    elseif life_prob_types == "subjective"
+        dir_name = "../data/ELSA/lifecycle_outputs/id_wave/"
+        if !isdir(dir_name)
[…]

### 2023-07-19 Julia/5__bequest_functions.jl

@@ -1,0 +1,150 @@
+
+
+RetirementLifecycleWithIncomeBequests = function (;
+    bequests=true,
+    age=65,
+    gender="male",
+    life_prob_types="objective",
+    year=2015,
+    id_wave="", 
+    c0 = 20000)
+
+    # two ways to run this function
+    # either with objective probs given by age, gender and year
+    # or subjective probs given by id_wave
+
+    if life_prob_types == "objective"
+        # set file name to save json to. 
+        birth_year = year - age
+        json_filename = string(
+            "birth_year_", birth_year,"_", bequests, "_",
+            gender, "_", life_prob_types, "_", c0,".json")
+
+        dir_name = "../data/ELSA/lifecycle_outputs/beq_testing/"
+        if !isdir(dir_name)
+            mkdir(dir_name)
+        end
+
+        filename_with_dir = dir_name * json_filename
+
+        death_probs = GetObjectiveDeathProbs(gender=gender, age=age, year=year)
+
+    elseif life_prob_types == "subjective"
+        dir_name = "../data/ELSA/lifecycle_outputs/id_wave"
+        if !isdir(dir_name)
+            mkdir(dir_name)
+        end
+        json_filename = string(age, "_beq", bequests, "_", gender, "_", life_prob_types, "_", year, "_", id_wave, ".json")
+        filename_with_dir = dir_name * json_filename
+
+        death_probs = GetSubjectiveDeathProbs(id_wave_to_get=id_wave)
+
+    end
+
+
+    # Check if file exists already
+    # if isfile(filename_with_dir)
+    #     return "Already done file"
+    # end
+
+
+    ## So now for each level of income and current assets we find best consumption next period
+    terminal_age = 110
+    start_age = age
+
+    # Dimensions are the state variables
+    # Income, current assets and age
+    value_array = ones(Float64, inc_grid_points, asset_grid_points, 1+ terminal_age - start_age)
+    policy_array = ones(Float64, inc_grid_points, asset_grid_points, terminal_age - start_age)
+
+    bequest_value = [bequest_lock(assets, c0 = c0, m = 0.94) for assets in asset_grid]
+
+    # for each income level find the optimal asset policy function
+    for inc in eachindex(income_grid)
+
+        utility_matrix = utility_array[:, :, inc]
+
+        # Initialize a vector to store value functions in
+        # it is age + 1 columns long becuase the terminal value is stored in age at 111
[…]

### 2023-07-19 Julia/6__run_bequest_functions.jl

@@ -1,0 +1,161 @@
+
+
+include("0__functions.jl")
+include("1__lifecycle_with_income_function.jl")
+include("5__bequest_functions.jl")
+
+years = cat(repeat([2005], 26) , [i for i in 2006:2015], dims =1 )
+ages = cat([i for i in 50:75], repeat([50], 10), dims =1 )
+
+year_ages_to_run_bequest_model = [years ages] 
+
+
+
+for row in eachrow(year_ages_to_run_bequest_model)
+    year = row[1]
+    age = row[2]
+    RetirementLifecycleWithIncomeBequests(
+            bequests=true,
+            age=age,
+            gender="male",
+            life_prob_types="objective",
+            year=year,
+            id_wave="", 
+            c0 = 20000)
+end 
+
+
+for row in eachrow(year_ages_to_run_bequest_model)
+    year = row[1]
+    age = row[2]
+    RetirementLifecycleWithIncomeBequests(
+            bequests=true,
+            age=age,
+            gender="female",
+            life_prob_types="objective",
+            year=year,
+            id_wave="", 
+            c0 = 20000)
+end 
+
+file
+
+
+# going to quickly change the names to birth-year gender 
+# as opposed to year and age
+
+pwd()
+
+cp() # copy files
+mv() # move and rename files
+
+json_files = readdir(
+    "../Data/ELSA/lifecycle_outputs/beq_testing")
+occursin("birth_year", "safasdfasfbirth_year")
+
+ChangeJsonName = function(
+        json_file_name, 
+        folder_name = "../Data/ELSA/lifecycle_outputs/beq_testing/")
+    
+    if occursin(".json", json_file_name) 
+        if occursin("birth_year", json_file_name)
+                return "done already"
+        else
+        println(json_file_name)
+        year_regex = r"((?<=objective_)\d{4})" # first occurence 
+        age_regex = r"^\d{2}"
+
+        year_out = match(year_regex, json_file_name).match
+        age_out = match(age_regex, json_file_name).match
+        age =  parse(Int, String(age_out))
+        year = parse(Int, String(year_out))
+        birth_year = year - age  
+
+        new_string = replace(json_file_name, year_regex => s"")
+        new_string = replace(new_string, age_regex => s"")
+        new_string = string("birth_year_", birth_year)* new_string
+        new_string = replace(new_string, "__" => "_")
+        
+        mv(folder_name * json_file_name, folder_name * new_string)
+        
+        return "done"
+        end 
+    else 
[…]

### 2023-07-19 Julia/old/1__lifecycle_functions.jl

@@ -14,9 +14,9 @@ const terminal_age = 110
 const asset_grid = [i * grid_gap for i in 1:grid_points]
 
 
-death_probabilities(gender="male", year=2010, age=63)
-death_probabilities(gender="female", year=2010, age=63)
-death_probabilities(gender="female", year=2015, age=63)
+GetObjectiveDeathProbs(gender="male", year=2010, age=63)
+GetObjectiveDeathProbs(gender="female", year=2010, age=63)
+GetObjectiveDeathProbs(gender="female", year=2015, age=63)
 
 
 
@@ -57,7 +57,7 @@ function LifeCycleSolve(
 
     # set death probabilities.
     if life_prob_types == "objective"
-        death_probs = death_probabilities(gender=gender, age=55, year=year)
+        death_probs = GetObjectiveDeathProbs(gender=gender, age=55, year=year)
     elseif life_prob_types == "subjective"
         death_probs = "" # come and define later
     end
@@ -195,7 +195,7 @@ end
 
 # This is starting wealth. For annuitisation we want to change this depending 
 # on how large the annuitisation is
-function asset_path_function(opt_policy_function, start_point)
+function AssetPathFunction(opt_policy_function, start_point)
 
     # Return asset paths given a set of optimal policy functions
     # and a starting value of wealth

### 2023-07-19 Julia/uk_rsp/src/uk_rsp.jl

@@ -5,7 +5,7 @@ using CSV, DataFrames, Plots
 include("death_probs.jl")
 include("lifecycle_solve.jl")
 
-export LifeCycleSolve, asset_path_function
+export LifeCycleSolve, AssetPathFunction
 
 end # module uk_rsp
 

### 2023-07-19 R/13__ELSA_eol.R

@@ -452,5 +452,5 @@ sub_life_dt[, expected_age := age + years_into_future]
 # ok great we now have subjective done
 fwrite(
     sub_life_dt,
-    "../../data/ELSA/subjective_tables/objective_survival_probs.csv"
+    "../../data/ELSA/subjective_tables/subjective_survival_probs.csv"
 )

### 2023-07-19 R/18__regression_with_lifecycle.R

@@ -1,0 +1,51 @@
+library(data.table)
+library(dplyr)
+library(ggplot2)
+library(jsonlite)
+
+
+# So I take the individuals who are untreated
+# i.e. early
+
+# estimate their bequest motive so that it fits the rate of 
+# annuitisation with forced annuitisation of DC pensions
+# then run regression on the difference as we do in real life
+
+# first read objective no bequests
+# need to price annuities
+# one annuity price for everyone
+# based on annuitant lifetables
+#   i should check whether they are predictions or if they are what actually happened
+
+# ------------ read in jsons ------------------
+
+## Ok now lets read in and see the rate of annuitisation we would expect
+dir("../../data/ELSA/lifecycle_outputs/", full.names = T)
+
+
+test = read_json("../../data/ELSA/lifecycle_outputs/80_beqfalse_male_objective_2011.json")
+
+# quite slow to read 
+# it was much faster in julia
+# perhaps i should do this kind of analysis in julia
+# then output a df and analyse that in R. 
+# first we need to compare asset paths with and without forced annuitisation 
+
+
+class(test)
+
+names(test)
+class(test$pol_func)
+names(test$pol_func) %>% head
+
+# 100 income point
+# 300 asset points
+# 110 - age , age points
+age_indexes = 110 - test$age
+
+
+test$pol_func %>% length()
+
+
+array(sapply(seq(5, 15, 5), function(i) vec[]), dim = c(5, 2, 3)) 
+

### 2023-07-22 Texfiles/Draft1.tex
Reorganisation of R

@@
 \usepackage{enumerate}
+\usepackage{booktabs}
 \setlength{\parindent}{0pt}
@@
 insure against the risk of late death \cite{yaari_65}. However, in developed countries rates of annuitization are far below the
-levels that theory predicts. In this paper I test two competing hypotheses for the annuity problem: bequests and
-pessimistic life expectancy. In most developed countries, individuals are forced to contribute to either a defined
-benefit or defined contribution pension. A defined benefit pension accumulates throughout your job tenure as
+levels that theory predicts.
 Under the coalition government in the UK the law regarding the use of private defined
 contribution pensions changed. Individuals were no longer forced to annuitise their pension pots and could access
-them in a variety of ways such as lump sum withdrawals or income drawdown and subsequently the number of annuities
+them in a variety of ways such as a lump sum withdrawal on retirement or income drawdown and subsequently the number of annuities
 sold in the UK dropped precipitously.
 
-Depending on the reason for the lack of annuitization in the UK, the consumption response of retirees to the pension
+I first use the policy reform as a discontintuity to measure the impact of forced annutisation on the consumption
+of individuals in the first few years of retirement. Given that the reform was implemented suddenly and without
+advanced notice before the Spring 2014 budget I claim that individuals with a retirement year in 2015, 2016 or 2017
+are otherwise similar to those with a retirement year in 2011, 2012 and 2013 but the do not need to annuitise their
+defined contribution pension pots.
+
+In this paper I test two competing hypotheses for the annuity problem: bequests and
+pessimistic life expectancy. Depending on the reason for the lack of annuitization in the UK, the consumption response of retirees to the pension
 reform will differ. If individuals do not annuitise because of pessimistic life expectancy I will show that their
@@
 retirees.
-
-The so called "pensions freedom act" received Royal Asset in December 2014 marked the end of a series
-of pension reforms carried out by the coalition government between 2010 and 2015. The reform was announced
-in the Spring budget and made it possible to withdraw money from a private pension pot subject to the
-marginal rate of income tax that an individual faced. In the June 2010 budget the government made a first
-reform to the annuitization rules, creating an minimum income requirement above which individuals would not need to
-annuitise more \cite{finance_act_hmt_2011}. However, this was set at £20,000 and therefore few individuals
-were eligible. The minimum income requirement was scrapped in the 2014 bill finally eliminating the
-compulsory annuities market. The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
-Using data from the Association of British Insurers they show that annuity demand dropped by 75\% from
-its maximum.
 
@@
 
-\section{Data}
+\section{Data and Policy reform}
+
+\subsection{Data}
+
 The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA repeatedly interviews
@@
 These track the sale of different financial products overtime including data on annuity sales.
+
 % Plot showing the drop of in annuity sales
@@
 
-\section{Methodology}
+\input{tables/r_summary_stats.tex}
+
+\subsection{Policy reform}
+
+\textbf{Add some other background to pensions in the UK. What does the Blundell etc have
+  to say}
+
+Successive governments in the UK have attempted to reform both the public and private pension system. Prior
+to 1987 participation in private schemes was limited to employees of firms that had offered one, and there
+were few alternatives to the public state pension or a define benefit scheme that a public sector employee
+would offer. \textbf{Were there pension reforms in 1990s and 00s}
+
+Announced in the Spring budge of 2014, the so called "pensions freedom act" received Royal Asset in December 2014 marked the end of a series
+of pension reforms carried out by the coalition government between 2010 and 2015. The reform was announced
+in the Spring budget and made it possible to withdraw money from a private pension pot subject to the
+marginal rate of income tax that an individual faced. In the June 2010 budget the government made a first
+reform to the annuitization rules, creating an minimum income requirement above which individuals would not need to
+annuitise more \cite{finance_act_hmt_2011}. However, this was set at £20,000 and therefore few individuals
+were eligible. The minimum income requirement was scrapped in the 2014 bill finally eliminating the
+compulsory annuities market. The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
+Using data from the Association of British Insurers they show that annuity demand dropped by 75\% from
+its maximum.
+
+After the announcement in spring 2014 sales of annuities dropped massively and income drawdown products increased.
+However, pension pots could also be taken out in one go and this became a much more common option for many people.
+
+Figure \ref{fig:annovertime} shows purchases of annuities overtime and demonstrates the sharp decrease
+in purchases that happened from 2014 to 2015. Likewise there was an increase in the number of pensions that were
+being accessed using an income drawdown product, but these do not fully account for the drop in annuitisation. So,
+either they are withdrawn at once or they are left alone.
+
+Figure \ref{fig:ann2122} shows the current distribution of how pension pots are first accessed at retirement.
+In 2021, 196,736 pots were fully withdrawn at retirement, accounting for over 50\% of pots, prior to the policy
+reform this was not the case and most defined contribution pensions were accessed through annuities.
+
+\begin{figure}[h]
+  \caption{Pension pots accessed}
+  \centering
+  \includegraphics[width=0.7\columnwidth]{figures/annuity_overtime.pdf}
+  \label{fig:annovertime}
+\end{figure}
+
+\begin{figure}[h]
+  \caption{How pension pots are accessed}
+  \centering
+  \includegraphics[width=0.7\columnwidth]{figures/annuity_pot_sizes.pdf}
+  \label{fig:ann2122}
+\end{figure}
+
+\subsection{Covariate Balance}
+
+I regress demographic and financial characteristics of individuals on year of birth and the treatment dummy.
+We can then see if the treatment groups differ on key characteristics such as financial wealth or retirement age.
+For a regression discontinuity to be valid we need the treatment and control groups to be similar along
+all other characteristics apart from the treatment variable.
+
+In particular I run:
+\begin{equation*}
+  Y_{i} = \alpha + \beta Post_{i} + \gamma YOB + \kappa (Post_{i} \times YOB_{i}) + \epsilon_{i}
+\end{equation*}
+
+\input{tables/cov_balance.tex}
+Table \ref{tab:cov_balance} shows these results. We can see ...
+
+Another threat to validity is manipulation of the running variable. This is probably quite a threat in our case
+as individuals could delay retirement and/or the purchase of an annuity.
+
+\section{Methodology and Econometric Strategy}
 
@@
 \begin{equation*}
-  c_{t} =a_{t}(1 +r) -  a_{t+1} + y
+  c_{t} =a_{
+  t}(1 +r) -  a_{t+1} + y
 \end{equation*}
@@
 
-The ta
+I also run the regression with a variety of consumption variables.
+
+I also simulate the expected change in consumption given forced annuitisation and no annuitisation.
+The benefit of simulating an individuals decision is that we can directly compare what they would have
+consumed with what they actually consumed.
 

### 2023-07-22 Texfiles/Notes.tex
Reorganisation of R

@@
 \subsection{ELSA notes}
+\subsection{\textbf{20-07}}
+\begin{itemize}
+  \item What other ELSA variables to use?
+  \item What will home ownership change?
+  \item How to deal with top wealth deciles? - at the moment they are excluded because of the grid.
+  \item Include other sources of income i.e. Got Dc pension
+  \item Run model just with pension wealth and not financial wealth
+  \item We could probably generate non annuitisation by adding some fixed cost.
+  \item OK, the financial wealth variable I use is quite comprehensive including all non-housing
+        wealth which is good.
+  \item Deflator?
+  \item
+
+  \item Pension wealth is what I need to look at now \textbf{DONE}
+  \item What
+  \item Do we look at just individuals? Or just households or both?
+\end{itemize}
 

### 2023-07-22 Julia/0__functions.jl

@@ -178,7 +178,7 @@ function AssetPathFunction(opt_policy_function, asset_start_point, income_grid_p
 
     # set intial assets
     asset_path[1] = asset_start_point
-    
+
     # Get optimal asset paths from optimal policy matrix
     for age in 1:(number_years)
         asset_path[age+1] = opt_policy_function[:, age][asset_path[age]]
@@ -291,11 +291,11 @@ function FindClosestGridPoint(x, grid_gap, num_grid_point)
     # Assume for now that no one actually is at 0
     if grid_point > 0
         return grid_point
-        
+
         # Also stop individuals considering points outside of the grid
     elseif grid_point > num_grid_point
         num_grid_point
-    else 
+    else
         return 1
     end
 end

### 2023-07-22 Julia/6__run_bequest_functions.jl

@@ -4,88 +4,54 @@ include("0__functions.jl")
 include("1__lifecycle_with_income_function.jl")
 include("5__bequest_functions.jl")
 
-years = cat(repeat([2005], 26) , [i for i in 2006:2015], dims =1 )
-ages = cat([i for i in 50:75], repeat([50], 10), dims =1 )
+years = cat(repeat([2005], 26), [i for i in 2006:2015], dims=1)
+ages = cat([i for i in 50:75], repeat([50], 10), dims=1)
 
-year_ages_to_run_bequest_model = [years ages] 
+year_ages_to_run_bequest_model = [years ages]
 
+birth_years = year_ages_to_run_bequest_model[:, 1] - year_ages_to_run_bequest_model[:, 2]
+
+birth_years
+
+println(year_ages_to_run_bequest_model)
+
+for row in eachrow(year_ages_to_run_bequest_model)
+    year = row[1]
+    age = row[2]
+    print(row)
+    # RetirementLifecycleWithIncomeBequests(
+    #         bequests=true,
+    #         age=age,
+    #         gender="male",
+    #         life_prob_types="objective",
+    #         year=year,
+    #         id_wave="", 
+    #         c0 = 20000)
+end
 
 
 for row in eachrow(year_ages_to_run_bequest_model)
     year = row[1]
     age = row[2]
     RetirementLifecycleWithIncomeBequests(
-            bequests=true,
-            age=age,
-            gender="male",
-            life_prob_types="objective",
-            year=year,
-            id_wave="", 
-            c0 = 20000)
-end 
-
-
-for row in eachrow(year_ages_to_run_bequest_model)
-    year = row[1]
-    age = row[2]
-    RetirementLifecycleWithIncomeBequests(
-            bequests=true,
-            age=age,
-            gender="female",
-            life_prob_types="objective",
-            year=year,
-            id_wave="", 
-            c0 = 20000)
-end 
-
-file
+        bequests=true,
+        age=age,
+        gender="female",
+        life_prob_types="objective",
+        year=year,
+        id_wave="",
+        c0=20000)
+end
 
 
 # going to quickly change the names to birth-year gender 
 # as opposed to year and age
 
-pwd()
+# pwd()
 
-cp() # copy files
-mv() # move and rename files
+# cp() # copy files
+# mv() # move and rename files
 
-json_files = readdir(
-    "../Data/ELSA/lifecycle_outputs/beq_testing")
[…]

### 2023-07-22 Julia/7__analyse_more_models.jl

@@ -1,0 +1,116 @@
+# analyse bequest runs
+
+# For each year-age-pension-wealth lets find if someone should annuitise
+
+include("0__functions.jl")
+include("1__lifecycle_with_income_function.jl")
+
+# first get data
+col_types = Dict(:id_wave => String, :age => Int128, :year => Int128,
+    :ever_dc_pen => String, :ever_db_pen => String, :public_pension => Float64,
+    :fin_wealth => Float64)
+
+idwave_df = CSV.File("../data/ELSA/elsa_to_use/for_julia.csv", types=col_types) |>
+            DataFrame
+idwave_df = idwave_df[Bool.(.!ismissing.(idwave_df.fin_wealth)) .& Bool.(.!ismissing.(idwave_df.public_pension)), :]
+      
+idwave_df.pen_grid = FindClosestGridPoint.(
+    idwave_df.public_pension, inc_grid_gap, inc_grid_points)
+
+idwave_df.ass_grid = FindClosestGridPoint.(
+    idwave_df.fin_wealth, asset_grid_gap, asset_grid_points)
+
+    sum(idwave_df.pen_grid .<= 100 )
+    sum(idwave_df.ass_grid. <=500  .& idwave_df.pen_grid .<= 100)
+
+
+good_assets = idwave_df.ass_grid .<=500   
+good_inc = idwave_df.pen_grid .<= 100
+
+# filter too wealthy and too high income
+idwave_df = idwave_df[good_assets .& good_inc, :]
+ 
+
+optimal_asset = []
+
+for row in eachrow(idwave_df)
+    println([row["pen_grid"] row["ass_grid"]])
+    output_val = ToAnnuitiseOrNot(
+        row["year"],
+        row["age"], 
+        row["gender"],
+        row["pen_grid"], 
+        row["ass_grid"])
+
+    push!(optimal_asset, output_val)
+end 
+
+
+
+sum(optimal_asset .==  "do not annuitise")
+
+not_worked_for = optimal_asset[optimal_asset .== "too poor to annuitise"]
+
+worked_for =  optimal_asset[optimal_asset .!= "too poor to annuitise"]
+
+ids_worked = idwave_df.id_wave[1:length(optimal_asset)][optimal_asset .!= "too poor to annuitise"]
+
+worked_dt = DataFrame(worked_for)
+
+worked_dt.id_wave = ids_worked
+
+idwave_df[:, "year"][1]
+
+joined_df = innerjoin(worked_dt, idwave_df, on = :id_wave)
+
[…]

### 2023-07-22 R/13__regression_with_lifecycle.R

@@ -7,7 +7,7 @@ library(jsonlite)
 # So I take the individuals who are untreated
 # i.e. early
 
-# estimate their bequest motive so that it fits the rate of 
+# estimate their bequest motive so that it fits the rate of
 # annuitisation with forced annuitisation of DC pensions
 # then run regression on the difference as we do in real life
 
@@ -23,29 +23,28 @@ library(jsonlite)
 dir("../../data/ELSA/lifecycle_outputs/", full.names = T)
 
 
-test = read_json("../../data/ELSA/lifecycle_outputs/80_beqfalse_male_objective_2011.json")
+test <- read_json("../../data/ELSA/lifecycle_outputs/80_beqfalse_male_objective_2011.json")
 
-# quite slow to read 
+# quite slow to read
 # it was much faster in julia
 # perhaps i should do this kind of analysis in julia
-# then output a df and analyse that in R. 
-# first we need to compare asset paths with and without forced annuitisation 
+# then output a df and analyse that in R.
+# first we need to compare asset paths with and without forced annuitisation
 
 
 class(test)
 
 names(test)
 class(test$pol_func)
-names(test$pol_func) %>% head
+names(test$pol_func) %>% head()
 
 # 100 income point
 # 300 asset points
 # 110 - age , age points
-age_indexes = 110 - test$age
+age_indexes <- 110 - test$age
 
 
 test$pol_func %>% length()
 
 
-array(sapply(seq(5, 15, 5), function(i) vec[]), dim = c(5, 2, 3)) 
-
+array(sapply(seq(5, 15, 5), function(i) vec[]), dim = c(5, 2, 3))

### 2023-07-22 R/2__ELSA_pension.R

@@ -1,0 +1,182 @@
+# Look at pension wealth for those who retired
+library(data.table)
+library(dplyr)
+library(stringr)
+library(readxl)
+
+source("__ELSA_functions.R")
+
+search_files <- function(string, ...) {
+    dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+        pattern = "wave",
+        full.names = TRUE
+    ) %>%
+        grep(string, ., value = T, ...)
+}
+
+# ---------------- Get Pension Wealth files ----------------
+pension_wealth_files <- search_files("(3|4|5)_pension_wealth") %>% 
+    lapply(fread) %>% 
+    Reduce(function(x, y) full_join(x, y,  by = "idauniq"), .)  %>% # bind together
+    setnames(str_to_lower)
+
+pw_long <- melt(
+    pension_wealth_files,
+    id.vars = c("idauniq", "file_wave"),
+    measure = patterns(
+        "totpenw_\\d{2}", "inreceipt_\\d{2}", "currentdc_\\d{2}",
+        "currentdb_\\d{2}", "retaineddc_\\d{2}", "retaineddb_\\d{2}",
+        "stpenw_\\d{2}", "bsp_\\d{2}", "addpen_\\d{2}", "pripenw_\\d{2}"),
+    value.name = c(
+        "total_pension_wealth","in_receipt",
+        "current_dc", "current_db", 
+        "retained_dc", "retained_db",
+        "state_pw", "bsp_wealth",
+        "add_pw", "private_pw"
+    ), 
+    variable.name = "file_index"
+)
+
+pw_long[, wave  := fcase(
+    file_index == 1, 3,
+    file_index == 2, 4, 
+    file_index == 3, 5)]
+
+# really we only use individuals from waves 3, 4, 5 
+# check file 1 pension wealth files are the same
+
+# Assume pension pots accrue at a rate of 3% a year
+# find last size of dc pension then
+# for every year after that just increase by three percent 
+accural_rate = 1.03
+
+
+id_date <- fread(
+    "../../data/ELSA/elsa_to_use/harmonised_data_long.csv",
+    select = c("idauniq", "int_month_date", "wave"))
+
+id_date <- id_date[wave == 5][, int_year := year(int_month_date)]
+
+wave_5_dt <- pw_long[wave == 5][, .(idauniq, init_total_dc = retained_dc + current_dc)]
+
[…]

### 2023-07-22 R/7__ELSA_pension.r

@@ -1,103 +1,0 @@
-# Look at pension wealth for those who retired
-
-library(data.table)
-library(dplyr)
-library(stringr)
-library(readxl)
-
-source("__ELSA_functions.R")
-
-pension_grid <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-    pattern = "wave",
-    full.names = TRUE
-) %>%
-    grep("pension(_grid|grid)", ., value = T) %>%
-    lapply(ReadAndSetWave)
-
-#  %>%
-# rbindlist(fill = T, use.names = T)
-
-mapply(search_names, pension_grid, "wpapu", ignore.case = TRUE)
-
-pension_grid
-
-head(pension_grid[[1]])
-
-pension_grid1 <- pension_grid[[1]]
-
-head(pension_grid1)
-
-# oh maybe this is when the pension mapping grid becomes useful.
-# get pension grid lookups
-grid_lookups <- read_excel(
-    "../../data/ELSA/elsa_unziped/UKDA-5050-tab/mrdoc/excel/5050_wave_2_pension_grid_corresponding_variables.xls"
-)
-
-names(grid_lookups)
-head(grid_lookups)
-# lets have a look at the excel files.
-# pension variables seem to be different in each
-
-search_files <- function(string, ...) {
-    dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-        pattern = "wave",
-        full.names = TRUE
-    ) %>%
-        grep(string, ., value = T, ...)
-}
-
-dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-    pattern = "wave",
-    full.names = TRUE
-) %>%
-    grep("pension(_wealth|wealth)", ., value = T)
-# Only up to wave 5
-
-search_files("fin")
-
-grid_5 <- fread("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_5_pension_wealth.tab")
-
-fin_7 <- fread("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_7_financial_derived_variables.tab")
-
-names(grid_5)[names(grid_5) %in% names(fin_7)]
-names(grid_5)
-names(fin_7)
-head(grid_5)
-
-fin_6 <- fread
-
-fread(pension_files[[1]]) %>%
-    names() %>%
-    grep("", ., value = T)
-
[…]

### 2023-07-22 R/9__ONS_lifeexpec.R

@@ -213,8 +213,7 @@ create_death_prob_matrix <- function(gender = "Males") {
             names_to = "birth_year"
         )
 
-    # birth years needed 1935 to 1955
-    # want death rates
+    # want year by year death rate
 
     ons_surv <- ons_surv %>%
         mutate(birth_year = as.numeric(birth_year)) %>%
@@ -245,6 +244,57 @@ create_death_prob_matrix <- function(gender = "Males") {
         ons_surv_matrix,
         str_to_lower(paste0("../../data/ONS/", gender, "_transition_probs.csv"))
     )
+    to_ret <- ons_surv_matrix %>%
+        mutate(gender = gender)
 }
 
-sapply(c("Males", "Females"), create_death_prob_matrix)
+data_for_life_exp <- lapply(c("Males", "Females"), create_death_prob_matrix) %>%
+    rbindlist()
+
+
+# Lets make an average life expectancy
+# Is it sum of prob*age?
+# I think so
+
+life_exps <- data_for_life_exp %>%
+    melt(c("age", "gender"),
+        variable.name = "birth_year",
+        value.name = "prob_of_death"
+    )
+
+gender <- "Males"
+
+life_exps %>%
+    filter(birth_year == 1920 & !is.na(prob_of_death))
+
+
+sheet_to_get <- paste0(gender, " cohort lx")
+
+ons_surv <- read_excel(
+    "../../data/ONS/ONS life tables 1841 onwards.xls",
+    sheet = sheet_to_get,
+    skip = 8
+) %>%
+    # rename_with(Vectorize(change_numeric_names)) %>%
+    rename(age = `age (years)`) %>%
+    tidyr::pivot_longer(
+        -age,
+        values_to = "alive",
+        names_to = "birth_year"
+    )
+
+# want year by year death rate
+
+ons_surv <- ons_surv %>%
+    mutate(birth_year = as.numeric(birth_year)) %>%
+    group_by(birth_year) %>%
+    arrange(age) %>%
+    mutate(deaths = lead(lag(alive) - alive)) %>% # alive yesterday minus alive today, shifted back to today
+    mutate(death_rate = deaths / alive) %>%
+    ungroup() %>%
+    filter(birth_year >= 1920, birth_year <= 1975) %>%
+    mutate(year = age + birth_year) %>%
+    filter(year >= 2005) %>%
+    filter(age >= 50, age <= 110) %>%
+    arrange(-birth_year)
+# prob of death??

### 2023-07-22 Texfiles/Draft1.tex
Diss edits and code updates

@@
 \usepackage{booktabs}
+\usepackage{pdflscape}
 \setlength{\parindent}{0pt}
@@
 individuals over the age of 50 and asks them a range of questions relating to their income and wealth as well as expectations about
-the future. Importantly it also includes detailed information pensions including the type of pension that an individual holds when
-working, thus I am able to distinguish between individuals who have defined benefit and defined contribution pensions. For ease of access
-I use Harmonized ELSA \footnote{"This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
+the future. One benefit of using ELSA is that it includes data on pension type for individuals who are working. Therefore,
+I can differentiate between individuals who have a defined contribution pension and those who have a defined benefit contribution.
+Importantly it also includes information on pension size that the Institute for Fiscal Studies calculated, this is only available
+up until wave 5 at which point I use a real return of 3\% to predict forward the pension wealth variable until retirement.
+ELSA also includes a measure of all non-housing financial wealth which I use in some specifications because of these issues with
+pension wealth. Because of slight differences between years in some of the ELSA questions I use 'Harmonized ELSA'
+\footnote{"This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
   July 2021 developed by the Gateway to Global Aging Data. The development of the Harmonized ELSA was funded by the National
@@
   please refer to https://g2aging.org/.”} which ensures that variables are comparable across waves. Since this only includes a
-subset of the questions in ELSA I also supplement it with variables taken directly from the data.
+subset of the questions in ELSA I also supplement it with variables taken directly from the data such as questions that deal
+with life expectancy.
 
-ELSA also includes questions on expenditures. In particular individuals are asked how much they consume on a range of broad categories.
-
-% Include a summary stats table here of the individuals selected for the regression
+ELSA also includes questions on expenditures. In particular individuals are asked how much they consume on a range of broad
+categories including in-house food consumption, out-of-house food consumption, leisure consumption, clothes consumption and
+consumption on utility bills and rent. I use also use two questions on life expectancy that I discuss in more detail below.
 
@@
 a given age since this is what is used in the life cycle simulations.
-%  include a table showing death probs
 
@@
 
-\input{tables/r_summary_stats.tex}
-
 \subsection{Policy reform}
-
-\textbf{Add some other background to pensions in the UK. What does the Blundell etc have
-  to say}
 
@@
 to 1987 participation in private schemes was limited to employees of firms that had offered one, and there
-were few alternatives to the public state pension or a define benefit scheme that a public sector employee
-would offer. \textbf{Were there pension reforms in 1990s and 00s}
+were few alternatives to the public state pension or defined benefit scheme that a public sector employee
+would offer. From 1989, individuals in the UK were allowed to open tax advantaged self-invested personal pensions
+alongside any pension their employer offered. The largest reform prior to 2014 came in the 2004 Finance Act which
+rationalised taxation rates on different types of pensions. However, defined contribution pension pots could still
+only be used to buy an annuity after taking out a maximum of 25\% as a tax free lump sum withdrawal.
 
@@
 
-Figure \ref{fig:ann2122} shows the current distribution of how pension pots are first accessed at retirement.
-In 2021, 196,736 pots were fully withdrawn at retirement, accounting for over 50\% of pots, prior to the policy
-reform this was not the case and most defined contribution pensions were accessed through annuities.
-
 \begin{figure}[h]
@@
 \end{figure}
+
+Figure \ref{fig:ann2122} shows the current distribution of how pension pots are first accessed at retirement.
+In 2021, 196,736 pots were fully withdrawn at retirement, accounting for over 50\% of pots, prior to the policy
+reform this was not the case and most defined contribution pensions were accessed through annuities.
 
@@
 
+Also of specific interest to the annuity market was the European Union's
+'Equal Treatment in Goods and Services Directive' of 2004. This prohibited discrimination
+in the provision and cost of good and services based on sex. Up until 2011 there was a clause that
+stated insurers were allowed to charge different premiums if it was based on evidence that sex is
+correlated with different amounts of risk. However, in March 2011 the European Court of Justice
+ruled that insurers were not allowed to charge different amounts and gave them until December 2012
+to implement the ruling. This change meant that annuity products could no longer be priced differently
+for men and women in the UK. However, given this change was implemented two years before the pension
+freedoms act, I can still identify the impact of the pensions freedom act on early retirement consumption.
+
 \subsection{Covariate Balance}
 
-I regress demographic and financial characteristics of individuals on year of birth and the treatment dummy.
+Table \ref{tab:sum_stats} shows various summary statistics for each variable of interest from ELSA and the ONS.
+First note that the second retirement group is small with just 728 non missing observations for gender as opposed
+to 941 individuals in the control group. The control group are less female, retired at a slightly younger age and expected to
+retire slightly earlier. This could be because this period was affected by the increase in the state pension age for
+women from age 60 in 2010 to age 65 in 2018. Since this change happened gradually and occured over the whole period
+I do not expect it to influence the results. The DifferenceAge row tracks the difference between expected and actual
+pension age.
+
+As required, retirement year is between 2011 and 2013 for the treatment group and 2015 and 2017 for the control group.
+However, as discussed above their interview date is after or the same as the retirement year because we are tracking
+consumption in retirement.
+
+The treatment group has higher financial wealth with a median of £65,000 at time of interview as opposed to £55,000
+in the control group. Likewise, the second group are more likely to have held at defined contribution pension at some point
+and have much more money in them. Both groups are equally likely to have a defined benefit pension with roughly 44\% of
+individuals across both samples having a defined benefit pension. Individuals are much more likely to have a DB pension
+than a DC pension which reflects the fact that retirees now are still more likely to have a DB pension although the number
+with DC pensions are growing. Home ownership is roughly equal across groups although the second group have slightly more
+housing wealth.
+
+Both groups are objectively similarly long-lived according when using ONS life tables to calculate life expectancy
+based on gender, age and the year the interview was carried out in. Subjective life expectancies are also similar
+across groups, with individuals expecting to live another 21 years as opposed to the 24 that the ONS would expect
+them to live. \textbf{Maybe add in prob of survival to age 85 or 90 or something}
+\textbf{Also add in an indicator for living alone or living with someone else}
+
+Unfortunately, some consumption data is missing for some individuals. The food categories have the least missing data
+and the leisure consumption category has the most missing data. \textbf{Check why total consumption adds up to so much
+  more than all the categories added together? Do they include rent and housing in consumption?}
+
+Overall, the groups do appear to be slightly different with the treatment group being richer and slightly more likely
+to have DC pension wealth. On key demographic characteristics though the groups are similar, the difference in retirement
+age between the two groups can be mostly explained by the difference in expected retirement age meaning that individuals
+have not on-mass decided to delay retirement and avoid annuitisation.
+
+\begin{landscape}
+
+  \input{tables/r_summary_stats.tex}
+
+\end{landscape}
+
+As a further test for balance I regress demographic and financial characteristics of individuals on year of birth and the treatment dummy.
 We can then see if the treatment groups differ on key characteristics such as financial wealth or retirement age.
@@
 \begin{equation*}
-  Y_{i} = \alpha + \beta Post_{i} + \gamma YOB + \kappa (Post_{i} \times YOB_{i}) + \epsilon_{i}
+  Y_{i} = \alpha + \beta Post_{i} + \gamma YOB_{i} + \kappa (Post_{i} \times YOB_{i}) + \epsilon_{i}
 \end{equation*}
@@
 \input{tables/cov_balance.tex}
-Table \ref{tab:cov_balance} shows these results. We can see ...
+Table \ref{tab:cov_balance} shows these results. We can see that being in the treatment group, con
 
-Another threat to validity is manipulation of the running variable. This is probably quite a threat in our case
-as individuals could delay retirement and/or the purchase of an annuity.
+A key threat to validity is manipulation of the running variable. This is probably quite a threat in our case
+as individuals could delay retirement and/or the purchase of an annuity until after the reform. To check whether
+this is occuring we see whether the difference between real retirement age and the first expected retirement age
+recorded in the data are different for those retiring before 2014 compared to those retiring after 2014. As shown
+above in Table \ref{tab:sum_stats} this not the case. However, since we do not have data on the exact purchase
+date of the annuity we cannot track whether annuity purchases were delayed because of this. However, we do still
+relatively stable annuity demand before the policy reform. If individuals delayed annuity purchases in expectation
+of the reform we would see steady declines in purchases before the reform was announced.
+
+\section{Empirical models}
+
+In this section I outline the key empirical models I run with both the simulated consumption data from the
+lifecycle models and with the real data from ELSA. I then see which lifecycle model better fits the consumption
+response that happened as a result of the pension reform.
+
+I use a regression discontinuity design where I compare the consumption of recent retirees after the policy reform
+to recent retirees before the policy reform. The key assumption implicit in regression discontinuities is
+that nothing else changes at the time of the jump apart from the policy of interest. And that the policy occurs
+without individuals predicting it and changing their behaviour. As I have argued above the demographic information in the
+data suggest that individuals did not delay retirement and that there was no delay in annuitisation as shown by the quick
+and sudden decline in annuity purchases. Moreover, anecdotal evidence from media and business sources at the time of the
+announcement were all surprised by the policy changes, with some financial planners even complaining that they had not been
+briefed.
+
+Retirement year is the running variable and individuals are treated if retirement year is greater than 2014 and less than
+or equal to 2017. I use consumption data of individuals up to 2 years into retirement so that the sample size is larger.
+So if someone retired in 2015 and had consumption data in 2015 and 2017 I include both values. An individual is considered
+not treated if they retire before 2013 and after 2011.  To account for the fact
+that the reform only impacted individuals who had accumulated defined contribution pension pots I interact the treatment dummy with
+an indicator variable signalling whether the individual had ever held a DC pension pot.
+
+Because of the differences in financial characteristics between the groups I add controls for financial wealth, housing wealth
+and whether an individual owns their own home. I also run regression discontintuity with s
+
+\begin{equation*}
+  Cons_{it} =  \gamma X_{it} + \beta PostReformDC_{it} + \epsilon
+\end{equation*}
+
+And $X_{it}$ is a set of controls.
 
@@
 
-\section{Empirical models}
-
-In this section I outline the key empirical models I run with both the simulated consumption data from the
-lifecycle models and with the real data from ELSA. I then see which lifecycle model better fits the consumption
-response that happened as a result of the pension reform.
-
-I use a regression discontinuity design where I compare the consumption of recent retirees after the policy reform
-to recent retirees before the policy reform. The individuals who retired in 2014 and later were not forced to
-annuitise their defined contribution pension. The key assumption implicit in regression discontinuities is
-that nothing else changes at the time of the jump apart from the policy of interest. And that the policy occurs
-without individuals predicting it. The policy change was widely seen as a surprise by the media and financial planners,
-some of whom in fact complained that they had not been consulted enough. \textbf{How can I justify no other jumps at the jump}
-
-Retirement year is the running variable and individuals are treated if retirement year is greater than 2013 and less than 2016.
-I consumption of individuals up to 2 years into retirement so that the sample size is larger. So if someone retired in 2014 and
-had consumption data in 2014 and 2016 I include both values. An individual is considered not treated if they retire before 2013 and after 2011.
-One benefit of using ELSA is that it includes data on pension type for individuals who are working. Therefore, I can differentiate
-between individuals who have a defined contribution pension and those who have a defined benefit contribution. To account for the fact
-that the reform only impacted individuals who had accumulated defined contribution pension pots I interact the treatment dummy with
-an indicator variable signalling whether the individual had ever held a DC pension pot.
-
-The estimating equation is therefore
-
-\begin{equation*}
-  Cons_{it} =  \gamma X_{it} + \beta PostReformDC_{it} + \epsilon
-\end{equation*}
-
-And $X_{it}$ is a set of controls including financial wealth etc
-
-I also run the regression with a variety of consumption variables.
-
 I also simulate the expected change in consumption given forced annuitisation and no annuitisation.
@@
                 \cite{banks_crawford_ar_2022}
+          \item https://www.moneysavingexpert.com/savings/pension-freedom/
         \end{enumerate}

### 2023-07-22 R/10__2_summay_stats.R

@@ -1,6 +1,11 @@
 library(knitr)
 library(kableExtra)
+library(data.table)
+library(dplyr)
+library(stringr)
 
+rm(list = ls())
+source("__ELSA_functions.R")
 
 # Covariate Balance
 # need to add home ownership
@@ -10,25 +15,44 @@ library(kableExtra)
 # lets hammer out the empirical work today and tomorrow
 # and then focus on the other models on sunday
 
-rm(list = ls())
 
 if (!dir.exists("../Texfiles/tables")) {
     dir.create("../Texfiles/tables")
 }
 
-data_for_regressions <- fread("../../data/ELSA/elsa_to_use/elsa_reg_data.csv")
+data_for_regressions <- fread("../../data/ELSA/elsa_to_use/elsa_reg_data.csv") %>%
+    mutate(
+        fin_wealth = fin_wealth / 1000,
+        house_value = house_value / 1000,
+        dc_pot = dc_pot / 1000
+    )
+data_for_regressions$year_of %>% table()
+
+search_names(data_for_regressions, "year")
 
 # List of names for the summary table
 clean_names <- list(
     "Gender" = "ragender",
+    "RetirementYear" = "retirement_year",
+    "InterviewYear" = "int_year",
     "YearsSinceRetirement" = "years_since_retirement",
     "RetiredAge" = "retired_age",
     "ExpectedRetiredAge" = "first_exp_ret_age",
     "DifferenceAge" = "exp_real_ret_age",
-    "FinancialWealth" = "fin_wealth",
+    "FinancialWealth (thousands)" = "fin_wealth",
     "DCPension" = "ever_dc_pen_bin",
+    "DCValue (thousands)" = "dc_pot",
     "DBPension" = "ever_db_pen_bin",
-    "HouseValue" = "house_value"
+    "OwnsHouse" = "owns_house",
+    "HouseValue (thousands)" = "house_value",
+    "ObjectiveLifeExp" = "obj_life_exps",
+    "SubjectiveLifeExp" = "sub_life_exps",
+    "TotalConsump" = "total_monthly_consumption",
+    "FoodConsumpInHouse" = "monthly_food_in",
+    "FoodConsumpOutHouse" = "monthly_food_out",
+    "ClothingConsump" = "monthly_clothing",
+    "LeisureConsump" = "monthly_leisure",
+    "UtilityConsump" = "monthly_utility"
 )
 
 
@@ -37,8 +61,9 @@ ReturnCleanName <- function(short_name) {
 }
 
[…]

### 2023-07-22 R/7__ELSA_harmonised.R

@@ -58,33 +58,39 @@ harm[, table(radyear)]
 harm[, unique(radyear)]
 harm[, sum(!is.na(radyear))]
 
-# H1ATOTF net value of non-housing financial wealth
-# H1ATOTH value of primary wealth
 
+# Get names of variables that we want to transition from wide to long
+names_patterns_list <- list(
+    "retired" = "r\\d{1}retemp",
+    "pension_type_job1" = "r\\d{1}ptyp1_e",
+    "pension_type_job2" = "r\\d{1}ptyp2_e",
+    "pension_type_job3" = "r\\d{1}ptyp3_e",
+    "retired_age" = "r\\d{1}retage",
+    "expected_retired_age" = "r\\d{1}wretage",
+    "age_at_interview" = "r\\d{1}agey",
+    "in_wave" = "inw\\d{1}$",
+    "date_month" = "r\\d{1}iwindm",
+    "date_year" = "r\\d{1}iwindy",
+    "total_monthly_consumption" = "hh\\d{1}ctot1m",
+    "monthly_food_in" = "hh\\d{1}cfoodi",
+    "monthly_food_out" = "hh\\d{1}cfoodo1m",
+    "monthly_food_total" = "hh\\d{1}cfood1m",
+    "monthly_clothing" = "hh\\d{1}cclo1m",
+    "monthly_leisure" = "hh\\d{1}clei1m",
+    "monthly_rent" = "hh\\d{1}crent",
+    "monthly_utility" = "hh\\d{1}cutil",
+    "fin_wealth" = "h\\d{1}atotf",
+    "house_value" = "h\\d{1}atoth",
+    "owns_house" = "h\\d{1}ahown",
+    "public_pension" = "r\\d{1}ipubpen"
+)
 
 
 harm_long <- melt(
     harm,
     id.vars = c("idauniq", "pn", "ragender", "rabyear", "radyear"),
-    measure = patterns(
-        "r\\d{1}retemp", "r\\d{1}ptyp1_e",
-        "r\\d{1}ptyp2_e", "r\\d{1}ptyp3_e",
-        "r\\d{1}retage", "r\\d{1}wretage",
-        "r\\d{1}agey", "inw\\d{1}$",
-        "r\\d{1}iwindm", "r\\d{1}iwindy",
-        "hh\\d{1}ctot1m", "h\\d{1}atotf",
-        "h\\d{1}atoth", "r\\d{1}ipubpen"
-    ),
-    value.name = c(
[…]

### 2023-07-22 R/8__1_ELSA_subjective_life.R

@@ -47,14 +47,6 @@ check_files_for_variable_name("eida")
 check_files_for_variable_name("exlo")
 # life expec
 
-search_names(
-    fread(
-        "../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/wave_6_elsa_data_v2.tab",
-        nrows = 1
-    ),
-    "exlo",
-    ignore.case = TRUE
-)
 
 # These are file names
 expectation_cols <- c(
@@ -74,7 +66,7 @@ data_with_exp <- lapply(files_with_exp, ReadAndSetWave) %>%
 
 
 # ok this is the data.
-# need to do some cleaning on it
+
 # get rid of negative values and set values of 0 and 1
 # to be just above and below respectively
 
@@ -84,8 +76,8 @@ data_with_exp <- lapply(files_with_exp, ReadAndSetWave) %>%
 # can get ons life exp from the table I made earlier
 # These are the parameters that I used to select the life probs
 
-start_year <- 2005
 
+# Get objective life probs
 life_exp_list <- list(
     female = fread("../../data/ONS/Females_transition_probs.csv"),
     male = fread("../../data/ONS/Males_transition_probs.csv")
@@ -110,13 +102,19 @@ ons_life_exp <- mapply(
     rbindlist()
 ## Calculate real prob of being alive at 110 using ONS lifetables
 
+ons_life_exp %>%
+    filter(birth_year == 1975) %>%
+    filter(gender == "male") %>%
+    arrange(-age) %>%
+    mutate(prob_110 = cumprod(1 - death_rate))
+
+
 ons_p110 <- ons_life_exp %>%
     mutate(surv_prob = 1 - death_rate) %>%
     filter(!is.na(surv_prob)) %>%
     group_by(gender, birth_year) %>%
     arrange(-age) %>%
     mutate(prob110 = lag(cumprod(surv_prob))) %>% # cumulative prov of being alive
-    mutate(prob110 = fifelse(age == 110, 1, prob110)) %>%
     select(age, gender, birth_year, prob110)
 
 
@@ -147,7 +145,9 @@ joined_harm <- left_join(
 # age in question depended on the age of the individual.
 # are expectations of death in here as well?
 
+
 ready_for_nls <- joined_harm %>%
+    ungroup() %>%
     select(idauniq, exlo80, exlo90, prob110, wave, age = age_at_interview) %>%
[…]

### 2023-07-22 R/8__2_ELSA_sub_lifeexps.R

@@ -1,0 +1,30 @@
+library(data.table)
+library(dplyr)
+
+
+sub_dt <- fread("../../data/ELSA/subjective_tables/subjective_survival_probs.csv")
+
+base_pop <- 100000
+
+sub_dt[, number_alive := base_pop * prob]
+
+setorder(sub_dt, -expected_age)
+
+sub_dt[, life_years_left := cumsum(number_alive),
+    by = .(id_wave)
+]
+
+sub_dt[, life_exp := life_years_left / number_alive]
+
+sub_life_exp <- sub_dt %>%
+    filter(age == expected_age - 1) %>% # just want life expectancy at interview date
+    mutate(
+        age = expected_age - 1,
+        subjective_life_exp = life_exp
+    ) %>%
+    select(id_wave, age, subjective_life_exp)
+
+fwrite(
+    sub_life_exp,
+    "../../data/ELSA/subjective_tables/subjective_life_expectancies.csv"
+)

### 2023-07-22 R/9__ONS_lifeexpec.R

@@ -195,7 +195,7 @@ source("__ELSA_functions.R")
 #### USE LIFE TABLES WITH AGE UNTIL 110 IN
 
 
-create_death_prob_matrix <- function(gender = "Males") {
+CreateDeathProbMatrix <- function(gender = "Males") {
     # Function that creates the transition matrix from ons data.
 
     sheet_to_get <- paste0(gender, " cohort lx")
@@ -248,53 +248,42 @@ create_death_prob_matrix <- function(gender = "Males") {
         mutate(gender = gender)
 }
 
-data_for_life_exp <- lapply(c("Males", "Females"), create_death_prob_matrix) %>%
+lapply(c("Males", "Females"), CreateDeathProbMatrix)
+
+
+
+CreateLifeExpectancyTables <- function(gender) {
+    sheet_to_get <- paste0(gender, " cohort lx")
+
+    ons_surv <- read_excel(
+        "../../data/ONS/ONS life tables 1841 onwards.xls",
+        sheet = sheet_to_get,
+        skip = 8
+    ) %>%
+        # rename_with(Vectorize(change_numeric_names)) %>%
+        rename(age = `age (years)`) %>%
+        tidyr::pivot_longer(
+            -age,
+            values_to = "alive",
+            names_to = "birth_year"
+        )
+
+    # life expectancy = sum(life years)/(alive currently)
+
+    # life exp is dividing all the remaining life years by the number of people at that age
+
+    life_exps <- ons_surv %>%
+        group_by(birth_year) %>%
+        arrange(-age) %>%
+        mutate(life_years_left = cumsum(alive)) %>%
+        mutate(life_exp = life_years_left / alive) %>%
+        mutate(gender = gender) %>%
+        select(birth_year, age, gender, life_exp)
+
+    return(life_exps)
+}
+
+life_exps_dt <- lapply(c("Males", "Females"), CreateLifeExpectancyTables) %>%
     rbindlist()
 
-
-# Lets make an average life expectancy
-# Is it sum of prob*age?
-# I think so
-
-life_exps <- data_for_life_exp %>%
-    melt(c("age", "gender"),
-        variable.name = "birth_year",
-        value.name = "prob_of_death"
-    )
-
-gender <- "Males"
-
-life_exps %>%
-    filter(birth_year == 1920 & !is.na(prob_of_death))
-
-
-sheet_to_get <- paste0(gender, " cohort lx")
-
[…]

### 2023-07-22 Texfiles/Draft1.tex
More (bad) words

@@
 \usepackage{booktabs}
+\usepackage{siunitx}
+\newcolumntype{d}{S[
+      input-open-uncertainty=,
+      input-close-uncertainty=,
+      parse-numbers = false,
+      table-align-text-pre=false,
+      table-align-text-post=false
+    ]}
 \usepackage{pdflscape}
@@
 
-\section{Methodology and Econometric Strategy}
+\begin{landscape}
+    \input{tables/elsa_results_only_dc.tex}
+\end{landscape}
 
-I first solve a modified retirement lifecycle model. The problem that retirees face is as follows.
+Table \ref{tab:DcOnlyRes} shows the results of this specification on the various consumption indicators.
+Column one has total monthly consumption as variable on the left hand side of the regression equation.
+We can see that being in the treatment group is associated with £83 more a week in spending
+overall across all categories of expenditure. This is not statistically significant using
+robust standard errors (find a reference for the SE that I use). Being older at the time of interview
+is also associated with lower consumption whilst those who retired later have lower consumption
+because years since the start of retirement is associated with higher total consumption.
+
+Columns 2 through 4 use food consumption as the outcome variable. Interestingly food consumption
+in the house does not increase as a result of the policy reform but food out side the house does
+increase. If individuals see their pension pot, that they previously could not access, as a
+shock to their wealth, like winning the lottery, then we would the consumption basket to
+include more goods that are perceived as luxuries. Likewise, consumption on clothing, shown in
+column 4 increases as a result of the policy change.
+
+I now run another specification which seeks to control for some of the cohort effect that
+could be driving these results instead. As the financial characteristics in \ref{tab:sum_stats}
+show, the treatment group has higher levels of wealth than the untreated group. To test
+for cohort effects I add individuals who have a defined benefit pension pot into the model
+and interact the policy change with a dummy for having a defined contribution pension pot.
+This makes the coefficient of interest the interacted term, which shows us the difference
+in the change in consumption between the DB and DC group. Since rules for defined benefit
+pensions did not change under the pensions freedom act, this group can be seen as a natural counterfactual
+group to judge the consumption of the defined contribution group against.
+
+\section{Life cycle theory}
+
+In the second part of the paper I first solve a modified retirement lifecycle model and then use data from the solved lifecycle
+model
+lifecycle consumption
+. The problem that retirees face is as follows.
 Every period retirees solve:

### 2023-07-22 R/10__2_summay_stats.R
Code updates

@@ -7,6 +7,7 @@ library(stringr)
 rm(list = ls())
 source("__ELSA_functions.R")
 
+
 # Covariate Balance
 # need to add home ownership
 # could have a look at the retirement paper that we did in s130
@@ -24,41 +25,15 @@ data_for_regressions <- fread("../../data/ELSA/elsa_to_use/elsa_reg_data.csv") %
     mutate(
         fin_wealth = fin_wealth / 1000,
         house_value = house_value / 1000,
-        dc_pot = dc_pot / 1000
+        dc_pot = dc_pot / 1000,
+        public_pension = public_pension / 1000
     )
-data_for_regressions$year_of %>% table()
 
+search_names(data_for_regressions, "pen")
 search_names(data_for_regressions, "year")
 
-# List of names for the summary table
-clean_names <- list(
-    "Gender" = "ragender",
-    "RetirementYear" = "retirement_year",
-    "InterviewYear" = "int_year",
-    "YearsSinceRetirement" = "years_since_retirement",
-    "RetiredAge" = "retired_age",
-    "ExpectedRetiredAge" = "first_exp_ret_age",
-    "DifferenceAge" = "exp_real_ret_age",
-    "FinancialWealth (thousands)" = "fin_wealth",
-    "DCPension" = "ever_dc_pen_bin",
-    "DCValue (thousands)" = "dc_pot",
-    "DBPension" = "ever_db_pen_bin",
-    "OwnsHouse" = "owns_house",
-    "HouseValue (thousands)" = "house_value",
-    "ObjectiveLifeExp" = "obj_life_exps",
-    "SubjectiveLifeExp" = "sub_life_exps",
-    "TotalConsump" = "total_monthly_consumption",
-    "FoodConsumpInHouse" = "monthly_food_in",
-    "FoodConsumpOutHouse" = "monthly_food_out",
-    "ClothingConsump" = "monthly_clothing",
-    "LeisureConsump" = "monthly_leisure",
-    "UtilityConsump" = "monthly_utility"
-)
 
 
-ReturnCleanName <- function(short_name) {
-    names(clean_names)[clean_names == short_name]
-}
 
 summary_cols <- unname(unlist(clean_names))
 consumption_cols <- summary_cols[grepl("monthly", summary_cols)]

### 2023-07-24 Texfiles/Draft1.tex
Lots of changes

--- 2023-07-22 Texfiles/Draft1.tex
@@
 \usepackage{booktabs}
+\usepackage{siunitx}
+\newcolumntype{d}{S[
+      input-open-uncertainty=,
+      input-close-uncertainty=,
+      parse-numbers = false,
+      table-align-text-pre=false,
+      table-align-text-post=false
+    ]}
+\usepackage{pdflscape}
+\usepackage{hyperref}
+
+\hypersetup{
+    colorlinks,
+    citecolor=black,
+    filecolor=black,
+    linkcolor=black,
+    urlcolor=black
+}
 \setlength{\parindent}{0pt}
@@
 \begin{document}
+
 \maketitle
+\tableofcontents
+\newpage
 
-Standard economic theory suggests that retirement annuities should be highly prized by individuals as a way to
-insure against the risk of late death \cite{yaari_65}. However, in developed countries rates of annuitization are far below the
-levels that theory predicts.
-Under the coalition government in the UK the law regarding the use of private defined
-contribution pensions changed. Individuals were no longer forced to annuitise their pension pots and could access
-them in a variety of ways such as a lump sum withdrawal on retirement or income drawdown and subsequently the number of annuities
-sold in the UK dropped precipitously.
+Standard economic theory suggests that retirement annuities, which provide a guaranteed income stream until the end of life,
+should be highly valued by individuals as a way to insure against late death \cite{yaari_65}. However, in developed countries,
+rates of annuitisation are far below the levels that theory predicts. The literature has suggested several reasons for
+this, but there is no consensus about which mechanism is dominant. I exploit early retirees' consumption responses to a
+reform to annuity policy in the UK to add to this literature.
+    [\textbf{Add something once I have a conclusion}].
 
-I first use the policy reform as a discontintuity to measure the impact of forced annutisation on the consumption
-of individuals in the first few years of retirement. Given that the reform was implemented suddenly and without
-advanced notice before the Spring 2014 budget I claim that individuals with a retirement year in 2015, 2016 or 2017
-are otherwise similar to those with a retirement year in 2011, 2012 and 2013 but the do not need to annuitise their
-defined contribution pension pots.
+In the UK, adults have three methods of funding retirement: the state pension; defined benefit (DB) employee pension schemes,
+which provide an income in retirement based on years of service and a function of wages; and defined contribution (DC) pensions,
+in which individuals save and invest in a tax-advantaged account that is usually supplemented with employer contributions.
+Under the 2010-2015 coalition government in the UK, the law regarding the use of private DC pensions at
+retirement changed. Individuals were no longer forced to annuitise, which provides a constant income until death,
+their pension pots and could access them in a variety of ways such as a lump sum withdrawal or income drawdown,
+which is a steady withdrawal of assets
+with the remainder left in the pot - common advice is to take 4\% a year.
+Subsequently the number of annuities sold in the UK dropped precipitously.
 
-In this paper I test two competing hypotheses for the annuity problem: bequests and
-pessimistic life expectancy. Depending on the reason for the lack of annuitization in the UK, the consumption response of retirees to the pension
-reform will differ. If individuals do not annuitise because of pessimistic life expectancy I will show that their
-consumption should increase. If, on the other hand, individuals do not annuitise because of a bequest motive, consumption
-should not change much as a result of the reform. I will solve lifecycle models for both of these cases and simulate
-consumption decisions with, and without, forced annuitization. I will then use a variety of empirical models to measure
-the consumption change in early retirement that resulted from the policy reform. The size and magnitude of this change
-will be indicative of the mechanism causing the annuitization problem.
+In this paper, I first use the policy reform as a discontinuity to measure the impact of forced annuitisation on the consumption
+levels of individuals in the first few years of retirement. Given that the reform was implemented suddenly and without
+advanced notice before the Spring 2014 budget, I claim that individuals who retired in 2015, 2016 or 2017
+are otherwise similar to those who retired year in 2011, 2012 and 2013 except for the fact the later retirees
+do not need to annuitise their DC pension pots. Therefore, I can use a regression discontinuity to identify
+the effect that forced annuitisation has on the consumption of retirees.
+
+I then test two competing hypotheses for the annuity puzzle: bequests and
+pessimistic life expectancy. The annuity puzzle was first identified by \cite{yaari_65} and is a classed as a problem
+since standard economic theory cannot explain the lack of annuitisation that is observed in the real world.
+Two possible explanations are bequests and life expectancy: individuals want to leave an estate for their heirs to
+inherit and they cannot do this if they annuitise their wealth; or individuals may believe they will not live
+as long as annuity providers think they will and therefore annuities that are on the market do not appear to be
+good value. Depending on which reason explains the lack of annuitisation in the UK, the consumption response of
+retirees to the pension
+reform will differ. If individuals do not annuitise because of pessimistic life expectancy their
+consumption should increase after the reform. If, on the other hand, individuals do not annuitise because of a bequest motive, consumption
+should not change much as a result of the reform.
+
+I will solve lifecycle models
+(\textbf{maybe add something about lifecycle models here})
+for both of these cases and simulate
+consumption decisions with and without forced annuitisation. I will then use a variety of empirical models to measure
+the consumption change in early retirement that resulted from the policy reform. The sign and magnitude of this change
+will be indicative of the mechanism causing the annuitisation problem.
 
 The importance of retirement policy to individuals in the UK is growing. The number of individuals of
-pensionable age is expected to grow from 11.9 million in 2020 to 15.2 million in
-2045 according to the latest ONS statistics and for every 1000 people of working age there will be 341 of pensionable
-age in 20145 compared to 280 in 2020 \cite{ons_population_predictions_2020}. The increase in absolute and relative
-numbers of elderly makes retirement policy more important. Moreover, private, defined contribution (DC), pensions are
+pensionable age is expected to increase from 11.9 million in 2020 to 15.2 million in
+2045 according to the latest Office for National Statistics (ONS) and for every 1000 people of working age there will be 341 of pensionable
+age in 2045 compared to 280 in 2020 \cite{ons_population_predictions_2020}. The increase in absolute and relative
+numbers of elderly people makes retirement policy more consequential. Moreover, private DC pensions are
 becoming increasingly common and are predicted to grow as current cohorts age \cite{cribb_karjalainen_ifs_2023}.
@@
 
+Moreover, understanding how retirees spend their money over retirement is an important policy question of its own.
+If retirees spend too much in early retirement, the state may need to provide for them towards the end of their
+lives. This has implications for government budgets, especially since population aging means more individuals will
+require expensive end of life care. On the other hand, if retirees over-save and do not spend, the economy may face
+a dynamic inefficiency whereby consumption per capita could be increased if savings were decreased. (Get dynamic inefficiency
+reference.)
+
 \subsection{Literature review}
-My paper draws on three main strands of literature. The annuity problem, the retirement saving
-problem and lifecycle models. \textbf{this intro is rubbish}
+My paper builds on three main strands of literature: that of the annuity puzzle, exploring reasons that some retirees choose not to annuitise;
+the retirement saving puzzle, looking at consumption and saving behaviour over retirement; and lifecycle models, whereby the above
+problems are explained using a tractable model of human decision making.
 
@@
 of annuities lower for the general population as opposed to the population of annuitants. However, they
-also find that theory would still predict annuitization.
+also find that theory would still predict annuitisation.
 
-\cite{friedman_warshawsky_qje_1990} show that annuitization decisions can be fully explained by a mixture of
+\cite{friedman_warshawsky_qje_1990} show that annuitisation decisions can be fully explained by a mixture of
 bequest motives and actuarily unfair annuities. They solve an augmented life-cycle model with a range of
@@
 \cite{friedman_warshawsky_chicago_1988} show that there is a significant difference between the life expectancy of
-annuitants and the general population in the American annuity market but this cannot fully explain the annuitization
-problem. Only when bequest motives are added to the model can annuitization rates be rationalized.
+annuitants and the general population in the American annuity market but this cannot fully explain the annuitisation
+puzzle. Only when bequest motives are added to the model can annuitisation rates be rationalized.
 
 \cite{lockwood_red_2012} builds on this and shows that a realistic bequest motive in lifecycle simulations achieves
-realistic annuitization rates. He solves a simple lifecycle model with bequest motives taken from several recent papers
+realistic annuitisation rates. He solves a simple lifecycle model with bequest motives taken from several recent papers
 in the literature. The bequest motives he picks therefore match other important aspects of the lifecycle model such
@@
 
-\textit{There are some more papers to include here. Time to go and have dinner. }
+\textbf{If I add 300 more words here then I will be at 5000!}
+
+\textbf{There are some more papers I will include here.}
 
@@
 
-The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA repeatedly interviews
-individuals over the age of 50 and asks them a range of questions relating to their income and wealth as well as expectations about
-the future. Importantly it also includes detailed information pensions including the type of pension that an individual holds when
-working, thus I am able to distinguish between individuals who have defined benefit and defined contribution pensions. For ease of access
-I use Harmonized ELSA \footnote{"This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
+The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA picks individuals
+over the age of 50 to survey every two years until death. If individuals leave the survey ELSA replaces them so that it is
+representative of the over 50 population in the UK. Individuals are asked a range of questions relating to their income
+and wealth as well as expectations about the future. One benefit of using ELSA is that it includes data on pension types
+for individuals who are working. Therefore, I can differentiate between individuals who have a DC pension and those who have a DB pension.
+There have been 9 waves of ELSA data collection, the first was in 2001 and collections happened every two years thereafter.
+
+Importantly ELSA also includes information on pension size that the Institute for Fiscal Studies (IFS) calculated, this is only available
+up until wave 5 (which was collected in 2010 and 2011) at which point I use a real return
+of 3\% to predict forward the pension wealth variable until retirement. ELSA also includes a measure of all
+non-housing financial wealth which I use in some specifications because of these issues with
+pension wealth. Due to slight differences in the ELSA questions between years, I use `Harmonized ELSA'
+\footnote{This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
     July 2021 developed by the Gateway to Global Aging Data. The development of the Harmonized ELSA was funded by the National
     Institute on Aging (R01 AG030153, RC2 AG036619, R03 AG043052). For more information,
-  please refer to https://g2aging.org/.”} which ensures that variables are comparable across waves. Since this only includes a
-subset of the questions in ELSA I also supplement it with variables taken directly from the data.
+    please refer to https://g2aging.org/.} which ensures that variables are comparable across waves. Since this only includes a
+subset of the questions in ELSA, I also supplement it with variables taken directly from the data such as questions that deal
+with life expectancy.
 
-ELSA also includes questions on expenditures. In particular individuals are asked how much they consume on a range of broad categories.
+ELSA also includes questions on expenditures. In particular, individuals are asked how much they consume across a range of broad
+categories including in-house food consumption, out-of-house food consumption, leisure consumption, clothes consumption and
+consumption on utility bills and rent. I use also use two variables related to life expectancy that I discuss in more detail below.
 
-% Include a summary stats table here of the individuals selected for the regression
-
-I also use life tables from the UK's Office for National Statistics. These give us risk of death for each age group. I adjust them
+I also use `life tables' from the UK's ONS. These provide me with risk of death for each age group. I adjust them
 to make death certain at age 110 as is common in the literature. I transform these so that I have risk of death conditional on being
-a given age since this is what is used in the life cycle simulations.
-%  include a table showing death probs
+a given age since this is what I use in the life cycle simulations. I also use these objective probabilities to
+calculate annuity prices for individuals.
 
 To illustrate the effect the reform had on sales of annuities in the UK I obtained product data from the Financial Conduct Authority.
-These track the sale of different financial products overtime including data on annuity sales.
+These track the sales of different financial products overtime including data annuities.
 
-% Plot showing the drop of in annuity sales
+Life expectancy impacts decision to retire for two reasons. Firstly, it directly impacts the price that an annuity will cost for
+individuals. Older individuals and those with pre-existing health conditions generally can buy an annuity that provides a greater
+income stream than individuals who are younger. Secondly, private information about life expectancy impacts the perceived
+value of an annuity. If an individual expects to outlive the general population an annuity, at a given price, will appear
+a much better deal to them. Likewise, an individuals life expectancy will change their decision on how to consume and save
+through retirement.
 
@@
 also asked “What are the chances that you will live to be age 85 or more?” if they were under 70. As most recent retirees are under 70 we
-therefore have two data points. I carry out the following procedures: drop any individuals who think it is more likely they reach an older
-age than a younger age since; add, as a third data point their objective chance of reaching 110 according to the ONS life tables; fit these
-three points to a Weibull distribution using non-linear least squares. Then I create subjective survival tables.
+therefore have two data points. I first drop from the data individuals who think it is more likely that they reach a higher age than a younger
+age since this shows a misunderstanding of the question. I then add as a third data point their objective chance of reaching 110 according to the ONS life tables.
+I fit these three points to a Weibull distribution, which is commonly used by demographers to estimate how populations
+will age, using non-linear least squares. Then I create subjective survival tables using parameters from the Weibull distribution.
 
@@
 
-\input{tables/r_summary_stats.tex}
-
 \subsection{Policy reform}
 
-\textbf{Add some other background to pensions in the UK. What does the Blundell etc have
-  to say}
+Successive governments in the UK have attempted to reform both the public and private pension system. Prior
+to 1987, participation in private schemes was limited to employees of firms that had offered them, and there
+were few alternatives to the public state pension or defined benefit scheme that a public sector employer
+would offer. From 1989, individuals in the UK were allowed to open tax advantaged self-invested personal pensions
+alongside any pension their employer offered. The 2004 Finance Act rationalised taxation rates on different types of pensions.
+But DC pension pots could still only be used to buy an annuity after taking out a maximum of 25\% as a tax free lump sum withdrawal
+from age 55 onwards. Most pension pots also required that individuals access them before age 75.
 
-Successive governments in the UK have attempted to reform both the public and private pension system. Prior
-to 1987 participation in private schemes was limited to employees of firms that had offered one, and there
-were few alternatives to the public state pension or a define benefit scheme that a public sector employee
-would offer. \textbf{Were there pension reforms in 1990s and 00s}
+Announced in the spring budget of 2014, the so called `pensions freedom act' received Royal Asset in December 2014 marked the end of a series
+of pension reforms carried out by the coalition government between 2010 and 2015. The reform made it possible
+to withdraw the first 25\% from a pension pot tax free and the rest subject to the marginal rate of income tax that an individual faced.
+In the June 2010 budget, the government made a first reform to the annuitisation rules, creating
+a minimum income requirement above which individuals would not need to
+annuitise more \cite{finance_act_hmt_2011} (this meant high income retirees would not need to annuitise their wealth).
+However, this was set at £20,000 and therefore few individuals
+were eligible. The minimum income requirement was scrapped in the 2014 bill, finally eliminating the
+compulsory annuities market, with one government minister saying pension pots `can be used to buy Lamborghinis'
+\cite{guardian_lambos}.
 
-Announced in the Spring budge of 2014, the so called "pensions freedom act" received Royal Asset in December 2014 marked the end of a series
-of pension reforms carried out by the coalition government between 2010 and 2015. The reform was announced
-in the Spring budget and made it possible to withdraw money from a private pension pot subject to the
-marginal rate of income tax that an individual faced. In the June 2010 budget the government made a first
-reform to the annuitization rules, creating an minimum income requirement above which individuals would not need to
-annuitise more \cite{finance_act_hmt_2011}. However, this was set at £20,000 and therefore few individuals
-were eligible. The minimum income requirement was scrapped in the 2014 bill finally eliminating the
-compulsory annuities market. The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
-Using data from the Association of British Insurers they show that annuity demand dropped by 75\% from
-its maximum.
-
-After the announcement in spring 2014 sales of annuities dropped massively and income drawdown products increased.
-However, pension pots could also be taken out in one go and this became a much more common option for many people.
-
-Figure \ref{fig:annovertime} shows purchases of annuities overtime and demonstrates the sharp decrease
-in purchases that happened from 2014 to 2015. Likewise there was an increase in the number of pensions that were
-being accessed using an income drawdown product, but these do not fully account for the drop in annuitisation. So,
-either they are withdrawn at once or they are left alone.
-
-Figure \ref{fig:ann2122} shows the current distribution of how pension pots are first accessed at retirement.
-In 2021, 196,736 pots were fully withdrawn at retirement, accounting for over 50\% of pots, prior to the policy
-reform this was not the case and most defined contribution pensions were accessed through annuities.
-
+The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
+Using data from the Association of British Insurers, they show that annuity demand dropped by 75\% from
+its maximum in 2011
+\textbf{Check}
+. Figure \ref{fig:annovertime} shows purchases of annuities overtime and demonstrates the sharp decrease
+in purchases that happened from 2014 to 2015. There was an increase in the number of pensions that were
+being accessed using an income drawdown product, but these do not fully account for the drop in annuitisation.
 \begin{figure}[h]
-  \caption{Pension pots accessed}
+    \caption{Pension pots accessed }
     \centering
@@
     \label{fig:annovertime}
+
 \end{figure}
@@
 \begin{figure}[h]
-  \caption{How pension pots are accessed}
+    \caption[Caption for LOF]{How pension pots are accessed\protect\footnotemark}
     \centering
@@
     \label{fig:ann2122}
+    \
 \end{figure}
+Figure \ref{fig:ann2122} shows the distribution of how pension pots were first accessed at retirement in 2021-22.
+In 2021, 196,736 pots were fully withdrawn at retirement, accounting for over 50\% of pots. Prior to the policy
+reform this was not the case -- most defined contribution pensions were accessed through annuities.
+\footnotetext{Partial UFPLS (uncrystallised funds pension lump sum) is similar to Partial drawdown but with a different tax schedule. With partial drawdown
+    you take the whole 25\% tax free amount at once whereas with UFPLS you only claim the 25\% on the amount you are taking out. This option is
+    better if you may want to buy another retirement product in the future such as an annuity. }
+
+Also of specific interest to the annuity market was the European Union's
+`Equal Treatment in Goods and Services Directive' of 2004. This prohibited discrimination
+in the provision and cost of goods and services based on sex. Up until 2011, there was a clause that
+stated insurers were allowed to charge different premiums if it was based on evidence that sex is
+correlated with different amounts of risk. However, in March 2011, the European Court of Justice
+ruled that insurers were not allowed to charge different amounts and gave them until December 2012
+to implement the ruling. This change meant that annuity products could no longer be priced differently
+for men and women in the UK. However, given that this change was implemented two years before the pension
+freedoms act, I can still identify the impact of the pensions freedom act on early retirement consumption.
 
@@
 
-I regress demographic and financial characteristics of individuals on year of birth and the treatment dummy.
+Table \ref{tab:sum_stats} shows various summary statistics for each variable of interest from ELSA and the ONS.
+
+As required, retirement year is between 2011 and 2013 for the control group and 2015 and 2017 for the treatment group.
+However, as discussed above their interview date is after or the same as the retirement year because we are tracking
+consumption in retirement.
+
+The second retirement group, those who retired post reform, is small with just 728 non missing observations for gender as opposed
+to 941 individuals in the control group. The control group are more female, retired at a slightly younger age and expected to
+retire slightly earlier. This could be because this period was affected by the increase in the state pension age for
+women from age 60 in 2010 to age 65 in 2018. Since this change happened gradually and occured over the whole period
+I do not expect it to influence the results. The DifferenceAge row tracks the difference between expected and actual
+pension age.
+
+The treatment group has higher financial wealth with a median of £66,000 at time of interview as opposed to £55,000
+in the control group. Likewise, the second group are more likely to have held a DC pension at some point
+and have much more money in them. Both groups are equally likely to have a defined benefit pension with roughly 44\% of
+individuals across both samples having a defined benefit pension. In general, DB pensions are more prevalent than
+DC pensions in the data, this reflects the same trend that is observed in population of UK retiree. Home ownership
+is roughly equal across groups although the second group have slightly higher housing wealth.
+\textbf{I have
+    not inflation adjusted but I am thinking about it?}
+
+Both groups are similarly long-lived according when using ONS life tables to calculate life expectancy
+based on gender, age and the year the interview was carried out in. Subjective life expectancies are also similar
+across groups, with individuals expecting to live another 21 years as opposed to the 24 that the ONS would expect
+them to live.
+\textbf{Maybe add in prob of survival to age 85 or 90 or something}
+\textbf{Also add in an indicator for living alone or living with someone else}
+
+Unfortunately, some consumption data is missing for some individuals. The food categories have the least missing data
+and the leisure consumption category has the most missing data.
+
+\textbf{Check why total consumption adds up to so much
+    more than all the categories added together? Do they include rent and housing in consumption?}
+
+Overall, the groups appear to have slightly different wealth profiles with the treatment group being richer and slightly more likely
+to have DC pension wealth. On key demographic characteristics, however, the groups are similar, the difference in retirement
+age between the two groups can be mostly explained by the difference in expected retirement age meaning that individuals
+have not en masse decided to delay retirement and avoid annuitisation.
+
+\begin{landscape}
+
+    \input{tables/r_summary_stats.tex}
+
+\end{landscape}
+
+As a further test for balance I regress demographic and financial characteristics of individuals on year of birth and the treatment dummy.
 We can then see if the treatment groups differ on key characteristics such as financial wealth or retirement age.
@@
 \begin{equation*}
-  Y_{i} = \alpha + \beta Post_{i} + \gamma YOB + \kappa (Post_{i} \times YOB_{i}) + \epsilon_{i}
+    Y_{i} = \alpha + \beta Post_{i} + \gamma YOB_{i} + \kappa (Post_{i} \times YOB_{i}) + \epsilon_{i}
 \end{equation*}
@@
 \input{tables/cov_balance.tex}
-Table \ref{tab:cov_balance} shows these results. We can see ...
+Table \ref{tab:cov_balance} shows these results. We can see that being in the treatment group.
+\textbf{I am still thinking about this.}
 
-Another threat to validity is manipulation of the running variable. This is probably quite a threat in our case
-as individuals could delay retirement and/or the purchase of an annuity.
+One threat to validity is manipulation of the running variable -- which is the variable that decides
+whether an individual is in the treatment group or not -- in our case retirement year.
+This is probably quite a threat in this setting as individuals could delay retirement and/or
+the purchase of an annuity until after the reform. To check whether
+this is happening we see whether the difference between real retirement age and the first expected retirement age
+recorded in the data are different for those retiring before 2014 compared to those retiring after 2014. As shown
+above in Table \ref{tab:sum_stats} this not the case.
 
-\section{Methodology and Econometric Strategy}
+However, it is possible that individuals do not buy an annuity at the time of retirement since the law prior to 2014 only
+required that they access the pot by age 75. They may decide to keep their DC pension untouched
+and live off other income and assets before accessing it later. Since we do not have data on the
+exact purchase date of annuities we cannot track whether
+annuity purchases were delayed because of this. However, there is relatively stable annuity demand
+before the policy reform. If individuals delayed annuity purchases in expectation of the reform we
+would observe declines in purchases before the reform was announced.
 
-I first solve a modified retirement lifecycle model. The problem that retirees face is as follows.
-Every period retirees solve:
+\section{Empirical models}
+
+In this section I outline the key empirical models I run with both the simulated consumption data from the
+lifecycle models and with the real data from ELSA. I then see which lifecycle model better fits the consumption
+response that occurred as a result of the pension reform.
+
+I use a regression discontinuity design for which I compare the consumption of early retirees after the policy reform
+to that of people who retired just before the policy reform. The key assumption implicit in regression discontinuities is
+that nothing else changes at the time of the jump apart from the policy of interest, and that the policy occurs
+without individuals predicting it and altering their behaviour. As I have argued above, the demographic information in the
+data suggest that individuals did not delay retirement and that there was no delay in annuitisation as shown by the quick
+and sudden decline in annuity purchases. Moreover, anecdotal evidence from media and business sources at the time of the
+announcement show that there was surprise the government had gone this far, with money marketing describing the change as a
+"bombshell" \cite{money_marketing_announcement}.
+
+Retirement year is the running variable and individuals are treated if retirement year is greater than 2014 and less than
+or equal to 2017. I use consumption data of individuals up to 2 years into retirement so that the sample size is larger.
+So if someone retired in 2015 and had consumption data in 2015 and 2017, I include both values. An individual is considered
+not treated if they retire before 2013 and after 2011.  To account for the fact
+that the reform only impacted individuals who had accumulated defined contribution pension pots I interact the treatment dummy with
+an indicator variable signalling whether the individual had ever held a DC pension pot.
+
+Because of the differences in financial characteristics between the groups, I add controls for financial wealth, housing wealth
+and whether an individual owns their own home.
+
+\begin{equation*}
+    Cons_{it} =  \gamma X_{it} + \beta PostReform_{it} + \epsilon
+\end{equation*}
+
+And $X_{it}$ is a set of controls.
+
+\begin{landscape}
+    \input{tables/elsa_results_only_dc.tex}
+\end{landscape}
+
+Table \ref{tab:DcOnlyRes} shows the results of this specification on the various consumption indicators.
+Column 1 has total monthly consumption on the left hand side of the regression equation.
+We can see that being in the treatment group is associated with £73 more a week in spending
+overall across all categories of expenditure. This is not statistically significant using
+robust standard errors \textbf{find a reference for the SE that I use}. Being further into retirement
+at the time of interview is also associated with lower consumption whilst those who retired later have lower consumption
+because years since the start of retirement is associated with higher total consumption.
+
+Columns 2 through 4 use food consumption as the outcome variable. Interestingly, food consumption
+inside the house does not increase as a result of the policy reform but food consumed away from the house does
+increase. If individuals see their pension pot, that they previously could not access, as a
+shock to their wealth, like winning the lottery, then we would expect the consumption basket to
+include more goods that are perceived as luxuries. Inline with this, consumption on clothing, shown in
+column 4, increases as a result of the policy change.
+
+One concern is that the group who retired after 2014 are just generally different to those
+who retired before and perhaps consume more outside the home anyway. The financial characteristic
+data presented in \ref{tab:sum_stats} would support this hypothesis. To test for these cohort
+effects I run a difference in difference regression to complement the regression discontinuity just shown.
+I add individuals who have a defined benefit pension pot into the model and interact the policy change
+with dummies for having different types of pension. Specifically I run:
+
+\begin{equation*}
+    Cons_{it} =  \gamma X_{it} + \beta_{1} PostReform_{it} + \beta_{2} PostReform*DC_{it} + \beta_{3} PostReform*DB_{it}  + \epsilon
+\end{equation*}
+
+This makes the coefficient of interest $\beta_{2}$, which shows us the difference
+in the change in consumption between the DC only group and the base group that has neither a DB or DC pension.
+Since rules for DB pensions did not change under the pensions freedom act, those with a DB pension or no pension
+can be seen as a natural counterfactual group to judge the consumption of the DC group against.
+
+\begin{landscape}
+    \input{tables/elsa_results_all_with_interact.tex}
+\end{landscape}
+
+Table \ref{tab:ElsaAllData} shows these results. The post reform variable
+now tells us that total consumption decreased but food consumption increased a little and clothing consumption
+decreased. The interactions at the bottom of the table show us what happened to individuals with a given pension
+type relative to the reference cohort which has no pension. Those with only a DC pension had an increase in total consumption
+which was lower than that for those with a DB pension but higher than those with no pension. Only for consumption on
+clothes did DC consumption increase more than the other groups. As above, the only variables that are significant
+at the 5\% level are financial wealth and whether the individual owns their house.
+
+We can also test whether treatment intensity is correlated with larger increases in spending. Those who have
+more money in a DC pension, and those with more financial wealth overall, would be impacted
+by the policy to a greater extent than those with smaller pensions. So, I interact the policy with the treatment
+variable. As in Table \ref{tab:DcOnlyRes}, this regression only uses those individuals who have a DC pension.
+
+To be precise I run:
+\begin{equation*}
+    Cons_{it} =  \gamma X_{it} + \beta PostReform_{it}*DCValue_{it} + \epsilon
+\end{equation*}
+
+\begin{landscape}
+    \input{tables/elsa_results_dc_pot_interact.tex}
+\end{landscape}
+
+Table \ref{tab:DcOnlyInteract} shows the results. Surprisingly, a larger DC pension
+pot is not associated with a greater increase in consumption. And although the reform is again associated with an
+increase in consumption, especially for food out of the house and clothing, this affect is decreasing with the
+size of the pension pot.
+
+The defined contribution pension pot size variable comes from work done by the IFS. This was discontinued
+after wave 5 so it is no longer possible to identify pension wealth in this way. For the above table pension
+wealth from the last available year for an individual was given a real return of 3\% and compounded until year of interview.
+Given that this is an imperfect measure, I also run the regression with financial wealth interacted with the treatment variable.
+
+\begin{landscape}
+    \input{tables/elsa_results_fin_wealth_interact.tex}
+\end{landscape}
+
+Table \ref{tab:DcOnlyFinWealthInteract} contains these results and shows a similar pattern to Table \ref{tab:DcOnlyInteract}.
+The reform is associated with an increase in consumption, but more financial wealth decreases the effect rather
+than making it stronger. This could be because those with higher levels of financial wealth have more assets in
+other savings apart from retirement savings \cite{cribb_karjalainen_ifs_2023} and therefore the DC pot is a smaller proportion of overall assets,
+and thus the policy change did not affect them as much.
+
+In conclusion, the results in the tables above show that in the basic regression discontinuity setup there is
+an impact of the policy reform on consumption, and this effect is stronger for consumption goods we
+would associate with a large increase in income. However, this is imprecisely estimated, and when we use a
+difference in difference identification method this effect disappears. Moreover, when we interact with a proxy for treatment strength
+(financial wealth or pension size) the effect is smaller the stronger the treatment.
+
+Theoretically, a null change in consumption is evidence for a bequest motive explaining a lack of annuitisation
+rather than pessimistic life expectancies being the cause. In the next section I simulate a lifecycle model
+for individuals in the ELSA dataset and
+\textbf{some thing that gives a taster of the results to come in the next section}
+
+\section{Life cycle theory}
+
+In the second part of the paper I first solve a modified retirement lifecycle model and then use data from the solved lifecycle
+model in similar regressions to the ones above.
+
+The problem that retirees face is as follows. Every period retirees solve:
+
 \begin{equation*}
@@
 \end{equation*}
+
 subject to their budget constraint
+
 \begin{equation*}
@@
 \end{equation*}
-where $a_{t}$ are asset holdings in time t, $y$ is constant income for all periods and . Income can come either
+
+where $a_{t}$ are asset holdings in time t, $y$ is constant income for all periods, $V_{t}$ is value in period $t$, $p_{t}$ is probability of
+death in period $t$ and $B()$ is a bequest function. Income can come either
 from state pensions, defined benefit pension plans or purchased annuities.
+
 \begin{equation*}
     u(c_{t}) = \frac{c_{t}^{1 - \sigma}}{1 - \sigma}
-\end{equation*} In some specifications retirees can leave bequests, I use the bequest function from
+\end{equation*}
+
+In some specifications retirees can leave bequests, I use the bequest function from
 \cite{lockwood_red_2012}.
+
 \begin{equation*}
-  B(a_{t}) = \bigl( \frac{m}{1 - m} \bigr)^{\sigma}  \frac{(\frac{m}{1 - m}c_{0} + a_{t})^{(1 - \sigma)}}{1 - \sigma}
+    B(a_{t+1}) = \bigl( \frac{m}{1 - m} \bigr)^{\sigma}  \frac{(\frac{m}{1 - m}c_{0} + a_{t+1})^{(1 - \sigma)}}{1 - \sigma}
 \end{equation*}
-Where $b_{t}$ is the amount left at death, $m$ is a measure of bequest motive strength and $c_{0}$ is
-a minimum amount of consumption that individuals want. \textbf{Check this}
+
+Where $a_{t+1}$ is the amount left at death, $m$ is a measure of bequest motive strength and $c_{0}$ is
+a minimum amount of consumption that individuals want. I pick values of $m = 0.95$ and $c_{0} = 20,000$ since
+these generate low rates of annuitisation, for comparison, Lockwood picks $m = 0.96$ and $c_{0} = 18,000$
+\textbf{Check all of this}
 
@@
 either 0 (if we do not allow a bequest motive) or the value of bequests. I then take this value function and solve
-an individuals final period problem, choosing assets next period (i.e. those to bequest) and how much to consume.
+an individual's final period problem, choosing assets next period (i.e. those to bequest) and how much to consume.
 
-Using the optimal policy function in the last period, I calculate the value of the last period, which the utility
-function and the value function evaluated at the maximum. This is then used in the problem the year before that.
+Using the optimal policy function in the last period, which is assets next period that maximise the utility
+function and the value function this period given all the potential income and asset states, I calculate the value of the last period.
+This is then used in the problem the year before that.
 I repeat this process back to the age of retirement to obtain optimal consumption amounts for each year of retirement
@@
 
-To simulate the ELSA data I solve this retirement problems for each new retiree in the data set dependent on their
-objective probability of death each period. I estimate with subjective life expectancies and objective life expectancies,
-I also estimate the model both with and without a bequest motive which was picked to fit the unforced real annuity
-rates seen in the data. I then estimate several empirical models with the simulated data.
+To simulate the ELSA data I solve this retirement problem for each new retiree in the data set dependent on their
+objective probability of death each period. I estimate with subjective life expectancies and objective life expectancies.
+I also estimate the model both with and without a bequest motive which was picked to roughly fit the real annuity
+rates seen in the data.
 
-In retirees first year of retirement I allow them to choose to annuitise some of their wealth. In practical terms
+In retiree's first year of retirement I allow them to choose to annuitise some of their wealth. In practical terms
 this is moving down the asset grid but up the income grid and seeing if the value of being in that position is better
-than where the individual is currently. To calculate this trade-off I calculate the annual annuity payment that follows
+than where the individual is currently. To assess this trade-off I calculate the annual annuity payment that follows
 from a given annuity cost. I calculate this using objective life tables from the ONS using the following equation:
@@
 
-Where $C$ is the one-off payment, $\delta$ is a factor that controls the 'money's worth' of annuity and $p_{t|Retage}$
+Where $C$ is the one-off payment, $\delta$ is a factor that controls the `money's worth' of annuity and $p_{t|Retage}$
 is the probability of death at age $t$ conditional on being age $Retage$. So individuals can move C on the asset grid
@@
 
-\section{Empirical models}
+I also simulate the expected change in consumption given different rates of annuitisation versus no annuitisation.
+The benefit of simulating an individual's decision is that we can directly compare what they would have
+consumed with what they actually consumed, and we can replicate the empirical analysis using the
+solution to the lifecycle model.
 
-In this section I outline the key empirical models I run with both the simulated consumption data from the
-lifecycle models and with the real data from ELSA. I then see which lifecycle model better fits the consumption
-response that happened as a result of the pension reform.
+I first round an individuals financial wealth to the nearest point on the discrete asset grid, and
+do likewise for an individuals state pension income. I then take the treatment group and evaluate
+their consumption in the year of interview with an annuity given that they annuitised some portion
+of their wealth in their retirement year. I also evaluate consumption without annuitisation. I then
+use the same specification as above, but now the treatment group is the values of consumption without
+annuitisation.
 
-I use a regression discontinuity design where I compare the consumption of recent retirees after the policy reform
-to recent retirees before the policy reform. The individuals who retired in 2014 and later were not forced to
-annuitise their defined contribution pension. The key assumption implicit in regression discontinuities is
-that nothing else changes at the time of the jump apart from the policy of interest. And that the policy occurs
-without individuals predicting it. The policy change was widely seen as a surprise by the media and financial planners,
-some of whom in fact complained that they had not been consulted enough. \textbf{How can I justify no other jumps at the jump}
+Because I discretised the state space, annuity prices are rounded down to the nearest income grid point.
+This means that for some individuals with low assets half of their financial wealth does not buy them
+at least one grid point of annuity income. I do not let these individuals annuitise, setting consumption
+with and without an annuity to the same amount. I also present the average implied loading factor
+$\delta$, although I set this to $0.9$ when calculating annuity prices, because of rounding down
+this is lower.
 
-Retirement year is the running variable and individuals are treated if retirement year is greater than 2013 and less than 2016.
-I consumption of individuals up to 2 years into retirement so that the sample size is larger. So if someone retired in 2014 and
-had consumption data in 2014 and 2016 I include both values. An individual is considered not treated if they retire before 2013 and after 2011.
-One benefit of using ELSA is that it includes data on pension type for individuals who are working. Therefore, I can differentiate
-between individuals who have a defined contribution pension and those who have a defined benefit contribution. To account for the fact
-that the reform only impacted individuals who had accumulated defined contribution pension pots I interact the treatment dummy with
-an indicator variable signalling whether the individual had ever held a DC pension pot.
+\begin{landscape}
 
-The estimating equation is therefore
+    \input{tables/standard_life_cycle.tex}
+\end{landscape}
 
-\begin{equation*}
-  Cons_{it} =  \gamma X_{it} + \beta PostReformDC_{it} + \epsilon
-\end{equation*}
+Table \ref{tab:StandardLifeCycle} shows the results from the bequest models.
 
-And $X_{it}$ is a set of controls including financial wealth etc
+\begin{landscape}
+    \input{tables/beq_life_cycle.tex}
+\end{landscape}
 
-I also run the regression with a variety of consumption variables.
+Table \ref{tab:BeqLifeCycle} shows the results from the bequest models.
 
-I also simulate the expected change in consumption given forced annuitisation and no annuitisation.
-The benefit of simulating an individuals decision is that we can directly compare what they would have
-consumed with what they actually consumed.
+Each column in the table shows a different model. Column 1 shows the model with no bequest motive where
+individuals use objective life probabilities from ONS, column 2 uses
 
-\subsection{Rough plan}
-\begin{itemize}
-  \item Intro
-        \begin{enumerate}
-          \item I think Eric French wrote something about population that I could use in then intro to say why it is important.
-          \item "Latest data from HM Revenue Customs (published in April) showed more than £45bn has been taken from pots since 2015."
-                https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
+\textbf{Also get some plots of consumption in early retirement
+    with and without annuities for the different model types}
 
-          \item Add bit about DC/DB pensions in the intro. Also talk about heterogeneity across countries.
-                Some countries want to move towards more annuitisation. This annual review is a good source of info
-                \cite{banks_crawford_ar_2022}
-        \end{enumerate}
-  \item Lit review
-  \item Models
-  \item Empirical
-        \begin{enumerate}
-          \item Diff in diff
-          \item RDD
-          \item Matching?
-          \item can I use anything I learnt in panel?
-                In some sense the decision to annuitise is a discrete choice problem so I could use something from there.
-        \end{enumerate}
-  \item Conclusion
-\end{itemize}
+% \subsection{Rough plan}
+% \begin{itemize}
+%   \item Intro
+%         \begin{enumerate}
+%           \item I think Eric French wrote something about population that I could use in then intro to say why it is important.
+%           \item "Latest data from HM Revenue Customs (published in April) showed more than £45bn has been taken from pots since 2015."
+%                 https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
 
-\bibliographystyle{chicago}
+%           \item Add bit about DC/DB pensions in the intro. Also talk about heterogeneity across countries.
+%                 Some countries want to move towards more annuitisation. This annual review is a good source of info
+%                 \cite{banks_crawford_ar_2022}
+%           \item https://www.moneysavingexpert.com/savings/pension-freedom/
+%         \end{enumerate}
+%   \item Lit review
+%   \item Models
+%   \item Empirical
+%         \begin{enumerate}
+%           \item Diff in diff
+%           \item RDD
+%           \item Matching?
+%           \item can I use anything I learnt in panel?
+%                 In some sense the decision to annuitise is a discrete choice problem so I could use something from there.
+%         \end{enumerate}
+%   \item Conclusion
+% \end{itemize}
+
+\bibliographystyle{apa}
 \bibliography{references}

### 2023-07-24 Texfiles/Notes.tex
Code updates
Lots of changes

@@
   \item Interesting point re annuities being low because of state pension.
-        I could test consumption response to forced annuitization if that is the
+        I could test consumption response to forced annuitisation if that is the
         case as well. I think it would just mean increased spending?
@@
           individuals but for double or even multiple lives."} This is interesting.
-        Could I compare annuitization decisions of couples and singles? Singles
+        Could I compare annuitisation decisions of couples and singles? Singles
         should be more likely to buy annuities if this is true.
@@
         of these other possible explanations." that is for other explanations for
-        lack of annuitization. Had a look at this and it wasn't super useful. Just
+        lack of annuitisation. Had a look at this and it wasn't super useful. Just
         seemed like a first draft of the paper.
@@
 
+        item
 \end{itemize}
@@
 \section{Personal notes}
+
+\subsection{Saturday before due date}
+\begin{itemize}
+  \item Thistlethwaite and Campbell, 1960 -- RD literature
+  \item Regression Discontinuity Inference with Specification Error -- Lee and Card. Seems like an interesing paper
+  \item Should I use financial wealth variables at retirement instead of at interview? I guess in alot of cases I wont
+        know this. I could check the differences for the people that I have pre and post for.
+  \item I also need to see what happens if I include family size.
+  \item Ok data work is done. Now I should need to work on julia data. I think I need to run some more subjectives
+\end{itemize}
 
@@
                 arent included in the normal elsa data.
-          \item generate correct rates of annutisation
+          \item generate correct rates of annuitisation
         \end{enumerate}

### 2023-07-24 Julia/1__lifecycle_with_income_function.jl

@@ -34,14 +34,7 @@ RetirementLifecycleWithIncome = function (;
     # either with objective probs given by age, gender and year
     # or subjective probs given by id_wave
 
-    if life_prob_types == "objective"
-        # set file name to save json to. 
-        json_filename = string(age, "_beq", bequests, "_", gender, "_", life_prob_types, "_", year, ".json")
-        filename_with_dir = "../data/ELSA/lifecycle_outputs/" * json_filename
-
-        death_probs = GetObjectiveDeathProbs(gender=gender, age=age, year=year)
-
-    elseif life_prob_types == "subjective"
+    if life_prob_types == "subjective"
         dir_name = "../data/ELSA/lifecycle_outputs/id_wave/"
         if !isdir(dir_name)
             mkdir(dir_name)
@@ -67,35 +60,32 @@ RetirementLifecycleWithIncome = function (;
     # Dimensions are the state variables
     # Income, current assets and age
     value_array = ones(Float64, inc_grid_points, asset_grid_points, terminal_age - start_age)
-    policy_array = ones(Float64, inc_grid_points, asset_grid_points, terminal_age - start_age)
+    policy_array = ones(Int, inc_grid_points, asset_grid_points, terminal_age - start_age)
 
     # for each income level find the optimal asset policy function
     for inc in eachindex(income_grid)
 
         utility_matrix = utility_array[:, :, inc]
 
-        # If bequests then we get utility at end of life from certain death
-        if bequests == true
-            terminal_value = [bequest_lock(final_assets) for final_assets in asset_grid]
-        else
-            terminal_value = [0 for final_assets in asset_grid]
-        end
+        terminal_value = [0 for final_assets in asset_grid]
+
 
         # Initialize a vector to store the terminal values for optimal consumption
-        temp_value = zeros(asset_grid_points, terminal_age - start_age)
+        temp_value = zeros(Float64, asset_grid_points, terminal_age - start_age)
         # Initialize a matrix to store temporary values
 
         temp_value[:, terminal_age-start_age] = terminal_value
         # Set the terminal values in the temporary value matrix
 
-        # each colum is an age
+        # each column is an age
         # each row amount of assets next period. 
-        opt_policy_function = zeros(asset_grid_points, terminal_age - start_age)
+        opt_policy_function = zeros(Int, asset_grid_points, terminal_age - start_age)
         # Initialize a matrix to store the optimal policy function
 
         # solve model backwards
         t = 1
[…]

### 2023-07-24 Julia/2__run_models.jl

@@ -1,37 +1,17 @@
 ### In this script we use the functions developed in 0 and 1
+using JSON3
 
 include("0__functions.jl")
 include("1__lifecycle_with_income_function.jl")
 
-
-# # This is the combo of demographic info that I solve the model for 
-# gender_age_year_combos = CSV.File("../data/ELSA/elsa_to_use/gender_age_year_unique.csv") |>
-#                          Tables.matrix
-
-
-# # This will take a while I think
-# # Depends on grid size 
-# for row in eachrow(gender_age_year_combos)
-#     println(row)
-#     RetirementLifecycleWithIncome(
-#         bequests=false,
-#         age=row[2],
-#         gender=row[1],
-#         life_prob_types="objective",
-#         year=row[3]
-#     )
-# end
-
-
-
 # ----------------- SUBJECTIVE --------------------
 
 # First read in the idwaves that we have flagged for running
-col_types = Dict(:id_wave => String, :age => Int128, :year => Int128,
-    :ever_dc_pen => String, :ever_db_pen => String, :public_pension => Float64)
+col_types = Dict(:id_wave => String, :age => Int128, :year => Int128)
 
 idwave_df = CSV.File("../data/ELSA/elsa_to_use/for_julia.csv", types=col_types) |>
             DataFrame
+
 
 
 # Subjective death comes from "0__functions.jl"
@@ -42,27 +22,66 @@ ids_we_have_probs = unique(subjective_death_df[!, :id_wave])
 ids_we_need_lifecycle = idwave_df[:, :id_wave]
 
 # we only want to run ids we care about and that we have subjective life probs for
-ids_to_run = intersect(ids_we_need_lifecycle, ids_we_have_probs)
+ids_to_run = intersect(
+    ids_we_need_lifecycle,
+    ids_we_have_probs)
 
+length(ids_to_run)
 
 # Now run all 
-for id_wave in ids_to_run
+RunIdsFunction = function (group_of_ids)
+    for id_wave in group_of_ids
 
-    id_wave_row = idwave_df[idwave_df.id_wave.==id_wave, :]
-    age, gender, year = id_wave_row[!, :age][1], id_wave_row[!, :gender][1], id_wave_row[!, :year][1]
+        id_wave_row = idwave_df[idwave_df.id_wave.==id_wave, :]
 
-    println(id_wave)
[…]

### 2023-07-24 Julia/3__analyse_models.jl

@@ -90,7 +90,7 @@ RetrieveValueFunction(false, "objective", 60, 2011, "male")[40, 200, 1] # withou
 
 
 # should get measure of wealth aswell
-idwave_df = idwave_df[Bool.(.!ismissing.(idwave_df.fin_wealth)) .& Bool.(.!ismissing.(idwave_df.public_pension)), :]
+idwave_df = idwave_df[Bool.(.!ismissing.(idwave_df.fin_wealth)).&Bool.(.!ismissing.(idwave_df.public_pension)), :]
 
 
 # so we want gender age year and grid points
@@ -102,7 +102,7 @@ OptimalAnnuityAmount = function (loading_factor=0.85; gender="male", age=65, yea
 
     # comparing different starting values on the income-asset array
     # and the loading factor gives the cost of the tradeoff. I.e. how far we move on each one
-    
+
     value_func = RetrieveValueFunction(false, "objective", age, year, gender)[:, :, 1]
     value_of_start = value_func[income_start_point, asset_start_point]
 
@@ -147,7 +147,7 @@ OptimalAnnuityAmount = function (loading_factor=0.85; gender="male", age=65, yea
         # return the optimal amount to spend on annuities and the
         return max_value_and_index
     elseif max_value_and_index[1] < value_of_start
-        return ("do not annuitise" ,value_of_start)
+        return ("do not annuitise", value_of_start)
     end
 
 end
@@ -171,18 +171,18 @@ end
 
 
 
-out = OptimalAnnuityAmount(0.85,  gender="male", age=65, year=2015, asset_start_point=200, income_start_point=50)
+out = OptimalAnnuityAmount(0.85, gender="male", age=65, year=2015, asset_start_point=200, income_start_point=50)
 typeof(out)
 
-df = DataFrame(A = Float64[], B = Int64[])
+df = DataFrame(A=Float64[], B=Int64[])
 
 
-potential_loading_factors = [i/100 for i in 50:100]
+potential_loading_factors = [i / 100 for i in 50:100]
 
 
-append!(df, out )
+append!(df, out)
 # now see what factor people stop demanding annuities at 
-for fac in potential_loading_factors 
+for fac in potential_loading_factors
     OptimalAnnuityAmount(fac; gender="male", age=65, year=2015, asset_start_point=200, income_start_point=50)
-end 
+end
 # why is this so slow

### 2023-07-24 Julia/5__run_bequest_functions.jl

@@ -1,0 +1,81 @@
+
+
+include("0__functions.jl")
+include("1__lifecycle_with_income_function.jl")
+include("4__bequest_functions.jl")
+
+years = cat(repeat([2005], 26), [i for i in 2006:2015], dims=1)
+ages = cat([i for i in 50:75], repeat([50], 10), dims=1)
+
+year_ages_to_run_bequest_model = [years ages]
+
+birth_years = year_ages_to_run_bequest_model[:, 1] - year_ages_to_run_bequest_model[:, 2]
+
+println(year_ages_to_run_bequest_model)
+
+# ------------- run bequest models ------------------
+
+RunAllBequests = function ()
+
+    for row in eachrow(year_ages_to_run_bequest_model)
+        year = row[1]
+        age = row[2]
+        print(row)
+        RetirementLifecycleWithIncomeBequests(
+            bequests=true,
+            age=age,
+            gender="male",
+            life_prob_types="objective",
+            year=year,
+            id_wave="",
+            c0=20000)
+    end
+
+
+    for row in eachrow(year_ages_to_run_bequest_model)
+        year = row[1]
+        age = row[2]
+        RetirementLifecycleWithIncomeBequests(
+            bequests=true,
+            age=age,
+            gender="female",
+            life_prob_types="objective",
+            year=year,
+            id_wave="",
+            c0=20000)
+    end
+
+end
+
+# ------------------ run non-bequest models ------------------
+
+RunAllNonBequest = function ()
+
+    for row in eachrow(year_ages_to_run_bequest_model)
+        year = row[1]
+        age = row[2]
+        print(row)
+        RetirementLifecycleWithIncomeBequests(
+            bequests=false,
+            age=age,
+            gender="male",
+            life_prob_types="objective",
+            year=year,
+            id_wave="",
+            c0=20000)
+    end
+
+
+    for row in eachrow(year_ages_to_run_bequest_model)
+        year = row[1]
+        age = row[2]
+        RetirementLifecycleWithIncomeBequests(
+            bequests=false,
+            age=age,
+            gender="female",
+            life_prob_types="objective",
+            year=year,
+            id_wave="",
+            c0=20000)
+    end
+end

### 2023-07-24 Julia/6__analyse_more_models.jl

@@ -4,6 +4,7 @@
 
 include("0__functions.jl")
 include("1__lifecycle_with_income_function.jl")
+include("4__bequest_models.jl")
 
 # first get data
 col_types = Dict(:id_wave => String, :age => Int128, :year => Int128,
@@ -12,24 +13,24 @@ col_types = Dict(:id_wave => String, :age => Int128, :year => Int128,
 
 idwave_df = CSV.File("../data/ELSA/elsa_to_use/for_julia.csv", types=col_types) |>
             DataFrame
-idwave_df = idwave_df[Bool.(.!ismissing.(idwave_df.fin_wealth)) .& Bool.(.!ismissing.(idwave_df.public_pension)), :]
-      
+idwave_df = idwave_df[Bool.(.!ismissing.(idwave_df.fin_wealth)).&Bool.(.!ismissing.(idwave_df.public_pension)), :]
+
 idwave_df.pen_grid = FindClosestGridPoint.(
     idwave_df.public_pension, inc_grid_gap, inc_grid_points)
 
 idwave_df.ass_grid = FindClosestGridPoint.(
     idwave_df.fin_wealth, asset_grid_gap, asset_grid_points)
 
-    sum(idwave_df.pen_grid .<= 100 )
-    sum(idwave_df.ass_grid. <=500  .& idwave_df.pen_grid .<= 100)
+sum(idwave_df.pen_grid .<= 100)
+sum((idwave_df.ass_grid .<= 500) & (idwave_df.pen_grid .<= 100))
 
 
-good_assets = idwave_df.ass_grid .<=500   
+good_assets = idwave_df.ass_grid .<= 500
 good_inc = idwave_df.pen_grid .<= 100
 
 # filter too wealthy and too high income
-idwave_df = idwave_df[good_assets .& good_inc, :]
- 
+idwave_df = idwave_df[good_assets.&good_inc, :]
+
 
 optimal_asset = []
 
@@ -37,23 +38,23 @@ for row in eachrow(idwave_df)
     println([row["pen_grid"] row["ass_grid"]])
     output_val = ToAnnuitiseOrNot(
         row["year"],
-        row["age"], 
+        row["age"],
         row["gender"],
-        row["pen_grid"], 
+        row["pen_grid"],
         row["ass_grid"])
 
     push!(optimal_asset, output_val)
-end 
+end
 
 
 
-sum(optimal_asset .==  "do not annuitise")
+sum(optimal_asset .== "do not annuitise")
 
[…]

### 2023-07-24 Julia/6__run_bequest_functions.jl

@@ -1,142 +1,0 @@
-
-
-include("0__functions.jl")
-include("1__lifecycle_with_income_function.jl")
-include("5__bequest_functions.jl")
-
-years = cat(repeat([2005], 26), [i for i in 2006:2015], dims=1)
-ages = cat([i for i in 50:75], repeat([50], 10), dims=1)
-
-year_ages_to_run_bequest_model = [years ages]
-
-birth_years = year_ages_to_run_bequest_model[:, 1] - year_ages_to_run_bequest_model[:, 2]
-
-birth_years
-
-println(year_ages_to_run_bequest_model)
-
-for row in eachrow(year_ages_to_run_bequest_model)
-    year = row[1]
-    age = row[2]
-    print(row)
-    # RetirementLifecycleWithIncomeBequests(
-    #         bequests=true,
-    #         age=age,
-    #         gender="male",
-    #         life_prob_types="objective",
-    #         year=year,
-    #         id_wave="", 
-    #         c0 = 20000)
-end
-
-
-for row in eachrow(year_ages_to_run_bequest_model)
-    year = row[1]
-    age = row[2]
-    RetirementLifecycleWithIncomeBequests(
-        bequests=true,
-        age=age,
-        gender="female",
-        life_prob_types="objective",
-        year=year,
-        id_wave="",
-        c0=20000)
-end
-
-
-# going to quickly change the names to birth-year gender 
-# as opposed to year and age
-
-# pwd()
-
-# cp() # copy files
-# mv() # move and rename files
-
-
-# Ok now we want to check optimal annuity rates in our real life data. 
-# Need a function that takes year, age, wealth and income 
-# and calculates optimal annuity amount, and consumption 
-
-# first step in this is getting the json
-# to do this we calc birth year
-# and then see which of the bequest files line up with a year-age that was allowed
-
-GetJson = function (
-    birth_year=1955,
-    gender="female",
-    bequests=true,
-    life_prob_types="subjective")
-
-    if bequests == true
-        folder_name = "../Data/ELSA/lifecycle_outputs/beq_testing/"
-        if life_prob_types == "subjective"
-            json_string = string("birth_year_", birth_year, "_true_", gender, "_objective_20000.json")
-            full_file_name = folder_name * json_string
-            println(full_file_name)
-            if isfile(full_file_name)
-                json_in = JSON3.read(full_file_name)
-            else
[…]

### 2023-07-24 Julia/7__analyse_subjective_models.jl

@@ -1,17 +1,16 @@
 
 json_idwave_folder = readdir("../data/ELSA/lifecycle_outputs/id_wave/")
 
-GetJSON_IDWave = function(id_wave)
+GetJSON_IDWave = function (id_wave)
     file_to_get = json_idwave_folder[findfirst(occursin.(id_wave, json_idwave_folder))]
-    
-    json_filename = "../data/ELSA/lifecycle_outputs/id_wave/" * 
-        file_to_get
-        
+
+    json_filename = "../data/ELSA/lifecycle_outputs/id_wave/" *
+                    file_to_get
+
     JSON3.read(json_filename)
-end 
+end
 
 # now we can use similar code to the bequest one
-
 
 # first get data
 col_types = Dict(:id_wave => String, :age => Int128, :year => Int128,
@@ -21,26 +20,26 @@ col_types = Dict(:id_wave => String, :age => Int128, :year => Int128,
 idwave_df = CSV.File("../data/ELSA/elsa_to_use/for_julia.csv", types=col_types) |>
             DataFrame
 
-idwave_df = idwave_df[Bool.(.!ismissing.(idwave_df.fin_wealth)) .& Bool.(.!ismissing.(idwave_df.public_pension)), :]
-      
+idwave_df = idwave_df[Bool.(.!ismissing.(idwave_df.fin_wealth)).&Bool.(.!ismissing.(idwave_df.public_pension)), :]
+
 idwave_df.pen_grid = FindClosestGridPoint.(
     idwave_df.public_pension, inc_grid_gap, inc_grid_points)
 
 idwave_df.ass_grid = FindClosestGridPoint.(
     idwave_df.fin_wealth, asset_grid_gap, asset_grid_points)
 
-    sum(idwave_df.pen_grid .<= 100 )
-    # sum(idwave_df.ass_grid. <=500  .& idwave_df.pen_grid .<= 100)
+sum(idwave_df.pen_grid .<= 100)
+# sum(idwave_df.ass_grid. <=500  .& idwave_df.pen_grid .<= 100)
 
-    
+
 json_idwave_folder
 
-good_assets = idwave_df.ass_grid .<=500   
+good_assets = idwave_df.ass_grid .<= 500
 good_inc = idwave_df.pen_grid .<= 100
 
 # filter too wealthy and too high income
-idwave_df = idwave_df[good_assets .& good_inc, :]
- 
+idwave_df = idwave_df[good_assets.&good_inc, :]
+
 # make sure we have run the id_wave
[…]

### 2023-07-24 Julia/old/4__bequest_models.jl

@@ -4,6 +4,10 @@
 # cb 17.1 (1000)
 # m 0.96
 
+
+# This is an old testing file.
+# I have copied functions I want into another file. Move this to old 
+
 include("0__functions.jl")
 
 ## Set grid points
@@ -27,7 +31,7 @@ utility_array = CreateUtilityArray()
 bequest_lock(b; c0, m=0.9) = (m / (1 - m))^σ * (((m / (1 - m)) * c0) + b)^(1 - σ) / (1 - σ)
 # but won't including bequests make utility lower
 
-bequest_lock.([1000, 5000, 50000], c0 = 5000, m = 0.9)
+bequest_lock.([1000, 5000, 50000], c0=5000, m=0.9)
 
 # lets try running bequest function for different values of c0 and m 
 
@@ -39,8 +43,8 @@ RetirementLifecycleWithIncome_beq_edit = function (;
     gender="male",
     life_prob_types="objective",
     year=2015,
-    id_wave="", 
-    c0 = 20000)
+    id_wave="",
+    c0=20000)
 
     # two ways to run this function
     # either with objective probs given by age, gender and year
@@ -48,7 +52,7 @@ RetirementLifecycleWithIncome_beq_edit = function (;
 
     if life_prob_types == "objective"
         # set file name to save json to. 
-        json_filename = string(age, "_beq", bequests, "_", gender, "_", life_prob_types, "_", year, "_", c0,".json")
+        json_filename = string(age, "_beq", bequests, "_", gender, "_", life_prob_types, "_", year, "_", c0, ".json")
 
         dir_name = "../data/ELSA/lifecycle_outputs/beq_testing/"
         if !isdir(dir_name)
@@ -84,10 +88,10 @@ RetirementLifecycleWithIncome_beq_edit = function (;
 
     # Dimensions are the state variables
     # Income, current assets and age
-    value_array = ones(Float64, inc_grid_points, asset_grid_points, 1+ terminal_age - start_age)
+    value_array = ones(Float64, inc_grid_points, asset_grid_points, 1 + terminal_age - start_age)
     policy_array = ones(Float64, inc_grid_points, asset_grid_points, terminal_age - start_age)
 
-    bequest_value = [bequest_lock(assets, c0 = c0, m = 0.94) for assets in asset_grid]
+    bequest_value = [bequest_lock(assets, c0=c0, m=0.94) for assets in asset_grid]
 
     # for each income level find the optimal asset policy function
     for inc in eachindex(income_grid)
[…]

### 2023-07-24 R/10__1_regressions_prep_data.R

@@ -21,7 +21,7 @@ pension_wealth_dt <- fread(
 )
 
 # ------- Life expectancy data -----------
-obj_life_exps_dt <- fread("../../data/ONS/objective_life_expectancies.csv")[, gender := fcase(
+obj_life_exps_dt <- fread("../../data/ONS/objective_life_expectancies.csv")[, gender_string := fcase(
     gender == "Males", "male",
     gender == "Females", "female"
 )]
@@ -37,27 +37,26 @@ harm_long <- fread(
 harm_long[, int_year := year(int_month_date)]
 harm_long[, birth_year := int_year - age_at_interview]
 harm_long[, id_wave := paste0(idauniq, "-", wave)]
-harm_long[, gender := fcase(
-    ragender == 1, "male",
-    ragender == 2, "female"
+harm_long[, ragender := fcase(
+    ragender == 1, 0,
+    ragender == 2, 1
+)]
+
+
+harm_long[, gender_string := fcase(
+    ragender == 0, "male",
+    ragender == 1, "female"
 )]
 
 harm_long[pension_wealth_dt, dc_pot := dc_pot, on = c("idauniq", "int_year" = "year")]
 
 harm_long[obj_life_exps_dt, obj_life_exps := life_exp,
-    on = c("age_at_interview" = "age", "gender", "birth_year")
+    on = c("age_at_interview" = "age", "gender_string", "birth_year")
 ]
 
 harm_long[sub_life_exps_dt, sub_life_exps := subjective_life_exp,
     on = c("id_wave")
 ]
-
-summary(harm_long$sub_life_exps, na.rm = T)
-mean(is.na(harm_long$sub_life_exps))
-
-harm_long %>%
-    filter(is.na(obj_life_exps)) %>%
-    select(age_at_interview, gender, birth_year)
 
 harm_long[, dc_pot := as.numeric(dc_pot)]
 
@@ -145,15 +144,8 @@ data_for_regressions <- harm_long[ret_in == TRUE & !is.na(pre_post_ref)]
 
 data_for_regressions[, id_wave := paste0(idauniq, "-", wave)]
 
-
-data_for_regressions[, ragender := ragender - 1] # make it binary rather than 1,2
-# Check that this is the case
-stopifnot(all(data_for_regressions$ragender %in% c(0, 1)))
-
-
-
-data_for_regressions[, ever_dc_pen_bin := ever_dc_pen == "yes"]
[…]

### 2023-07-24 R/10__2_summay_stats.R

@@ -32,6 +32,7 @@ data_for_regressions <- fread("../../data/ELSA/elsa_to_use/elsa_reg_data.csv") %
 search_names(data_for_regressions, "pen")
 search_names(data_for_regressions, "year")
 
+data_for_regressions$ragender %>% table()
 
 
 
@@ -50,8 +51,9 @@ ss_stat_out <- data_for_regressions %>%
             sd = \(x) sd(x, na.rm = T),
             min = \(x) min(x, na.rm = T),
             max = \(x) max(x, na.rm = T),
-            median = \(x) median(x, na.rm = T),
-            non_missing = \(x) sum(!is.na(x))
+            median = \(x) median(x, na.rm = T)
+            # ,
+            # non_missing = \(x) sum(!is.na(x))
         ),
         .names = "{.col}__{.fn}"
     )) %>%

### 2023-07-24 R/10__3_run_regressions.R

@@ -2,6 +2,8 @@ library(dplyr)
 library(ggplot2)
 library(modelsummary)
 
+source("./__ELSA_functions.R")
+
 data_for_regressions <- fread(
     "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
 ) %>%
@@ -10,11 +12,15 @@ data_for_regressions <- fread(
         house_value = house_value / 1000,
         dc_pot = dc_pot / 1000,
         public_pension = public_pension / 1000
+    ) %>%
+    mutate(
+        pre_post_ref_bin = fcase(
+            pre_post_ref == "treat", 1,
+            pre_post_ref == "control", 0
+        )
     )
 
-
-
-
+clean_names <- append(clean_names, c("PostReform" = "pre_post_ref_bin"), 0)
 
 # #### Plot consumption
 # consump_time_in <- harm_long[ret_in == TRUE,
@@ -43,41 +49,29 @@ data_for_regressions <- fread(
 # similar apart from for the reform (after controlling for vars)
 
 
-lm(total_monthly_consumption ~ pre_post_ref,
+lm(total_monthly_consumption ~ pre_post_ref_bin,
     data = data_for_regressions
 )
 
-lm(total_monthly_consumption ~ pre_post_ref + age_at_interview,
+lm(total_monthly_consumption ~ pre_post_ref_bin + age_at_interview,
     data = data_for_regressions
 )
 
 lm(
-    total_monthly_consumption ~ pre_post_ref + age_at_interview + fin_wealth +
-        gender + ever_dc_pen + ever_dc_pen * pre_post_ref + years_since_retirement + owns_house,
+    total_monthly_consumption ~ pre_post_ref_bin + age_at_interview + fin_wealth +
+        ragender + ever_dc_pen_bin + ever_dc_pen_bin * pre_post_ref_bin + years_since_retirement + owns_house,
     data = data_for_regressions
 )
 
 
 lm(
-    total_monthly_consumption ~ pre_post_ref + age_at_interview + fin_wealth +
-        gender + dc_pot + years_since_retirement + owns_house + public_pension,
+    total_monthly_consumption ~ pre_post_ref_bin + age_at_interview + fin_wealth +
+        ragender + dc_pot + years_since_retirement + owns_house + public_pension,
     data = data_for_regressions[ever_dc_pen == "yes"]
 )
 
-# ------- Other consumption variables -------
-
-rhs <- "pre_post_ref + retired_age + fin_wealth +
[…]

### 2023-07-24 R/13__regression_with_lifecycle.R

@@ -1,50 +1,240 @@
 library(data.table)
 library(dplyr)
 library(ggplot2)
-library(jsonlite)
+library(stringr)
+library(modelsummary)
+
+source("./__ELSA_functions.R")
+clean_names <- append(clean_names, c("PostReform" = "pre_post_ref_bin"), 0)
+
+# Reading data for regressions
+
+bequest_consump_dt <- fread("../../data/ELSA/lifecycle_outputs/bequest_consumption_simulation.csv") %>%
+    rename(
+        consump_beq_ann = annuity_consump,
+        consump_beq_no_ann = no_annuity_consump
+    )
+
+standard_consump_dt <-
+    fread("../../data/ELSA/lifecycle_outputs/standard_model_consumption_simulation.csv") %>%
+    rename(
+        consump_stan_ann = annuity_consump,
+        consump_stan_no_ann = no_annuity_consump
+    ) %>%
+    select(id_wave, consump_stan_ann, consump_stan_no_ann)
+
+# standard_consump_dt[, .(annuity_cost, annuity_payment)]
+# bequest_consump_dt[, .(annuity_cost, annuity_payment)]
+
+reg_dt <- fread(
+    "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
+) %>%
+    mutate(
+        fin_wealth = fin_wealth / 1000,
+        house_value = house_value / 1000,
+        dc_pot = dc_pot / 1000,
+        public_pension = public_pension / 1000
+    ) %>%
+    mutate(
+        pre_post_ref_bin = fcase(
+            pre_post_ref == "treat", 1,
+            pre_post_ref == "control", 0
+        )
+    )
 
 
-# So I take the individuals who are untreated
-# i.e. early
-
-# estimate their bequest motive so that it fits the rate of
-# annuitisation with forced annuitisation of DC pensions
-# then run regression on the difference as we do in real life
-
-# first read objective no bequests
-# need to price annuities
-# one annuity price for everyone
-# based on annuitant lifetables
-#   i should check whether they are predictions or if they are what actually happened
-
-# ------------ read in jsons ------------------
-
-## Ok now lets read in and see the rate of annuitisation we would expect
-dir("../../data/ELSA/lifecycle_outputs/", full.names = T)
+# change so that we have c(unclean_name = clean_name)
[…]

### 2023-07-24 Julia/2__run_models.jl
Latest changes

@@ -26,7 +26,6 @@ ids_to_run = intersect(
     ids_we_need_lifecycle,
     ids_we_have_probs)
 
-length(ids_to_run)
 
 # Now run all 
 RunIdsFunction = function (group_of_ids)
@@ -49,59 +48,7 @@ RunIdsFunction = function (group_of_ids)
     end
 end
 
-# mmm it uses loads of ram 
 
-groups_to_split = round.(Int, rand(length(ids_to_run)) * num_groups)
+rev_ids_to_run = reverse(ids_to_run)
 
-
-
-
-v = [i for i in 1:10]
-splits = 3
-
-
-
-V = [Vector{eltype(v)}(undef, 1) for _ in 1:splits]
-
-V[1] = [1, 2, 3]
-V
-
-
-
-# SplitVector = function (v, splits)
-
-# check if not perfect length 
-
-if !isinteger(length(v) / (splits))
-    len_out = floor(Int, length(v) / (splits - 1))
-    remainder = length(v) - len_out * (splits - 1)
-end
-
-
-groups = round.(Int, rand(1000) * 6)
-
-
-# just want to split a vector into 6 parts
-
-# GetSubjectiveDeathProbs(id_wave_to_get="100021-5")
-
-
-
-# Nice this is working now!!
-# different life expectancies & different years and genders (these are kinda the same)
-
-# get a unique df with all the types we will need
-
-
-# ################################# TO DO #################################
-# # 
-
-# # then take individuals from elsa and estimate what consumption should be
-# # under the different model types i.e. conditional on assets, age, object live probs and 
-# # then both bequests and subjective life probs
-
-# # then we can run regressions on the impact of forced annuitisation
-
-# # there is annuity price data vs income drawdown on fca website which is cool 
-
-
+RunIdsFunction(rev_ids_to_run)

### 2023-07-24 R/13__regression_with_lifecycle.R

@@ -234,7 +234,23 @@ mean(joint_dt$id_wave %in% ids_done)
 
 ids_not_done <- joint_dt$id_wave[!joint_dt$id_wave %in% ids_done]
 
-# ok these are not done
+ids_that_we_can_do <- dir("../../data/ELSA/subjective_tables/jsons/") %>%
+    str_extract("\\d{6}-\\d{1}")
 
 
-"107746-8"
+length(intersect(ids_not_done, ids_that_we_can_do))
+
+
+all_sub_jsons <- dir("../../data/ELSA/subjective_tables/jsons/")
+
+files_to_copy <- paste0(
+    "../../data/ELSA/subjective_tables/jsons/",
+    ids_not_done,
+    ".json"
+)
+
+new_names <- paste0("../../../../temp_data_dir/", ids_not_done, ".json")
+
+file.copy(files_to_copy, new_names, overwrite = TRUE)
+
+length(files_to_copy) == length(new_names)

### 2023-07-25 Texfiles/Draft1.tex
More words finally

@@
 
+\textbf{I don't actually know why I do this}
 In retiree's first year of retirement I allow them to choose to annuitise some of their wealth. In practical terms
@@
 
-I also simulate the expected change in consumption given different rates of annuitisation versus no annuitisation.
-The benefit of simulating an individual's decision is that we can directly compare what they would have
-consumed with what they actually consumed, and we can replicate the empirical analysis using the
-solution to the lifecycle model.
+\section{Lifecycle Results}
+
+Having found optimal policy functions and value functions I can now simulate the change in consumption given
+different rates of annuitisation versus no annuitisation.
+There are multiple benefits to building a model of individuals decisions. Firstly, we can directly compare
+what a given individual would have consumed with what they actually consumed, conditional on our model being
+correct. In other words we have the counterfactual world in which individuals were not forced to annuitise.
+Secondly, we can simulate the impact of different policy changes on the consumption, savings and bequest behaviour.
+For example, I can change the proportion of wealth that will be annuitised and see the corresponding changes in
+consumption indicators for all individuals.
+
+I start this section by plotting the retirement path of consumption for a median individual in
+the ELSA data set. I plot consumption with different proportions of starting wealth annuitised
+and the different lifecycle model types.
+
+\begin{figure}[h]
+    \caption{Consumption in retirement}
+    \centering
+    \includegraphics[width=0.7\columnwidth]{figures/consumption_plot_median_retiree.pdf}
+    \label{fig:ConsumpPlot}
+\end{figure}
+
+\begin{figure}[h]
+    \caption{Assets in retirement}
+    \centering
+    \includegraphics[width=0.7\columnwidth]{figures/asset_plot_median_retiree.pdf}
+    \label{fig:AssetPlot}
+\end{figure}
+
+Figures \ref{fig:ConsumpPlot} and \ref{fig:AssetPlot} show the retirement
+path of consumption and assets for a median retiree. This hypothetical women has
+120,000 of assets at age 65 when they retired in 2012 with a state pension income of 6000.
+The dashed lines show consumption and assets with an annuity that costs half of their wealth.
+This buys them 3500 of annuity income, which, given their objective life expectancies is a money's
+worth of about 90\%.
+
+We can see that the asset path with the bequest motive keeps more for longer but still decreases
+and by their late 90s they have completely run down their wealth. Interestingly the asset path
+with the quickest run down is the standard lifecycle model using objective life expectancies,
+which points towards this individual being infact optimistic about their life chances.
+Also rather surprisingly the asset paths of all three types are similar without annuities.
+The paths only diverge after age 75 at which point the bequest path stays high and the
+other paths start to decrease faster. This is contrary to the path with annuities where
+the standard path diverges from the bequest and subjective paths immediately.
+
+The consumption paths all start at around 1400. The three models without an annuity start at
+exactly the same amount and stay that way until the individual reaches her mid seventies -- showing
+that even quite different lifecycle models can generate similar paths. As expected, the consumption
+path with the bequest motive with an annuity is lower at the start of retirement than the bequest path 
+without an annuity as individuals seek to build their savings back up.
+But it is also lower for the subjective life expectancy case,
+again pointing toward this individual being optimistic. However, the subjective consumption
+line rises at a younger age than the line with bequests, showing that the models align 
+with the prediction that subjective lifecycle agents consume more in early retirement.
+
+The period of interest in this study is the first three years of retirement. For this period 
+we do see substantially differences in consumption for the diffe
 
@@
 \begin{landscape}
-
     \input{tables/standard_life_cycle.tex}
@@
 
-Table \ref{tab:StandardLifeCycle} shows the results from the bequest models.
+Table \ref{tab:StandardLifeCycle} shows the results from the standard model lifecycle when individuals use ONS
+for risk of death and have no bequest function. In this case .
 

### 2023-07-25 Julia/0__functions.jl

@@ -168,7 +168,8 @@ GetJson = function (
         vector_of_matches = .!(isnothing.(match.(regex_search, files_in_folder)))
 
         # there should be a unique match
-        if sum(vector_of_matches) != 1
+        if sum(vector_of_matches) == 0
+
             error("Not a match for this json string")
         end
 
@@ -218,7 +219,7 @@ RetrievePolicyFunction = function (
 end
 
 
-function AssetPathFunction(
+AssetPathFunction = function (
     opt_policy_function,
     asset_start_point,
     income_grid_point)

### 2023-07-25 Julia/2__run_models.jl

@@ -21,10 +21,26 @@ ids_we_have_probs = unique(subjective_death_df[!, :id_wave])
 # these are ids that we will use in the regression
 ids_we_need_lifecycle = idwave_df[:, :id_wave]
 
+
+folder_matches = match.(
+    r"(?<=_)(\d{6}-\d{1})",
+    readdir("../Data/ELSA/lifecycle_outputs/id_wave/"))
+
+ids_we_have_done = [""]
+for i in 1:length(folder_matches)
+    append!(ids_we_have_done, [folder_matches[i].match])
+end
+ids_we_have_done
+
+
 # we only want to run ids we care about and that we have subjective life probs for
-ids_to_run = intersect(
+ids_we_can_and_we_want = intersect(
     ids_we_need_lifecycle,
     ids_we_have_probs)
+
+# we only want to run ids we havent run before 
+ids_to_run = ids_we_can_and_we_want[ids_we_can_and_we_want.∉Ref(ids_we_have_done)]
+
 
 
 # Now run all 
@@ -48,6 +64,7 @@ RunIdsFunction = function (group_of_ids)
     end
 end
 
+# RunIdsFunction(ids_to_run)
 
 rev_ids_to_run = reverse(ids_to_run)
 

### 2023-07-25 Julia/8__analyse_all_models.jl

@@ -30,6 +30,119 @@ names_vec = vec(["id_wave",
 "annuity_payment"])
 
 
+
+# ------------------ Life cycle plots ------------------
+# ret_year = 2012
+# int_year = 2014
+# fin_wealth = 120,000 (mean) => 120 grid 
+# pension = 6,000 => 12
+# id_wave 107132-7
+# gender = "female"
+# ret_age = 65
+# birth_year = 1947
+
+# No bequests, objective life probs
+stan_pol_func = RetrievePolicyFunction(
+    false, "objective", 1947, "female",  "107132-7"
+)
+
+beq_pol_func = RetrievePolicyFunction(
+    true, "objective", 1947, "female",  "107132-7"
+)
+
+sub_pol_func = RetrievePolicyFunction(
+    false, "subjective", 1947, "female",  "107132-7"
+)
+
+# the other policy functions are too long so we trim them down so that it starts at 
+# retirement
+
+# maybe also trim them a few years
+beq_pol_func = beq_pol_func[:, :, 10:end]
+stan_pol_func = stan_pol_func[:, :, 10:end]
+
+# identical in every other respect
+using Plots
+
+plot([AssetPathFunction(stan_pol_func, 120, 12).cons AssetPathFunction(beq_pol_func, 120, 12).cons AssetPathFunction(sub_pol_func, 120, 12).cons]) 
+
+
+Annuity_GeneralDeathProbs(
+    60000,
+    0.90, 
+    GetObjectiveDeathProbs(gender = "female", year = 2012, age = 65)    
+)
+# So move down 60 asset grids
+# and move up 7 income grids
+
+# ---- Consumption Plot ----------
+consump_paths = hcat(
+    AssetPathFunction(stan_pol_func, 120, 12).cons, 
+    AssetPathFunction(stan_pol_func, 60, 19).cons,
+    AssetPathFunction(beq_pol_func, 120, 12).cons,
+    AssetPathFunction(beq_pol_func, 60, 19).cons, 
+    AssetPathFunction(sub_pol_func, 120, 12).cons, 
+    AssetPathFunction(sub_pol_func, 60, 19).cons
+)
+lables = ["Standard" "Standard with annuity" "Bequest" "Bequest with annuity" "Subjective" "Subjective with annuity"]
+
+all_ages = [64 + i for i in 1:size(sub_pol_func, 3)]
+
[…]

### 2023-07-26 Texfiles/Draft1.tex
latest edits

@@
 \usepackage{siunitx}
-\newcolumntype{d}{S[
-      input-open-uncertainty=,
-      input-close-uncertainty=,
-      parse-numbers = false,
-      table-align-text-pre=false,
-      table-align-text-post=false
-    ]}
+\newcolumntype{d}{S[ input-open-uncertainty=, input-close-uncertainty=,
+      parse-numbers = false, table-align-text-pre=false,
+      table-align-text-post=false ]}
 \usepackage{pdflscape}
@@
 
-\hypersetup{
-    colorlinks,
-    citecolor=black,
-    filecolor=black,
-    linkcolor=black,
-    urlcolor=black
-}
+\hypersetup{ colorlinks, citecolor=black, filecolor=black, linkcolor=black,
+    urlcolor=black }
 \setlength{\parindent}{0pt}
 \usepackage[parfill]{parskip}
-\date{Lent 2023}
-\title{Dissertation draft}
-\author{Tobias Leigh-Wood}
+\date{July 2023}
+\title{Retirement Consumption: Evidence from UK Pension Reform}
+\author{INSERT NUMBER HERE}
 \usepackage[letterpaper, margin=1.1in]{geometry}
-\linespread{1.1}
+\linespread{1.25}
 
@@
 
+\newpage
 \maketitle
@@
 
-Standard economic theory suggests that retirement annuities, which provide a guaranteed income stream until the end of life,
-should be highly valued by individuals as a way to insure against late death \cite{yaari_65}. However, in developed countries,
-rates of annuitisation are far below the levels that theory predicts. The literature has suggested several reasons for
-this, but there is no consensus about which mechanism is dominant. I exploit early retirees' consumption responses to a
-reform to annuity policy in the UK to add to this literature.
-    [\textbf{Add something once I have a conclusion}].
+Classical economic theory suggests that retirement annuities, which provide a
+guaranteed income stream until the end of life, should be highly valued by
+individuals as a way to insure against late death \cite{yaari_65}. However, in
+developed countries, rates of annuitisation are far below the levels that theory
+predicts - this phenomenon has been termed the "annuity puzzle". The literature
+has suggested several reasons for this, but there is no consensus about which
+mechanism is dominant. I exploit early retirees' consumption responses to a
+reform to annuity policy in the UK to add to this literature. [\textbf{Add
+            something once I have a conclusion}].
 
-In the UK, adults have three methods of funding retirement: the state pension; defined benefit (DB) employee pension schemes,
-which provide an income in retirement based on years of service and a function of wages; and defined contribution (DC) pensions,
-in which individuals save and invest in a tax-advantaged account that is usually supplemented with employer contributions.
-Under the 2010-2015 coalition government in the UK, the law regarding the use of private DC pensions at
-retirement changed. Individuals were no longer forced to annuitise, which provides a constant income until death,
-their pension pots and could access them in a variety of ways such as a lump sum withdrawal or income drawdown,
-which is a steady withdrawal of assets
-with the remainder left in the pot - common advice is to take 4\% a year.
-Subsequently the number of annuities sold in the UK dropped precipitously.
+In the UK, adults have three methods of funding retirement: the state pension;
+defined benefit (DB) employee pension schemes, which provide an income in
+retirement based on years of service and a function of wages; and defined
+contribution (DC) pensions, under which individuals save and invest in a
+tax-advantaged account that is usually supplemented with employer contributions.
+Under the 2010-2015 coalition government in the UK, the law regarding the use of
+private DC pensions at retirement changed. Individuals were no longer forced to
+annuitise their pension pots and could access them in a variety of ways, such as
+a lump sum withdrawal or income drawdown, involving a steady withdrawal of
+assets from the pension pot -- common advice is to take 4\% a year. Subsequently
+the number of annuities sold in the UK dropped precipitously.
 
-In this paper, I first use the policy reform as a discontinuity to measure the impact of forced annuitisation on the consumption
-levels of individuals in the first few years of retirement. Given that the reform was implemented suddenly and without
-advanced notice before the Spring 2014 budget, I claim that individuals who retired in 2015, 2016 or 2017
-are otherwise similar to those who retired year in 2011, 2012 and 2013 except for the fact the later retirees
-do not need to annuitise their DC pension pots. Therefore, I can use a regression discontinuity to identify
-the effect that forced annuitisation has on the consumption of retirees.
+In this paper, I first use the policy reform as a discontinuity to measure the
+impact of forced annuitisation on the consumption levels of individuals in the
+first few years of retirement. Given that the reform was implemented suddenly
+and without advanced notice before the Spring 2014 budget, I claim that
+individuals who retired in 2015, 2016 or 2017 are otherwise similar to those who
+retired in 2011, 2012 and 2013 except for the fact the later retirees do not
+need to annuitise their DC pension pots. Therefore, I can use a regression
+discontinuity to identify the effect that forced annuitisation has on the
+consumption of retirees.
 
 I then test two competing hypotheses for the annuity puzzle: bequests and
-pessimistic life expectancy. The annuity puzzle was first identified by \cite{yaari_65} and is a classed as a problem
-since standard economic theory cannot explain the lack of annuitisation that is observed in the real world.
-Two possible explanations are bequests and life expectancy: individuals want to leave an estate for their heirs to
-inherit and they cannot do this if they annuitise their wealth; or individuals may believe they will not live
-as long as annuity providers think they will and therefore annuities that are on the market do not appear to be
-good value. Depending on which reason explains the lack of annuitisation in the UK, the consumption response of
-retirees to the pension
-reform will differ. If individuals do not annuitise because of pessimistic life expectancy their
-consumption should increase after the reform. If, on the other hand, individuals do not annuitise because of a bequest motive, consumption
+pessimistic life expectancy. The annuity puzzle was first identified by
+\cite{yaari_65} and is a classed as a problem since standard economic theory
+cannot explain the lack of annuitisation that is observed in the real world. Two
+possible explanations are bequests and life expectancy: individuals want to
+leave an estate for their heirs to inherit and they cannot do this if they
+annuitise their wealth; or individuals may believe they will not live as long as
+annuity providers think they will and therefore annuities that are on the market
+do not appear to be good value. Depending on the reason for the lack of
+annuitisation in the UK, the consumption response of retirees to the pension
+reform will differ. If individuals do not annuitise because of pessimistic life
+expectancy their consumption should increase after the reform. If, on the other
+hand, individuals do not annuitise \ because of a bequest motive, consumption
 should not change much as a result of the reform.
 
-I will solve lifecycle models
-(\textbf{maybe add something about lifecycle models here})
-for both of these cases and simulate
-consumption decisions with and without forced annuitisation. I will then use a variety of empirical models to measure
-the consumption change in early retirement that resulted from the policy reform. The sign and magnitude of this change
-will be indicative of the mechanism causing the annuitisation problem.
+I will solve lifecycle models (\textbf{maybe add something about lifecycle
+    models here}) for both of these cases and simulate consumption decisions with
+and without forced annuitisation. I will then use a variety of empirical models
+to measure the consumption change in early retirement that resulted from the
+policy reform. The sign and magnitude of this change will be indicative of the
+mechanism causing the annuitisation problem.
 
-The importance of retirement policy to individuals in the UK is growing. The number of individuals of
-pensionable age is expected to increase from 11.9 million in 2020 to 15.2 million in
-2045 according to the latest Office for National Statistics (ONS) and for every 1000 people of working age there will be 341 of pensionable
-age in 2045 compared to 280 in 2020 \cite{ons_population_predictions_2020}. The increase in absolute and relative
-numbers of elderly people makes retirement policy more consequential. Moreover, private DC pensions are
-becoming increasingly common and are predicted to grow as current cohorts age \cite{cribb_karjalainen_ifs_2023}.
-Therefore, policies regarding how private pensions can be accessed will have a larger impact on overall welfare for
-retirees.
+The importance of retirement policy to individuals in the UK is growing. The
+number of individuals of pensionable age is expected to increase from 11.9
+million in 2020 to 15.2 million in 2045 according to the latest projections from
+the Office for National Statistics (ONS) and for every 1000 people of working
+age there will be 341 of pensionable age in 2045 compared to 280 in 2020
+\cite{ons_population_predictions_2020}. The increase in absolute and relative
+numbers of elderly people makes retirement policy more consequential. Moreover,
+private DC pensions are becoming increasingly common and are predicted to grow
+as current cohorts age \cite{cribb_karjalainen_ifs_2023}. Therefore, policies
+regarding how private pensions can be accessed will have a larger impact on
+overall welfare for retirees.
 
-Moreover, understanding how retirees spend their money over retirement is an important policy question of its own.
-If retirees spend too much in early retirement, the state may need to provide for them towards the end of their
-lives. This has implications for government budgets, especially since population aging means more individuals will
-require expensive end of life care. On the other hand, if retirees over-save and do not spend, the economy may face
-a dynamic inefficiency whereby consumption per capita could be increased if savings were decreased. (Get dynamic inefficiency
-reference.)
+Moreover, understanding how retirees spend their money over retirement is an
+important policy question of its own. If retirees spend too much in early
+retirement, the state may need to provide for them towards the end of their
+lives. This has implications for government budgets, especially since population
+aging means more individuals will require expensive end of life care. On the
+other hand, if retirees over-save and do not spend, the economy may be
+dynamically inefficient whereby the capital stock is too high because
+individuals save too much for retirement and for bequests. In this case
+consumption per capita could be increased if savings were decreased.
 
 \subsection{Literature review}
-My paper builds on three main strands of literature: that of the annuity puzzle, exploring reasons that some retirees choose not to annuitise;
-the retirement saving puzzle, looking at consumption and saving behaviour over retirement; and lifecycle models, whereby the above
-problems are explained using a tractable model of human decision making.
+My paper builds on three main strands of literature: that of the annuity puzzle,
+exploring reasons that some retirees choose not to annuitise; the retirement
+saving puzzle, which seeks to explain why retirees drawdown assets slowly, ; and
+lifecycle models, whereby the above problems are explained using a tractable
+model of human decision making.
 
-\cite{yaari_65} was the first to show that under standard assumptions we would expect individuals to
-annuitise all of their wealth at retirement to insure against the risk of long life. Since then there
-has been much literature discussing possible reasons that people do not annuitise. \cite{finkelstein_porteba_2002}
-and \cite{finkelstein_porteba_2004} find evidence of adverse selection, thereby making the 'money's worth'
-of annuities lower for the general population as opposed to the population of annuitants. However, they
-also find that theory would still predict annuitisation.
+\cite{yaari_65} showed that under standard assumptions we would expect
+individuals to annuitise all of their wealth at retirement to insure against the
+risk of long life. Since then there has been much literature discussing possible
+reasons that people do not annuitise. \cite{finkelstein_porteba_2002} and
+\cite{finkelstein_porteba_2004} find evidence of adverse selection, thereby
+making the 'money's worth' of annuities lower for the general population as
+opposed to the population of annuitants. However, they also find that theory
+would still predict annuitisation.
 
-\cite{friedman_warshawsky_qje_1990} show that annuitisation decisions can be fully explained by a mixture of
-bequest motives and actuarily unfair annuities. They solve an augmented life-cycle model with a range of
-parameters on how severe the rate of return is on the annuity versus market rates. For plausible values
-they find that individuals would optimally not annuitise much wealth. Similarly to \cite{finkelstein_porteba_2004},
-\cite{friedman_warshawsky_chicago_1988} show that there is a significant difference between the life expectancy of
-annuitants and the general population in the American annuity market but this cannot fully explain the annuitisation
-puzzle. Only when bequest motives are added to the model can annuitisation rates be rationalized.
+\cite{friedman_warshawsky_qje_1990} show that annuitisation decisions can be
+fully explained by a mixture of bequest motives and actuarily unfair annuities.
+They solve an augmented life-cycle model with a range of parameters on how
+severe the rate of return is on the annuity versus market rates. For plausible
+values they find that individuals would optimally not annuitise much wealth.
+Similarly to \cite{finkelstein_porteba_2004},
+\cite{friedman_warshawsky_chicago_1988} show that there is a significant
+difference between the life expectancy of annuitants and the general population
+in the American annuity market but this cannot fully explain the annuitisation
+puzzle. Only when bequest motives are added to the model can annuitisation rates
+be rationalized.
 
-\cite{lockwood_red_2012} builds on this and shows that a realistic bequest motive in lifecycle simulations achieves
-realistic annuitisation rates. He solves a simple lifecycle model with bequest motives taken from several recent papers
-in the literature. The bequest motives he picks therefore match other important aspects of the lifecycle model such
-as how much individuals actually bequest and how rich individuals are when they bequest.
+\cite{lockwood_red_2012} builds on this and shows that a realistic bequest
+motive in lifecycle simulations achieves realistic annuitisation rates. He
+solves a simple lifecycle model with bequest motives taken from several recent
+papers in the literature. The bequest motives he picks therefore match other
+important aspects of the lifecycle model such as how much individuals actually
+bequest and how rich individuals are when they bequest.
 
@@
 
-\cite{vidalmelia_lejarragagarcia_munich_2004} have some interesting results. Need to talk about that.
-
-\textbf{If I add 300 more words here then I will be at 5000!}
+\cite{vidalmelia_lejarragagarcia_munich_2004} have some interesting results.
+Need to talk about that.
 
@@
 
-The main data set I use is the English Longitudinal Study of Ageing (ELSA) \cite{main_elsa_citation}. ELSA picks individuals
-over the age of 50 to survey every two years until death. If individuals leave the survey ELSA replaces them so that it is
-representative of the over 50 population in the UK. Individuals are asked a range of questions relating to their income
-and wealth as well as expectations about the future. One benefit of using ELSA is that it includes data on pension types
-for individuals who are working. Therefore, I can differentiate between individuals who have a DC pension and those who have a DB pension.
-There have been 9 waves of ELSA data collection, the first was in 2001 and collections happened every two years thereafter.
+The main dataset I use is the English Longitudinal Study of Ageing (ELSA)
+\cite{main_elsa_citation}. ELSA picks individuals over the age of 50 to survey
+every two years until death. If individuals leave the survey, ELSA replaces them
+so that it is representative of the over 50 population in the UK. Individuals
+are asked a range of questions relating to their income and wealth as well as
+expectations about the future. One benefit of using ELSA is that it includes
+data on pension types for individuals who are working. Therefore, I can
+differentiate between individuals who have a DC pension and those who have a DB
+pension. There have been 9 waves of ELSA data collection -- the first was in
+2001 and collections happened every two years thereafter.
 
-Importantly ELSA also includes information on pension size that the Institute for Fiscal Studies (IFS) calculated, this is only available
-up until wave 5 (which was collected in 2010 and 2011) at which point I use a real return
-of 3\% to predict forward the pension wealth variable until retirement. ELSA also includes a measure of all
-non-housing financial wealth which I use in some specifications because of these issues with
-pension wealth. Due to slight differences in the ELSA questions between years, I use `Harmonized ELSA'
-\footnote{This analysis uses data or information from the Harmonized ELSA dataset and Codebook, Version G.2 as of
-    July 2021 developed by the Gateway to Global Aging Data. The development of the Harmonized ELSA was funded by the National
-    Institute on Aging (R01 AG030153, RC2 AG036619, R03 AG043052). For more information,
-    please refer to https://g2aging.org/.} which ensures that variables are comparable across waves. Since this only includes a
-subset of the questions in ELSA, I also supplement it with variables taken directly from the data such as questions that deal
-with life expectancy.
+Importantly, ELSA also includes information on pension size calculated by the
+Institute for Fiscal Studies (IFS). This is only available up to and including
+wave 5 (which was collected in 2010 and 2011) at which point I use a real return
+of 3\% to predict forward the pension wealth variable until retirement. ELSA
+also includes a measure of all non-housing financial wealth which I use in some
+specifications because of these issues with pension wealth data. Due to slight
+differences in the ELSA questions between years, I use `Harmonized ELSA'
+\footnote{This analysis uses data or information from the Harmonized ELSA
+    dataset and Codebook, Version G.2 as of July 2021 developed by the Gateway to
+    Global Aging Data. The development of the Harmonized ELSA was funded by the
+    National Institute on Aging (R01 AG030153, RC2 AG036619, R03 AG043052). For more
+    information, please refer to https://g2aging.org/.} which ensures that variables
+are comparable across waves. Since this only includes a subset of the questions
+in ELSA, I supplement it with variables taken directly from the data such as
+questions that deal with life expectancy.
 
-ELSA also includes questions on expenditures. In particular, individuals are asked how much they consume across a range of broad
-categories including in-house food consumption, out-of-house food consumption, leisure consumption, clothes consumption and
-consumption on utility bills and rent. I use also use two variables related to life expectancy that I discuss in more detail below.
+ELSA also includes questions on expenditures. In particular, individuals are
+asked how much they consume across a range of broad categories including
+in-house food consumption, out-of-house food consumption, leisure consumption,
+clothes consumption and consumption on utility bills and rent.
 
-I also use `life tables' from the UK's ONS. These provide me with risk of death for each age group. I adjust them
-to make death certain at age 110 as is common in the literature. I transform these so that I have risk of death conditional on being
-a given age since this is what I use in the life cycle simulations. I also use these objective probabilities to
-calculate annuity prices for individuals.
+I also use `life tables' from the UK's ONS. These provide me with the risk of
+death for each age group. Since the tables are produced until age 120, I adjust
+them to make death certain at age 110 as is common in the literature
+\cite{odea_sturrock_rest_2023}. I transform these so that I have risk of death
+conditional on being a given age since this is what I use in the life cycle
+simulations. I also use these objective probabilities to calculate annuity
+prices for individuals.
 
-To illustrate the effect the reform had on sales of annuities in the UK I obtained product data from the Financial Conduct Authority.
-These track the sales of different financial products overtime including data annuities.
+To illustrate the effect that the reform had on sales of annuities in the UK I
+obtained product data from the Financial Conduct Authority. These track the
+sales of different financial products overtime including data on annuities.
 
-Life expectancy impacts decision to retire for two reasons. Firstly, it directly impacts the price that an annuity will cost for
-individuals. Older individuals and those with pre-existing health conditions generally can buy an annuity that provides a greater
-income stream than individuals who are younger. Secondly, private information about life expectancy impacts the perceived
-value of an annuity. If an individual expects to outlive the general population an annuity, at a given price, will appear
-a much better deal to them. Likewise, an individuals life expectancy will change their decision on how to consume and save
-through retirement.
+Life expectancy impacts the decision to annuitise for two reasons. Firstly, it
+directly impacts the price that an annuity will cost for individuals. Older
+individuals and those with pre-existing health conditions generally can buy an
+annuity that provides a greater income stream than individuals who are younger.
+Secondly, private information about life expectancy impacts the perceived value
+of an annuity. If an individual expects to outlive the general population, an
+annuity, at a given price, will appear a much better deal to them. Likewise, an
+individual's life expectancy will change their decision on how to consume and
+save through retirement.
 
-To calculate subjective life expectancies I follow \cite{odea_sturrock_rest_2023}. Individuals are asked “What are the chances that you will
-live to be age X or more?” where X changes depending on the age of the interviewee. If individuals were under 65 then X was 75, if individuals
-were 66 and older they were asked the age that was 11 to 15 years older than them and is a multiple of 5. From wave three respondents were
-also asked “What are the chances that you will live to be age 85 or more?” if they were under 70. As most recent retirees are under 70 we
-therefore have two data points. I first drop from the data individuals who think it is more likely that they reach a higher age than a younger
-age since this shows a misunderstanding of the question. I then add as a third data point their objective chance of reaching 110 according to the ONS life tables.
-I fit these three points to a Weibull distribution, which is commonly used by demographers to estimate how populations
-will age, using non-linear least squares. Then I create subjective survival tables using parameters from the Weibull distribution.
+To calculate subjective life expectancies from ELSA data, I follow
+\cite{odea_sturrock_rest_2023}. Individuals were asked “What are the chances
+that you will live to be age X or more?” where X changed depending on the age of
+the interviewee. If individuals were under 65 then X was 75, if individuals were
+66 and older they were asked the age that was 11 to 15 years older than them and
+a multiple of 5. From wave 3 respondents were also asked “What are the chances
+that you will live to be age 85 or more?” if they were under 70. As most recent
+retirees are under 70 we therefore have two data points. I first drop from the
+data individuals who think it is more likely that they reach a higher age than a
+younger age since this shows a misunderstanding of the question. I then add as a
+third data point their objective chance of reaching 110 according to the ONS
+life tables. I fit these three points to a Weibull distribution, which is
+commonly used by demographers to estimate how populations will age, using
+non-linear least squares. Then I create subjective survival tables using
+parameters from the Weibull distribution.
 
@@
 
-Successive governments in the UK have attempted to reform both the public and private pension system. Prior
-to 1987, participation in private schemes was limited to employees of firms that had offered them, and there
-were few alternatives to the public state pension or defined benefit scheme that a public sector employer
-would offer. From 1989, individuals in the UK were allowed to open tax advantaged self-invested personal pensions
-alongside any pension their employer offered. The 2004 Finance Act rationalised taxation rates on different types of pensions.
-But DC pension pots could still only be used to buy an annuity after taking out a maximum of 25\% as a tax free lump sum withdrawal
-from age 55 onwards. Most pension pots also required that individuals access them before age 75.
+Successive governments in the UK have attempted to reform both the public and
+private pension system. Prior to 1987, participation in private schemes was
+limited to employees of firms that had offered them, and there were few
+alternatives to the public state pension or DB scheme that a public sector
+employer would offer. From 1989, individuals in the UK were allowed to open tax
+advantaged self-invested personal pensions alongside any pension their employer
+offered. The 2004 Finance Act rationalised taxation rates on DB and DC pensions.
+The rules for DC pension pots were such that individuals had to buy an annuity
+after optionally taking a maximum of 25\% of the pot as a tax free lump sum
+withdrawal. DC pension pots were accessible from age 55 and most required that
+they be accessed before the age of 75.
 
-Announced in the spring budget of 2014, the so called `pensions freedom act' received Royal Asset in December 2014 marked the end of a series
-of pension reforms carried out by the coalition government between 2010 and 2015. The reform made it possible
-to withdraw the first 25\% from a pension pot tax free and the rest subject to the marginal rate of income tax that an individual faced.
-In the June 2010 budget, the government made a first reform to the annuitisation rules, creating
-a minimum income requirement above which individuals would not need to
-annuitise more \cite{finance_act_hmt_2011} (this meant high income retirees would not need to annuitise their wealth).
-However, this was set at £20,000 and therefore few individuals
-were eligible. The minimum income requirement was scrapped in the 2014 bill, finally eliminating the
-compulsory annuities market, with one government minister saying pension pots `can be used to buy Lamborghinis'
-\cite{guardian_lambos}.
+In the June 2010 budget, the coalition government made the first of two key
+reforms it would make to the pension system between 2010 and 2015. The 2010
+policy reform created a minimum income requirement above which individuals would
+not need to annuitise more \cite{finance_act_hmt_2011}, this meant high income
+retirees would not need to annuitise their wealth. The income requirement was
+set at £20,000. This meant, for example, that an individual who was a member of
+a DB scheme paying them £20,000 a year would not need to annuitise their DC
+pension pot. However, the relatively high minimum income requirement meant that
+most individuals still had to purchase an annuity.
 
-The impact of the reforms on annuity demand has been documented by \cite{cannon_et_al_nier_2016}.
-Using data from the Association of British Insurers, they show that annuity demand dropped by 75\% from
-its maximum in 2011
-\textbf{Check}
-. Figure \ref{fig:annovertime} shows purchases of annuities overtime and demonstrates the sharp decrease
-in purchases that happened from 2014 to 2015. There was an increase in the number of pensions that were
-being accessed using an income drawdown product, but these do not fully account for the drop in annuitisation.
+In spring 2014, George Osbourne announced the so called `pensions freedom act'
+scrapping the minimum income requirement and eliminating the compulsory
+annuities market \cite{pen_freedoms_hmt_2014}. One government minister
+infamously brought the message home by saying pension pots `can be used to buy
+Lamborghinis' if retirees wanted to \cite{guardian_lambos}. The government also
+announced plans to make switches between DB and DC pension schemes possible
+although this was only realised in the next parliament -- future research could
+explore the impact of this second reform on the consumption and saving decisions
+of retirees.
+
+The impact of the reforms on annuity demand has been documented by
+\cite{cannon_et_al_nier_2016}. Using data from the Association of British
+Insurers, they show that annuity demand dropped by 75\% from its maximum in
+2011. Figure \ref{fig:annovertime} shows purchases of annuities overtime and
+demonstrates the sharp decrease in purchases that happened from 2014 to 2015.
+There was an increase in the number of pensions that were being accessed using
+an income drawdown product, but these only partly account for the drop in
+annuitisation as is shown by the graph.
 \begin{figure}[h]
-    \caption{Pension pots accessed }
+    \caption{New pension pots accessed }
     \centering
@@
     \label{fig:annovertime}
-
 \end{figure}
@@
 \end{figure}
-Figure \ref{fig:ann2122} shows the distribution of how pension pots were first accessed at retirement in 2021-22.
-In 2021, 196,736 pots were fully withdrawn at retirement, accounting for over 50\% of pots. Prior to the policy
-reform this was not the case -- most defined contribution pensions were accessed through annuities.
-\footnotetext{Partial UFPLS (uncrystallised funds pension lump sum) is similar to Partial drawdown but with a different tax schedule. With partial drawdown
-    you take the whole 25\% tax free amount at once whereas with UFPLS you only claim the 25\% on the amount you are taking out. This option is
-    better if you may want to buy another retirement product in the future such as an annuity. }
+Figure \ref{fig:ann2122} shows the distribution methods to access pension pots
+at retirement in 2021-22. In 2021, 196,736 pots were fully withdrawn at
+retirement, accounting for over 50\% of pots. Prior to the policy reform this
+was not the case -- most defined contribution pensions were accessed through
+annuities.
 
-Also of specific interest to the annuity market was the European Union's
-`Equal Treatment in Goods and Services Directive' of 2004. This prohibited discrimination
-in the provision and cost of goods and services based on sex. Up until 2011, there was a clause that
-stated insurers were allowed to charge different premiums if it was based on evidence that sex is
-correlated with different amounts of risk. However, in March 2011, the European Court of Justice
-ruled that insurers were not allowed to charge different amounts and gave them until December 2012
-to implement the ruling. This change meant that annuity products could no longer be priced differently
-for men and women in the UK. However, given that this change was implemented two years before the pension
-freedoms act, I can still identify the impact of the pensions freedom act on early retirement consumption.
+Also of specific interest to the annuity market was the European Union's `Equal
+Treatment in Goods and Services Directive' of 2004. This prohibited
+discrimination in the provision and cost of goods and services based on sex. Up
+until 2011, there was a clause that stated insurers were allowed to charge
+different premiums if it was based on evidence that sex is correlated with
+different amounts of risk. However, in March 2011, the European Court of Justice
+ruled that insurers were not allowed to charge different amounts and gave them
+until December 2012 to implement the ruling. This change meant that annuity
+products could no longer be priced differently for men and women in the UK.
+However, given that this change was implemented two years before the pension
+freedoms act it will not impact my results.
+
+\footnotetext{Partial UFPLS (uncrystallised funds pension lump sum) is similar
+    to Partial drawdown but with a different tax schedule. With partial drawdown
+    you take the whole 25\% tax free amount at once whereas with UFPLS you only
+    claim the tax relief on 25\% of the amount you are taking out. This option
+    is better if you may want to buy another retirement product in the future
+    such as an annuity. }
 
@@
 
-Table \ref{tab:sum_stats} shows various summary statistics for each variable of interest from ELSA and the ONS.
+Table \ref{tab:sum_stats} shows various summary statistics for each variable of
+interest from ELSA and the ONS.
 
-As required, retirement year is between 2011 and 2013 for the control group and 2015 and 2017 for the treatment group.
-However, as discussed above their interview date is after or the same as the retirement year because we are tracking
-consumption in retirement.
+As required, retirement year is between 2011 and 2013 for the control group and
+2015 and 2017 for the treatment group. However, as discussed above their ELSA
+interview date is after or the same as the retirement year because we are
+tracking consumption in retirement.
 
-The second retirement group, those who retired post reform, is small with just 728 non missing observations for gender as opposed
-to 941 individuals in the control group. The control group are more female, retired at a slightly younger age and expected to
-retire slightly earlier. This could be because this period was affected by the increase in the state pension age for
-women from age 60 in 2010 to age 65 in 2018. Since this change happened gradually and occured over the whole period
-I do not expect it to influence the results. The DifferenceAge row tracks the difference between expected and actual
-pension age.
+The second retirement group, those who retired post-reform, is small with just
+728 non missing observations for gender as opposed to 941 individuals in the
+control group. The control group are more female, retired at a slightly younger
+age and expected to retire slightly earlier. This could be because this period
+was affected by the increase in the state pension age for women from age 60 in
+2010 to age 65 in 2018. Since this change happened gradually and ocurred over
+the whole period I do not expect it to influence the results. The DifferenceAge
+row tracks the difference between expected and actual retirement age.
 
-The treatment group has higher financial wealth with a median of £66,000 at time of interview as opposed to £55,000
-in the control group. Likewise, the second group are more likely to have held a DC pension at some point
-and have much more money in them. Both groups are equally likely to have a defined benefit pension with roughly 44\% of
-individuals across both samples having a defined benefit pension. In general, DB pensions are more prevalent than
-DC pensions in the data, this reflects the same trend that is observed in population of UK retiree. Home ownership
-is roughly equal across groups although the second group have slightly higher housing wealth.
-\textbf{I have
-    not inflation adjusted but I am thinking about it?}
+The treatment group has higher financial wealth with a median of £66,000 at the
+time of interview as opposed to £55,000 in the control group. Likewise, the
+second group are more likely to have held a DC pension at some point and have
+much more money in them. Both groups are equally likely to have a DB pension
+with roughly 44\% of individuals across both samples having a DB pension. In
+general, DB pensions are more prevalent than DC pensions in the data, this
+reflects the same trend that is observed in the general UK population. Home
+ownership is roughly equal across groups although the second group have slightly
+higher housing wealth.
 
-Both groups are similarly long-lived according when using ONS life tables to calculate life expectancy
-based on gender, age and the year the interview was carried out in. Subjective life expectancies are also similar
-across groups, with individuals expecting to live another 21 years as opposed to the 24 that the ONS would expect
-them to live.
+Both groups are similarly long-lived when using ONS life tables to calculate
+life expectancy based on gender, age and the year the interview was carried out
+in. Subjective life expectancies are also similar across groups, with
+individuals expecting to live another 21 years as opposed to the 24 that the ONS
+would expect them to live.
+
+Unfortunately, some consumption data is missing for some individuals. The food
+categories have the least missing data and the leisure consumption category has
+the most missing data.
+
 \textbf{Maybe add in prob of survival to age 85 or 90 or something}
@@
 
-Unfortunately, some consumption data is missing for some individuals. The food categories have the least missing data
-and the leisure consumption category has the most missing data.
-
-\textbf{Check why total consumption adds up to so much
-    more than all the categories added together? Do they include rent and housing in consumption?}
-
-Overall, the groups appear to have slightly different wealth profiles with the treatment group being richer and slightly more likely
-to have DC pension wealth. On key demographic characteristics, however, the groups are similar, the difference in retirement
-age between the two groups can be mostly explained by the difference in expected retirement age meaning that individuals
-have not en masse decided to delay retirement and avoid annuitisation.
-
+Overall, the groups appear to have slightly different wealth profiles with the
+treatment group being richer and slightly more likely to have DC pension wealth.
+On key demographic characteristics, however, the groups are similar. The
+difference in retirement age between the two groups can be mostly explained by
+the difference in expected retirement age meaning that individuals have not en
+masse decided to delay retirement and avoid annuitisation. Moreover, for the
+regression discontinuity models I only use individuals who have a DC pension so
+it does not matter that there are more individuals with DC pensions in the
+second group.
 \begin{landscape}
-
+    \tiny
     \input{tables/r_summary_stats.tex}
-
+    \normalsize
 \end{landscape}
 
-As a further test for balance I regress demographic and financial characteristics of individuals on year of birth and the treatment dummy.
-We can then see if the treatment groups differ on key characteristics such as financial wealth or retirement age.
-For a regression discontinuity to be valid we need the treatment and control groups to be similar along
-all other characteristics apart from the treatment variable.
+As a further test for balance I regress demographic and financial
+characteristics of individuals on the treatment dummy. We can then see if the
+treatment groups differ on key characteristics such as financial wealth or
+retirement age. For a regression discontinuity to be valid we need the treatment
+and control groups to be similar along all other characteristics apart from the
+treatment variable.
 
@@
 \begin{equation*}
-    Y_{i} = \alpha + \beta Post_{i} + \gamma YOB_{i} + \kappa (Post_{i} \times YOB_{i}) + \epsilon_{i}
+    Y_{i} = \alpha + \beta PostRef_{i}  + \epsilon_{i}
 \end{equation*}
 
+Where $Y_{i}$ are the different demographic and financial characteristics.
+
 \input{tables/cov_balance.tex}
-Table \ref{tab:cov_balance} shows these results. We can see that being in the treatment group.
-\textbf{I am still thinking about this.}
 
-One threat to validity is manipulation of the running variable -- which is the variable that decides
-whether an individual is in the treatment group or not -- in our case retirement year.
-This is probably quite a threat in this setting as individuals could delay retirement and/or
-the purchase of an annuity until after the reform. To check whether
-this is happening we see whether the difference between real retirement age and the first expected retirement age
-recorded in the data are different for those retiring before 2014 compared to those retiring after 2014. As shown
-above in Table \ref{tab:sum_stats} this not the case.
+Table \ref{tab:cov_balance} shows these results. As expected, retirement year
+and interview year are greater for the treated group. The difference between
+expected and real retirement age is 0 which shows that individuals are not
+delaying retirement in order to be part of the treatment group although
+retirement age is higher post reform. DC pension value is higher, as is
+financial wealth but these are both imprecisely estimated. Because of the
+differences in these variables I add them to the regression model, following
 
-However, it is possible that individuals do not buy an annuity at the time of retirement since the law prior to 2014 only
-required that they access the pot by age 75. They may decide to keep their DC pension untouched
-and live off other income and assets before accessing it later. Since we do not have data on the
-exact purchase date of annuities we cannot track whether
-annuity purchases were delayed because of this. However, there is relatively stable annuity demand
-before the policy reform. If individuals delayed annuity purchases in expectation of the reform we
-would observe declines in purchases before the reform was announced.
+One threat to validity is manipulation of the running variable, retirement year.
+This is probably quite a threat in this setting as individuals could delay
+retirement and/or the purchase of an annuity until after the reform. As observed
+above in Tables \ref{tab:sum_stats} and \ref{tab:cov_balance}, it does not
+appear like individuals are delaying retirement. However, it is possible that
+individuals do not buy an annuity at the time of retirement since the law prior
+to 2014 only required that they access the pot by age 75. They may decide to
+keep their DC pension untouched and live off other income and assets before
+accessing it later. Since we do not have data on the exact purchase date of
+annuities we cannot track whether annuity purchases were delayed because of
+this. However, there is relatively stable annuity demand before the policy
+reform. If individuals delayed annuity purchases in expectation of the reform,
+so they would not have to buy an annuity at all, we would observe declines in
+purchases before the reform was announced. Likewise, if individuals bought an
+annuity early in expectation that the end of the compulsory market would cause
+prices to go up, we would observe a big spike in annuity purchases prior to
+2014.
 
@@
 
-In this section I outline the key empirical models I run with both the simulated consumption data from the
-lifecycle models and with the real data from ELSA. I then see which lifecycle model better fits the consumption
-response that occurred as a result of the pension reform.
+In this section, I outline the key empirical models I run with both the simulated
+consumption data from the lifecycle models and with the real data from ELSA. I
+then see which lifecycle model better fits the consumption response that
+occurred as a result of the pension reform.
 
-I use a regression discontinuity design for which I compare the consumption of early retirees after the policy reform
-to that of people who retired just before the policy reform. The key assumption implicit in regression discontinuities is
-that nothing else changes at the time of the jump apart from the policy of interest, and that the policy occurs
-without individuals predicting it and altering their behaviour. As I have argued above, the demographic information in the
-data suggest that individuals did not delay retirement and that there was no delay in annuitisation as shown by the quick
-and sudden decline in annuity purchases. Moreover, anecdotal evidence from media and business sources at the time of the
-announcement show that there was surprise the government had gone this far, with money marketing describing the change as a
-"bombshell" \cite{money_marketing_announcement}.
+I use a fuzzy regression discontinuity design for which I compare the
+consumption of individuals who retired after the policy reform to that of people
+who retired just before the reform. The key assumption implicit in regression
+discontinuities is that nothing else changes at the time of the jump apart from
+the policy of interest, and that the policy occurs without individuals
+predicting it and altering their behaviour. As I have argued above, the
+demographic information in the data suggests that individuals did not delay
+retirement and that there was no delay in annuitisation as shown by the quick
+and sudden decline in annuity purchases. Moreover, anecdotal evidence from media
+and business sources at the time of the announcement show that there was
+surprise the government put forward the reform, with Money Marketing describing the
+change as a ``bombshell" \cite{money_marketing_announcement}.
 
-Retirement year is the running variable and individuals are treated if retirement year is greater than 2014 and less than
-or equal to 2017. I use consumption data of individuals up to 2 years into retirement so that the sample size is larger.
-So if someone retired in 2015 and had consumption data in 2015 and 2017, I include both values. An individual is considered
-not treated if they retire before 2013 and after 2011.  To account for the fact
-that the reform only impacted individuals who had accumulated defined contribution pension pots I interact the treatment dummy with
-an indicator variable signalling whether the individual had ever held a DC pension pot.
+Retirement year is the running variable and individuals are treated if
+retirement year is greater than 2014 and less than or equal to 2017. I use
+consumption data of individuals up to 2 years into retirement so that the sample
+size is larger. So if someone retired in 2015 and had consumption data in 2015
+and 2017, I include both values. An individual is considered not treated if they
+retire before 2013 and after 2011.
 
-Because of the differences in financial characteristics between the groups, I add controls for financial wealth, housing wealth
-and whether an individual owns their own home.
+Because of the differences in financial characteristics between the groups, I
+add controls for financial wealth, housing wealth, whether an individual owns
+their own home, their gender and the size of their state pension. The following
+regression equation is used.
 
@@
 
-And $X_{it}$ is a set of controls.
+Where $Cons_{it}$ is one of five different consumption variables and $X_{it}$ is
+a set of controls.
+
+Table \ref{tab:DcOnlyRes} shows the results of this specification on the various
+consumption indicators. Column 1 has total monthly consumption on the left hand
+side of the regression equation. We can see that being in the treatment group is
+associated with £74 more a week in spending overall across all categories of
+expenditure. This is not statistically significant using robust standard errors.
+Being further into retirement at the time of interview is associated with higher
+total consumption whilst those who retired later have lower consumption because
+years since the start of retirement is associated with higher total consumption.
+
+Columns 2 through 4 use food consumption as the outcome variable. Interestingly,
+food consumption inside the house does not increase as a result of the policy
+reform but food consumed away from the house does increase. A shock to income,
+caused by no longer needing to annuitise, may cause spending on luxury goods to
+increase and for spending on their substitutes, like food in the house relative
+to food away from home, to decrease. This income effect may be driving the
+results we observe in Table \ref{tab:DcOnlyRes} since food consumption outside the
+home increases as does expenditure on clothing -- seen in column 5.
+
+One concern is that the group who retired after 2014 are just generally
+different to those who retired before and perhaps consume more outside the home
+anyway. The financial characteristic data presented in \ref{tab:sum_stats} would
+support this hypothesis as it shows all retirees are richer post reform. To test
+for these cohort effects, I run a difference in difference regression to
+complement the regression discontinuity just shown. I add individuals who have a
+DB pension and those with no private pension into the model and interact the
+policy change with dummies for having different types of pension. Specifically, I
+run:
+
+\begin{equation*}
+    Cons_{it} =  \gamma X_{it} + \beta_{1} PostReform_{it} + \beta_{2} PostReform*DC_{it} + \beta_{3} PostReform*DB_{it}  + \epsilon
+\end{equation*}
+
+This makes the coefficient of interest $\beta_{2}$, which shows us the
+difference in the change in consumption between the DC-only group and the base
+group that has neither a DB or DC pension. Since rules for DB pensions did not
+change under the pensions freedom act, those with a DB pension or no pension can
+be seen as a natural counterfactual group to judge the consumption of the DC
+group against. Likewise, group with no private pension are unaffected by the
+reform.
+
+The key identifying assumption for difference in difference regressions is the
+parallel trends assumption. In this context, it means that consumption changes
+overtime between the DB, DC and no pension group would have been the same were
+it not for the policy change. Given that there were no other reforms to pension
+schemes at this time that would have only impacted one of these groups in
+particular I claim that it holds.
+
+Table \ref{tab:ElsaAllData} shows these results. The post reform variable now
+tells us that total and food consumption increased but clothing consumption
+decreased. The interactions at the bottom of the table show us what happened to
+individuals with a given pension type relative to the reference cohort which has
+no pension. Those with only a DC pension had an increase in total consumption
+which was higher than those with a DB pension and higher than those with no
+pension. Consumption on food decreased slightly for the DC group relative to no
+pension but the decrease was less than that for the DB group. Expenditure on
+clothing also increased for the DC group compared to no pension and DB pensions.
+As above, the only variables that are significant at the 5\% level are financial
+wealth and whether the individual owns their house. However, the fact that using
+a difference in difference setup also gives us similarly sized estimates is
+encouraging and validates the regression discontinuity approach.
+
+We can also test whether treatment intensity is correlated with larger increases
+in spending. Those who have more money in a DC pension, and those with more
+financial wealth overall, would be impacted by the policy to a greater extent
+than those with smaller pensions, since they would be forced to annuitise a
+larger amount. So, I interact the policy with financial wealth and DC wealth. As
+in Table \ref{tab:DcOnlyRes}, this regression only uses those individuals who
+have a DC pension.
+
+To be precise, I run:
+\begin{equation*}
+    Cons_{it} =  \gamma X_{it} + \beta PostReform_{it}*DCValue_{it} + \epsilon_{it}
+\end{equation*}
 
@@
 
-Table \ref{tab:DcOnlyRes} shows the results of this specification on the various consumption indicators.
-Column 1 has total monthly consumption on the left hand side of the regression equation.
-We can see that being in the treatment group is associated with £73 more a week in spending
-overall across all categories of expenditure. This is not statistically significant using
-robust standard errors \textbf{find a reference for the SE that I use}. Being further into retirement
-at the time of interview is also associated with lower consumption whilst those who retired later have lower consumption
-because years since the start of retirement is associated with higher total consumption.
-
-Columns 2 through 4 use food consumption as the outcome variable. Interestingly, food consumption
-inside the house does not increase as a result of the policy reform but food consumed away from the house does
-increase. If individuals see their pension pot, that they previously could not access, as a
-shock to their wealth, like winning the lottery, then we would expect the consumption basket to
-include more goods that are perceived as luxuries. Inline with this, consumption on clothing, shown in
-column 4, increases as a result of the policy change.
-
-One concern is that the group who retired after 2014 are just generally different to those
-who retired before and perhaps consume more outside the home anyway. The financial characteristic
-data presented in \ref{tab:sum_stats} would support this hypothesis. To test for these cohort
-effects I run a difference in difference regression to complement the regression discontinuity just shown.
-I add individuals who have a defined benefit pension pot into the model and interact the policy change
-with dummies for having different types of pension. Specifically I run:
-
-\begin{equation*}
-    Cons_{it} =  \gamma X_{it} + \beta_{1} PostReform_{it} + \beta_{2} PostReform*DC_{it} + \beta_{3} PostReform*DB_{it}  + \epsilon
-\end{equation*}
-
-This makes the coefficient of interest $\beta_{2}$, which shows us the difference
-in the change in consumption between the DC only group and the base group that has neither a DB or DC pension.
-Since rules for DB pensions did not change under the pensions freedom act, those with a DB pension or no pension
-can be seen as a natural counterfactual group to judge the consumption of the DC group against.
-
 \begin{landscape}
@@
 \end{landscape}
-
-Table \ref{tab:ElsaAllData} shows these results. The post reform variable
-now tells us that total consumption decreased but food consumption increased a little and clothing consumption
-decreased. The interactions at the bottom of the table show us what happened to individuals with a given pension
-type relative to the reference cohort which has no pension. Those with only a DC pension had an increase in total consumption
-which was lower than that for those with a DB pension but higher than those with no pension. Only for consumption on
-clothes did DC consumption increase more than the other groups. As above, the only variables that are significant
-at the 5\% level are financial wealth and whether the individual owns their house.
-
-We can also test whether treatment intensity is correlated with larger increases in spending. Those who have
-more money in a DC pension, and those with more financial wealth overall, would be impacted
-by the policy to a greater extent than those with smaller pensions. So, I interact the policy with the treatment
-variable. As in Table \ref{tab:DcOnlyRes}, this regression only uses those individuals who have a DC pension.
-
-To be precise I run:
-\begin{equation*}
-    Cons_{it} =  \gamma X_{it} + \beta PostReform_{it}*DCValue_{it} + \epsilon
-\end{equation*}
 
@@
 
-Table \ref{tab:DcOnlyInteract} shows the results. Surprisingly, a larger DC pension
-pot is not associated with a greater increase in consumption. And although the reform is again associated with an
-increase in consumption, especially for food out of the house and clothing, this affect is decreasing with the
-size of the pension pot.
-
-The defined contribution pension pot size variable comes from work done by the IFS. This was discontinued
-after wave 5 so it is no longer possible to identify pension wealth in this way. For the above table pension
-wealth from the last available year for an individual was given a real return of 3\% and compounded until year of interview.
-Given that this is an imperfect measure, I also run the regression with financial wealth interacted with the treatment variable.
-
 \begin{landscape}
@@
 
-Table \ref{tab:DcOnlyFinWealthInteract} contains these results and shows a similar pattern to Table \ref{tab:DcOnlyInteract}.
-The reform is associated with an increase in consumption, but more financial wealth decreases the effect rather
-than making it stronger. This could be because those with higher levels of financial wealth have more assets in
-other savings apart from retirement savings \cite{cribb_karjalainen_ifs_2023} and therefore the DC pot is a smaller proportion of overall assets,
-and thus the policy change did not affect them as much.
+Table \ref{tab:DcOnlyInteract} shows the results. Surprisingly, a larger DC
+pension pot is not associated with a greater increase in consumption. And
+although the reform is again associated with an increase in consumption,
+especially for food out of the house and clothing, this effect is decreasing
+with the size of the pension pot.
 
-In conclusion, the results in the tables above show that in the basic regression discontinuity setup there is
-an impact of the policy reform on consumption, and this effect is stronger for consumption goods we
-would associate with a large increase in income. However, this is imprecisely estimated, and when we use a
-difference in difference identification method this effect disappears. Moreover, when we interact with a proxy for treatment strength
-(financial wealth or pension size) the effect is smaller the stronger the treatment.
+The DC pension pot size variable comes from work done by the IFS. The data after
+wave 5 has not been released so for individuals later than this I have no DC
+wealth variable. For Table \ref{tab:DcOnlyInteract}, pension wealth from the
+last available year for an individual was given a real return of 3\% and
+compounded until year of interview. Given that this is an imperfect measure, I
+also run the regression with financial wealth interacted with the treatment
+variable.
 
-Theoretically, a null change in consumption is evidence for a bequest motive explaining a lack of annuitisation
-rather than pessimistic life expectancies being the cause. In the next section I simulate a lifecycle model
-for individuals in the ELSA dataset and
-\textbf{some thing that gives a taster of the results to come in the next section}
+Table \ref{tab:DcOnlyFinWealthInteract} contains these results and shows a
+similar pattern to Table \ref{tab:DcOnlyInteract}. The reform is associated with
+an increase in consumption, but more financial wealth decreases the effect
+rather than making it stronger. This could be because those with higher levels
+of financial wealth have more assets in other savings apart from retirement
+savings \cite{cribb_karjalainen_ifs_2023} and therefore the DC pot is a smaller
+proportion of overall assets meaning the policy change did not affect them as
+much.
+
+In conclusion, the results in the tables above show that in the basic regression
+discontinuity setup there is an impact of the policy reform on consumption, and
+this effect is stronger for consumption goods we would associate with a large
+increase in income. However, this is imprecisely estimated, and when we use a
+difference in difference identification method this effect disappears. Moreover,
+when we interact with a proxy for treatment strength (financial wealth or
+pension size) the effect is smaller the stronger the treatment.
+
+Theoretically, a null change in consumption is evidence for a bequest motive
+explaining a lack of annuitisation rather than pessimistic life expectancies
+being the cause. In the next section I solve and simulate a lifecycle model for
+individuals in the ELSA dataset. I find that the bequest motive most closely
+matches the patterns observed in the empirical models above.
 
@@
 
-In the second part of the paper I first solve a modified retirement lifecycle model and then use data from the solved lifecycle
-model in similar regressions to the ones above.
+In the second part of the paper I first solve a modified retirement lifecycle
+model and then use data from the solved lifecycle model in similar regressions
+to the ones above.
 
-The problem that retirees face is as follows. Every period retirees solve:
+In recursive form, the problem that retirees face is as follows. Every period
+retirees solve:
 
 \begin{equation*}
-    V_{t}(a_{t}, y) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V_{t+1}(a_{t+1}, y) \}
+    V(a_{t}, y, t) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V(a_{t+1}, y, t) \}
 \end{equation*}
 
-subject to their budget constraint
+subject to their budget constraint:
 
@@
 
-where $a_{t}$ are asset holdings in time t, $y$ is constant income for all periods, $V_{t}$ is value in period $t$, $p_{t}$ is probability of
-death in period $t$ and $B()$ is a bequest function. Income can come either
-from state pensions, defined benefit pension plans or purchased annuities.
+where $a_{t}$ are asset holdings in time t, $y$ is constant income for all
+periods, $V_{t}$ is value in period $t$, $p_{t}$ is probability of death in
+period $t$ and $B()$ is a bequest function. Income comes from the state pension
+and purchased annuities. $\beta = 0.97$ is an individuals discount rate and
+$r = 1/0.97$ is the return on savings. I enter $t$ directly into the value
+function as a state variable to make it clear that value is also conditional on
+age -- through the death probability and the total number of periods left until
+terminal age.
 
+As is common in the literature \cite{lockwood_red_2012}, I use a constant relative
+risk aversion utility function.
 \begin{equation*}
@@
 
-In some specifications retirees can leave bequests, I use the bequest function from
-\cite{lockwood_red_2012}.
+Where $\sigma = 2.5$.
+
+In some specifications retirees can leave bequests. I use the bequest function
+from \cite{lockwood_red_2012}.
 
@@
 
-Where $a_{t+1}$ is the amount left at death, $m$ is a measure of bequest motive strength and $c_{0}$ is
-a minimum amount of consumption that individuals want. I pick values of $m = 0.95$ and $c_{0} = 20,000$ since
-these generate low rates of annuitisation, for comparison, Lockwood picks $m = 0.96$ and $c_{0} = 18,000$
-\textbf{Check all of this}
+Where $a_{t+1}$ is the amount left at death. $c_{0}$ is the amount of
+consumption below which individuals will not leave bequests. If $c_{0} >0$ then
+bequests are a luxury good and the wealthier the individual the more they will
+bequest. $m$ is the marginal propensity to bequeath after consuming at least
+$c_{0}$ and a higher value of $m$ means that individuals will leave a greater
+share of wealth to their heirs. I pick values of $m = 0.95$ and $c_{0} = 20,000$
+since these generate low rates of annuitisation, for comparison,
+\cite{lockwood_aer_2018} finds $c_{0}$ is in the range of $\$12,000$ to
+$\$30,000$ and $m$ is between $0.93$ and $0.96$ in a variety of models.
 
-First, I discretise the state space. I create a grid from 500 to 50,000 incrementing by 500 for income and 1000 to
-500,000 incrementing by 1000 for financial assets. I solve the retirees problem using backward induction. At age 110 there is certainty of death so any leftover assets
-are carried over into the next period and bequested. This means that the value at the end of the final period is
-either 0 (if we do not allow a bequest motive) or the value of bequests. I then take this value function and solve
-an individual's final period problem, choosing assets next period (i.e. those to bequest) and how much to consume.
+To solve this problem, I discretise the state space. I create a grid from £500
+to £50,000 incrementing by £500 for income and £1,000 to £500,000 incrementing by
+£1,000 for financial assets. I solve the retirees problem using backward
+induction. At age 110 there is certainty of death so any leftover assets are
+bequested. This means that the value at the end of the final period is either 0
+(if we do not allow a bequest motive) or the value of bequests. I then take this
+value function and solve an individual's final period problem, choosing assets
+next period (i.e. those to bequest) and how much to consume. This will be all
+assets for individuals with no bequest motive or assets below £20,000 ($c_{0}$).
 
-Using the optimal policy function in the last period, which is assets next period that maximise the utility
-function and the value function this period given all the potential income and asset states, I calculate the value of the last period.
-This is then used in the problem the year before that.
-I repeat this process back to the age of retirement to obtain optimal consumption amounts for each year of retirement
-and associated value functions.
+Using the optimal policy function in the last period, which is assets next
+period that maximise the utility function and the value function this period
+given all the potential income and asset states, I calculate the value of the
+last period. This is then used in the problem the year before that. I repeat
+this process back to the age of retirement to obtain optimal consumption amounts
+for each year of retirement and associated value functions.
 
-To simulate the ELSA data I solve this retirement problem for each new retiree in the data set dependent on their
-objective probability of death each period. I estimate with subjective life expectancies and objective life expectancies.
-I also estimate the model both with and without a bequest motive which was picked to roughly fit the real annuity
-rates seen in the data.
+To simulate the ELSA data I solve this retirement problem for each new retiree
+in the dataset dependent on their objective probability of death each period. I
+estimate with subjective life expectancies and objective life expectancies. I
+also estimate the model both with and without a bequest motive which was picked
+to roughly fit the real annuity rates seen in the data.
 
-\textbf{I don't actually know why I do this}
-In retiree's first year of retirement I allow them to choose to annuitise some of their wealth. In practical terms
-this is moving down the asset grid but up the income grid and seeing if the value of being in that position is better
-than where the individual is currently. To assess this trade-off I calculate the annual annuity payment that follows
-from a given annuity cost. I calculate this using objective life tables from the ONS using the following equation:
+In retiree's first year of retirement I somtimes annuitise some of their wealth.
+In practical terms, this is moving down the asset grid but up the permanent
+income grid and seeing if the value of being in that position is better than
+where the individual is currently. To assess this trade-off I calculate the
+annual annuity payment that follows from a given annuity cost. I calculate this
+using objective life tables from the ONS using the following equation:
 
@@
 
-Where $C$ is the one-off payment, $\delta$ is a factor that controls the `money's worth' of annuity and $p_{t|Retage}$
-is the probability of death at age $t$ conditional on being age $Retage$. So individuals can move C on the asset grid
-for gaining Ann on the income grid for the rest of their lives.
+Where $C$ is the one-off payment, $\delta$ is the `money's worth' of an annuity
+and $p_{t|Retage}$ is the probability of death at age $t$ conditional on being
+age $Retage$. So individuals can move C on the asset grid for gaining Ann on the
+income grid for the rest of their lives. The `money's worth' of an annuity is
+the ratio of expected present discounted value to the price of an annuity that
+an individual can purchase. Generally, for annuities, they have been found to be
+between $0.8$ and $0.9$ for the population as a whole in the UK,
+\cite{finkelstein_porteba_2002} \cite{finkelstein_porteba_2004}.
 
@@
 
-Having found optimal policy functions and value functions I can now simulate the change in consumption given
-different rates of annuitisation versus no annuitisation.
-There are multiple benefits to building a model of individuals decisions. Firstly, we can directly compare
-what a given individual would have consumed with what they actually consumed, conditional on our model being
-correct. In other words we have the counterfactual world in which individuals were not forced to annuitise.
-Secondly, we can simulate the impact of different policy changes on the consumption, savings and bequest behaviour.
-For example, I can change the proportion of wealth that will be annuitised and see the corresponding changes in
-consumption indicators for all individuals.
+Having found optimal policy functions and value functions I can now simulate the
+change in consumption given different rates of annuitisation versus no
+annuitisation. There are multiple benefits to building a model of individual's
+decisions. Firstly, we can directly compare what a given individual would have
+consumed with what they actually consumed, conditional on our model being
+correct. In other words, we have the counterfactual world in which individuals
+were not forced to annuitise. Secondly, I can simulate the impact of different
+policy changes on the consumption, savings and bequest behaviour. For example, I
+could change the proportion of wealth that will be annuitised and see the
+corresponding changes in consumption indicators for all individuals.
 
-I start this section by plotting the retirement path of consumption for a median individual in
-the ELSA data set. I plot consumption with different proportions of starting wealth annuitised
-and the different lifecycle model types.
+I now plot the retirement path of consumption for a median individual in the
+ELSA data set. I plot consumption with different proportions of starting wealth
+annuitised and the different lifecycle model types.
 
@@
 
-Figures \ref{fig:ConsumpPlot} and \ref{fig:AssetPlot} show the retirement
-path of consumption and assets for a median retiree. This hypothetical women has
-120,000 of assets at age 65 when they retired in 2012 with a state pension income of 6000.
-The dashed lines show consumption and assets with an annuity that costs half of their wealth.
-This buys them 3500 of annuity income, which, given their objective life expectancies is a money's
-worth of about 90\%.
+Figures \ref{fig:ConsumpPlot} and \ref{fig:AssetPlot} show the retirement path
+of consumption and assets for a median retiree. This hypothetical woman has
+£120,000 of assets at age 65 when they retired in 2012 with a state pension
+income of £6000. The dashed lines show consumption and assets with an annuity
+that costs half of their wealth. This buys them £3500 of annuity income, which,
+given their objective life expectancy is a money's worth of about 90\%.
 
-We can see that the asset path with the bequest motive keeps more for longer but still decreases
-and by their late 90s they have completely run down their wealth. Interestingly the asset path
-with the quickest run down is the standard lifecycle model using objective life expectancies,
-which points towards this individual being infact optimistic about their life chances.
-Also rather surprisingly the asset paths of all three types are similar without annuities.
-The paths only diverge after age 75 at which point the bequest path stays high and the
-other paths start to decrease faster. This is contrary to the path with annuities where
-the standard path diverges from the bequest and subjective paths immediately.
+We can see that the asset path with the bequest motive keeps more for longer but
+still decreases, and by their late 90s they have completely run down their
+wealth. Interestingly, the asset path with the quickest run down is the standard
+lifecycle model using objective life expectancies, which points towards this
+individual being optimistic about their life chances. Also rather surprisingly
+the asset paths of all three types are similar without annuities. The paths only
+diverge after age 75 at which point the bequest path stays high and the other
+paths start to decrease faster. This is contrary to the path with annuities
+where the standard path diverges from the bequest and subjective paths
+immediately.
 
-The consumption paths all start at around 1400. The three models without an annuity start at
-exactly the same amount and stay that way until the individual reaches her mid seventies -- showing
-that even quite different lifecycle models can generate similar paths. As expected, the consumption
-path with the bequest motive with an annuity is lower at the start of retirement than the bequest path 
-without an annuity as individuals seek to build their savings back up.
-But it is also lower for the subjective life expectancy case,
-again pointing toward this individual being optimistic. However, the subjective consumption
-line rises at a younger age than the line with bequests, showing that the models align 
-with the prediction that subjective lifecycle agents consume more in early retirement.
+The consumption paths all start at around £14,000 a year. The three models without an
+annuity start at exactly the same amount and stay that way until the individual
+reaches her mid seventies -- showing that even quite different lifecycle models
+can generate similar paths. As expected, the consumption path with the bequest
+motive with an annuity is lower at the start of retirement than the bequest path
+without an annuity as individuals seek to build their savings back up. But it is
+also lower for the subjective life expectancy case, again pointing toward this
+individual being optimistic. However, the subjective consumption line rises at a
+younger age than the line with bequests, showing that the models align with the
+prediction that subjective lifecycle agents consume more in early retirement
+compared to the other lifecycle models.
 
-The period of interest in this study is the first three years of retirement. For this period 
-we do see substantially differences in consumption for the diffe
+It is worth noting that the period of interest of this study is early retirement
+and although differences across the lifecycle are not large, there are
+significant differences at the start of retirement. This shows that different
+lifecycle models will produce substantive differences in predictions, so knowing
+which model is correct is important for policy.
 
-I first round an individuals financial wealth to the nearest point on the discrete asset grid, and
-do likewise for an individuals state pension income. I then take the treatment group and evaluate
-their consumption in the year of interview with an annuity given that they annuitised some portion
-of their wealth in their retirement year. I also evaluate consumption without annuitisation. I then
-use the same specification as above, but now the treatment group is the values of consumption without
-annuitisation.
+I first round an ELSA individual's financial wealth to the nearest point on the
+discrete asset grid, and do likewise for an individual's state pension income. I
+then take the treatment group and evaluate their consumption in the year of
+interview with an annuity, given that they annuitised 50\% of their financial
+wealth in their retirement year. I also evaluate consumption without
+annuitisation. The treatment group did not need to annuitise their wealth so I
+use their predicted consumption from the lifecycle models with their starting
+values of assets and pension income taken from their values in the ELSA dataset.
+To be clear, the only individuals for whom I annuitise wealth are those with DC
+pension pots who retired before 2014.
 
-Because I discretised the state space, annuity prices are rounded down to the nearest income grid point.
-This means that for some individuals with low assets half of their financial wealth does not buy them
-at least one grid point of annuity income. I do not let these individuals annuitise, setting consumption
-with and without an annuity to the same amount. I also present the average implied loading factor
-$\delta$, although I set this to $0.9$ when calculating annuity prices, because of rounding down
-this is lower.
+Because I discretised the state space, annuity prices are rounded down to the
+nearest income grid point. I set the money's worth of annuities to $0.9$ when
+calculating annuity prices but because I round down to the nearest income grid
+point the real money's worth an individual receives is lower. I assume that
+those who have a DC pension have half of their financial wealth in it since this
+matches roughly what is seen in the summary statistics, so, for the untreated
+group -- those who retired pre-reform -- I annuitise half of wealth. However,
+since some individuals cannot buy a £500 annuity, which relates to one grid
+point, with half of their financial wealth I only annuitise when an individual
+has financial assets large enough to cross this threshold. I do not let these
+individuals, with low assets, annuitise, setting consumption with and without an
+annuity to the same amount.
+
+\input{tables/simulation_summary_stats.tex}
+
+Table \ref{tab:simulation_prediction} shows the mean of consumption that each
+lifecycle model predicts for the treated and control groups as well as the
+original mean of consumption from the ELSA data for the individuals with
+non-missing values. The sample size of those with a DC pension is greatly
+reduced. This is mainly because not everyone in the ELSA data answers the life
+expectancy questions and of those who do answer some give answers that we cannot
+use for reasons listed in the data section.
+
+We can see the means for the control group are higher than the treated mean for
+all models but there is significant difference between the original level of
+consumption in the control group. The bequest and standard model both predict
+lower consumption than the subjective model which is not surprising given the
+subjective lifecycle individuals generally think their lives will be shorter and
+therefore can consume more early in retirement. Likewise, it makes sense the
+bequest model is lower than both other models since there is extra utility to be
+gained from saving so that one can bequest to their heirs.
+
+I now run the empirical models from above on this simulated data. This lets us
+see which lifecycle model gives us the closest match to the coefficients seen in the
+real data. We can therefore predict how other policy changes will impact
+behaviour as well as predict the full path for assets and consumption later into
+retirement. Matching a full model with method of simulated moments, though
+possible using the entire ELSA data like \textbf{cite paper eric sent me}, does
+not allow us to use incorporate policy changes into model.
 
@@
 
-Table \ref{tab:StandardLifeCycle} shows the results from the standard model lifecycle when individuals use ONS
-for risk of death and have no bequest function. In this case .
-
 \begin{landscape}
@@
 
-Table \ref{tab:BeqLifeCycle} shows the results from the bequest models.
+\begin{landscape}
+    \input{tables/subjective_life_cycle.tex}
+\end{landscape}
 
-Each column in the table shows a different model. Column 1 shows the model with no bequest motive where
-individuals use objective life probabilities from ONS, column 2 uses
+Tables \ref{tab:StandardLifeCycle}, \ref{tab:BeqLifeCycle} and
+\ref{tab:SubjectiveLifeCycle} show the results. Each column in the table shows a
+different empirical model that was applied to the data in the first part of this
+paper. The first column just takes those with DC pensions, the second looks at
+all individuals, whilst the third and fourth have interactions with financial
+wealth and DC pension wealth.
 
-\textbf{Also get some plots of consumption in early retirement
-    with and without annuities for the different model types}
+Comparing the DC-only column across tables we can see that simulated bequest
+model gives us the highest increase in consumption as a result of the reform.
+This could be because this type of individual was previously saving out of the
+annuity income to regain assets, so when they do not need to buy an annuity
+anymore their consumption actually increases. Surprisingly, the predictions from
+the lifecycle model using subjective life expectancies actually has a decreasing
+coefficient on the reform even though on average for a given individual
+consumption should increase as shown by the simple comparison in means. However,
+this difference could be the result of the post-reform group having higher
+subjective life expectancies, and therefore their consumption would be lower
+than the control group.
 
-% \subsection{Rough plan}
-% \begin{itemize}
-%   \item Intro
-%         \begin{enumerate}
-%           \item I think Eric French wrote something about population that I could use in then intro to say why it is important.
-%           \item "Latest data from HM Revenue Customs (published in April) showed more than £45bn has been taken from pots since 2015."
-%                 https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
+The lifecycle model that is closest to the empirical models shown in Table
+\ref{tab:DcOnlyRes} is the bequest model which gives an increase of 17 in
+monthly consumption. However, this coefficient is just under half of the change
+estimated in Table \ref{tab:DcOnlyRes} and reflects only a 0.6\% rise in
+consumption after the reform.
 
-%           \item Add bit about DC/DB pensions in the intro. Also talk about heterogeneity across countries.
-%                 Some countries want to move towards more annuitisation. This annual review is a good source of info
-%                 \cite{banks_crawford_ar_2022}
-%           \item https://www.moneysavingexpert.com/savings/pension-freedom/
-%         \end{enumerate}
-%   \item Lit review
-%   \item Models
-%   \item Empirical
-%         \begin{enumerate}
-%           \item Diff in diff
-%           \item RDD
-%           \item Matching?
-%           \item can I use anything I learnt in panel?
-%                 In some sense the decision to annuitise is a discrete choice problem so I could use something from there.
-%         \end{enumerate}
-%   \item Conclusion
-% \end{itemize}
+In the second column, the coefficient of interest is the DC pension interaction
+with the treatment variable since I use the difference in difference setup
+again. The bequest model has a coefficient of 25.6, the standard model one of
+17.7 and the standard model with subjective life expectancies has one of -30.9.
+These coefficients are higher than the coefficient seen in Table
+\ref{tab:ElsaAllData} with the predictions from the standard model matching the
+coefficient in the empirical model best. The subjective model again has the
+opposite sign to the one seen in the empirical model.
 
-\bibliographystyle{apa}
+The interaction models in columns 3 and 4 both have correct signs for the main
+reform effect in the standard and bequest models, although in both the
+interaction with financial wealth has a negative coefficient showing that the
+increase in consumption caused by the reform was progressively lower for those
+who would have had to annuitise more wealth. In the subjective model table the
+signs are again wrong apart from for the main treatment effect in the financial
+interaction column. The lifecycle model that appears closest to the empirical
+results is again the bequest model, which matches the signs from Tables
+\ref{tab:DcOnlyInteract} and \ref{tab:DcOnlyFinWealthInteract}, but the
+coefficients are smaller.
+
+Across all regressions, the lifecycle model using subjective life expectancies
+does not match the change in consumption that is seen when using ELSA data. It
+is surprising that the consumption response is also very different to the other
+lifecycle models used. One possible explanation for this could be that the
+control and treatment groups have quite different sets of objective life
+expectancies which therefore cause different consumption paths in early
+retirement. The structure that the Weibull distribution imposes on the
+transition probabilities could also be causing discontinuities in a given
+individuals subjective probability of death between two periods, which would
+impact how they consume across that part of retirement.
+
+The evidence just presented points towards the bequest model being best able to
+replicate the consumption response in the data. As highlighted above, the
+benefit of having a theoretical model is that we can trace out future values of
+consumption for individuals, only needing to know their starting point. So using
+the bequest solution to the retirement problem, I plot the whole path of average
+consumption and assets for the ELSA treated cohort. This plot is shown in the
+figure below.
+
+\textbf{Also get some plots of consumption in early retirement with and without
+    annuities for the different model types}
+
+\section{Conclusion}
+
+\subsection{Rough plan}
+\begin{itemize}
+    \item Intro
+          \begin{enumerate}
+              \item "Latest data from HM Revenue Customs (published in April)
+                    showed more than £45bn has been taken from pots since 2015."
+                    https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
+              \item what portion of savings are because of retirement savings?
+              \item Also talk about heterogeneity across countries. Some
+                    countries want to move towards more annuitisation. This
+                    annual review is a good source of info
+                    \cite{banks_crawford_ar_2022}
+              \item https://www.moneysavingexpert.com/savings/pension-freedom/
+          \end{enumerate}
+
+          \textbf{Follow up by changing chat because the parameter estimates for
+              the diff in diff have changed a bit -- DONE but check again when I
+              have gone through it -- haven't checked the comparisons }
+
+    \item \textbf{Abstract}
+    \item \textbf{Talk about covariate balance table? We could just set it for
+              the DC groups. }
+    \item \textbf{Add bit about testing just those people who retired when they
+              expected to retire.}
+    \item \textbf{Could add consumption and retirement year plot?}
+    \item \textbf{Maybe a small discussion of controls for the main regression
+              tables as well}
+    \item Conclusion
+\end{itemize}
+
+\bibliographystyle{chicago}
 \bibliography{references}

### 2023-07-26 R/10__1_regressions_prep_data.R

@@ -139,7 +139,9 @@ harm_long[ret_in & ret_equals, .(retired_age, age_at_interview, ret_in)]
 
 
 # We want recent retirees in the three years after they have retired.
-data_for_regressions <- harm_long[ret_in == TRUE & !is.na(pre_post_ref)]
+data_for_regressions <- harm_long[ret_in == TRUE & !is.na(pre_post_ref)] %>%
+    # One value of 450,000 for consumption
+    filter(total_monthly_consumption < 3500)
 
 
 data_for_regressions[, id_wave := paste0(idauniq, "-", wave)]

### 2023-07-26 R/13__regression_with_lifecycle.R

@@ -23,6 +23,15 @@ standard_consump_dt <-
     ) %>%
     select(id_wave, consump_stan_ann, consump_stan_no_ann)
 
+subjective_consump_dt <-
+    fread("../../data/ELSA/lifecycle_outputs/subjective_model_consumption_simulation.csv") %>%
+    rename(
+        consump_sub_ann = annuity_consump,
+        consump_sub_no_ann = no_annuity_consump
+    ) %>%
+    select(id_wave, consump_sub_ann, consump_sub_no_ann)
+
+
 # standard_consump_dt[, .(annuity_cost, annuity_payment)]
 # bequest_consump_dt[, .(annuity_cost, annuity_payment)]
 
@@ -59,6 +68,10 @@ joint_dt <- reg_dt %>%
         standard_consump_dt,
         by = "id_wave"
     ) %>%
+    merge.data.table(
+        subjective_consump_dt,
+        by = "id_wave"
+    ) %>%
     # for 'treated' consumption is without annuity
     # for 'control' consumption is with annuity
     mutate(bequest_consumption = fifelse(
@@ -70,6 +83,11 @@ joint_dt <- reg_dt %>%
         consump_stan_ann,
         consump_stan_no_ann
     )) %>%
+    mutate(subjective_consumption = fifelse(
+        pre_post_ref == "control" & ever_dc_pen == "yes",
+        consump_sub_ann,
+        consump_sub_no_ann
+    )) %>%
     mutate(rich_enough = fifelse(
         unlist(mapply(
             function(x, y) all.equal(x, y) == TRUE,
@@ -77,7 +95,12 @@ joint_dt <- reg_dt %>%
             consump_beq_no_ann
         )),
         "no", "yes"
-    ))
+    )) %>%
+    mutate(
+        bequest_consumption = bequest_consumption / 12,
+        standard_consumption = standard_consumption / 12,
+        subjective_consumption = subjective_consumption / 12
+    )
 
 
 joint_dt %>%
@@ -88,8 +111,71 @@ joint_dt %>%
     filter(public_pension > 5) %>%
     select(id_wave)
 
+# --------------- Save some summary statistics ---------------
+
+mean_consumption_out <- joint_dt %>%
+    filter(ever_dc_pen == "yes") %>%
+    group_by(pre_post_ref) %>%
+    summarise(across(
+        c("bequest_consumption", "standard_consumption", "subjective_consumption", "total_monthly_consumption"),
+        list(
[…]

### 2023-07-26 R/__ELSA_functions.R

@@ -61,13 +61,13 @@ clean_names <- list(
     "AgeAtInterview" = "age_at_interview",
     "ExpectedRetiredAge" = "first_exp_ret_age",
     "DifferenceAge" = "exp_real_ret_age",
-    "FinancialWealth (thousands)" = "fin_wealth",
+    "FinWealth(£000s)" = "fin_wealth",
     "DCPension" = "ever_dc_pen_bin",
-    "DCValue (thousands)" = "dc_pot",
+    "DCValue(£000s)" = "dc_pot",
     "DBPension" = "ever_db_pen_bin",
     "StatePension" = "public_pension",
     "OwnsHouse" = "owns_house",
-    "HouseValue (thousands)" = "house_value",
+    "HouseValue(£000s)" = "house_value",
     "ObjectiveLifeExp" = "obj_life_exps",
     "SubjectiveLifeExp" = "sub_life_exps",
     "TotalConsump" = "total_monthly_consumption",

### 2023-07-26 Texfiles/Draft1.tex
updates

@@
 \title{Retirement Consumption: Evidence from UK Pension Reform}
-\author{INSERT NUMBER HERE}
 \usepackage[letterpaper, margin=1.1in]{geometry}
-\linespread{1.25}
+\linespread{1.5}
 
 \begin{document}
+\begin{titlepage}
+    \begin{center}
 
+        \normalsize
+        {MPhil in Economic Research}
+        \vfill
+
+        \huge
+        Retirement Consumption: Evidence from UK Pension Reform
+        \vfill
+        \normalsize
+        Candidate number: 5348H
+
+        Date: 27th July, 2023
+
+        Word count: [Insert]
+        \vfill
+        I confirm that this is entirely my own work and has not previously been submitted for assessment, and I have read and understood the University’s and Faculty’s definition of Plagiarism
+    \end{center}
+\end{titlepage}
 \newpage
 \maketitle
+\begin{abstract}
+    I find the consumption response of early retirees to the `pensions freedom
+    act', which meant defined contribution pensions could be used for other
+    options than just purchasing an annuity. I use a discontinuity in year of
+    retirement and a difference in difference approach using those who were
+    unaffected by the policy reform as a natural control group. I then solve
+    three sets of lifecycle models from age of retirement until an agent reaches
+    110. I simulate their consumption with and without annuities, i.e. a the
+    policy reform, and match their consumption response to what was seen in the
+    empirical models. I find that a lifecycle model with a bequest motive best
+    fits this consumption response.
+\end{abstract}
+\newpage
 \tableofcontents
@@
 
+\section{Introduction}
 Classical economic theory suggests that retirement annuities, which provide a
@@
 
-\cite{yaari_65} showed that under standard assumptions we would expect
-individuals to annuitise all of their wealth at retirement to insure against the
-risk of long life. Since then there has been much literature discussing possible
-reasons that people do not annuitise. \cite{finkelstein_porteba_2002} and
-\cite{finkelstein_porteba_2004} find evidence of adverse selection, thereby
-making the 'money's worth' of annuities lower for the general population as
-opposed to the population of annuitants. However, they also find that theory
-would still predict annuitisation.
+\cite{yaari_65} showed that under standard assumptions with risk averse agents
+we would expect individuals to annuitise all of their wealth at retirement to
+insure against the risk of long life. And more recently
+\cite{davidoff_brown_diamond_aer_2005} show that complete annuitisation is
+optimal under a broader range of assumptions.
+Since then there has been much literature discussing possible reasons that
+people do not annuitise. \cite{finkelstein_porteba_2002,
+    finkelstein_porteba_2004} for the UK market and \cite{mitchell_et_al_1999} for
+the US market and  find evidence of adverse selection, thereby making the
+'money's worth' of annuities lower for the general population as opposed to the
+population of annuitants. However, they also find that theory would still
+predict annuitisation because the money's worth of annuities is not that far
+from actuarily fair.
 
@@
 
-\cite{lockwood_red_2012} builds on this and shows that a realistic bequest
-motive in lifecycle simulations achieves realistic annuitisation rates. He
-solves a simple lifecycle model with bequest motives taken from several recent
-papers in the literature. The bequest motives he picks therefore match other
-important aspects of the lifecycle model such as how much individuals actually
-bequest and how rich individuals are when they bequest.
+More recently, \cite{lockwood_red_2012} builds on this and shows that a
+realistic bequest motive in lifecycle simulations achieves realistic
+annuitisation rates. He solves a simple lifecycle model with bequest motives
+taken from several recent papers in the retirement savings literature. The
+bequest motives he picks therefore match other important aspects of the
+lifecycle model such as how much individuals actually bequest and how rich
+individuals are when they bequest. Also, \cite{odea_sturrock_rest_2023} find that subjective life expectancies can
+explain a large part of the annuitisation puzzle especially if agents are less
+risk averse.
 
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{figures/annuity_overtime.pdf}
+    \includegraphics[width=0.9\columnwidth]{figures/annuity_overtime.pdf}
     \label{fig:annovertime}
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{figures/annuity_pot_sizes.pdf}
+    \includegraphics[width=0.9\columnwidth]{figures/annuity_pot_sizes.pdf}
     \label{fig:ann2122}
@@
 
-Table \ref{tab:sum_stats} shows various summary statistics for each variable of
-interest from ELSA and the ONS.
+Table \ref{tab:sum_stats}\footnote{There are hyperlinks on the table number so,
+    if your pdf viewer supports this, you can click to go to the table or
+    figure} shows various summary statistics for each variable of interest from
+ELSA and the ONS.
 
@@
 \begin{landscape}
-    \tiny
+    \linespread{1.25}
     \input{tables/r_summary_stats.tex}
@@
 
+\linespread{1.25}
 \input{tables/cov_balance.tex}
+\linespread{1.5}
 
@@
 \begin{landscape}
+    \linespread{1}
+
     \input{tables/elsa_results_only_dc.tex}
@@
 \begin{landscape}
+    \linespread{1}
+
     \input{tables/elsa_results_all_with_interact.tex}
@@
 \begin{landscape}
+    \linespread{1}
+
     \input{tables/elsa_results_dc_pot_interact.tex}
@@
 \begin{landscape}
+    \linespread{1}
     \input{tables/elsa_results_fin_wealth_interact.tex}
@@
 of financial wealth have more assets in other savings apart from retirement
-savings \cite{cribb_karjalainen_ifs_2023} and therefore the DC pot is a smaller
+savings \citep{cribb_karjalainen_ifs_2023} and therefore the DC pot is a smaller
 proportion of overall assets meaning the policy change did not affect them as
 much.
+
+\subsection{Robustness Checks}
+
+I carry out two checks for robustness. Firstly, I exclude those individuals who
+retired year was greater than a year away from their expected retirement year in
+the first interview they gave with ELSA. This checks whether the effects shown
+above are being driven by individuals either retiring early to buy cheaper
+annuities or retiring later and avoiding annuitisation altogether.
+
+I also check whether the EU reform to annuity pricing is changing the results in
+some way. Since the reform only came into affect in December 2012 there are some
+retirees in our sample who may have been able to purchase an annuity before
+annuity pricing was made the same for men and women. I drop anyone with a
+retirement year in 2012 or 2011 and repeat the analysis from Table \ref{tab:DcOnlyRes}.
+
+\begin{landscape}
+    \linespread{1}
+    \input{tables/rob_elsa_results_only_dc_only_exp.tex}
+\end{landscape}
+
+\begin{landscape}
+    \linespread{1}
+    \input{tables/rob_elsa_results_only_dc_only_late_ret_year.tex}
+\end{landscape}
+
+Tables \ref{tab:DcOnlyExpOnlyRes} and \ref{tab:DcOnlyNot2012} show these
+results. In Table \ref{tab:DcOnlyExpOnlyRes} we can see the sign is reversed on
+PostReform in all consumption categories apart from clothing, however the sample size is a third of what it
+was prior to excluding individuals so there is even more uncertainty in these
+estimates. Table \ref{tab:DcOnlyNot2012} the sample size is again much smaller
+and the sign has flipped on all consumption categories apart from clothing. The
+implication of this regression is that those who retired in 2013 had a higher
+level of consumption in early retirement than those who retired post 2014.
+Although these results give
 
@@
 this process back to the age of retirement to obtain optimal consumption amounts
-for each year of retirement and associated value functions.
+for each year of retirement and associated value functions. I write code in
+Julia that solves this maximisation problem.
 
@@
 also estimate the model both with and without a bequest motive which was picked
-to roughly fit the real annuity rates seen in the data.
+to roughly fit the rates of annuitisation that are observed in the real world,
+which are around 6\%.
 
@@
 and $p_{t|Retage}$ is the probability of death at age $t$ conditional on being
-age $Retage$. So individuals can move C on the asset grid for gaining Ann on the
+age $Retage$. So individuals can move down $C$ on the asset grid for gaining $Ann$ on the
 income grid for the rest of their lives. The `money's worth' of an annuity is
 the ratio of expected present discounted value to the price of an annuity that
-an individual can purchase. Generally, for annuities, they have been found to be
-between $0.8$ and $0.9$ for the population as a whole in the UK,
-\cite{finkelstein_porteba_2002} \cite{finkelstein_porteba_2004}.
+an individual can purchase. Generally they have been found to be
+between $0.8$ and $0.9$ for the population as a whole in the UK
+\citep{finkelstein_porteba_2002, finkelstein_porteba_2004, mitchell_et_al_1999}.
 
@@
 \begin{landscape}
+    \linespread{1.25}
+
     \input{tables/standard_life_cycle.tex}
@@
 \begin{landscape}
+    \linespread{1.25}
+
     \input{tables/beq_life_cycle.tex}
@@
 \begin{landscape}
+    \linespread{1.25}
+
     \input{tables/subjective_life_cycle.tex}
@@
 
+I have shown that consumption in early retirement increased overall and
+particularly on
+
 \subsection{Rough plan}
@@
               have gone through it -- haven't checked the comparisons }
-
     \item \textbf{Abstract}
-    \item \textbf{Talk about covariate balance table? We could just set it for
-              the DC groups. }
     \item \textbf{Add bit about testing just those people who retired when they
@@
 
+\section{Appendix}
+\begin{landscape}
+    \input{tables/elsa_results_only_dc_only_exp.tex}
+\end{landscape}
+
 \bibliographystyle{chicago}

### 2023-07-26 R/10__3_run_regressions.R
Code and tex updates

@@ -85,6 +85,29 @@ unclean_names <- unname(unlist(clean_names))
 unclean_name_vector <- names(clean_names)
 names(unclean_name_vector) <- unclean_names
 
+# select stats to have at the bottom
+bottom_stats <- modelsummary::gof_map %>%
+    filter(raw %in% c(
+        "nobs", "r.squared", "adj.r.squared", "std.errors",
+        "vcov.type"
+    ))
+
+align_string_ms <- function(col_names, type_alignment = "d") {
+    # Left align first col
+    # right align others
+    multiplier <- 1
+
+    right <- rep(
+        type_alignment,
+        (length(col_names) * multiplier)
+    )
+    # first collapse the vector
+    # then join with l
+    full_string <- paste0("l", paste0(right, collapse = ""))
+
+    return(full_string)
+}
+
 
 # --------- dc only models ---------
 
@@ -111,14 +134,18 @@ names(out_models_only_dc) <- sapply(lhs, ReturnCleanName)
 # and then chat about it!
 
 
+
+
 modelsummary(
     out_models_only_dc,
     output = "../Texfiles/tables/elsa_results_only_dc.tex",
     vcov = "robust",
     title = "DC Only \\label{tab:DcOnlyRes}",
     coef_rename = unclean_name_vector,
-    estimate = "{estimate}\\quad   ({std.error})",
-    statistic = NULL
+    estimate = "{estimate}",
+    statistic = "({std.error})",
+    align = align_string_ms(names(out_models_only_dc)),
+    gof_map = bottom_stats
 )
 
 
@@ -148,8 +175,10 @@ modelsummary(
     vcov = "robust",
     title = "All individuals with interaction \\label{tab:ElsaAllData}",
     coef_rename = unclean_name_vector,
-    estimate = "{estimate}\\quad   ({std.error})",
-    statistic = NULL
+    estimate = "{estimate}",
+    statistic = "({std.error})",
+    align = align_string_ms(names(out_models_all_data)),
+    gof_map = bottom_stats
 )
 
 
@@ -178,8 +207,10 @@ modelsummary(
     vcov = "robust",
     title = "DC Pension Size interaction \\label{tab:DcOnlyInteract}",
     coef_rename = unclean_name_vector,
-    estimate = "{estimate}\\quad   ({std.error})",
-    statistic = NULL
+    estimate = "{estimate}",
+    statistic = "({std.error})",
[…]

### 2023-07-26 R/13__regression_with_lifecycle.R

@@ -58,6 +58,31 @@ unclean_name_vector <- names(clean_names)
 names(unclean_name_vector) <- unclean_names
 
 
+# select stats to have at the bottom
+bottom_stats <- modelsummary::gof_map %>%
+    filter(raw %in% c(
+        "nobs", "r.squared", "adj.r.squared", "std.errors",
+        "se_type"
+    ))
+
+align_string_ms <- function(col_names, type_alignment = "d") {
+    # Left align first col
+    # right align others
+    multiplier <- 1
+
+    right <- rep(
+        type_alignment,
+        (length(col_names) * multiplier)
+    )
+    # first collapse the vector
+    # then join with l
+    full_string <- paste0("l", paste0(right, collapse = ""))
+
+    return(full_string)
+}
+
+
+
 #  ------------ join data together ------------
 joint_dt <- reg_dt %>%
     merge.data.table(
@@ -239,7 +264,9 @@ modelsummary(
     vcov = "robust",
     title = "Simulated bequest lifecycle models \\label{tab:BeqLifeCycle}",
     coef_rename = unclean_name_vector,
-    statistic = NULL
+    statistic = NULL, 
+    align = align_string_ms(names(bequest_outmodels)), 
+    gof_map = bottom_stats
 )
 
 
@@ -261,8 +288,11 @@ modelsummary(
     vcov = "robust",
     title = "Simulated standard lifecycle models \\label{tab:StandardLifeCycle}",
     coef_rename = unclean_name_vector,
-    statistic = NULL
+    statistic = NULL, 
+    align = align_string_ms(names(standard_outmodels)), 
+    gof_map = bottom_stats
 )
+
 
 
 
@@ -283,7 +313,9 @@ modelsummary(
     vcov = "robust",
     title = "Simulated subjective lifecycle models \\label{tab:SubjectiveLifeCycle}",
     coef_rename = unclean_name_vector,
-    statistic = NULL
+    statistic = NULL, 
+    align = align_string_ms(names(subjective_outmodels)), 
+    gof_map = bottom_stats
 )
 
 

### 2023-07-26 R/14__create_lifecycle_plot.R

@@ -1,0 +1,50 @@
+library(data.table)
+library(dplyr)
+library(ggplot2)
+library(stringr)
+library(modelsummary)
+
+source("./__ELSA_functions.R")
+
+# want to read in all of those strings
+
+vec_dt <- fread(
+    "../../data/ELSA/lifecycle_outputs/model_ass_cons_paths/100006-6.csv",
+    header = F
+)
+
+min_length <- vec_dt[1, V2] %>%
+    strsplit(., ",", fixed = T) %>%
+    unlist() %>%
+    length()
+
+lapply(vec_dt, strsplit, ", ") %>%
+    unlist(recursive = F) %>%
+    lapply(function(x) x[seq(1, min_length)]) %>%
+    lapply(function(x) stringi::stri_replace(x, "", "[", regex = T, mode = "first"))
+
+
+
+# start cut the first bits off the bequest and standard lifecycle models
+# until they are same length as other ones
+
+lapply(vec_dt, strsplit, ",") %>%
+    unlist(recursive = F) %>%
+    lapply(length) %>%
+    unlist() %>%
+    min()
+# take last of this string
+
+
+
+
+
+
+strsplit()
+# three rows,
+# sub
+# stan
+# beq
+
+
+# should i start from age or retirement year?

### 2023-07-26 Texfiles/Draft1.tex
Latest edits

@@
 guaranteed income stream until the end of life, should be highly valued by
-individuals as a way to insure against late death \cite{yaari_65}. However, in
+individuals as a way to insure against late death \citep{yaari_65}. However, in
 developed countries, rates of annuitisation are far below the levels that theory
-predicts - this phenomenon has been termed the "annuity puzzle". The literature
+predicts -- this phenomenon has been termed the ``annuity puzzle". The literature
 has suggested several reasons for this, but there is no consensus about which
@@
 a lump sum withdrawal or income drawdown, involving a steady withdrawal of
-assets from the pension pot -- common advice is to take 4\% a year. Subsequently
-the number of annuities sold in the UK dropped precipitously.
+assets from the pension pot -- common advice is to take 4\% a year -- whilst the
+rest remains invested. Subsequently, the number of annuities sold in the UK
+dropped precipitously.
 
@@
 and without advanced notice before the Spring 2014 budget, I claim that
-individuals who retired in 2015, 2016 or 2017 are otherwise similar to those who
-retired in 2011, 2012 and 2013 except for the fact the later retirees do not
-need to annuitise their DC pension pots. Therefore, I can use a regression
-discontinuity to identify the effect that forced annuitisation has on the
-consumption of retirees.
+individuals who retired in 2015, 2016 and 2017 are otherwise similar to those
+who retired in 2011, 2012 and 2013 except for the fact that the later retirees
+do not need to annuitise their DC pension pots. Therefore, I can use a
+regression discontinuity to identify the effect that forced annuitisation has on
+the consumption of retirees.
 
@@
 pessimistic life expectancy. The annuity puzzle was first identified by
-\cite{yaari_65} and is a classed as a problem since standard economic theory
-cannot explain the lack of annuitisation that is observed in the real world. Two
+\cite{yaari_65} and is a classed as such since standard economic theory cannot
+explain the lack of annuitisation that is observed in the real world. Two
 possible explanations are bequests and life expectancy: individuals want to
 leave an estate for their heirs to inherit and they cannot do this if they
-annuitise their wealth; or individuals may believe they will not live as long as
-annuity providers think they will and therefore annuities that are on the market
-do not appear to be good value. Depending on the reason for the lack of
-annuitisation in the UK, the consumption response of retirees to the pension
+annuitise their wealth; or individuals may believe that they will not live as
+long as annuity providers think they will, and therefore annuities that are on
+the market do not appear to be good value. Depending on the reason for the lack
+of annuitisation in the UK, the consumption response of retirees to the pension
 reform will differ. If individuals do not annuitise because of pessimistic life
-expectancy their consumption should increase after the reform. If, on the other
-hand, individuals do not annuitise \ because of a bequest motive, consumption
-should not change much as a result of the reform.
-
-I will solve lifecycle models (\textbf{maybe add something about lifecycle
-    models here}) for both of these cases and simulate consumption decisions with
-and without forced annuitisation. I will then use a variety of empirical models
-to measure the consumption change in early retirement that resulted from the
-policy reform. The sign and magnitude of this change will be indicative of the
-mechanism causing the annuitisation problem.
+expectancy, their consumption should increase after the reform. If, on the other
+hand, individuals do not annuitise because of a bequest motive, consumption
+should not change as much as a result of the reform.
+
+I will solve lifecycle models for both of these cases and simulate consumption
+decisions with and without forced annuitisation. I will then use a variety of
+empirical models to measure the consumption change in early retirement that
+resulted from the policy reform. The sign and magnitude of this change will be
+indicative of the mechanism causing the annuitisation problem.
 
@@
 million in 2020 to 15.2 million in 2045 according to the latest projections from
-the Office for National Statistics (ONS) and for every 1000 people of working
+the Office for National Statistics (ONS), and for every 1,000 people of working
 age there will be 341 of pensionable age in 2045 compared to 280 in 2020
-\cite{ons_population_predictions_2020}. The increase in absolute and relative
+\citep{ons_population_predictions_2020}. The increase in absolute and relative
 numbers of elderly people makes retirement policy more consequential. Moreover,
 private DC pensions are becoming increasingly common and are predicted to grow
-as current cohorts age \cite{cribb_karjalainen_ifs_2023}. Therefore, policies
+as current cohorts age \citep{cribb_karjalainen_ifs_2023}. Therefore, policies
 regarding how private pensions can be accessed will have a larger impact on
@@
 other hand, if retirees over-save and do not spend, the economy may be
-dynamically inefficient whereby the capital stock is too high because
+dynamically inefficient -- when the capital stock is too high -- because
 individuals save too much for retirement and for bequests. In this case
@@
 \subsection{Literature review}
-My paper builds on three main strands of literature: that of the annuity puzzle,
-exploring reasons that some retirees choose not to annuitise; the retirement
-saving puzzle, which seeks to explain why retirees drawdown assets slowly, ; and
+My paper draws on three main strands of literature: that of the annuity puzzle,
+exploring reasons why retirees choose not to annuitise; the retirement
+saving puzzle, which seeks to explain why retirees drawdown assets slowly; and
 lifecycle models, whereby the above problems are explained using a tractable
@@
 \cite{davidoff_brown_diamond_aer_2005} show that complete annuitisation is
-optimal under a broader range of assumptions.
-Since then there has been much literature discussing possible reasons that
-people do not annuitise. \cite{finkelstein_porteba_2002,
-    finkelstein_porteba_2004} for the UK market and \cite{mitchell_et_al_1999} for
-the US market and  find evidence of adverse selection, thereby making the
-'money's worth' of annuities lower for the general population as opposed to the
-population of annuitants. However, they also find that theory would still
+optimal under a broader range of assumptions. Since then there has been much
+discussion of possible reasons for why people do not annuitise.
+\cite{finkelstein_porteba_2002, finkelstein_porteba_2004} find evidence of
+adverse selection into the UK annuity market and \cite{mitchell_et_al_1999} find
+evidence of the same in the US. This means longer lived individuals select into
+buying annuities which drives up annuity prices because providers need to pay
+individuals for longer than they originally thought they would need to. This
+makes the `money's worth' of an annuity lower for the general population as
+opposed to the population of annuitants, putting off individuals from buying an
+annuity. However, \cite{mitchell_et_al_1999} also find that theory would still
 predict annuitisation because the money's worth of annuities is not that far
-from actuarily fair.
+from `actuarially fair' (this is the where the present discounted value of annuities
+equals the price that an individual pays for it).
 
 \cite{friedman_warshawsky_qje_1990} show that annuitisation decisions can be
-fully explained by a mixture of bequest motives and actuarily unfair annuities.
-They solve an augmented life-cycle model with a range of parameters on how
-severe the rate of return is on the annuity versus market rates. For plausible
-values they find that individuals would optimally not annuitise much wealth.
-Similarly to \cite{finkelstein_porteba_2004},
-\cite{friedman_warshawsky_chicago_1988} show that there is a significant
-difference between the life expectancy of annuitants and the general population
-in the American annuity market but this cannot fully explain the annuitisation
-puzzle. Only when bequest motives are added to the model can annuitisation rates
-be rationalized.
+fully explained by a mixture of bequest motives and `actuarially unfair'
+annuities. They solve an augmented life-cycle model and calculate annuity demand
+for a range of differences between rate of return assets would get in the stock
+market and the rate of return annuities give an individual. For plausible values
+they find that individuals would optimally not annuitise much wealth. Similarly
+to \cite{finkelstein_porteba_2004}, \cite{friedman_warshawsky_chicago_1988} show
+that there is a significant difference between the life expectancy of annuitants
+and the general population in the American annuity market, which therefore
+implies a degree of adverse selection. But they find this cannot fully explain
+the annuitisation puzzle and only when bequest motives are added to the model
+can annuitisation rates be rationalized.
 
@@
 taken from several recent papers in the retirement savings literature. The
-bequest motives he picks therefore match other important aspects of the
-lifecycle model such as how much individuals actually bequest and how rich
-individuals are when they bequest. Also, \cite{odea_sturrock_rest_2023} find that subjective life expectancies can
-explain a large part of the annuitisation puzzle especially if agents are less
-risk averse.
-
-\cite{lockwood_aer_2018}
-
-\cite{vidalmelia_lejarragagarcia_munich_2004} have some interesting results.
-Need to talk about that.
-
-\textbf{There are some more papers I will include here.}
+bequest motives he picks matches other important aspects of empirical
+distribution of bequests such as how much individuals give and the financial
+characteristics of these individuals. Also, \cite{odea_sturrock_rest_2023} find
+that subjective life expectancies can explain a large part of the annuitisation
+puzzle especially if agents are less risk averse.
+
+Apart from bequests and annuity prices, a range of other reasons for the annuity
+puzzle have been explored. Pre-cautionary savings for unexpected end-of-life
+medical expenses might cause individuals to prefer holding assets rather than a
+an annuity. Moreover, \cite{vidalmelia_lejarragagarcia_munich_2004} suggests
+that couples might share the risk of living for a long-time between them and
+therefore couples may be less likely to annuitise their wealth.
+
+Given that uncertain end-of-life medical expenses are not such an issue in the
+UK as opposed to the US, and that the focus of recent literature has been on the
+bequest model and uncertain life expectancy, I choose also to focus on bequest
+and objective life expectancy in the UK context.
 
@@
 The main dataset I use is the English Longitudinal Study of Ageing (ELSA)
-\cite{main_elsa_citation}. ELSA picks individuals over the age of 50 to survey
+\citep{main_elsa_citation}. ELSA picks individuals over the age of 50 to survey
 every two years until death. If individuals leave the survey, ELSA replaces them
@@
 pension. There have been 9 waves of ELSA data collection -- the first was in
-2001 and collections happened every two years thereafter.
+2002/03, and collections happened every two years thereafter.
 
@@
 Institute for Fiscal Studies (IFS). This is only available up to and including
-wave 5 (which was collected in 2010 and 2011) at which point I use a real return
+wave 5 (which was collected in 2010 and 2011), after which point I use a real return
 of 3\% to predict forward the pension wealth variable until retirement. ELSA
@@
 
-I also use `life tables' from the UK's ONS. These provide me with the risk of
-death for each age group. Since the tables are produced until age 120, I adjust
-them to make death certain at age 110 as is common in the literature
-\cite{odea_sturrock_rest_2023}. I transform these so that I have risk of death
+I also use `life tables' from the UK's ONS from 2014. These provide me with the
+risk of death for each age group. The tables are produced until age 120 and I
+adjust them to make death certain at age 110 as is common in the literature
+\citep{odea_sturrock_rest_2023}. I transform these so that I have risk of death
 conditional on being a given age since this is what I use in the life cycle
@@
 
-To illustrate the effect that the reform had on sales of annuities in the UK I
-obtained product data from the Financial Conduct Authority. These track the
-sales of different financial products overtime including data on annuities.
+To illustrate the effect that the reform had on sales of annuities in the UK, I
+obtained product data from the Financial Conduct Authority (FCA). These track the
+sales of different financial products over time including data on annuities.
 
@@
 annuity, at a given price, will appear a much better deal to them. Likewise, an
-individual's life expectancy will change their decision on how to consume and
+individual's life expectancy will affect their decisions over how to consume and
 save through retirement.
@@
 that you will live to be age X or more?” where X changed depending on the age of
-the interviewee. If individuals were under 65 then X was 75, if individuals were
+the interviewee. If individuals were under 65 then X was 75; if individuals were
 66 and older they were asked the age that was 11 to 15 years older than them and
-a multiple of 5. From wave 3 respondents were also asked “What are the chances
+a multiple of 5. From wave 3, respondents were also asked “What are the chances
 that you will live to be age 85 or more?” if they were under 70. As most recent
@@
 offered. The 2004 Finance Act rationalised taxation rates on DB and DC pensions.
-The rules for DC pension pots were such that individuals had to buy an annuity
-after optionally taking a maximum of 25\% of the pot as a tax free lump sum
-withdrawal. DC pension pots were accessible from age 55 and most required that
+The rules for DC pension pots were such that an individual had to buy an annuity
+after optionally taking a maximum of 25\% of the pot as a tax-free lump sum
+withdrawal. DC pension pots were accessible from age 55, and most required that
 they be accessed before the age of 75.
 
-In the June 2010 budget, the coalition government made the first of two key
+In the June 2010 budget, the coalition government announced the first of two key
 reforms it would make to the pension system between 2010 and 2015. The 2010
 policy reform created a minimum income requirement above which individuals would
-not need to annuitise more \cite{finance_act_hmt_2011}, this meant high income
+not need to annuitise \citep{finance_act_hmt_2011}, meaning high income
 retirees would not need to annuitise their wealth. The income requirement was
@@
 
-In spring 2014, George Osbourne announced the so called `pensions freedom act'
-scrapping the minimum income requirement and eliminating the compulsory
-annuities market \cite{pen_freedoms_hmt_2014}. One government minister
-infamously brought the message home by saying pension pots `can be used to buy
-Lamborghinis' if retirees wanted to \cite{guardian_lambos}. The government also
-announced plans to make switches between DB and DC pension schemes possible
-although this was only realised in the next parliament -- future research could
-explore the impact of this second reform on the consumption and saving decisions
-of retirees.
+In spring 2014, Osbourne, the chancellor at the time, announced the so called
+`pensions freedom act', scrapping the minimum income requirement and eliminating
+the compulsory annuities market \citep{pen_freedoms_hmt_2014}. One government
+minister infamously brought the message home by saying pension pots `can be used
+to buy Lamborghinis' if retirees wanted to \citep{guardian_lambos}. The
+government also announced plans to make switches between DB and DC pension
+schemes possible although this was only realised in the next parliament.
+
+\textbf{For conclusion: future research could explore the impact of this second reform on the
+    consumption and saving decisions of retirees.}
 
@@
 Insurers, they show that annuity demand dropped by 75\% from its maximum in
-2011. Figure \ref{fig:annovertime} shows purchases of annuities overtime and
-demonstrates the sharp decrease in purchases that happened from 2014 to 2015.
-There was an increase in the number of pensions that were being accessed using
-an income drawdown product, but these only partly account for the drop in
-annuitisation as is shown by the graph.
+2011 to 2016. Figure \ref{fig:annovertime} shows
+purchases of annuities over time and demonstrates the sharp decrease in purchases
+that ocurred from 2014 to 2015. There was an increase in the number of pensions
+that were being accessed using an income drawdown product, but these only partly
+account for the drop in annuitisation as is shown by the graph.
 \begin{figure}[h]
-    \caption{New pension pots accessed }
+
     \centering
@@
     \label{fig:annovertime}
+    \caption{New pension pots accessed}
 \end{figure}
@@
 \begin{figure}[h]
-    \caption[Caption for LOF]{How pension pots are accessed\protect\footnotemark}
+
     \centering
     \includegraphics[width=0.9\columnwidth]{figures/annuity_pot_sizes.pdf}
+    \caption[Caption for LOF]{How pension pots are accessed\protect\footnotemark}
     \label{fig:ann2122}
+
     \
 \end{figure}
-Figure \ref{fig:ann2122} shows the distribution methods to access pension pots
+Figure \ref{fig:ann2122} shows the distribution of methods used to access pension pots
 at retirement in 2021-22. In 2021, 196,736 pots were fully withdrawn at
 retirement, accounting for over 50\% of pots. Prior to the policy reform this
-was not the case -- most defined contribution pensions were accessed through
+was not the case -- most DC pensions were accessed through
 annuities.
@@
 until 2011, there was a clause that stated insurers were allowed to charge
-different premiums if it was based on evidence that sex is correlated with
+different premiums if this was based on evidence that sex is correlated with
 different amounts of risk. However, in March 2011, the European Court of Justice
@@
     to Partial drawdown but with a different tax schedule. With partial drawdown
-    you take the whole 25\% tax free amount at once whereas with UFPLS you only
+    you take the whole 25\% tax-free amount at once whereas with UFPLS you only
     claim the tax relief on 25\% of the amount you are taking out. This option
@@
     if your pdf viewer supports this, you can click to go to the table or
-    figure} shows various summary statistics for each variable of interest from
+    figure.} shows various summary statistics for each variable of interest from
 ELSA and the ONS.
@@
 2015 and 2017 for the treatment group. However, as discussed above their ELSA
-interview date is after or the same as the retirement year because we are
+interview date are after or the same as the retirement year because we are
 tracking consumption in retirement.
@@
 
-The treatment group has higher financial wealth with a median of £66,000 at the
+The treatment group has higher financial wealth with a median of £67,000 at the
 time of interview as opposed to £55,000 in the control group. Likewise, the
 second group are more likely to have held a DC pension at some point and have
-much more money in them. Both groups are equally likely to have a DB pension
-with roughly 44\% of individuals across both samples having a DB pension. In
-general, DB pensions are more prevalent than DC pensions in the data, this
-reflects the same trend that is observed in the general UK population. Home
-ownership is roughly equal across groups although the second group have slightly
-higher housing wealth.
+more money in them. Both groups are equally likely to have a DB pension with
+roughly 45\% of individuals across both samples having a DB pension. In general,
+DB pensions are more prevalent than DC pensions in the data -- this reflects the
+trend that is observed in the general UK population. Home ownership is roughly
+equal across groups although the second group has slightly higher housing
+wealth.
 
@@
 the most missing data.
-
-\textbf{Maybe add in prob of survival to age 85 or 90 or something}
-\textbf{Also add in an indicator for living alone or living with someone else}
 
@@
 masse decided to delay retirement and avoid annuitisation. Moreover, for the
-regression discontinuity models I only use individuals who have a DC pension so
+regression discontinuity models I only use individuals who have a DC pension, so
 it does not matter that there are more individuals with DC pensions in the
 second group.
+
+\textbf{conclusion: chat about couples, chat about better data on at what
+    stage of retirement individuals buy annuities}
+
 \begin{landscape}
@@
 characteristics of individuals on the treatment dummy. We can then see if the
-treatment groups differ on key characteristics such as financial wealth or
-retirement age. For a regression discontinuity to be valid we need the treatment
-and control groups to be similar along all other characteristics apart from the
-treatment variable.
+treatment and control groups differ on key characteristics such as financial
+wealth or retirement age. For a regression discontinuity to be valid we need the
+groups to be similar along all other characteristics apart from the treatment
+variable.
 
@@
 
-Where $Y_{i}$ are the different demographic and financial characteristics.
-
-\linespread{1.25}
+Where $Y_{i}$ are the different demographic and financial characteristics for
+only those individuals with a DC pension.
+
+\linespread{1}
 \input{tables/cov_balance.tex}
@@
 expected and real retirement age is 0 which shows that individuals are not
-delaying retirement in order to be part of the treatment group although
-retirement age is higher post reform. DC pension value is higher, as is
-financial wealth but these are both imprecisely estimated. Because of the
-differences in these variables I add them to the regression model, following
+delaying retirement, beyond their original expectation,  in order to be part of
+the treatment group although retirement age is higher post reform. DC pension
+value is higher, as is financial wealth but these are both imprecisely
+estimated. Because of the differences in these variables I add them to the
+regression model.
 
@@
 above in Tables \ref{tab:sum_stats} and \ref{tab:cov_balance}, it does not
-appear like individuals are delaying retirement. However, it is possible that
-individuals do not buy an annuity at the time of retirement since the law prior
-to 2014 only required that they access the pot by age 75. They may decide to
-keep their DC pension untouched and live off other income and assets before
+appear as though individuals are delaying retirement. However, it is possible
+that individuals do not buy an annuity at the time of retirement since the law
+prior to 2014 only required that they access the pot by age 75. They may decide
+to keep their DC pension untouched and live off other income and assets before
 accessing it later. Since we do not have data on the exact purchase date of
-annuities we cannot track whether annuity purchases were delayed because of
+annuities, we cannot track whether annuity purchases were delayed because of
 this. However, there is relatively stable annuity demand before the policy
 reform. If individuals delayed annuity purchases in expectation of the reform,
-so they would not have to buy an annuity at all, we would observe declines in
-purchases before the reform was announced. Likewise, if individuals bought an
+so that they would not have to buy an annuity at all, we would observe declines
+in purchases before the reform was announced. Likewise, if individuals bought an
 annuity early in expectation that the end of the compulsory market would cause
-prices to go up, we would observe a big spike in annuity purchases prior to
-2014.
+prices to go up, we would observe a spike in annuity purchases prior to 2014.
 
@@
 
-In this section, I outline the key empirical models I run with both the simulated
-consumption data from the lifecycle models and with the real data from ELSA. I
-then see which lifecycle model better fits the consumption response that
-occurred as a result of the pension reform.
-
-I use a fuzzy regression discontinuity design for which I compare the
-consumption of individuals who retired after the policy reform to that of people
-who retired just before the reform. The key assumption implicit in regression
+In this section, I outline the key empirical models I run with first the real
+data from ELSA, and then the simulated data from the lifecycle models. I then
+see which lifecycle model better fits the consumption response that occurred as
+a result of the pension reform.
+
+I use a fuzzy, since some individuals will probably still annuitise a little,
+regression discontinuity design for which I compare the consumption of
+individuals who retired after the policy reform to that of people who retired
+just before the reform. The key assumption implicit in regression
 discontinuities is that nothing else changes at the time of the jump apart from
@@
 and business sources at the time of the announcement show that there was
-surprise the government put forward the reform, with Money Marketing describing the
-change as a ``bombshell" \cite{money_marketing_announcement}.
+surprise the government put forward the reform, with Money Marketing describing
+the change as a ``bombshell" \citep{money_marketing_announcement}.
 
@@
 consumption data of individuals up to 2 years into retirement so that the sample
-size is larger. So if someone retired in 2015 and had consumption data in 2015
+size is larger. So, if someone retired in 2015 and had consumption data in 2015
 and 2017, I include both values. An individual is considered not treated if they
-retire before 2013 and after 2011.
+retire in or before 2013.
 
@@
 side of the regression equation. We can see that being in the treatment group is
-associated with £74 more a week in spending overall across all categories of
+associated with £36 more a week in spending overall across all categories of
 expenditure. This is not statistically significant using robust standard errors.
 Being further into retirement at the time of interview is associated with higher
-total consumption whilst those who retired later have lower consumption because
-years since the start of retirement is associated with higher total consumption.
+total consumption whilst those who retired later have lower consumption. This is
+surprising since lifecycle models predict consumption decreases over retirement
+and this has also been documented empirically \citep{hurd_rohwedder_nber_2003}.
 
@@
 different to those who retired before and perhaps consume more outside the home
-anyway. The financial characteristic data presented in \ref{tab:sum_stats} would
-support this hypothesis as it shows all retirees are richer post reform. To test
-for these cohort effects, I run a difference in difference regression to
+anyway. The financial characteristic data presented in Table \ref{tab:sum_stats}
+would support this hypothesis as it shows all retirees are richer post-reform.
+To test for these cohort effects, I run a difference-in-difference (DD) regression to
 complement the regression discontinuity just shown. I add individuals who have a
 DB pension and those with no private pension into the model and interact the
-policy change with dummies for having different types of pension. Specifically, I
-run:
+policy change with dummies for having different types of pension. Specifically,
+I run:
 
@@
 be seen as a natural counterfactual group to judge the consumption of the DC
-group against. Likewise, group with no private pension are unaffected by the
-reform.
-
-The key identifying assumption for difference in difference regressions is the
-parallel trends assumption. In this context, it means that consumption changes
-overtime between the DB, DC and no pension group would have been the same were
-it not for the policy change. Given that there were no other reforms to pension
+group against.
+
+The key identifying assumption for DD is the parallel trends assumption. In this
+context, it means that consumption differences between the DB, DC and no pension
+group across cohorts would have been the same were it not for the policy change.
+For DD to not be valid we would need a reason to expect that the individuals
+with a DB and no pension who are retiring post 2014 are impacted by some factor
+that did not equally impact the DC group and that would impact the consumption
+of all groups differently. Given that there were no other reforms to pension
 schemes at this time that would have only impacted one of these groups in
-particular I claim that it holds.
-
-Table \ref{tab:ElsaAllData} shows these results. The post reform variable now
+particular, I claim that the parallel trends assumption holds holds.
+
+Table \ref{tab:ElsaAllData} shows these results. The post-reform variable now
 tells us that total and food consumption increased but clothing consumption
@@
 pension. Consumption on food decreased slightly for the DC group relative to no
-pension but the decrease was less than that for the DB group. Expenditure on
+pension, but the decrease was less than that for the DB group. Expenditure on
 clothing also increased for the DC group compared to no pension and DB pensions.
-As above, the only variables that are significant at the 5\% level are financial
-wealth and whether the individual owns their house. However, the fact that using
-a difference in difference setup also gives us similarly sized estimates is
-encouraging and validates the regression discontinuity approach.
+The increase in sample size also brings decreases in the standard errors of
+these variables -- the interactions are now all significant a the 5\% level as
+is the main coefficient on retiring post-reform.
+
+However, it is worth noting that there is a significant change between the DB
+and the no pension group. A natural test of the parallel trends assumption is to
+see whether two different, untreated, units have different outcomes
+post-treatment. If they are different post treatment then there could be
+something else driving the differences between the treatment group and the
+non-treated groups.
 
 We can also test whether treatment intensity is correlated with larger increases
-in spending. Those who have more money in a DC pension, and those with more
-financial wealth overall, would be impacted by the policy to a greater extent
-than those with smaller pensions, since they would be forced to annuitise a
-larger amount. So, I interact the policy with financial wealth and DC wealth. As
-in Table \ref{tab:DcOnlyRes}, this regression only uses those individuals who
-have a DC pension.
-
-To be precise, I run:
+in spending. Those who have more money in a DC pension would be impacted by the
+policy to a greater extent than those with smaller pensions, since they would be
+forced to annuitise a larger amount. So, I interact the policy with financial
+wealth and DC wealth. As in Table \ref{tab:DcOnlyRes}, this regression only uses
+those individuals who have a DC pension.
+
+Specifically, I run:
 \begin{equation*}
@@
 
-Table \ref{tab:DcOnlyInteract} shows the results. Surprisingly, a larger DC
-pension pot is not associated with a greater increase in consumption. And
-although the reform is again associated with an increase in consumption,
-especially for food out of the house and clothing, this effect is decreasing
-with the size of the pension pot.
+Table \ref{tab:DcOnlyInteract} shows the results. The interaction with DC
+pension size is positive, meaning that a larger DC pension pot associated with a
+greater increase in total monthly consumption. But, although the reform is again
+associated with an increase in consumption, especially for food out of the house
+and clothing, this effect is decreasing with the size of the pension pot for
+these consumption goods.
 
 The DC pension pot size variable comes from work done by the IFS. The data after
-wave 5 has not been released so for individuals later than this I have no DC
-wealth variable. For Table \ref{tab:DcOnlyInteract}, pension wealth from the
-last available year for an individual was given a real return of 3\% and
-compounded until year of interview. Given that this is an imperfect measure, I
-also run the regression with financial wealth interacted with the treatment
+wave 5 has not been released yet, so for individuals who retire after 2011 I
+have no DC wealth variable. For Table \ref{tab:DcOnlyInteract}, pension wealth
+from the last available year for an individual was given a real return of 3\%
+and compounded until year of interview. Given that this is an imperfect measure,
+I also run the regression with financial wealth interacted with the treatment
 variable.
@@
 similar pattern to Table \ref{tab:DcOnlyInteract}. The reform is associated with
-an increase in consumption, but more financial wealth decreases the effect
-rather than making it stronger. This could be because those with higher levels
-of financial wealth have more assets in other savings apart from retirement
-savings \citep{cribb_karjalainen_ifs_2023} and therefore the DC pot is a smaller
-proportion of overall assets meaning the policy change did not affect them as
-much.
+an increase in total monthly consumption and more financial wealth increases.
 
@@
 I carry out two checks for robustness. Firstly, I exclude those individuals who
-retired year was greater than a year away from their expected retirement year in
-the first interview they gave with ELSA. This checks whether the effects shown
-above are being driven by individuals either retiring early to buy cheaper
-annuities or retiring later and avoiding annuitisation altogether.
+retired more than one year away from their expected retirement year -- I define
+an individuals expected retirement year as their expected retirement age in
+their first interview with ELSA, added to the year that the interview was
+carried out in. This checks whether the effects shown above are being driven by
+individuals either retiring early in order to buy cheaper annuities or retiring
+later and avoiding annuitisation altogether.
 
@@
 annuity pricing was made the same for men and women. I drop anyone with a
-retirement year in 2012 or 2011 and repeat the analysis from Table \ref{tab:DcOnlyRes}.
+retirement year in 2012 or 2011 and repeat the analysis from Table
+\ref{tab:DcOnlyRes}.
 
@@
 Tables \ref{tab:DcOnlyExpOnlyRes} and \ref{tab:DcOnlyNot2012} show these
-results. In Table \ref{tab:DcOnlyExpOnlyRes} we can see the sign is reversed on
-PostReform in all consumption categories apart from clothing, however the sample size is a third of what it
-was prior to excluding individuals so there is even more uncertainty in these
-estimates. Table \ref{tab:DcOnlyNot2012} the sample size is again much smaller
-and the sign has flipped on all consumption categories apart from clothing. The
-implication of this regression is that those who retired in 2013 had a higher
-level of consumption in early retirement than those who retired post 2014.
-Although these results give
-
-In conclusion, the results in the tables above show that in the basic regression
-discontinuity setup there is an impact of the policy reform on consumption, and
-this effect is stronger for consumption goods we would associate with a large
-increase in income. However, this is imprecisely estimated, and when we use a
-difference in difference identification method this effect disappears. Moreover,
-when we interact with a proxy for treatment strength (financial wealth or
-pension size) the effect is smaller the stronger the treatment.
-
-Theoretically, a null change in consumption is evidence for a bequest motive
-explaining a lack of annuitisation rather than pessimistic life expectancies
-being the cause. In the next section I solve and simulate a lifecycle model for
-individuals in the ELSA dataset. I find that the bequest motive most closely
-matches the patterns observed in the empirical models above.
+results. In Table \ref{tab:DcOnlyExpOnlyRes}, we can see the sign is reversed on
+PostReform in all consumption categories apart from clothing, however the sample
+size is a third of what it was prior to excluding individuals so there is more
+uncertainty in these estimates. In Table \ref{tab:DcOnlyNot2012}, the sample
+size is again much smaller and the sign has flipped on all consumption
+categories apart from clothing and food consumed in the house. These regressions
+show that there are potentially identification issues with the regression
+discontinuity and DD strategies that I have presented thus far. However, since
+the sample sizes for both the robustness tests and main results are small it is
+difficult to asses what the impact of the reform was.
+
+In conclusion, the main results tables show that the reform could have had a
+positive impact on consumption in early retirement, and this increase is clearer
+for more `luxury' consumption goods, such as expenditure on clothing and eating
+out. And, when we interact with a proxy for treatment strength (financial wealth
+or pension size) the affect of the policy is larger the stronger the treatment.
+However, the robust checks point in the other direction, and there are
+significant differences between the DB and no pension groups in the DD
+regression, perhaps highlighting that other, unobserved, variables are driving
+changes in consumption. For example, consumer confidence between years could be
+changing, or, because of the small sample size in the ELSA dataset, there is too
+much noise in the consumption values each year.
+
+Theoretically, a null change or small change in consumption is evidence for a
+bequest motive explaining a lack of annuitisation rather than pessimistic life
+expectancies being the cause. In the next section I solve and simulate a
+lifecycle model for individuals in the ELSA dataset. I find that the bequest
+motive most closely matches the patterns observed in the empirical models above.
 
@@
 
-In the second part of the paper I first solve a modified retirement lifecycle
+In the second part of the paper, I first solve a modified retirement lifecycle
 model and then use data from the solved lifecycle model in similar regressions
@@
 \begin{equation*}
-    V(a_{t}, y, t) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V(a_{t+1}, y, t) \}
+    V(a_{t}, y, t) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V(a_{t+1}, y, t+1) \}
 \end{equation*}
@@
 period $t$ and $B()$ is a bequest function. Income comes from the state pension
-and purchased annuities. $\beta = 0.97$ is an individuals discount rate and
-$r = 1/0.97$ is the return on savings. I enter $t$ directly into the value
-function as a state variable to make it clear that value is also conditional on
-age -- through the death probability and the total number of periods left until
-terminal age.
-
-As is common in the literature \cite{lockwood_red_2012}, I use a constant relative
+and purchased annuities. I follow the literature and use $\beta = 0.97$ as an
+individual's discount rate and $r = 1/0.97$ as the interest rate. I enter $t$
+directly into the value function as a state variable to make it clear that value
+is also conditional on age -- through the chance of death and the total number
+of periods left until terminal age.
+
+As is common in the literature \citep{lockwood_red_2012}, I use a constant relative
 risk aversion utility function.
@@
 In some specifications retirees can leave bequests. I use the bequest function
-from \cite{lockwood_red_2012}.
+from \cite{lockwood_red_2012,lockwood_aer_2018}.
 
@@
 Where $a_{t+1}$ is the amount left at death. $c_{0}$ is the amount of
-consumption below which individuals will not leave bequests. If $c_{0} >0$ then
+consumption below which individuals will not leave bequests. If $c_{0} >0$, then
 bequests are a luxury good and the wealthier the individual the more they will
 bequest. $m$ is the marginal propensity to bequeath after consuming at least
-$c_{0}$ and a higher value of $m$ means that individuals will leave a greater
-share of wealth to their heirs. I pick values of $m = 0.95$ and $c_{0} = 20,000$
-since these generate low rates of annuitisation, for comparison,
+$c_{0}$, and a higher value of $m$ means that individuals will leave a greater
+share of wealth to their heirs. I pick values of $m = 0.95$ and $c_{0} = \pounds
+    20,000$ since these generate low rates of annuitisation. For comparison,
 \cite{lockwood_aer_2018} finds $c_{0}$ is in the range of $\$12,000$ to
@@
 To solve this problem, I discretise the state space. I create a grid from £500
-to £50,000 incrementing by £500 for income and £1,000 to £500,000 incrementing by
-£1,000 for financial assets. I solve the retirees problem using backward
+to £50,000 incrementing by £500 for income and £1,000 to £500,000 incrementing
+by £1,000 for financial assets. I solve the retirees' problem using backward
 induction. At age 110 there is certainty of death so any leftover assets are
-bequested. This means that the value at the end of the final period is either 0
-(if we do not allow a bequest motive) or the value of bequests. I then take this
-value function and solve an individual's final period problem, choosing assets
-next period (i.e. those to bequest) and how much to consume. This will be all
-assets for individuals with no bequest motive or assets below £20,000 ($c_{0}$).
-
-Using the optimal policy function in the last period, which is assets next
-period that maximise the utility function and the value function this period
-given all the potential income and asset states, I calculate the value of the
-last period. This is then used in the problem the year before that. I repeat
-this process back to the age of retirement to obtain optimal consumption amounts
-for each year of retirement and associated value functions. I write code in
-Julia that solves this maximisation problem.
-
-To simulate the ELSA data I solve this retirement problem for each new retiree
+bequested. This means that the value of assets at age 111 is either 0 (if we do
+not allow a bequest motive) or the value of bequests. I then take this value
+function and solve an individual's problem at age 110, choosing assets for next
+period (i.e. those to bequest) and how much to consume. This will be all assets
+for individuals with no bequest motive or assets below £20,000 ($c_{0}$).
+
+Using the policy function at age 110, which we just found by doing the
+optimisation at age 110, I calculate the value for all income and asset states
+at age 110. This value is then used to solve the problem an individual at age
+109 faces, specifically, for each income and asset state find the assets next
+period that maximise: utility from consumption; bequests multiplied by the
+probability of death; and the discounted, by $\beta$ value at age 110 multiplied
+by the probability of being alive. I repeat this process back to the age of
+retirement to obtain optimal consumption amounts for each year of retirement and
+associated value functions. I write code in Julia that solves this maximisation
+problem.
+
+To simulate the ELSA data, I solve this retirement problem for each new retiree
 in the dataset dependent on their objective probability of death each period. I
 estimate with subjective life expectancies and objective life expectancies. I
-also estimate the model both with and without a bequest motive which was picked
-to roughly fit the rates of annuitisation that are observed in the real world,
-which are around 6\%.
-
-In retiree's first year of retirement I somtimes annuitise some of their wealth.
-In practical terms, this is moving down the asset grid but up the permanent
-income grid and seeing if the value of being in that position is better than
-where the individual is currently. To assess this trade-off I calculate the
-annual annuity payment that follows from a given annuity cost. I calculate this
-using objective life tables from the ONS using the following equation:
+also estimate the model both with and without a bequest motive.
+
+In a retiree's first year of retirement, I sometimes annuitise a portion of
+their wealth. In practical terms, this is moving down the asset grid but up the
+permanent income grid. To assess this trade-off I calculate the annual annuity
+payment that follows from a given annuity cost. I calculate this, using
+objective life tables from the ONS, with the following equation:
 
@@
 and $p_{t|Retage}$ is the probability of death at age $t$ conditional on being
-age $Retage$. So individuals can move down $C$ on the asset grid for gaining $Ann$ on the
-income grid for the rest of their lives. The `money's worth' of an annuity is
-the ratio of expected present discounted value to the price of an annuity that
-an individual can purchase. Generally they have been found to be
+age $Retage$. So individuals can move down $C$ on the asset grid for gaining
+$Ann$ on the permanent income grid for the rest of their lives. The `money's
+worth' of an annuity is the ratio of expected present discounted value to the
+price of an annuity that an individual can purchase. An actuarially fair annuity
+will have a money's worth of 1 whilst an annuity that pays more than it costs
+will have a money's worth greater than 1. Generally they have been found to be
 between $0.8$ and $0.9$ for the population as a whole in the UK
@@
 
-Having found optimal policy functions and value functions I can now simulate the
+Having found optimal policy functions and value functions, I can now simulate the
 change in consumption given different rates of annuitisation versus no
@@
 were not forced to annuitise. Secondly, I can simulate the impact of different
-policy changes on the consumption, savings and bequest behaviour. For example, I
-could change the proportion of wealth that will be annuitised and see the
-corresponding changes in consumption indicators for all individuals.
-
-I now plot the retirement path of consumption for a median individual in the
-ELSA data set. I plot consumption with different proportions of starting wealth
-annuitised and the different lifecycle model types.
-
+policy changes on the consumption, savings and bequest behaviour of individuals.
+For example, I could change the proportion of wealth that will be annuitised and
+see the corresponding changes in consumption indicators for all individuals.
+
+I now plot the retirement path of consumption for a median individual in my
+subsample of ELSA. I plot consumption with different proportions of starting wealth
+annuitised (either 0\% or 50\%) and the different lifecycle model types.
+
+\begin{figure}[h]
+    \caption{Assets in retirement}
+    \centering
+    \includegraphics[width=0.7\columnwidth]{figures/asset_plot_median_retiree.pdf}
+    \label{fig:AssetPlot}
+\end{figure}
 \begin{figure}[h]
@@
 
-\begin{figure}[h]
-    \caption{Assets in retirement}
-    \centering
-    \includegraphics[width=0.7\columnwidth]{figures/asset_plot_median_retiree.pdf}
-    \label{fig:AssetPlot}
-\end{figure}
-
-Figures \ref{fig:ConsumpPlot} and \ref{fig:AssetPlot} show the retirement path
-of consumption and assets for a median retiree. This hypothetical woman has
-£120,000 of assets at age 65 when they retired in 2012 with a state pension
+Figures \ref{fig:AssetPlot} and \ref{fig:ConsumpPlot} show the retirement path
+of consumption and assets for a median retiree. This hypothetical female has
+£120,000 of assets at age 65 when she retired in 2012 with a state pension
 income of £6000. The dashed lines show consumption and assets with an annuity
-that costs half of their wealth. This buys them £3500 of annuity income, which,
-given their objective life expectancy is a money's worth of about 90\%.
-
-We can see that the asset path with the bequest motive keeps more for longer but
-still decreases, and by their late 90s they have completely run down their
+that costs half of her wealth. This buys her £3500 of annuity income, which,
+given her objective life expectancy is a money's worth of about 90\%.
+
+We can see that the asset path with the bequest motive remains higher for longer
+but still decreases, and by her late 90s she has completely run down her
 wealth. Interestingly, the asset path with the quickest run down is the standard
 lifecycle model using objective life expectancies, which points towards this
-individual being optimistic about their life chances. Also rather surprisingly
+individual being optimistic about her life chances. Also rather surprisingly,
 the asset paths of all three types are similar without annuities. The paths only
-diverge after age 75 at which point the bequest path stays high and the other
+diverge after age 75, at which point the bequest path stays high and the other
 paths start to decrease faster. This is contrary to the path with annuities
-where the standard path diverges from the bequest and subjective paths
+which show the standard path diverging from the bequest and subjective paths
 immediately.
@@
 prediction that subjective lifecycle agents consume more in early retirement
-compared to the other lifecycle models.
+compared to the other lifecycle models because people underpredict lifespan.
 
@@
 and although differences across the lifecycle are not large, there are
-significant differences at the start of retirement. This shows that different
+significant differences at the start of retirement. This means that different
 lifecycle models will produce substantive differences in predictions, so knowing
@@
 
-I first round an ELSA individual's financial wealth to the nearest point on the
-discrete asset grid, and do likewise for an individual's state pension income. I
+I now simulate consumption and asset paths for all the individuals in my ELSA
+dataset. I first round an individual's financial wealth to the nearest point on the
+discrete asset grid and do likewise for an individual's state pension income. I
 then take the treatment group and evaluate their consumption in the year of
@@
 wealth in their retirement year. I also evaluate consumption without
-annuitisation. The treatment group did not need to annuitise their wealth so I
+annuitisation. The treatment group did not need to annuitise their wealth, so I
 use their predicted consumption from the lifecycle models with their starting
@@
 nearest income grid point. I set the money's worth of annuities to $0.9$ when
-calculating annuity prices but because I round down to the nearest income grid
+calculating annuity prices, but because I round down to the nearest income grid
 point the real money's worth an individual receives is lower. I assume that
@@
 reduced. This is mainly because not everyone in the ELSA data answers the life
-expectancy questions and of those who do answer some give answers that we cannot
+expectancy questions, and of those who do answer some give answers that we cannot
 use for reasons listed in the data section.
 
-We can see the means for the control group are higher than the treated mean for
-all models but there is significant difference between the original level of
-consumption in the control group. The bequest and standard model both predict
+We can see the means for the treatment group are higher than the control group
+means for all models but there is significant difference between the consumption
+levels for the different models. The bequest and standard model both predict
 lower consumption than the subjective model which is not surprising given the
 subjective lifecycle individuals generally think their lives will be shorter and
-therefore can consume more early in retirement. Likewise, it makes sense the
-bequest model is lower than both other models since there is extra utility to be
-gained from saving so that one can bequest to their heirs.
-
-I now run the empirical models from above on this simulated data. This lets us
+therefore can consume more early in retirement. Likewise, it makes sense that
+the bequest model predicts lower consumption, since there is utility to be
+gained from saving so that an individual can bequest to their heirs.
+
+I now run the empirical models on this simulated data. This lets us
 see which lifecycle model gives us the closest match to the coefficients seen in the
-real data. We can therefore predict how other policy changes will impact
-behaviour as well as predict the full path for assets and consumption later into
-retirement. Matching a full model with method of simulated moments, though
-possible using the entire ELSA data like \textbf{cite paper eric sent me}, does
-not allow us to use incorporate policy changes into model.
+real data.
 
@@
 paper. The first column just takes those with DC pensions, the second looks at
-all individuals, whilst the third and fourth have interactions with financial
-wealth and DC pension wealth.
-
-Comparing the DC-only column across tables we can see that simulated bequest
+all individuals and uses the DD set up, whilst the third and fourth have
+interactions with financial wealth and DC pension wealth.
+
+Comparing the DC-only column across tables we can see that the simulated bequest
 model gives us the highest increase in consumption as a result of the reform.
@@
 annuity income to regain assets, so when they do not need to buy an annuity
-anymore their consumption actually increases. Surprisingly, the predictions from
-the lifecycle model using subjective life expectancies actually has a decreasing
-coefficient on the reform even though on average for a given individual
-consumption should increase as shown by the simple comparison in means. However,
-this difference could be the result of the post-reform group having higher
-subjective life expectancies, and therefore their consumption would be lower
-than the control group.
+anymore their consumption actually increases. Surprisingly, when using the predictions from
+subjective mode, the coefficient on PostReform is negative, even though the mean
+increased for this model as shown in Table \ref{tab:simulation_prediction}.
 
@@
 In the second column, the coefficient of interest is the DC pension interaction
-with the treatment variable since I use the difference in difference setup
-again. The bequest model has a coefficient of 25.6, the standard model one of
-17.7 and the standard model with subjective life expectancies has one of -30.9.
-These coefficients are higher than the coefficient seen in Table
-\ref{tab:ElsaAllData} with the predictions from the standard model matching the
-coefficient in the empirical model best. The subjective model again has the
-opposite sign to the one seen in the empirical model.
+with the treatment variable since I use the DD setup again. The bequest model
+has a coefficient of 25.5, the standard model one of 17.8 and the standard model
+with subjective life expectancies has one of -30.9. The bequest and subjective
+coefficients are
+higher than the coefficient seen in Table \ref{tab:ElsaAllData} with the
+predictions from the standard model matching the empirical model best. The
+subjective model again has the opposite sign to the one seen in the empirical
+model.
 
@@
 retirement. The structure that the Weibull distribution imposes on the
-transition probabilities could also be causing discontinuities in a given
+death probabilities could also be causing discontinuities in a given
 individuals subjective probability of death between two periods, which would
@@
 
-The evidence just presented points towards the bequest model being best able to
-replicate the consumption response in the data. As highlighted above, the
-benefit of having a theoretical model is that we can trace out future values of
-consumption for individuals, only needing to know their starting point. So using
-the bequest solution to the retirement problem, I plot the whole path of average
-consumption and assets for the ELSA treated cohort. This plot is shown in the
-figure below.
+The evidence just presented points towards predictions from the bequest model
+being best able to replicate the consumption response in the data. As
+highlighted above, the benefit of having a theoretical model is that we can
+trace out future values of consumption for individuals, only needing to know
+their starting point. So using the bequest solution to the retirement problem, I
+plot the whole path of average consumption and assets for the ELSA treated
+cohort. This plot is shown in the figure below.
 
@@
 
-\section{Appendix}
-\begin{landscape}
-    \input{tables/elsa_results_only_dc_only_exp.tex}
-\end{landscape}
-
 \bibliographystyle{chicago}

### 2023-07-26 R/10__2_summay_stats.R

@@ -86,10 +86,10 @@ header_to_add <- c(1, rep(2, length(function_names) - 1))
 
 names(header_to_add) <- function_names
 
-align_string <- function(col_names) {
+align_string <- function(col_names, alignment = "r") {
     # Left align first col
     # right align others
-    right <- rep("r", length(col_names) - 1)
+    right <- rep(alignment, length(col_names) - 1)
     # first collapse the vector
     # then join with l
     full_string <- paste0("l", paste0(right, collapse = ""))
@@ -198,7 +198,7 @@ coef_table <- kbl(cov_balance,
     booktabs = TRUE,
     format = "latex",
     caption = "Covariate Balance \\label{tab:cov_balance}",
-    align = align_string(col_names)
+    align = align_string(col_names, alignment = "d")
 )
 
 writeLines(

### 2023-07-26 R/6__fca_annuities_data.R

@@ -8,6 +8,8 @@ library(lubridate)
 
 dir("../../data/FCA_annuities", full.names = T)
 
+base_size_for_plots <- 16
+
 annuity_data_2018 <- read_excel("../../data/FCA_annuities/psd-ri-2018.xlsx",
     sheet = "Quarterly ",
     range = "B9:BD28"
@@ -73,7 +75,7 @@ annuity_plot <- ggplot(joined_data) +
     geom_point(aes(x = date, y = num_sales, colour = product_type)) +
     geom_line(aes(x = date, y = num_sales, colour = product_type)) +
     scale_x_date(date_breaks = "years", date_labels = "%Y") +
-    theme_bw() +
+    theme_bw(base_size = base_size_for_plots) +
     labs(
         x = "", y = "Sales", title = "Select Product Sales Data",
         colour = "Product type"
@@ -133,7 +135,7 @@ annuity_sizes_plot <- ggplot(
 ) +
     geom_col() +
     labs(title = "Number of pots accessed by size, 2021-22") +
-    theme_bw() +
+    theme_bw(base_size = base_size_for_plots) +
     scale_y_continuous(labels = scales::comma) +
     coord_flip()
 

### 2023-07-27 Texfiles/Draft1.tex
Potential finish

@@
 \begin{abstract}
-    I find the consumption response of early retirees to the `pensions freedom
-    act', which meant defined contribution pensions could be used for other
-    options than just purchasing an annuity. I use a discontinuity in year of
-    retirement and a difference in difference approach using those who were
-    unaffected by the policy reform as a natural control group. I then solve
-    three sets of lifecycle models from age of retirement until an agent reaches
-    110. I simulate their consumption with and without annuities, i.e. a the
-    policy reform, and match their consumption response to what was seen in the
-    empirical models. I find that a lifecycle model with a bequest motive best
-    fits this consumption response.
+    I identify the consumption responses of retirees to the UK's `pensions
+    freedom act', a landmark law change that meant defined contribution pension
+    holders were no longer forced to annuitise their funds and could instead
+    access their pensions in a variety of ways. I use a discontinuity in year of
+    retirement and a difference-in-difference approach, harnessing those who
+    were unaffected by the policy reform as a natural control group. I then
+    solve three sets of lifecycle models, tracking assets from retirement until an
+    agent reaches age 110. I simulate their consumption with and without
+    annuities, mimicking the policy reform, and match their consumption
+    responses to what was seen in the empirical models. I find that a lifecycle
+    model with a bequest motive best fits the identified consumption response —
+    a result that could inform policymakers in designing optimal pension
+    policies for an ageing population.
 \end{abstract}
@@
 mechanism is dominant. I exploit early retirees' consumption responses to a
-reform to annuity policy in the UK to add to this literature. [\textbf{Add
-            something once I have a conclusion}].
+reform to annuity policy in the UK to add to this literature.
 
@@
 defined benefit (DB) employee pension schemes, which provide an income in
-retirement based on years of service and a function of wages; and defined
+retirement based on years of service and wages; and defined
 contribution (DC) pensions, under which individuals save and invest in a
@@
 we would expect individuals to annuitise all of their wealth at retirement to
-insure against the risk of long life. And more recently
+insure against the risk of long life. More recently,
 \cite{davidoff_brown_diamond_aer_2005} show that complete annuitisation is
-optimal under a broader range of assumptions. Since then there has been much
-discussion of possible reasons for why people do not annuitise.
+optimal under a broader range of assumptions. Since then, there has been much
+discussion of possible reasons for why people choose not to annuitise.
 \cite{finkelstein_porteba_2002, finkelstein_porteba_2004} find evidence of
 adverse selection into the UK annuity market and \cite{mitchell_et_al_1999} find
-evidence of the same in the US. This means longer lived individuals select into
+evidence of the same in the US. This means longer-lived individuals select into
 buying annuities which drives up annuity prices because providers need to pay
-individuals for longer than they originally thought they would need to. This
+individuals for longer than they originally predicted they would need to. This
 makes the `money's worth' of an annuity lower for the general population as
@@
 predict annuitisation because the money's worth of annuities is not that far
-from `actuarially fair' (this is the where the present discounted value of annuities
+from `actuarially fair' (meaning the present discounted value of annuities
 equals the price that an individual pays for it).
@@
 fully explained by a mixture of bequest motives and `actuarially unfair'
-annuities. They solve an augmented life-cycle model and calculate annuity demand
-for a range of differences between rate of return assets would get in the stock
-market and the rate of return annuities give an individual. For plausible values
+annuities. They solve an augmented lifecycle model and calculate annuity demand
+for a range of differences between the rate of return assets would get in the stock
+market and the rate of return annuities give an individual. For plausible values,
 they find that individuals would optimally not annuitise much wealth. Similarly
@@
 and the general population in the American annuity market, which therefore
-implies a degree of adverse selection. But they find this cannot fully explain
-the annuitisation puzzle and only when bequest motives are added to the model
-can annuitisation rates be rationalized.
+implies a degree of adverse selection. But, they find this cannot fully explain
+the annuitisation puzzle, and only when bequest motives are added to the model
+can annuitisation rates be rationalised.
 
@@
 taken from several recent papers in the retirement savings literature. The
-bequest motives he picks matches other important aspects of empirical
-distribution of bequests such as how much individuals give and the financial
+bequest motives he picks match other important aspects of the empirical
+distribution of bequests, such as how much individuals bequeath and the financial
 characteristics of these individuals. Also, \cite{odea_sturrock_rest_2023} find
 that subjective life expectancies can explain a large part of the annuitisation
-puzzle especially if agents are less risk averse.
+puzzle, especially if agents are less risk averse.
 
 Apart from bequests and annuity prices, a range of other reasons for the annuity
-puzzle have been explored. Pre-cautionary savings for unexpected end-of-life
-medical expenses might cause individuals to prefer holding assets rather than a
-an annuity. Moreover, \cite{vidalmelia_lejarragagarcia_munich_2004} suggests
-that couples might share the risk of living for a long-time between them and
+puzzle have been explored. Precautionary savings for unexpected end-of-life
+medical expenses might cause individuals to prefer holding assets rather than
+an annuity. Moreover, \cite{vidalmelia_lejarragagarcia_munich_2004} suggest
+that couples might share the risk of long life by pooling their resources, and
 therefore couples may be less likely to annuitise their wealth.
@@
 bequest model and uncertain life expectancy, I choose also to focus on bequest
-and objective life expectancy in the UK context.
+and objective life expectancy in the UK context. To my knowledge, there has been
+no attempt to analyse the consumption response to 2014 reform, nor evaluate what
+common lifecycle models predict for the policy change.
 
@@
 second group.
-
-\textbf{conclusion: chat about couples, chat about better data on at what
-    stage of retirement individuals buy annuities}
 
@@
 
-I now run the empirical models on this simulated data. This lets us
-see which lifecycle model gives us the closest match to the coefficients seen in the
-real data.
+I now run the empirical models on this simulated data. This lets us see which
+lifecycle model gives us the closest match to the coefficients seen in the real
+data. In the structural modelling literature, it is often common to run policy
+tests on estimated models to see if simulated responses match real world
+responses, for example see \cite{mcgee_2021}. Due to the time constraints of
+this project, I did not estimate a full structural model but I do check which
+lifecycle model fits the response seen in the ELSA data.
 
@@
 with subjective life expectancies has one of -30.9. The bequest and subjective
-coefficients are
-higher than the coefficient seen in Table \ref{tab:ElsaAllData} with the
-predictions from the standard model matching the empirical model best. The
-subjective model again has the opposite sign to the one seen in the empirical
-model.
+coefficients are higher than the coefficient seen in Table \ref{tab:ElsaAllData}
+with the predictions from the standard model matching the empirical model best.
+The subjective model again has the opposite sign to the one seen in the
+empirical model.
 
@@
 reform effect in the standard and bequest models, although in both the
-interaction with financial wealth has a negative coefficient showing that the
+interaction with financial wealth has a negative coefficient, showing that the
 increase in consumption caused by the reform was progressively lower for those
-who would have had to annuitise more wealth. In the subjective model table the
+who would have had to annuitise more wealth. In the subjective model table, the
 signs are again wrong apart from for the main treatment effect in the financial
@@
 lifecycle models used. One possible explanation for this could be that the
-control and treatment groups have quite different sets of objective life
-expectancies which therefore cause different consumption paths in early
-retirement. The structure that the Weibull distribution imposes on the
-death probabilities could also be causing discontinuities in a given
-individuals subjective probability of death between two periods, which would
-impact how they consume across that part of retirement.
-
-The evidence just presented points towards predictions from the bequest model
-being best able to replicate the consumption response in the data. As
-highlighted above, the benefit of having a theoretical model is that we can
-trace out future values of consumption for individuals, only needing to know
-their starting point. So using the bequest solution to the retirement problem, I
-plot the whole path of average consumption and assets for the ELSA treated
-cohort. This plot is shown in the figure below.
-
-\textbf{Also get some plots of consumption in early retirement with and without
-    annuities for the different model types}
+control and treatment groups have quite different sets of subjective life
+expectancies, which therefore cause different consumption paths in early
+retirement. This heterogeneity in death probabilities would not occur in the
+with the other models because the objective death probabilities at a given age
+for different birth cohorts are not that large. The structure that the Weibull
+distribution imposes on the death probabilities could also be causing
+discontinuities in a given individuals subjective probability of death between
+two periods, which would impact how they consume across that part of retirement.
+
+The evidence just presented points towards simulated data using the bequest
+model being best able to replicate the consumption response observed in the
+empirical models. As highlighted above, the benefit of having a theoretical
+model is that we can trace out the asset path of consumption for individuals,
+only needing to know their retirement wealth.
 
@@
 
-I have shown that consumption in early retirement increased overall and
-particularly on
-
-\subsection{Rough plan}
-\begin{itemize}
-    \item Intro
-          \begin{enumerate}
-              \item "Latest data from HM Revenue Customs (published in April)
-                    showed more than £45bn has been taken from pots since 2015."
-                    https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
-              \item what portion of savings are because of retirement savings?
-              \item Also talk about heterogeneity across countries. Some
-                    countries want to move towards more annuitisation. This
-                    annual review is a good source of info
-                    \cite{banks_crawford_ar_2022}
-              \item https://www.moneysavingexpert.com/savings/pension-freedom/
-          \end{enumerate}
-
-          \textbf{Follow up by changing chat because the parameter estimates for
-              the diff in diff have changed a bit -- DONE but check again when I
-              have gone through it -- haven't checked the comparisons }
-    \item \textbf{Abstract}
-    \item \textbf{Add bit about testing just those people who retired when they
-              expected to retire.}
-    \item \textbf{Could add consumption and retirement year plot?}
-    \item \textbf{Maybe a small discussion of controls for the main regression
-              tables as well}
-    \item Conclusion
-\end{itemize}
+The results from the empirical section of this paper point toward there being
+some impact of the `pensions freedom act' on the consumption of retirees early
+in retirement. The specifications that used the full available data set show a
+clear increase in total monthly consumption, and particularly an increase in food
+consumption away from home and on clothes. Although, this increase of 36 a month
+only equates to 432 a year, over a longer period it may mean some retirees end
+up with fewer resources than they require toward the end of their life.
+
+Moreover, the results from the lifecycle section show not just that the models
+create different consumption predictions, but they also add to the evidence that
+lifecycle models with bequests can do a good job of explaining the consumption
+patterns of retirees. That simulated data from a bequest model matches the
+consumption response also points towards bequests being able to explain the
+annuitisation puzzle.
+
+Since the bequest motive model fits the consumption response best, we can worry
+about the increase in early retirement consumption less. This is because,
+as highlighted in the analysis of the median individual, assets will diverge
+from the objective and subjective lifecycle models during retirement, reducing
+at a slower rate. Therefore, the government should not worry about spending
+early in retirement being linked to low assets at the end of an individuals
+life.
+
+Future research is still needed to understand the full impact the reform has had
+on consumption across the whole of retirement, especially as more data on
+retirees impacted by the reform becomes available. Moreover, as the number of
+individuals retiring with a defined contribution pension grows, understanding
+the mechanisms behind their choices will matter more to policy makers. For
+example, the advice that is provided to individuals on retirement.
+
+\textbf{conclusion: chat about couples, chat about better data on at what
+    stage of retirement individuals buy annuities}
 

### 2023-07-27 Texfiles/Draft1.tex
almost final commit

@@
 
-        Word count: [Insert]
+        Word count: 8335\footnote{This is an estimate that my LaTeX editor gave me.} + 1750 (figure and table word equivalent)
         \vfill
@@
 I then test two competing hypotheses for the annuity puzzle: bequests and
-pessimistic life expectancy. The annuity puzzle was first identified by
-\cite{yaari_65} and is a classed as such since standard economic theory cannot
-explain the lack of annuitisation that is observed in the real world. Two
-possible explanations are bequests and life expectancy: individuals want to
-leave an estate for their heirs to inherit and they cannot do this if they
-annuitise their wealth; or individuals may believe that they will not live as
-long as annuity providers think they will, and therefore annuities that are on
-the market do not appear to be good value. Depending on the reason for the lack
-of annuitisation in the UK, the consumption response of retirees to the pension
-reform will differ. If individuals do not annuitise because of pessimistic life
-expectancy, their consumption should increase after the reform. If, on the other
-hand, individuals do not annuitise because of a bequest motive, consumption
-should not change as much as a result of the reform.
+pessimistic life expectancy. A bequest motive means that individuals want to
+leave an estate for their heirs to inherit -- which they cannot do if they
+annuitise their wealth. On the other hand, individuals may believe that they
+will not live as long as annuity providers think they will, and therefore
+annuities that are on the market do not appear to be good value. Depending on
+the reason for the lack of annuitisation in the UK, the consumption response of
+retirees to the pension reform will differ. If individuals do not annuitise
+because of pessimistic life expectancy, their consumption should increase after
+the reform. If, on the other hand, individuals do not annuitise because of a
+bequest motive, consumption should not change as much as a result of the reform.
 
 I will solve lifecycle models for both of these cases and simulate consumption
-decisions with and without forced annuitisation. I will then use a variety of
-empirical models to measure the consumption change in early retirement that
-resulted from the policy reform. The sign and magnitude of this change will be
-indicative of the mechanism causing the annuitisation problem.
-
-The importance of retirement policy to individuals in the UK is growing. The
+decisions with and without annuitisation. I will then apply the empirical models
+to this simulated data and measure the consumption change in early retirement
+that results from annuitisation. The sign and magnitude of these coefficients
+relative to the coefficients from the real data will be indicative of the
+mechanism causing the annuitisation puzzle.
+
+The importance of retirement policy to the UK is growing. The
 number of individuals of pensionable age is expected to increase from 11.9
@@
 lives. This has implications for government budgets, especially since population
-aging means more individuals will require expensive end of life care. On the
-other hand, if retirees over-save and do not spend, the economy may be
-dynamically inefficient -- when the capital stock is too high -- because
-individuals save too much for retirement and for bequests. In this case
-consumption per capita could be increased if savings were decreased.
+aging means more individuals may require expensive end of life care. On the
+other hand, if retirees over-save and do not spend, perhaps in order to bequest
+wealth, the economy may be dynamically inefficient -- a state in which the
+capital stock is too high. In this case consumption, per capita could be
+increased if savings were decreased.
 
@@
 saving puzzle, which seeks to explain why retirees drawdown assets slowly; and
-lifecycle models, whereby the above problems are explained using a tractable
-model of human decision making.
+the lifecycle model, the predictions of which generate the above puzzles.
 
 \cite{yaari_65} showed that under standard assumptions with risk averse agents
-we would expect individuals to annuitise all of their wealth at retirement to
-insure against the risk of long life. More recently,
+we would expect individuals to annuitise all of their wealth at retirement, so
+that they are insured against the risk of long life. More recently,
 \cite{davidoff_brown_diamond_aer_2005} show that complete annuitisation is
@@
 evidence of the same in the US. This means longer-lived individuals select into
-buying annuities which drives up annuity prices because providers need to pay
-individuals for longer than they originally predicted they would need to. This
+buying annuities, driving up annuity prices since providers need to pay
+annuitants for longer than they originally predicted. This
 makes the `money's worth' of an annuity lower for the general population as
@@
 from `actuarially fair' (meaning the present discounted value of annuities
-equals the price that an individual pays for it).
-
-\cite{friedman_warshawsky_qje_1990} show that annuitisation decisions can be
-fully explained by a mixture of bequest motives and `actuarially unfair'
-annuities. They solve an augmented lifecycle model and calculate annuity demand
-for a range of differences between the rate of return assets would get in the stock
-market and the rate of return annuities give an individual. For plausible values,
-they find that individuals would optimally not annuitise much wealth. Similarly
+equals the price that an individual pays for one).
+
+However, \cite{friedman_warshawsky_qje_1990} show that annuitisation decisions
+can be fully explained by social security and `actuarially unfair' annuities, at
+least for decisions in early retirement. They solve an augmented lifecycle model
+and calculate annuity demand for a range of money's worths. For plausible values,
+they find that individuals would optimally not annuitise much wealth. Similarly,
 to \cite{finkelstein_porteba_2004}, \cite{friedman_warshawsky_chicago_1988} show
@@
 
-More recently, \cite{lockwood_red_2012} builds on this and shows that a
+More recently, \cite{lockwood_red_2012} builds on this by showing that a
 realistic bequest motive in lifecycle simulations achieves realistic
@@
 bequest motives he picks match other important aspects of the empirical
-distribution of bequests, such as how much individuals bequeath and the financial
-characteristics of these individuals. Also, \cite{odea_sturrock_rest_2023} find
-that subjective life expectancies can explain a large part of the annuitisation
+distribution of bequests, such as how much individuals bequeath and the
+financial characteristics of these individuals. \cite{lockwood_aer_2018} also
+finds that a model that treats bequests as luxury goods fits the lack of demand
+for long-term care insurance well. \cite{odea_sturrock_rest_2023} find that
+subjective life expectancies can explain a large part of the annuitisation
 puzzle, especially if agents are less risk averse.
@@
 that couples might share the risk of long life by pooling their resources, and
-therefore couples may be less likely to annuitise their wealth.
+therefore couples may be less likely to annuitise their wealth. Another
+explanation is that individuals already have enough guaranteed income from DB
+pension schemes and state pensions that they do not need to annuitise more.
 
 Given that uncertain end-of-life medical expenses are not such an issue in the
-UK as opposed to the US, and that the focus of recent literature has been on the
-bequest model and uncertain life expectancy, I choose also to focus on bequest
-and objective life expectancy in the UK context. To my knowledge, there has been
-no attempt to analyse the consumption response to 2014 reform, nor evaluate what
-common lifecycle models predict for the policy change.
-
-\section{Data and Policy reform}
+UK, as opposed to the US, and that the focus of the recent literature has been
+on the bequest model and subjective life expectancy, I choose also to focus on
+these explanations in the UK context. To my knowledge, there has been no attempt
+to analyse the consumption response to the 2014 reform, nor evaluate predictions
+from common lifecycle models for this particular policy change.
+
+\section{Data and policy reform}
 
@@
 
-\textbf{I could go into more detail here? Maybe I should. Add equation etc}
-
 \subsection{Policy reform}
 
-Successive governments in the UK have attempted to reform both the public and
+Successive governments in the UK have reformed both the public and
 private pension system. Prior to 1987, participation in private schemes was
@@
 
-\textbf{For conclusion: future research could explore the impact of this second reform on the
-    consumption and saving decisions of retirees.}
-
 The impact of the reforms on annuity demand has been documented by
-\cite{cannon_et_al_nier_2016}. Using data from the Association of British
-Insurers, they show that annuity demand dropped by 75\% from its maximum in
-2011 to 2016. Figure \ref{fig:annovertime} shows
+\cite{cannon_et_al_nier_2016}. Figure \ref{fig:annovertime} shows
 purchases of annuities over time and demonstrates the sharp decrease in purchases
@@
 account for the drop in annuitisation as is shown by the graph.
+
 \begin{figure}[h]
@@
     \includegraphics[width=0.9\columnwidth]{figures/annuity_overtime.pdf}
+
+    \caption{Product Sales Data}
     \label{fig:annovertime}
-    \caption{New pension pots accessed}
 \end{figure}
@@
 retirement, accounting for over 50\% of pots. Prior to the policy reform this
-was not the case -- most DC pensions were accessed through
-annuities.
+was not the case -- most DC pension pots were accessed through
+annuitisation.
 
@@
 until December 2012 to implement the ruling. This change meant that annuity
-products could no longer be priced differently for men and women in the UK.
-However, given that this change was implemented two years before the pension
-freedoms act it will not impact my results.
+products could no longer be priced differently for men and women in the UK --
+changing the money's worth of an annuity. I run a robustness test in the
+empirical section to check whether this reform biases my results.
 
@@
 
-\subsection{Covariate Balance}
-
-Table \ref{tab:sum_stats}\footnote{There are hyperlinks on the table number so,
-    if your pdf viewer supports this, you can click to go to the table or
-    figure.} shows various summary statistics for each variable of interest from
-ELSA and the ONS.
-
-As required, retirement year is between 2011 and 2013 for the control group and
-2015 and 2017 for the treatment group. However, as discussed above their ELSA
-interview date are after or the same as the retirement year because we are
-tracking consumption in retirement.
-
-The second retirement group, those who retired post-reform, is small with just
-728 non missing observations for gender as opposed to 941 individuals in the
-control group. The control group are more female, retired at a slightly younger
-age and expected to retire slightly earlier. This could be because this period
-was affected by the increase in the state pension age for women from age 60 in
-2010 to age 65 in 2018. Since this change happened gradually and ocurred over
-the whole period I do not expect it to influence the results. The DifferenceAge
-row tracks the difference between expected and actual retirement age.
+\subsection{Covariate balance}
+
+Table \ref{tab:sum_stats} shows various summary statistics for each variable of
+interest from ELSA and the ONS.
+
+An individual is defined as treated if they retired between and including 2015
+and 2017, and an individual is in the control group if they retired in 2014 or
+earlier. Since I am interested in consumption in early retirement, an
+individuals interview date needs to be within the first three years of them
+retiring.
+
+The treated retirement group is smaller with just 303 non missing observations
+for gender as opposed to 765 individuals in the control group. The control group
+are more male, retired at a slightly younger age and expected to retire slightly
+earlier. This could be because this period was affected by the increase in the
+state pension age for women from age 60 in 2010 to age 65 in 2018, so early
+cohorts had slightly earlier state pension eligibility ages. But, this change
+happened gradually and ocurred over the whole period, so I do not expect it to
+influence the results. The DifferenceAge row tracks the difference between
+expected and actual retirement age.
 
 The treatment group has higher financial wealth with a median of £67,000 at the
-time of interview as opposed to £55,000 in the control group. Likewise, the
+time of interview as opposed to £54,300 in the control group. Likewise, the
 second group are more likely to have held a DC pension at some point and have
@@
 and interview year are greater for the treated group. The difference between
-expected and real retirement age is 0 which shows that individuals are not
+expected and real retirement age is 0, which shows that individuals are not
 delaying retirement, beyond their original expectation, in order to be part of
-the treatment group although retirement age is higher post reform. DC pension
-value is higher, as is financial wealth but these are both imprecisely
-estimated. Because of the differences in these variables I add them to the
-regression model.
+the treatment. DC pension value is higher, as is financial wealth but these are
+both imprecisely estimated. Because of the differences in these variables I add
+them to the regression model.
 
 One threat to validity is manipulation of the running variable, retirement year.
-This is probably quite a threat in this setting as individuals could delay
-retirement and/or the purchase of an annuity until after the reform. As observed
-above in Tables \ref{tab:sum_stats} and \ref{tab:cov_balance}, it does not
-appear as though individuals are delaying retirement. However, it is possible
-that individuals do not buy an annuity at the time of retirement since the law
-prior to 2014 only required that they access the pot by age 75. They may decide
-to keep their DC pension untouched and live off other income and assets before
-accessing it later. Since we do not have data on the exact purchase date of
-annuities, we cannot track whether annuity purchases were delayed because of
-this. However, there is relatively stable annuity demand before the policy
+As observed above in Tables \ref{tab:sum_stats} and \ref{tab:cov_balance}, it
+does not appear as though individuals are delaying retirement. However, it is
+possible that individuals do not buy an annuity at the time of retirement since
+the law prior to 2014 only required that they access the pot by age 75. They may
+decide to keep their DC pension untouched and live off other income and assets
+before accessing it later. Since I do not have data for when individuals
+purchase annuities, I cannot track whether annuity purchases were delayed for
+people in the treatment group. However, there is relatively stable annuity
+demand before the policy
 reform. If individuals delayed annuity purchases in expectation of the reform,
@@
 
+Moreover, if individuals retired and then delayed purchasing an annuity, the
+control group would have lower rates of annuitisation as well. This would bias
+my results downward, since then individuals who retired in prior to the policy
+change would also be consuming more.
+
 \section{Empirical models}
 
-In this section, I outline the key empirical models I run with first the real
-data from ELSA, and then the simulated data from the lifecycle models. I then
+In this section, I outline the key empirical models I run, first with real
+data from ELSA, and then with simulated data from the lifecycle models. I then
 see which lifecycle model better fits the consumption response that occurred as
@@
 
-I use a fuzzy, since some individuals will probably still annuitise a little,
-regression discontinuity design for which I compare the consumption of
-individuals who retired after the policy reform to that of people who retired
-just before the reform. The key assumption implicit in regression
+I use a fuzzy regression discontinuity design for which I compare the
+consumption of individuals who retired after the policy reform to that of people
+who retired just before the reform. The key assumption implicit in regression
 discontinuities is that nothing else changes at the time of the jump apart from
@@
 
-Retirement year is the running variable and individuals are treated if
-retirement year is greater than 2014 and less than or equal to 2017. I use
-consumption data of individuals up to 2 years into retirement so that the sample
-size is larger. So, if someone retired in 2015 and had consumption data in 2015
-and 2017, I include both values. An individual is considered not treated if they
-retire in or before 2013.
+I use consumption data of individuals up to 3 years into their retirement so
+that the sample size is larger. So, if someone retired in 2015 and had
+consumption data in 2015 and 2017, I include both values.
 
@@
 side of the regression equation. We can see that being in the treatment group is
-associated with £36 more a week in spending overall across all categories of
-expenditure. This is not statistically significant using robust standard errors.
-Being further into retirement at the time of interview is associated with higher
-total consumption whilst those who retired later have lower consumption. This is
-surprising since lifecycle models predict consumption decreases over retirement
-and this has also been documented empirically \citep{hurd_rohwedder_nber_2003}.
+associated with £36 more a week in spending overall. This is statistically
+significant using robust standard errors clustered at the treatment level. Being
+further into retirement at the time of interview is associated with higher total
+consumption -- this is surprising since lifecycle models predict consumption
+decreases over retirement and this has also been documented empirically
+\citep{hurd_rohwedder_nber_2003}.
 
@@
 increase and for spending on their substitutes, like food in the house relative
-to food away from home, to decrease. This income effect may be driving the
+to food away from home, to decrease. This wealth effect may be driving the
 results we observe in Table \ref{tab:DcOnlyRes} since food consumption outside the
@@
 would support this hypothesis as it shows all retirees are richer post-reform.
-To test for these cohort effects, I run a difference-in-difference (DD) regression to
-complement the regression discontinuity just shown. I add individuals who have a
-DB pension and those with no private pension into the model and interact the
-policy change with dummies for having different types of pension. Specifically,
-I run:
+To test for these cohort effects, I run a difference-in-difference (DD)
+regression to complement the regression discontinuity just shown. I add
+individuals who have a DB pension and those with no private pension into the
+model and interact the policy change with dummies for having different types of
+pension. Specifically, I run:
 
@@
 group across cohorts would have been the same were it not for the policy change.
-For DD to not be valid we would need a reason to expect that the individuals
-with a DB and no pension who are retiring post 2014 are impacted by some factor
-that did not equally impact the DC group and that would impact the consumption
-of all groups differently. Given that there were no other reforms to pension
-schemes at this time that would have only impacted one of these groups in
-particular, I claim that the parallel trends assumption holds holds.
-
-Table \ref{tab:ElsaAllData} shows these results. The post-reform variable now
-tells us that total and food consumption increased but clothing consumption
-decreased. The interactions at the bottom of the table show us what happened to
-individuals with a given pension type relative to the reference cohort which has
-no pension. Those with only a DC pension had an increase in total consumption
-which was higher than those with a DB pension and higher than those with no
-pension. Consumption on food decreased slightly for the DC group relative to no
-pension, but the decrease was less than that for the DB group. Expenditure on
-clothing also increased for the DC group compared to no pension and DB pensions.
-The increase in sample size also brings decreases in the standard errors of
-these variables -- the interactions are now all significant a the 5\% level as
-is the main coefficient on retiring post-reform.
+Therefore for DD to be invalid, we would need a reason to expect that the
+individuals with a DB and no pension who are retiring post 2014 are impacted by
+some factor that did not equally impact the DC group. Given that there were no
+other pension reforms at this time that differently affected these groups, I
+claim that the parallel trends assumption holds.
+
+Table \ref{tab:ElsaAllData} shows these results. The interactions at the bottom
+of the table show us what happened to individuals with a given pension type
+relative to the reference cohort, which has no pension. Those with only a DC
+pension had an increase in total consumption which was higher than those with a
+DB pension and higher than those with no pension. Consumption on food decreased
+slightly for the DC group relative to no pension, but the decrease was less than
+that for the DB group. Expenditure on clothing increased for the DC group
+relative to the no pension and DB pension groups. The increase in sample size
+also brings decreases in the standard errors of these variables -- the
+interactions are now all significant a the 5\% level as is the main coefficient
+on retiring post-reform.
 
@@
 something else driving the differences between the treatment group and the
-non-treated groups.
+non-treated groups. This therefore raises questions about the validity of the DD
+regression in this setting.
 
@@
 policy to a greater extent than those with smaller pensions, since they would be
-forced to annuitise a larger amount. So, I interact the policy with financial
-wealth and DC wealth. As in Table \ref{tab:DcOnlyRes}, this regression only uses
-those individuals who have a DC pension.
+forced to annuitise a larger amount. So, I interact the policy with DC wealth.
+As in Table \ref{tab:DcOnlyRes}, this regression only uses those individuals who
+have a DC pension.
 
@@
 
-\begin{landscape}
-    \linespread{1}
-
-    \input{tables/elsa_results_only_dc.tex}
-\end{landscape}
-
-\begin{landscape}
-    \linespread{1}
-
-    \input{tables/elsa_results_all_with_interact.tex}
-\end{landscape}
-
-\begin{landscape}
-    \linespread{1}
-
-    \input{tables/elsa_results_dc_pot_interact.tex}
-\end{landscape}
-
-\begin{landscape}
-    \linespread{1}
-    \input{tables/elsa_results_fin_wealth_interact.tex}
-\end{landscape}
-
 Table \ref{tab:DcOnlyInteract} shows the results. The interaction with DC
-pension size is positive, meaning that a larger DC pension pot associated with a
-greater increase in total monthly consumption. But, although the reform is again
-associated with an increase in consumption, especially for food out of the house
-and clothing, this effect is decreasing with the size of the pension pot for
-these consumption goods.
-
-The DC pension pot size variable comes from work done by the IFS. The data after
+pension size is positive, meaning that a larger DC pension pot is associated with a
+greater increase in total monthly consumption. For the all other categories,
+more pension wealth is associated with lower consumption, though for food out
+the house and for clothing the main coefficients are still positive and
+significant.
+
+As mentioned above, the DC pension pot value data after
 wave 5 has not been released yet, so for individuals who retire after 2011 I
@@
 similar pattern to Table \ref{tab:DcOnlyInteract}. The reform is associated with
-an increase in total monthly consumption and more financial wealth increases.
-
-\subsection{Robustness Checks}
-
-I carry out two checks for robustness. Firstly, I exclude those individuals who
-retired more than one year away from their expected retirement year -- I define
-an individuals expected retirement year as their expected retirement age in
-their first interview with ELSA, added to the year that the interview was
-carried out in. This checks whether the effects shown above are being driven by
-individuals either retiring early in order to buy cheaper annuities or retiring
-later and avoiding annuitisation altogether.
-
-I also check whether the EU reform to annuity pricing is changing the results in
-some way. Since the reform only came into affect in December 2012 there are some
-retirees in our sample who may have been able to purchase an annuity before
-annuity pricing was made the same for men and women. I drop anyone with a
-retirement year in 2012 or 2011 and repeat the analysis from Table
-\ref{tab:DcOnlyRes}.
-
+an increase in total monthly consumption and more financial wealth increases the
+size of this effect. As for the DC pension value variable, the interactions
+in the regressions of the other consumption variables have negative signs.
+
+\begin{landscape}
+    \linespread{1}
+    \input{tables/elsa_results_only_dc.tex}
+\end{landscape}
+
+\begin{landscape}
+    \linespread{1}
+    \input{tables/elsa_results_all_with_interact.tex}
+\end{landscape}
+\begin{landscape}
+    \linespread{1}
+    \input{tables/elsa_results_dc_pot_interact.tex}
+\end{landscape}
+
+\begin{landscape}
+    \linespread{1}
+    \input{tables/elsa_results_fin_wealth_interact.tex}
+\end{landscape}
 \begin{landscape}
@@
 
+\subsection{Robustness checks}
+
+I carry out two checks for robustness. Firstly, I exclude those individuals who
+retired more than one year away from their expected retirement year -- I define
+an individual's expected retirement year as their expected retirement age in
+their first interview with ELSA, added to the year that the interview was
+carried out in. This checks whether the effects shown above are being driven by
+individuals either retiring early in order to buy an annuity, since one may be
+cheaper pre-reform because the compulsory market reduces adverse selection
+\cite{finkelstein_porteba_2002}, or retiring later to avoid annuitisation
+altogether.
+
+I also check whether the EU reform to annuity pricing is changing the results in
+some way. Since the reform only came into effect in December 2012, there are some
+retirees in the sample who may have purchased an annuity before
+annuity pricing was made the same for men and women. I drop anyone with a
+retirement year in 2012 or 2011 and repeat the analysis from Table
+\ref{tab:DcOnlyRes}.
+
 Tables \ref{tab:DcOnlyExpOnlyRes} and \ref{tab:DcOnlyNot2012} show these
-results. In Table \ref{tab:DcOnlyExpOnlyRes}, we can see the sign is reversed on
+results. In Table \ref{tab:DcOnlyExpOnlyRes}, we see that the sign is reversed on
 PostReform in all consumption categories apart from clothing, however the sample
@@
 the sample sizes for both the robustness tests and main results are small it is
-difficult to asses what the impact of the reform was.
+difficult to assess what the true impact of the reform was.
+
+In a third set of robustness checks, I would have varied which years of
+retirement I assign to treatment or not, and varied the years of retirement that
+I collect consumption data for. But, due to space constraints I do not present
+these. The consumption response should diminish the longer the retirement year
+period either side of 2014 that we take data from, and we should not see
+positive results if we pick a `fake' treatment year, such as 2009, and run the
+same set of regressions using this year.
 
@@
 out. And, when we interact with a proxy for treatment strength (financial wealth
-or pension size) the affect of the policy is larger the stronger the treatment.
-However, the robust checks point in the other direction, and there are
+or pension size) the impact of the policy is larger the stronger the treatment.
+However, the robustness checks point in the other direction, and there are
 significant differences between the DB and no pension groups in the DD
@@
 
-Theoretically, a null change or small change in consumption is evidence for a
+Theoretically, a null or small change in consumption is evidence for a
 bequest motive explaining a lack of annuitisation rather than pessimistic life
-expectancies being the cause. In the next section I solve and simulate a
+expectancies being the cause. In the next, section I solve and simulate a
 lifecycle model for individuals in the ELSA dataset. I find that the bequest
@@
 In the second part of the paper, I first solve a modified retirement lifecycle
-model and then use data from the solved lifecycle model in similar regressions
-to the ones above.
+model, then simulate the ELSA data and run the empirical models from
+above on the implied consumption data.
 
@@
 retirees solve:
-
 \begin{equation*}
-    V(a_{t}, y, t) = \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V(a_{t+1}, y, t+1) \}
+    \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V(a_{t+1}, y, t+1) \}
 \end{equation*}
-
-subject to their budget constraint:
-
+Where the maximised value of this objective function is denoted: $V(a_{t}, y,
+    t)$. This is what I refer to as an individuals value function.
+This is maximised subject to their budget constraint:
 \begin{equation*}
-    c_{t} =a_{
-    t}(1 +r) -  a_{t+1} + y
+    c_{t} =a_{t}(1 +r) -  a_{t+1} + y
 \end{equation*}
@@
 where $a_{t}$ are asset holdings in time t, $y$ is constant income for all
-periods, $V_{t}$ is value in period $t$, $p_{t}$ is probability of death in
-period $t$ and $B()$ is a bequest function. Income comes from the state pension
-and purchased annuities. I follow the literature and use $\beta = 0.97$ as an
-individual's discount rate and $r = 1/0.97$ as the interest rate. I enter $t$
-directly into the value function as a state variable to make it clear that value
-is also conditional on age -- through the chance of death and the total number
-of periods left until terminal age.
-
-As is common in the literature \citep{lockwood_red_2012}, I use a constant relative
-risk aversion utility function.
+periods, $p_{t}$ is probability of death in period $t$ conditional on being
+alive at the start of the period. And $B()$ is a bequest function. Income comes
+from the state pension and purchased annuities. I follow the literature and use
+$\beta = 0.97$ as an individual's discount rate and $r = 1/0.97$ as the interest
+rate. I enter $t$ directly into the value function as a state variable to make
+it clear that value is also conditional on age -- through the chance of death
+and the total number of periods left until terminal age.
+
+Solving the value function gives us two `policy functions': $a_{t+1}(a_{t},
+    y,t)$ for assets, and $c_{t}(a_{t}, y, t)$ for consumption. Policy
+functions are functions of the state variables that maximise the value function.
+
+As is common in the literature \citep{lockwood_red_2012,de_nardi_et_al_2010}, I
+use a constant relative risk aversion utility function.
 \begin{equation*}
@@
 
-Where $\sigma = 2.5$.
+Where $\sigma = 2.5$ which is around the median found in the retirement savings
+literature.
 
@@
 from \cite{lockwood_red_2012,lockwood_aer_2018}.
-
 \begin{equation*}
@@
 \cite{lockwood_aer_2018} finds $c_{0}$ is in the range of $\$12,000$ to
-$\$30,000$ and $m$ is between $0.93$ and $0.96$ in a variety of models.
-
-To solve this problem, I discretise the state space. I create a grid from £500
+$\$30,000$ and $m$ is between $0.93$ and $0.96$ in a variety of estimated
+models.
+
+To solve this problem, I first discretise the state space. I create a grid from £500
 to £50,000 incrementing by £500 for income and £1,000 to £500,000 incrementing
 by £1,000 for financial assets. I solve the retirees' problem using backward
-induction. At age 110 there is certainty of death so any leftover assets are
-bequested. This means that the value of assets at age 111 is either 0 (if we do
+induction. At age 110, there is certainty of death so any leftover assets are
+bequested. This means that the value of any remaining assets at age 111 is either 0 (if we do
 not allow a bequest motive) or the value of bequests. I then take this value
@@
 period (i.e. those to bequest) and how much to consume. This will be all assets
-for individuals with no bequest motive or assets below £20,000 ($c_{0}$).
+for individuals with no bequest motive or with assets below £20,000 ($c_{0}$).
 
@@
 at age 110. This value is then used to solve the problem an individual at age
-109 faces, specifically, for each income and asset state find the assets next
-period that maximise: utility from consumption; bequests multiplied by the
-probability of death; and the discounted, by $\beta$ value at age 110 multiplied
-by the probability of being alive. I repeat this process back to the age of
-retirement to obtain optimal consumption amounts for each year of retirement and
-associated value functions. I write code in Julia that solves this maximisation
-problem.
+109 faces, specifically, for each income and asset state, they need to find the
+assets next period that maximise: utility from consumption; bequests multiplied
+by the probability of death; and thevalue function at age 110, discounted by
+$\beta$, multiplied by the probability of being alive. I repeat this process
+back to the age of retirement to obtain optimal consumption amounts for each
+year of retirement and the associated value functions. I write code in Julia
+that solves this maximisation problem.
 
 To simulate the ELSA data, I solve this retirement problem for each new retiree
-in the dataset dependent on their objective probability of death each period. I
-estimate with subjective life expectancies and objective life expectancies. I
-also estimate the model both with and without a bequest motive.
+in the dataset. I solve the problem with subjective and objective life
+expectancies, and with and without a bequest motive.
 
@@
 
-\section{Lifecycle Results}
+To see that $\delta$ is the money's worth, I rearrange the above equation:
+\begin{equation*}
+    \delta =  \frac{Ann*\sum_{t = Retage}^{110}\frac{1 - p_{t|Retage}}{(1 + r)^{t - Retage}}}{C}
+\end{equation*}
+
+\section{Lifecycle results}
 
@@
 
-I now simulate consumption and asset paths for all the individuals in my ELSA
-dataset. I first round an individual's financial wealth to the nearest point on the
-discrete asset grid and do likewise for an individual's state pension income. I
-then take the treatment group and evaluate their consumption in the year of
-interview with an annuity, given that they annuitised 50\% of their financial
-wealth in their retirement year. I also evaluate consumption without
-annuitisation. The treatment group did not need to annuitise their wealth, so I
-use their predicted consumption from the lifecycle models with their starting
-values of assets and pension income taken from their values in the ELSA dataset.
-To be clear, the only individuals for whom I annuitise wealth are those with DC
-pension pots who retired before 2014.
+To simulate consumption and asset paths for the individuals in my ELSA dataset.
+I round financial wealth and state pension income to the closest point on the
+discrete asset and pension grids respectively. I then take the treatment group
+and evaluate their consumption in the year of interview with an annuity, given
+that they annuitised 50\% of their financial wealth in their retirement year. I
+also evaluate consumption without annuitisation. The treatment group did not
+need to annuitise their wealth, so I use their predicted consumption from the
+lifecycle models with their starting values of assets and pension income taken
+from their values in the ELSA dataset. To be clear, the only individuals for
+whom I annuitise wealth are those with DC pension pots who retired before 2014.
 
@@
 matches roughly what is seen in the summary statistics, so, for the untreated
-group -- those who retired pre-reform -- I annuitise half of wealth. However,
-since some individuals cannot buy a £500 annuity, which relates to one grid
-point, with half of their financial wealth I only annuitise when an individual
-has financial assets large enough to cross this threshold. I do not let these
-individuals, with low assets, annuitise, setting consumption with and without an
-annuity to the same amount.
+group -- those who retired pre-reform -- I annuitise half of wealth. I also only
+annuitise when an individual has financial assets large enough to move up at
+least one unit of income (£500). I do not let these individuals, with low
+assets, annuitise, setting consumption with and without an annuity to the same
+amount.
 
@@
 reduced. This is mainly because not everyone in the ELSA data answers the life
-expectancy questions, and of those who do answer some give answers that we cannot
-use for reasons listed in the data section.
-
-We can see the means for the treatment group are higher than the control group
-means for all models but there is significant difference between the consumption
+expectancy questions, and of those who do answer, some give answers that I
+cannot use for reasons listed in the data section.
+
+We can see that the means for the treatment group are higher than the control
+for all models, but there is significant difference between the consumption
 levels for the different models. The bequest and standard model both predict
-lower consumption than the subjective model which is not surprising given the
+lower consumption than the subjective model. This is not surprising given the
 subjective lifecycle individuals generally think their lives will be shorter and
@@
 \begin{landscape}
-    \linespread{1.25}
-
-    \input{tables/standard_life_cycle.tex}
+    \linespread{1.5}
+
+    \input{tables/all_simulated_models.tex }
 \end{landscape}
 
-\begin{landscape}
-    \linespread{1.25}
-
-    \input{tables/beq_life_cycle.tex}
-\end{landscape}
-
-\begin{landscape}
-    \linespread{1.25}
-
-    \input{tables/subjective_life_cycle.tex}
-\end{landscape}
-
-Tables \ref{tab:StandardLifeCycle}, \ref{tab:BeqLifeCycle} and
-\ref{tab:SubjectiveLifeCycle} show the results. Each column in the table shows a
-different empirical model that was applied to the data in the first part of this
-paper. The first column just takes those with DC pensions, the second looks at
-all individuals and uses the DD set up, whilst the third and fourth have
-interactions with financial wealth and DC pension wealth.
-
-Comparing the DC-only column across tables we can see that the simulated bequest
-model gives us the highest increase in consumption as a result of the reform.
-This could be because this type of individual was previously saving out of the
-annuity income to regain assets, so when they do not need to buy an annuity
-anymore their consumption actually increases. Surprisingly, when using the predictions from
-subjective mode, the coefficient on PostReform is negative, even though the mean
-increased for this model as shown in Table \ref{tab:simulation_prediction}.
+Table \ref{tab:SubjectiveLifeCycle} shows all these results. Each set of colums shows the
+results from the empirical models using a different lifecycle model to simulate
+the data. The sub columns show the empirical modes: the first runs the
+regression discontinuity just on those with DC pensions, the second looks at
+all individuals and uses the DD set up, whilst the third interacts financial
+wealth with the treatment indicator for those with DC pensions only.
+
+Comparing the DC-only column across lifecycle models, we can see that the
+simulated bequest model gives us the highest increase in consumption as a result
+of forced annuitisation. This could be because this type of individual was
+previously saving out of their annuity income to regain assets, so when they do
+not need to buy an annuity anymore their consumption increases. Surprisingly,
+when using the predictions from subjective models, the coefficient on PostReform
+is negative, even though the mean increased for this model as shown in Table
+\ref{tab:simulation_prediction}.
 
@@
 
-The interaction models in columns 3 and 4 both have correct signs for the main
-reform effect in the standard and bequest models, although in both the
+The interaction model in column 3 has the sign for the main
+reform effect in the standard and bequest models. Although in both the
 interaction with financial wealth has a negative coefficient, showing that the
@@
 retirement. This heterogeneity in death probabilities would not occur in the
-with the other models because the objective death probabilities at a given age
-for different birth cohorts are not that large. The structure that the Weibull
-distribution imposes on the death probabilities could also be causing
-discontinuities in a given individuals subjective probability of death between
-two periods, which would impact how they consume across that part of retirement.
-
-The evidence just presented points towards simulated data using the bequest
+other models because the difference in objective death probabilities at a given
+age for different birth cohorts is not that large, whereas differences in
+subjective life expectancies between individuals can be big. The structure that
+the Weibull distribution imposes on the death probabilities could also be
+causing discontinuities in a given individuals subjective probability of death
+between two periods, which would impact how they consume across that part of
+retirement.
+
+The evidence just presented points towards data simulated using the bequest
 model being best able to replicate the consumption response observed in the
@@
 model is that we can trace out the asset path of consumption for individuals,
-only needing to know their retirement wealth.
+only needing to know their retirement wealth as we were able to do in
+Figure \ref{fig:AssetPlot}.
 
@@
 
-The results from the empirical section of this paper point toward there being
-some impact of the `pensions freedom act' on the consumption of retirees early
-in retirement. The specifications that used the full available data set show a
-clear increase in total monthly consumption, and particularly an increase in food
-consumption away from home and on clothes. Although, this increase of 36 a month
-only equates to 432 a year, over a longer period it may mean some retirees end
-up with fewer resources than they require toward the end of their life.
+The results from the empirical section of this paper point towards there being
+some impact of the UK's pensions freedom act on the consumption of retirees
+early in retirement. The specifications that used the full available dataset
+show a clear increase in total monthly consumption, and particularly an increase
+in food consumption away from home and on spending on clothes. Although the
+increase of £36 in total monthly consumption only equates to £432 a year, over
+a longer period it may mean some retirees end up with fewer resources than they
+require toward the end of their life.
 
@@
 lifecycle models with bequests can do a good job of explaining the consumption
-patterns of retirees. That simulated data from a bequest model matches the
-consumption response also points towards bequests being able to explain the
+patterns of retirees in the UK. That simulated data from a bequest model matches
+the consumption response also points towards bequests being able to explain the
 annuitisation puzzle.
@@
 Since the bequest motive model fits the consumption response best, we can worry
-about the increase in early retirement consumption less. This is because,
-as highlighted in the analysis of the median individual, assets will diverge
-from the objective and subjective lifecycle models during retirement, reducing
-at a slower rate. Therefore, the government should not worry about spending
-early in retirement being linked to low assets at the end of an individuals
-life.
+about the increase in early retirement consumption less. This is because, as
+highlighted in the analysis of the median individual, assets in the bequest model
+diverge from the objective and subjective lifecycle models during retirement,
+reducing at a slower rate. Therefore, the government should not worry about
+spending early in retirement being linked to low assets at the end of an
+individual's life. The bequest explanation also has implications for the
+macroeconomic impact of other public policy --- for example it means an
+inheritance tax could impact the savings of retirees.
 
@@
 on consumption across the whole of retirement, especially as more data on
-retirees impacted by the reform becomes available. Moreover, as the number of
-individuals retiring with a defined contribution pension grows, understanding
-the mechanisms behind their choices will matter more to policy makers. For
-example, the advice that is provided to individuals on retirement.
-
-\textbf{conclusion: chat about couples, chat about better data on at what
-    stage of retirement individuals buy annuities}
-
-\bibliographystyle{chicago}
+retirees impacted by the reform becomes available. The effect that reforms since
+this one, such as allowing individuals on DB plans to switch to DC plans, has
+had on consumption could also be studied. Moreover, as the number of individuals
+retiring with a DC pension grows, understanding the mechanisms
+behind their choices will matter more to policy makers. For example, if the
+annuitisation puzzle is caused by a bequest motive rather than low money's worth
+or short subjective life expectancies, then policies that force annuitisation or
+tax assets to fund state pensions will not be welfare enhancing.
+
+\bibliographystyle{apa}
 \bibliography{references}

### 2023-07-27 R/10__1_regressions_prep_data.R

@@ -45,7 +45,7 @@ harm_long[, ragender := fcase(
 
 harm_long[, gender_string := fcase(
     ragender == 0, "male",
-    ragender == 1, "female"
+    ragender == 1, "female" # one is female
 )]
 
 harm_long[pension_wealth_dt, dc_pot := dc_pot, on = c("idauniq", "int_year" = "year")]
@@ -160,8 +160,8 @@ data_for_regressions <- data_for_regressions[monthly_food_in < 500 | is.na(month
 data_for_regressions <- data_for_regressions[monthly_utility < 800 | is.na(monthly_utility)] # 1 row
 
 table(data_for_regressions$pre_post_ref)
-
-
+# drop early retirees
+data_for_regressions <- data_for_regressions[retired_age >= 55]
 
 fwrite(
     data_for_regressions,

### 2023-07-27 R/10__2_summay_stats.R

@@ -85,7 +85,7 @@ col_names <- names(ss_stat_out) %>%
 header_to_add <- c(1, rep(2, length(function_names) - 1))
 
 names(header_to_add) <- function_names
-
+stop()
 align_string <- function(col_names, alignment = "r") {
     # Left align first col
     # right align others

### 2023-07-27 R/13__regression_with_lifecycle.R

@@ -3,6 +3,7 @@ library(dplyr)
 library(ggplot2)
 library(stringr)
 library(modelsummary)
+library(kableExtra)
 
 source("./__ELSA_functions.R")
 clean_names <- append(clean_names, c("PostReform" = "pre_post_ref_bin"), 0)
@@ -264,8 +265,8 @@ modelsummary(
     vcov = "robust",
     title = "Simulated bequest lifecycle models \\label{tab:BeqLifeCycle}",
     coef_rename = unclean_name_vector,
-    statistic = NULL, 
-    align = align_string_ms(names(bequest_outmodels)), 
+    statistic = NULL,
+    align = align_string_ms(names(bequest_outmodels)),
     gof_map = bottom_stats
 )
 
@@ -288,8 +289,8 @@ modelsummary(
     vcov = "robust",
     title = "Simulated standard lifecycle models \\label{tab:StandardLifeCycle}",
     coef_rename = unclean_name_vector,
-    statistic = NULL, 
-    align = align_string_ms(names(standard_outmodels)), 
+    statistic = NULL,
+    align = align_string_ms(names(standard_outmodels)),
     gof_map = bottom_stats
 )
 
@@ -313,12 +314,71 @@ modelsummary(
     vcov = "robust",
     title = "Simulated subjective lifecycle models \\label{tab:SubjectiveLifeCycle}",
     coef_rename = unclean_name_vector,
-    statistic = NULL, 
-    align = align_string_ms(names(subjective_outmodels)), 
+    statistic = NULL,
+    align = align_string_ms(names(subjective_outmodels)),
     gof_map = bottom_stats
 )
 
+# -------------- all models one table
 
+all_models <- list(
+    bequest_outmodels[!grepl("PensionIn", names(bequest_outmodels))],
+    standard_outmodels[!grepl("PensionIn", names(standard_outmodels))],
+    subjective_outmodels[!grepl("PensionIn", names(subjective_outmodels))]
+) %>%
+    unlist(recursive = F)
+
+new_names <- names(all_models) %>%
+    str_replace("AllDataDCInteraction", "DCInt") %>%
+    str_replace("DCOnyFinancialInt", "FinInt")
+
+names(all_models) <- new_names
+
+# two layers of names
+# top layer is model type
+# bottom layer is
+
+top_names <- c(" " = 1, "Bequest" = 3, "Standard" = 3, "Subjective" = 3)
+tab_out <- modelsummary(
+    all_models,
[…]

### 2023-07-27 Texfiles/Draft1.tex
last before name change

@@
 
-        Word count: 8335\footnote{This is an estimate that my LaTeX editor gave me.} + 1750 (figure and table word equivalent)
+        Word count: 8336\footnote{This is an estimate that my LaTeX editor gave me.} + 1650 (figure and table word equivalent)
         \vfill
@@
 to food away from home, to decrease. This wealth effect may be driving the
-results we observe in Table \ref{tab:DcOnlyRes} since food consumption outside the
+results we observe in Table \ref{tab:DcOnlyRes}, since food consumption outside the
 home increases as does expenditure on clothing -- seen in column 5.
@@
 context, it means that consumption differences between the DB, DC and no pension
-group across cohorts would have been the same were it not for the policy change.
+groups across cohorts would have been the same were it not for the policy change.
 Therefore for DD to be invalid, we would need a reason to expect that the
@@
 at age 110. This value is then used to solve the problem an individual at age
-109 faces, specifically, for each income and asset state, they need to find the
+109 faces. Specifically, for each income and asset state, they need to find the
 assets next period that maximise: utility from consumption; bequests multiplied
-by the probability of death; and thevalue function at age 110, discounted by
+by the probability of death; and the value function at age 110, discounted by
 $\beta$, multiplied by the probability of being alive. I repeat this process
@@
 will have a money's worth of 1 whilst an annuity that pays more than it costs
-will have a money's worth greater than 1. Generally they have been found to be
+will have a money's worth greater than 1. Generally, they have been found to be
 between $0.8$ and $0.9$ for the population as a whole in the UK
@@
 or short subjective life expectancies, then policies that force annuitisation or
-tax assets to fund state pensions will not be welfare enhancing.
+tax assets to fund state pensions will not be welfare-enhancing.
 

### 2023-07-27 Texfiles/5348H.tex
final commit

--- 2023-07-27 Texfiles/Draft1.tex
@@
 
-        Word count: 8336\footnote{This is an estimate that my LaTeX editor gave me.} + 1650 (figure and table word equivalent)
+        Word count: 8336\footnote{This is an estimate that my \LaTeX editor gave me.} + 1650 (figure and table word equivalent)
         \vfill
@@
 groups across cohorts would have been the same were it not for the policy change.
-Therefore for DD to be invalid, we would need a reason to expect that the
-individuals with a DB and no pension who are retiring post 2014 are impacted by
+Therefore, for DD to be invalid, we would need a reason to expect that the
+individuals with a DB and no pension who are retiring post-2014 are impacted by
 some factor that did not equally impact the DC group. Given that there were no
@@
 bequest motive explaining a lack of annuitisation rather than pessimistic life
-expectancies being the cause. In the next, section I solve and simulate a
+expectancies being the cause. In the next section, I solve and simulate a
 lifecycle model for individuals in the ELSA dataset. I find that the bequest
@@
 
-\section{Life cycle theory}
+\section{Lifecycle theory}
 
@@
 change in consumption given different rates of annuitisation versus no
-annuitisation. There are multiple benefits to building a model of individual's
+annuitisation. There are multiple benefits to building a model of individuals'
 decisions. Firstly, we can directly compare what a given individual would have
@@
 £120,000 of assets at age 65 when she retired in 2012 with a state pension
-income of £6000. The dashed lines show consumption and assets with an annuity
-that costs half of her wealth. This buys her £3500 of annuity income, which,
-given her objective life expectancy is a money's worth of about 90\%.
+income of £6,000. The dashed lines show consumption and assets with an annuity
+that costs half of her wealth. This buys her £3,500 of annuity income, which,
+given her objective life expectancy, is a money's worth of about 90\%.
 
 We can see that the asset path with the bequest motive remains higher for longer
-but still decreases, and by her late 90s she has completely run down her
+but still decreases, and by the woman's late 90s she has completely run down her
 wealth. Interestingly, the asset path with the quickest run down is the standard
@@
 
-To simulate consumption and asset paths for the individuals in my ELSA dataset.
+To simulate consumption and asset paths for the individuals in my ELSA dataset,
 I round financial wealth and state pension income to the closest point on the

### 2023-11-16 Texfiles/Notes.tex
Potential finish
post submit edits

@@
 
+\subsection{2023-11-15}
+\begin{itemize}
+      \item Making a list of things to change and update.
+      \item Regressions -- I need to add the RV into the regression equations. I
+            guess I am actually a bit confused at what I think the RV is -- did I use
+            expected retirement age? I used retirement year.
+
+      \item I need to make the case that retirement is sticky?
+      \item I need to add retirement year as a RV -- this is a first order
+            mistake I think.
+\end{itemize}
+
 \subsection{Saturday before due date}
@@
 
+% \subsection{Rough plan}
+% \begin{itemize}
+%     \item Intro
+%           \begin{enumerate}
+%               \item "Latest data from HM Revenue Customs (published in April)
+%                     showed more than £45bn has been taken from pots since 2015."
+%                     https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
+%               \item what portion of savings are because of retirement savings?
+%               \item Also talk about heterogeneity across countries. Some
+%                     countries want to move towards more annuitisation. This
+%                     annual review is a good source of info
+%                     \cite{banks_crawford_ar_2022}
+%               \item https://www.moneysavingexpert.com/savings/pension-freedom/
+
 \bibliographystyle{chicago}

### 2023-11-16 R/10__3_run_regressions.R

@@ -1,6 +1,7 @@
 library(dplyr)
 library(ggplot2)
 library(modelsummary)
+library(data.table)
 
 source("./__ELSA_functions.R")
 
@@ -48,10 +49,45 @@ clean_names <- append(clean_names, c("PostReform" = "pre_post_ref_bin"), 0)
 ## No this is a diff in diff basically says that individuals are
 # similar apart from for the reform (after controlling for vars)
 
+names(data_for_regressions)
 
-lm(total_monthly_consumption ~ pre_post_ref_bin,
-    data = data_for_regressions
+head(data_for_regressions$idauniq)
+data_for_regressions[, .(count = .N),
+    by = .(idauniq)
+]
+
+data_for_regressions[, .(count = .N),
+    by = .(retirement_year)
+] %>%
+    ggplot(aes(
+        x = retirement_year,
+        y = count
+    )) +
+    geom_col()
+
+data_for_regressions[, .(count = .N),
+    by = .(retirement_year)
+] %>%
+    ggplot(aes(
+        x = retirement_year,
+        y = count
+    )) +
+    geom_col()
+data_for_regressions[, rv := retirement_year - 2014]
+
+
+
+
+# Ok this shold be the spec
+lm(total_monthly_consumption ~ pre_post_ref_bin + rv + rv:pre_post_ref_bin,
+    data = data_for_regressions[retirement_year >= 2012 & retirement_year <= 2016]
 )
+
+# Then I should IV buying an annuity with this
+
+data_for_regressions[retirement_year >= 2012 & retirement_year <= 2016, mean(total_monthly_consumption)]
+
+
 
 lm(total_monthly_consumption ~ pre_post_ref_bin + age_at_interview,
     data = data_for_regressions

### 2023-11-16 R/13__regression_with_lifecycle.R

@@ -84,6 +84,8 @@ align_string_ms <- function(col_names, type_alignment = "d") {
 
 
 
+
+
 #  ------------ join data together ------------
 joint_dt <- reg_dt %>%
     merge.data.table(
@@ -136,6 +138,18 @@ joint_dt %>%
     filter(fin_wealth < 150) %>%
     filter(public_pension > 5) %>%
     select(id_wave)
+
+
+
+# which ids do we NOT want
+full_directory <- dir("../../data/ELSA/lifecycle_outputs/id_wave/", full.names = T)
+files_to_keep <- joint_dt$id_wave %>%
+    lapply(function(idwave) grep(idwave, full_directory, value = T)) %>%
+    lapply(function(x) x[1]) %>%
+    unlist()
+
+files_to_remove <- full_directory[!full_directory %in% files_to_keep]
+file.remove(files_to_remove)
 
 # --------------- Save some summary statistics ---------------
 

### 2023-11-16 R/10__2_summay_stats.R
Couple of changes

@@ -116,6 +116,7 @@ writeLines(
 
 
 
+
 # table not separated by treated and control
 # all_ss_stat <- data_for_regressions %>%
 #     mutate(across(all_of(summary_cols), ~ as.numeric(.x))) %>%
@@ -143,7 +144,7 @@ writeLines(
 CreateCovBalanceFormulas <- function(variable_name) {
     string_linear <- paste0(
         variable_name,
-        " ~ pre_post_ref"
+        " ~ pre_post_ref + rv + rv:pre_post_ref "
     )
 
     formula_to_use <- formula(string_linear)
@@ -157,7 +158,7 @@ RunCovariateBalance <- function(variable_name) {
     print(variable_name)
     formula_to_use <- CreateCovBalanceFormulas(variable_name)
 
-    output <- lm(formula_to_use, data_for_regressions[ever_dc_pen == "yes"])
+    output <- lm(formula_to_use, data_for_regressions)
 
     sum_output <- summary(output)
 
@@ -175,7 +176,6 @@ RunCovariateBalance <- function(variable_name) {
     ))
 }
 
-
 cov_balance <- lapply(non_consump_cols, RunCovariateBalance) %>%
     rbindlist() %>%
     mutate(
@@ -183,7 +183,6 @@ cov_balance <- lapply(non_consump_cols, RunCovariateBalance) %>%
         std_error = sprintf("%.3f", std_error)
     ) %>%
     mutate(var_name = sapply(var_name, ReturnCleanName))
-
 
 cov_balance <- cov_balance[order(match(
     cov_balance$var_name,
@@ -201,6 +200,12 @@ coef_table <- kbl(cov_balance,
     align = align_string(col_names, alignment = "d")
 )
 
+??systemfit()
+library(data.table, verbose = T)
+
+
+install.packages("systemfit")
+installed.packages()
 writeLines(
     coef_table,
     "../Texfiles/tables/cov_balance.tex"

### 2023-11-25 Texfiles/Notes.tex
Couple of changes
trying own fstat

--- 2023-07-24 Texfiles/Notes.tex
@@
 
-\hypersetup{
-    colorlinks,
-    citecolor=black,
-    filecolor=black,
-    linkcolor=black,
-    urlcolor=black
-}
+\hypersetup{ colorlinks, citecolor=black, filecolor=black, linkcolor=black,
+    urlcolor=black }
 \setlength{\parindent}{0pt}
@@
 \begin{itemize}
-  \item Compare wealth decumulation in retirement in the US and the UK using HRS and ELSA data.
-        ELSA was modelled on HRS so they are comparable. Both follow elderly people and have similar questions
-        and methodologies. Data from 2002 to 2010.
+      \item Compare wealth decumulation in retirement in the US and the UK using
+            HRS and ELSA data. ELSA was modelled on HRS so they are comparable.
+            Both follow elderly people and have similar questions and
+            methodologies. Data from 2002 to 2010.
 
-  \item Median wealth at ages 70-74 is similar. But UK individuals tend to have slightly higher income.
-        They also have lower medical expenditures because of the NHS. But UK households unexpectedly decumulate assets more slowly.
-        This is partly due to the rise in house prices in the UK during that period.
+      \item Median wealth at ages 70-74 is similar. But UK individuals tend to
+            have slightly higher income. They also have lower medical
+            expenditures because of the NHS. But UK households unexpectedly
+            decumulate assets more slowly. This is partly due to the rise in
+            house prices in the UK during that period.
 
@@
 \begin{itemize}
-  \item Wealth in the US has more of a right-skew at the start of retirement. Medians are similar.
-  \item Primary housing wealth accounts for 2/3rds of UK retirement wealth but only 1/3 of US retirement wealth.
-  \item Life expectancies are pretty similar. US household income is higher than the UKs but it drops with age
-        whereas UK income rises with age -- this could be down to different indexation practices between the UK and US. UK pensions
-        are often indexed above prices.
+      \item Wealth in the US has more of a right-skew at the start of
+            retirement. Medians are similar.
+      \item Primary housing wealth accounts for 2/3rds of UK retirement wealth
+            but only 1/3 of US retirement wealth.
+      \item Life expectancies are pretty similar. US household income is higher
+            than the UKs but it drops with age whereas UK income rises with age
+            -- this could be down to different indexation practices between the
+            UK and US. UK pensions are often indexed above prices.
 
-  \item Median net wealth grew in the UK but fell in the US. Even once we strip out housing wealth assets are decumulated faster
-        in America. Non-housing assets are much larger in the US so this could be because in the UK people want to keep enough assets to
-        finance funeral costs. But this is still surprising.
+      \item Median net wealth grew in the UK but fell in the US. Even once we
+            strip out housing wealth assets are decumulated faster in America.
+            Non-housing assets are much larger in the US so this could be
+            because in the UK people want to keep enough assets to finance
+            funeral costs. But this is still surprising.
 
@@
             \begin{itemize}
-          \item Empirical literature in the US points to this being a channel. Less of a literature in the UK but
-                surveys have shown that people want to pass on wealth.
+                  \item Empirical literature in the US points to this being a
+                        channel. Less of a literature in the UK but surveys have
+                        shown that people want to pass on wealth.
             \end{itemize}
@@
             \begin{itemize}
-          \item \cite{de_nardi_et_al_2010} have shown that pre-cautionary saving for medical expenses explain alot of the slow decumulation
-                of assets in America.
+                  \item \cite{de_nardi_et_al_2010} have shown that
+                        pre-cautionary saving for medical expenses explain alot
+                        of the slow decumulation of assets in America.
 
-          \item There is not comparable data for late life medical expenses in the UK. And although the NHS covers some medical
-                expenses they do not cover care home costs which can be a large proportion of medical expenses.
-          \item \textit{Note to self: } \cite{french_jones_jae_2004} estimate the stochastic process for medical expenses in retirement.
+                  \item There is not comparable data for late life medical
+                        expenses in the UK. And although the NHS covers some
+                        medical expenses they do not cover care home costs which
+                        can be a large proportion of medical expenses.
+                  \item \textit{Note to self: } \cite{french_jones_jae_2004}
+                        estimate the stochastic process for medical expenses in
+                        retirement.
 
-          \item But because there is not enough data we cannot compare the US and the UK.
+                  \item But because there is not enough data we cannot compare
+                        the US and the UK.
             \end{itemize}
@@
             \begin{itemize}
-          \item \cite{Nakajima_Telyukova_ier_2020} make a life-cycle model that takes into account the extra utility from housing
-                and the fact that it is difficult to borrow against. Without housing assets they find that assets would up to 55\% lower.
-                Combination of utility from housing and housing being difficult to get rid of.
+                  \item \cite{Nakajima_Telyukova_ier_2020} make a life-cycle
+                        model that takes into account the extra utility from
+                        housing and the fact that it is difficult to borrow
+                        against. Without housing assets they find that assets
+                        would up to 55\% lower. Combination of utility from
+                        housing and housing being difficult to get rid of.
 
-          \item Later work has shown that housing is not a pre-cautionary saving tool.
-          \item In the US there is a government administered equity release scheme, the UK has several private operators. In 2011 2\%
-                of homeowners had some reverse mortgage.
+                  \item Later work has shown that housing is not a
+                        pre-cautionary saving tool.
+                  \item In the US there is a government administered equity
+                        release scheme, the UK has several private operators. In
+                        2011 2\% of homeowners had some reverse mortgage.
 
-          \item Nature of housing as an asset impacts decumulation. Empirical work has not been able to distinguish mechanisms from each
-                other -- consumption value, transaction costs or risk-return matrix.
+                  \item Nature of housing as an asset impacts decumulation.
+                        Empirical work has not been able to distinguish
+                        mechanisms from each other -- consumption value,
+                        transaction costs or risk-return matrix.
 
@@
 \begin{itemize}
-  \item Housing can explain difference between the UK and US. But so far there is not a satisfactory life-cycle model
-        that can account for the difference in asset decumulation.
+      \item Housing can explain difference between the UK and US. But so far
+            there is not a satisfactory life-cycle model that can account for
+            the difference in asset decumulation.
 \end{itemize}
@@
 \begin{itemize}
-  \item 1/3rd of US wealth is held by those over 65. Standard life-cycle models with known time of death cannot account for the
-        slow decumulation of these assets, particularly for those who are wealthy.
+      \item 1/3rd of US wealth is held by those over 65. Standard life-cycle
+            models with known time of death cannot account for the slow
+            decumulation of these assets, particularly for those who are
+            wealthy.
 
-  \item Two groups of mechanisms that could explain this. 1. \textbf{Risks} -- uncertainty over time of death and medical expenses.
-        2. \textbf{Bequests}.
+      \item Two groups of mechanisms that could explain this. 1. \textbf{Risks}
+            -- uncertainty over time of death and medical expenses. 2.
+            \textbf{Bequests}.
 
@@
 \begin{itemize}
-  \item Three main sources of retirement wealth -- pension, social security and private savings.
-  \item To understand how assets are decumulated across retirement \citeauthor*{de_nardi_et_al_are_2016} use
-        data from AHEAD and HRS. They split up the data by postretirement permanent income (PI). This is the average nonasset,
-        non means tested social security income over retirement.
+      \item Three main sources of retirement wealth -- pension, social security
+            and private savings.
+      \item To understand how assets are decumulated across retirement
+            \citeauthor*{de_nardi_et_al_are_2016} use data from AHEAD and HRS.
+            They split up the data by postretirement permanent income (PI). This
+            is the average nonasset, non means tested social security income
+            over retirement.
 
-  \item Mortality bias -- assets in the unbalanced panel are 50\% of assets in the balanced panel at the start of the
-        period.
+      \item Mortality bias -- assets in the unbalanced panel are 50\% of assets
+            in the balanced panel at the start of the period.
 
-  \item Show an interesting graph of decomposed medical expenses by age and by source of spending. Social spending
-        programs make up the difference in out-of-pocked medical spending as opposed to differences in the use of medical services.
+      \item Show an interesting graph of decomposed medical expenses by age and
+            by source of spending. Social spending programs make up the
+            difference in out-of-pocked medical spending as opposed to
+            differences in the use of medical services.
 
-  \item Medical spending has a massive right skew. Top 5\% of the distribution account for over a third of the total spend.
+      \item Medical spending has a massive right skew. Top 5\% of the
+            distribution account for over a third of the total spend.
 
-  \item Wealthy and the rich have significantly lower mortality and live longer.
-  \item The literature is very mixed on bequests -- some papers find that lots of people leave bequests whereas others find no bequests
-        for a majority of the population and then large bequests for the wealthiest.
+      \item Wealthy and the rich have significantly lower mortality and live
+            longer.
+      \item The literature is very mixed on bequests -- some papers find that
+            lots of people leave bequests whereas others find no bequests for a
+            majority of the population and then large bequests for the
+            wealthiest.
 
@@
       \item Look at the role of housing equity in the retirement saving problem.
-  \item Find, contrary to \cite{de_nardi_et_al_2010}, that medical expenses do not explain the decumulation problem.
+      \item Find, contrary to \cite{de_nardi_et_al_2010}, that medical expenses
+            do not explain the decumulation problem.
       \item This is the closest in content to the paper that I want to write.
@@
 \begin{itemize}
-  \item Three categories of explanation for the retirement saving problem (RSP) -- uncertain mortality, bequests and precautionary saving.
-  \item Parallel literature looking at housing equity in retirement (HER), find that home-owners are unlikely to draw down
-        housing equity because they like living in their home, and partial drawdowns are costly.
+      \item Three categories of explanation for the retirement saving problem
+            (RSP) -- uncertain mortality, bequests and precautionary saving.
+      \item Parallel literature looking at housing equity in retirement (HER),
+            find that home-owners are unlikely to draw down housing equity
+            because they like living in their home, and partial drawdowns are
+            costly.
 
-  \item \cite{Nakajima_Telyukova_ier_2020} augment the model from \cite{de_nardi_et_al_2010}. The NT model treats housing
-        as a different asset. They find that home ownership interacts with the bequest motive.
+      \item \cite{Nakajima_Telyukova_ier_2020} augment the model from
+            \cite{de_nardi_et_al_2010}. The NT model treats housing as a
+            different asset. They find that home ownership interacts with the
+            bequest motive.
 \end{itemize}
@@
 \begin{itemize}
-  \item Three strands to the RSP literature. The first looks at the impact of lifetime uncertainty on retirement savings.
-        Initiated by \cite{yaari_65} with recent contributions by \cite{de_nardi_et_al_2010}. The risk of outliving their wealth leads
-        individuals to decumulate more slowly than they would if time of death was certain.
+      \item Three strands to the RSP literature. The first looks at the impact
+            of lifetime uncertainty on retirement savings. Initiated by
+            \cite{yaari_65} with recent contributions by
+            \cite{de_nardi_et_al_2010}. The risk of outliving their wealth leads
+            individuals to decumulate more slowly than they would if time of
+            death was certain.
 
-  \item The second set of literature studies bequest motives. Altruistic, egoistic, and strategic motives.
-        There is also a small literature on inter-vivo.
+      \item The second set of literature studies bequest motives. Altruistic,
+            egoistic, and strategic motives. There is also a small literature on
+            inter-vivo.
 
-  \item More recently, the impact that uncertain medical expenses can have on retirement savings has been studied.
-        \cite{de_nardi_et_al_2010} show that this can explain a large part of the savings of singles.
+      \item More recently, the impact that uncertain medical expenses can have
+            on retirement savings has been studied. \cite{de_nardi_et_al_2010}
+            show that this can explain a large part of the savings of singles.
 \end{itemize}
@@
 \begin{itemize}
-  \item Housing is special because of its role as a consumption and investment good, and because it is infrequently traded due to
-        high transaction costs.
+      \item Housing is special because of its role as a consumption and
+            investment good, and because it is infrequently traded due to high
+            transaction costs.
 
-  \item There is at most a modest decline in housing wealth among older people. \cite{venti_wise_nber_2004} have shown that this is
-        caused by the subset of households that experience widowhood or a nursing home event. They conclude that
-        housing wealth is generally not used for consumption; this explains the low demand for reverse mortgages and implies we should
-        not count housing wealth as retirement wealth.
+      \item There is at most a modest decline in housing wealth among older
+            people. \cite{venti_wise_nber_2004} have shown that this is caused
+            by the subset of households that experience widowhood or a nursing
+            home event. They conclude that housing wealth is generally not used
+            for consumption; this explains the low demand for reverse mortgages
+            and implies we should not count housing wealth as retirement wealth.
 
-  \item Less moving in the UK than the US. Higher taxes and lower variation in climate and institutional choice.
-        Confirmed by \cite{Blundell_et_al_2016_fs}.
+      \item Less moving in the UK than the US. Higher taxes and lower variation
+            in climate and institutional choice. Confirmed by
+            \cite{Blundell_et_al_2016_fs}.
 
-  \item Few international studies. A couple of dutch papers have shown also that older people tend not to move or use reverse
-        mortgages.
+      \item Few international studies. A couple of dutch papers have shown also
+            that older people tend not to move or use reverse mortgages.
 
@@
 \begin{itemize}
-  \item First key contribution comes from \cite{yang_red_2009}. She builds a model that explicity models housing and its features.
-        Owner occupied housing counts contributes to utility and can be used as collateral.
+      \item First key contribution comes from \cite{yang_red_2009}. She builds a
+            model that explicity models housing and its features. Owner occupied
+            housing counts contributes to utility and can be used as collateral.
 
-  \item Households can either be owners or renters. They can become renters exogenously or endogenously. She matches the model
-        to parameter estimates from the literature and uses it to explain features of the life-cycle pattern of housing and non-housing
-        consumption.
+      \item Households can either be owners or renters. They can become renters
+            exogenously or endogenously. She matches the model to parameter
+            estimates from the literature and uses it to explain features of the
+            life-cycle pattern of housing and non-housing consumption.
 
-  \item She finds that the transaction cost associated with moving makes older households much less likely to sell their house.
-        If transaction costs are 0 then households would sell houses at a faster rate.
+      \item She finds that the transaction cost associated with moving makes
+            older households much less likely to sell their house. If
+            transaction costs are 0 then households would sell houses at a
+            faster rate.
 
-  \item But the model leaves out key things features that are in the RSP literature; death uncertainty, medical expense uncertainty and
+      \item But the model leaves out key things features that are in the RSP
+            literature; death uncertainty, medical expense uncertainty and
             bequests which could also explain the pattern.
 
-  \item \cite{Nakajima_Telyukova_ier_2020} adds housing consumption to the model from \cite{de_nardi_et_al_2010}.
-        Quite a complicated model but I don't see why I could not copy it. \citeauthor*{Nakajima_Telyukova_ier_2020} have a single health
-        shock as opposed to a medical expense shock.
+      \item \cite{Nakajima_Telyukova_ier_2020} adds housing consumption to the
+            model from \cite{de_nardi_et_al_2010}. Quite a complicated model but
+            I don't see why I could not copy it.
+            \citeauthor*{Nakajima_Telyukova_ier_2020} have a single health shock
+            as opposed to a medical expense shock.
 
-  \item NT find that bequest motives and housing are the big drivers of retirement saving. But \cite{de_nardi_et_al_2010} only use
-        singles and they do not match the housing distribution in their method of simulated moments. (Is this process, in general, robust
-        to choosing different moments?)
+      \item NT find that bequest motives and housing are the big drivers of
+            retirement saving. But \cite{de_nardi_et_al_2010} only use singles
+            and they do not match the housing distribution in their method of
+            simulated moments. (Is this process, in general, robust to choosing
+            different moments?)
 
@@
       \item Authors consider three different motivations for bequests.
-  \item \textbf{Altruistic} NT model bequest motvie as an egoistic motive. Households just want to have positive
-        wealth at death no matter who they are passing on the bequests to. An altruistic bequester models the utility of the heirs and
-        seeks to maximise this. With concave utility, bequests are decreasing in the life time income of the heirs because their marginal
-        utility from money is lower.
+      \item \textbf{Altruistic} NT model bequest motvie as an egoistic motive.
+            Households just want to have positive wealth at death no matter who
+            they are passing on the bequests to. An altruistic bequester models
+            the utility of the heirs and seeks to maximise this. With concave
+            utility, bequests are decreasing in the life time income of the
+            heirs because their marginal utility from money is lower.
 
-  \item \textbf{Strategic}. Bequests should be seen in the context of inter-generational transfers where parents care about
-        what their children do for them in old age. Re-write utility with a parameter which increases parents utility but decreases the heirs.
-        Some evidence for this channel but I think it is a bit shaky. There is also some evidence that inter-vivo transfers increase if someone
-        is a caregiver.
+      \item \textbf{Strategic}. Bequests should be seen in the context of
+            inter-generational transfers where parents care about what their
+            children do for them in old age. Re-write utility with a parameter
+            which increases parents utility but decreases the heirs. Some
+            evidence for this channel but I think it is a bit shaky. There is
+            also some evidence that inter-vivo transfers increase if someone is
+            a caregiver.
 
@@
 \begin{itemize}
-  \item Freedom and choice in pensions announced in March \cite{pen_freedoms_hmt_2014}. Gains royal
-        assent in December becoming the Taxation of Pensions Act. \cite{tax_pensions_hmg_2014}
+      \item Freedom and choice in pensions announced in March
+            \cite{pen_freedoms_hmt_2014}. Gains royal assent in December
+            becoming the Taxation of Pensions Act. \cite{tax_pensions_hmg_2014}
 
@@
 \begin{itemize}
-  \item Interesting point re annuities being low because of state pension.
-        I could test consumption response to forced annuitisation if that is the
-        case as well. I think it would just mean increased spending?
+      \item Interesting point re annuities being low because of state pension. I
+            could test consumption response to forced annuitisation if that is
+            the case as well. I think it would just mean increased spending?
 
-  \item Make a good point that other explanations for the AP are to do with unfairly priced
-        annuities. This could be subjectively or not and it would have the same impact (or
-        a mixture of the two). So I don't think there is massive need in estimating subjective
-        survival curves.
+      \item Make a good point that other explanations for the AP are to do with
+            unfairly priced annuities. This could be subjectively or not and it
+            would have the same impact (or a mixture of the two). So I don't
+            think there is massive need in estimating subjective survival
+            curves.
 
-  \item I think there has been some work by cannon and tokes on estimating this in the
-        early 2010s so I am not sure whether there is point me trying to do this again using the
-        CMI annuitants tables.
+      \item I think there has been some work by cannon and tokes on estimating
+            this in the early 2010s so I am not sure whether there is point me
+            trying to do this again using the CMI annuitants tables.
 
-  \item Also test model with both bequests and unfair annuities. This seems most obvious tbh.
-        Maybe whole idea is rubbish because of this.
+      \item Also test model with both bequests and unfair annuities. This seems
+            most obvious tbh. Maybe whole idea is rubbish because of this.
 
-  \item Could I compare how a mixture of load factors vs a mixture of subjective death probs
-        vs a mixture of adverse selection interacts. Disentangling the extent to which each causes
-        people not to annuitise would be difficult though. But can use evidence from ELSA and CMI for
-        subjective and adverse selection amounts respectively.
+      \item Could I compare how a mixture of load factors vs a mixture of
+            subjective death probs vs a mixture of adverse selection interacts.
+            Disentangling the extent to which each causes people not to
+            annuitise would be difficult though. But can use evidence from ELSA
+            and CMI for subjective and adverse selection amounts respectively.
 
-  \item \begin{equation*}
-          \mathbb{E}(U) = \sum_{0}^{w - x - 1}[p_{t}U_{t}(c_{t}) + p_{t}q_{t}V_{t+1}(G_{t+1})]
+      \item \begin{equation*} \mathbb{E}(U) = \sum_{0}^{w - x -
+                  1}[p_{t}U_{t}(c_{t}) + p_{t}q_{t}V_{t+1}(G_{t+1})]
             \end{equation*}
-        Where $V_{t+1}(G_{t+1})$ is the utility from bequests next period and $q_{t}$ is the probability of death.
+            Where $V_{t+1}(G_{t+1})$ is the utility from bequests next period
+            and $q_{t}$ is the probability of death.
 
-  \item Does some simulations showing how bequests and acturarial unfairness interact with each other.
-        This type of table could be a good thing to include in my paper. Then a section showing implied consumption
-        changes and then the data work. I think I am kinda close to being able to make that anyway.
+      \item Does some simulations showing how bequests and acturarial unfairness
+            interact with each other. This type of table could be a good thing
+            to include in my paper. Then a section showing implied consumption
+            changes and then the data work. I think I am kinda close to being
+            able to make that anyway.
 
@@
       \item \textit{" but investigating it seriously would require data on the
-          degree of actuarial unfairness in the pricing of annuities not just for
-          individuals but for double or even multiple lives."} This is interesting.
-        Could I compare annuitisation decisions of couples and singles? Singles
-        should be more likely to buy annuities if this is true.
+                  degree of actuarial unfairness in the pricing of annuities not
+                  just for individuals but for double or even multiple lives."}
+            This is interesting. Could I compare annuitisation decisions
+            of couples and singles? Singles should be more likely to buy
+            annuities if this is true.
 
@@
             two person models? Probably? But I think joint annuities would face
-        more than double the problem with actuarially unfair annuities worse --
-        impact would be multiplied and not summed.
+            more than double the problem with actuarially unfair annuities worse
+            -- impact would be multiplied and not summed.
 
-  \item Have not seen this before? Could be interesting. I think Eric solved a life-cycle model
-        with a couple in so I could try and look at that paper. (No I think it was another paper from S130)
+      \item Have not seen this before? Could be interesting. I think Eric solved
+            a life-cycle model with a couple in so I could try and look at that
+            paper. (No I think it was another paper from S130)
 
@@
 \begin{itemize}
-  \item Solve life cycle models again. The variable of interest they use is the
-        loading factor.
+      \item Solve life cycle models again. The variable of interest they use is
+            the loading factor.
 
       \item Need bequest motives to get annuity rates that are observed.
-  \item "See \cite{friedman_warshawsky_nber_1985} for a brief discussion of several
-        of these other possible explanations." that is for other explanations for
-        lack of annuitisation. Had a look at this and it wasn't super useful. Just
-        seemed like a first draft of the paper.
+      \item "See \cite{friedman_warshawsky_nber_1985} for a brief discussion of
+            several of these other possible explanations." that is for other
+            explanations for lack of annuitisation. Had a look at this and it
+            wasn't super useful. Just seemed like a first draft of the paper.
 
@@
 \begin{itemize}
-  \item Literature (Abel 2004?? Find when landed) claims that bequest motives
-        can only explain part of the AP. But Lockwood shows that it can explain why people
-        annuitise no wealth given loading factors (prices etc) of annuities.
+      \item Literature (Abel 2004?? Find when landed) claims that bequest
+            motives can only explain part of the AP. But Lockwood shows that it
+            can explain why people annuitise no wealth given loading factors
+            (prices etc) of annuities.
 
-  \item  (Davidoff et al., 2005??) is another paper on annuities I should read.
+      \item  (Davidoff et al., 2005??) is another paper on annuities I should
+            read.
 
@@
 
+\section{Lee and Lemiuex}
+\begin{itemize}
+      \item Hahn, Todd, and van der Klaauw (2001) -- more theoretical
+      \item RD should be seen as a data generating process rather than a method
+            (i.e IV is a method, RCT is a data generating process)
+
+      \item We want to test for covariates around the cutoff, and we want to
+            test for density. In my setting we can't really test for density since
+            this will be impacted by sample sizes, but we can test for
+            covariates which I haven't really done poperly.
+
+      \item McCrary (2008) looks like I should read this as well.
+\end{itemize}
+
 \section{Personal notes}
+\subsection{2023-11-25}
+\begin{itemize}
+      \item RV is retirement year
+      \item First thing is to run a proper covariate balance test.
+      \item the re-run all specs
+      \item then rewrite the results
+      \item add in some discussion of cohort affects
+      \item then think about using an instrument for retirement year
+      \item then write out the two period models for standard lifecycle,
+            pessimisitic survival and bequests
+      \item
+\end{itemize}
+
+\subsection{2023-11-15}
+\begin{itemize}
+      \item Making a list of things to change and update.
+      \item Regressions -- I need to add the RV into the regression equations. I
+            guess I am actually a bit confused at what I think the RV is -- did
+            I use expected retirement age? I used retirement year.
+
+      \item I need to make the case that retirement is sticky?
+      \item I need to add retirement year as a RV -- this is a first order
+            mistake I think.
+
+      \item Can I IV for retirement as well? What would this look like? using
+            predicted retirement year as the running variable.
+
+\end{itemize}
 
@@
       \item Thistlethwaite and Campbell, 1960 -- RD literature
-  \item Regression Discontinuity Inference with Specification Error -- Lee and Card. Seems like an interesing paper
-  \item Should I use financial wealth variables at retirement instead of at interview? I guess in alot of cases I wont
-        know this. I could check the differences for the people that I have pre and post for.
+      \item Regression Discontinuity Inference with Specification Error -- Lee
+            and Card. Seems like an interesing paper
+      \item Should I use financial wealth variables at retirement instead of at
+            interview? I guess in alot of cases I wont know this. I could check
+            the differences for the people that I have pre and post for.
       \item I also need to see what happens if I include family size.
-  \item Ok data work is done. Now I should need to work on julia data. I think I need to run some more subjectives
+      \item Ok data work is done. Now I should need to work on julia data. I
+            think I need to run some more subjectives
 \end{itemize}
@@
       \item What will home ownership change?
-  \item How to deal with top wealth deciles? - at the moment they are excluded because of the grid.
+      \item How to deal with top wealth deciles? - at the moment they are
+            excluded because of the grid.
       \item Include other sources of income i.e. Got Dc pension
       \item Run model just with pension wealth and not financial wealth
-  \item We could probably generate non annuitisation by adding some fixed cost.
-  \item OK, the financial wealth variable I use is quite comprehensive including all non-housing
-        wealth which is good.
+      \item We could probably generate non annuitisation by adding some fixed
+            cost.
+      \item OK, the financial wealth variable I use is quite comprehensive
+            including all non-housing wealth which is good.
       \item Deflator?
@@
 \begin{itemize}
-  \item Pension wealth variables from IFS -- I could illustrate
-        the method with just financial wealth and then use pension wealth calculations
-        as well. Pension wealth gives us retirement income. This leaves financial wealth to be
-        invested in annuities.
+      \item Pension wealth variables from IFS -- I could illustrate the method
+            with just financial wealth and then use pension wealth calculations
+            as well. Pension wealth gives us retirement income. This leaves
+            financial wealth to be invested in annuities.
 
-  \item Apparently there are pension wealth variables for DC pensions?
-        Cant find in harmonized but perhaps they are in the pension files?
+      \item Apparently there are pension wealth variables for DC pensions? Cant
+            find in harmonized but perhaps they are in the pension files?
 
-  \item Could subjective survival probs also explain slow asset decumulation. And then by the time they are old
-        they have medical risks.
+      \item Could subjective survival probs also explain slow asset
+            decumulation. And then by the time they are old they have medical
+            risks.
 
-  \item "Allowing individuals to solve optimisation problems using hueristics"
+      \item "Allowing individuals to solve optimisation problems using
+            hueristics"
 \end{itemize}
@@
                   \item Write behaviourel model code need to do
-          \item Run lifecycle models with different starting points - this is to do.
-                Do we have size of pension though? No. Only size of income.
+                  \item Run lifecycle models with different starting points -
+                        this is to do. Do we have size of pension though? No.
+                        Only size of income.
 
-          \item Do I want to add defined pension income to the model? Yes I think so.
-          \item Can also check the consumption responses of individuals who think they are going to leave a bequest.
+                  \item Do I want to add defined pension income to the model?
+                        Yes I think so.
+                  \item Can also check the consumption responses of individuals
+                        who think they are going to leave a bequest.
             \end{enumerate}
@@
 \begin{itemize}
-  \item Good idea about couples. Good idea about table showing \% of annuititants
-        with param values and then also consumption changes.
+      \item Good idea about couples. Good idea about table showing \% of
+            annuititants with param values and then also consumption changes.
 
-  \item How to split consumption? Guess have more grids? -- Don't need to since only one
-        continuation value and there is no heterogeneity between partners.
+      \item How to split consumption? Guess have more grids? -- Don't need to
+            since only one continuation value and there is no heterogeneity
+            between partners.
       \item How do behaviourel models change the DP problem. Could present bias
@@
 \begin{itemize}
-  \item Trying to understand the weibull code from \cite{odea_sturrock_rest_2023}. Non-linear
-        least squares
+      \item Trying to understand the weibull code from
+            \cite{odea_sturrock_rest_2023}. Non-linear least squares
       \item Big docs have questions in that we can use.
@@
 \begin{itemize}
-  \item Do I have death data? check UKDS.
-        Harmonized data does have year of death but only for very few
-        people. It doesn't have it for the majority of exits.
+      \item Do I have death data? check UKDS. Harmonized data does have year of
+            death but only for very few people. It doesn't have it for the
+            majority of exits.
 
-  \item I can change pension for individuals and asset levels and subjective life expectancy.
+      \item I can change pension for individuals and asset levels and subjective
+            life expectancy.
       \item
@@
       \item Also check
-  \item should think about the impact of housing on the effects of the change
-        -- will their be heterogeneity
+      \item should think about the impact of housing on the effects of the
+            change -- will their be heterogeneity
 
@@
                   \item add gender into life cycle
-          \item get subjective life expectancies -- have got the code from
-                david and cormac but I think they have birth dates and death dates which
-                arent included in the normal elsa data.
+                  \item get subjective life expectancies -- have got the code
+                        from david and cormac but I think they have birth dates
+                        and death dates which arent included in the normal elsa
+                        data.
                   \item generate correct rates of annuitisation
@@
 
-  \item Continue working on retirement lifecycle code.
-        Once we have a working version of a simple model we can then attempt
-        to add the extensions.
+      \item Continue working on retirement lifecycle code. Once we have a
+            working version of a simple model we can then attempt to add the
+            extensions.
 
-  \item I think I should also check the \cite{odea_sturrock_rest_2023} paper,
-        I think they just use a simple model without bequests.
+      \item I think I should also check the \cite{odea_sturrock_rest_2023}
+            paper, I think they just use a simple model without bequests.
 
@@
 
-  \item Plot consumption - done. Also sorted out some problems with the code.
-  \item Then I need to augment with annuity products -
-        I should be able to do this. At age 65 they can make a decision to buy an annuity or not.
-        Check expected utility from consumption path.
-        Then do the same with bequests.
-        Shall I asked ODea and Sturrock for codes?
+      \item Plot consumption - done. Also sorted out some problems with the
+            code.
+      \item Then I need to augment with annuity products - I should be able to
+            do this. At age 65 they can make a decision to buy an annuity or
+            not. Check expected utility from consumption path. Then do the same
+            with bequests. Shall I asked ODea and Sturrock for codes?
 
-  \item Lockwood makes a discrete space that is exponential.
-        Seems a bit strange tbh. I'd like to look at Odea's code.
-        How to draft the email?
+      \item Lockwood makes a discrete space that is exponential. Seems a bit
+            strange tbh. I'd like to look at Odea's code. How to draft the
+            email?
 
-  \item So the grid of assets is created from the consumption choices?
-        -- Yes I think that is how it works and then consumption should always be in the grid.
+      \item So the grid of assets is created from the consumption choices? --
+            Yes I think that is how it works and then consumption should always
+            be in the grid.
 
       \item Also should compare to income drawdown. Most pensioners have income
-        below the taxable amount, and therefore the optimal amount of drawdown
-        should not be impacted by taxes.
+            below the taxable amount, and therefore the optimal amount of
+            drawdown should not be impacted by taxes.
       \item And state pension?
@@
 \begin{itemize}
-  \item Not sure what I should work ok. I could do more data work or work on the simulation part.
-        I think simulation part. Data is probably not far from being done tbh.
+      \item Not sure what I should work ok. I could do more data work or work on
+            the simulation part. I think simulation part. Data is probably not
+            far from being done tbh.
 
-  \item Aim is to get a small working version just with risky retirement life expectancy.
+      \item Aim is to get a small working version just with risky retirement
+            life expectancy.
       \item
@@
 \begin{itemize}
-  \item Pretty bad week of work. Understand the data a bit better.
-        But have made little progress elsewhere. I want to work on my Julia code
-        to replicate the analysis in \cite{odea_sturrock_rest_2023} and \cite{lockwood_red_2012}.
+      \item Pretty bad week of work. Understand the data a bit better. But have
+            made little progress elsewhere. I want to work on my Julia code to
+            replicate the analysis in \cite{odea_sturrock_rest_2023} and
+            \cite{lockwood_red_2012}.
 
-  \item Lets have a look at the matlab code from Lockwood. Is actually kinda hard but I think doable.
-        That would be the bit I enjoy the most.
-        Why do people keep on coming up with new life-cycle models though.
+      \item Lets have a look at the matlab code from Lockwood. Is actually kinda
+            hard but I think doable. That would be the bit I enjoy the most. Why
+            do people keep on coming up with new life-cycle models though.
       \item
-  \item What are the implications for the reason no annuities being the returns in the stockmarket.
-        - could someone look at actual returns that retirees get. Presumably many switch to portfolios with bonds.
+      \item What are the implications for the reason no annuities being the
+            returns in the stockmarket. - could someone look at actual returns
+            that retirees get. Presumably many switch to portfolios with bonds.
 
       \item I think the implication would be increased
-  \item \cite{inkman_et_al_rfs_2011} -- reading this now. Say that there is an annuity question in the
-        Income and Assets section of the ELSA data set. "IAAIM" this variable. Makes sense now how they got
-        such a small amount. No way to tell whether people are using income drawdown products or not.
-        Guess we could see if non-annuity income increases after.  "IAPPEN" then this variable is the amount of
-        pension income which includes annuity income. Not sure whether we should see this go up or down around 2013.
+      \item \cite{inkman_et_al_rfs_2011} -- reading this now. Say that there is
+            an annuity question in the Income and Assets section of the ELSA
+            data set. "IAAIM" this variable. Makes sense now how they got such a
+            small amount. No way to tell whether people are using income
+            drawdown products or not. Guess we could see if non-annuity income
+            increases after.  "IAPPEN" then this variable is the amount of
+            pension income which includes annuity income. Not sure whether we
+            should see this go up or down around 2013.
 
@@
 \begin{itemize}
-  \item Get familiar with the pension data and the different pension variables. \cite{odea_sturrock_rest_2023} -- in this paper they calculate money's
-        worth using subjective mortality risk as opposed to normal life tables. No simulations, just
+      \item Get familiar with the pension data and the different pension
+            variables. \cite{odea_sturrock_rest_2023} -- in this paper they
+            calculate money's worth using subjective mortality risk as opposed
+            to normal life tables. No simulations, just
       \item
@@
 \begin{itemize}
-  \item Dear Eric, thanks for your advice -- I'll stay clear of trying to estimate a savings model. That paper looks
-        very comprehensive so I am not sure what I could add to the topic in four weeks.
+      \item Dear Eric, thanks for your advice -- I'll stay clear of trying to
+            estimate a savings model. That paper looks very comprehensive so I
+            am not sure what I could add to the topic in four weeks.
 
-        I've come back to the ideas around the reform to annuities and how it fits with the implications that different models
-        of retirement saving would predict. Models that have been estimated, such as the Lockwood paper, give quite sharp
-        predictions about what should happen to savings and consumption in retirement when the annuity constraint is lifted.
+            I've come back to the ideas around the reform to annuities and how
+            it fits with the implications that different models of retirement
+            saving would predict. Models that have been estimated, such as the
+            Lockwood paper, give quite sharp predictions about what should
+            happen to savings and consumption in retirement when the annuity
+            constraint is lifted.
 
-        My analysis could take several of the most recent models in the lifecycle literature that explain retirement saving
-        and estimate the expected changes in consumption of retirees in the first few years of retirement conditional on not
-        being forced to annuitise pension wealth
+            My analysis could take several of the most recent models in the
+            lifecycle literature that explain retirement saving and estimate the
+            expected changes in consumption of retirees in the first few years
+            of retirement conditional on not being forced to annuitise pension
+            wealth
 
-  \item If Lockwood paper is correct and bequests are the driver of retirement savings then we should see a decrease in savings
-        from income. If individuals dont buy annuities because they aren't priced fairly for them then we should see a rise in consumption
-        when annuities are no longer compulsory. One section of the paper could fit these two simple models of non-annuitisation using
-        the drop in demand for annuities.
+      \item If Lockwood paper is correct and bequests are the driver of
+            retirement savings then we should see a decrease in savings from
+            income. If individuals dont buy annuities because they aren't priced
+            fairly for them then we should see a rise in consumption when
+            annuities are no longer compulsory. One section of the paper could
+            fit these two simple models of non-annuitisation using the drop in
+            demand for annuities.
 
-  \item ok found quite an interesting new paper \cite{inkman_et_al_rfs_2011}. Look
-        at annuity demand and characteristics that are associated with it. Also,
-        estimate a life-cycle model that can explain the portfolio choices
-        of individuals in the data.
+      \item ok found quite an interesting new paper
+            \cite{inkman_et_al_rfs_2011}. Look at annuity demand and
+            characteristics that are associated with it. Also, estimate a
+            life-cycle model that can explain the portfolio choices of
+            individuals in the data.
 
-  \item Mmm could I also compare those with a DB pension and those with a DC pension.
-        Nothing changed for DB pension pots. Could I do a diff in diff using the DB pension group
-        as the control? I think that would work.
+      \item Mmm could I also compare those with a DB pension and those with a DC
+            pension. Nothing changed for DB pension pots. Could I do a diff in
+            diff using the DB pension group as the control? I think that would
+            work.
 
-  \item Have a look at the pension data today. Can we distinguish the point at which individuals retire?
-        When they convert pension to annuity income? Or when they keep on running it down as normal?
+      \item Have a look at the pension data today. Can we distinguish the point
+            at which individuals retire? When they convert pension to annuity
+            income? Or when they keep on running it down as normal?
 
@@
 
-  \item Does consumption increase with bequest motive as well? Only if people under forced bequests were
-        saving more at the start of retirement to regain what they were forced to annuitise.
+      \item Does consumption increase with bequest motive as well? Only if
+            people under forced bequests were saving more at the start of
+            retirement to regain what they were forced to annuitise.
 
@@
       \item But I still need a better idea. Read Lee Lockwood RED paper.
-  \item Also figure out pre-retirement pension DC pots. i.e. what they can spend
-        on annuities.
+      \item Also figure out pre-retirement pension DC pots. i.e. what they can
+            spend on annuities.
 
-  \item I think a basic easier question would be what impact has the change had on consumption profiles.
+      \item I think a basic easier question would be what impact has the change
+            had on consumption profiles.
       \item Could I estimate a small model somehow?
 
-  \item New idea is to estimate reduced form estimates of the impact compulsory annuitisation
-        had on wealth.
+      \item New idea is to estimate reduced form estimates of the impact
+            compulsory annuitisation had on wealth.
 
@@
 \begin{itemize}
-  \item There is data on consumption and spending in ELSA. The technical reports have good information on what is available
-        in the data sets.
+      \item There is data on consumption and spending in ELSA. The technical
+            reports have good information on what is available in the data sets.
 
-  \item Questions: Is health expenditure data available? There is some social care cost data but there is not any medical expense data.
-        Or I cannot find it at least.
+      \item Questions: Is health expenditure data available? There is some
+            social care cost data but there is not any medical expense data. Or
+            I cannot find it at least.
 
@@
             \begin{enumerate}
-          \item Minimum distance estimation and method of simulated moments in Julia? Solving the model in Julia?
+                  \item Minimum distance estimation and method of simulated
+                        moments in Julia? Solving the model in Julia?
                         \begin{enumerate}
-                  \item Minimum distance theory vs method of simulated moments -- check Hansen.
-                  \item Perhaps get a small working example going and then play around with it. There are some on quant econ that I can use \hyperref{text}
-                        https://github.com/fediskhakov/CompEcon -- this has some nice python notebooks that I could try and replicate.
-                        I do not know how to start this though? Should I just start trying to bang away at it?
-                        Do I even understand how the method works? Do I solve for optimal policy functions for a given
-                  \item Good textbook treatments of SMM are found in Adda and Cooper (2003, pp. 87-100) and Davidson and MacKinnon (2004, pp. 383-394).
+                              \item Minimum distance theory vs method of
+                                    simulated moments -- check Hansen.
+                              \item Perhaps get a small working example going
+                                    and then play around with it. There are some
+                                    on quant econ that I can use \hyperref{text}
+                                    https://github.com/fediskhakov/CompEcon --
+                                    this has some nice python notebooks that I
+                                    could try and replicate. I do not know how
+                                    to start this though? Should I just start
+                                    trying to bang away at it? Do I even
+                                    understand how the method works? Do I solve
+                                    for optimal policy functions for a given
+                              \item Good textbook treatments of SMM are found in
+                                    Adda and Cooper (2003, pp. 87-100) and Davidson
+                                    and MacKinnon (2004, pp. 383-394).
                         \end{enumerate}
@@
                   \item Medical expense data?
-          \item Read \cite{banks_et_al_aej_2019} -- this looks at differences in consumption across retirement in the US and the UK.
-          \item Read \cite{gourinchas_parker_ec_2003} -- big contribution to the field it seems.
+                  \item Read \cite{banks_et_al_aej_2019} -- this looks at
+                        differences in consumption across retirement in the US and the
+                        UK.
+                  \item Read \cite{gourinchas_parker_ec_2003} -- big
+                        contribution to the field it seems.
                   \item \textbf{Scrap this idea. Eric advised against it.}
@@
 
-  \item I am back to the annuity idea. I could document changes in annuities in ELSA and FES and also
-        see if consumption changed because of this. I like the idea of fitting a bequests model to the change in annuitisation
-        rates like Lee Lockwood did but I do not know if that is possible. What moments would I match it to?
+      \item I am back to the annuity idea. I could document changes in annuities
+            in ELSA and FES and also see if consumption changed because of this.
+            I like the idea of fitting a bequests model to the change in
+            annuitisation rates like Lee Lockwood did but I do not know if that
+            is possible. What moments would I match it to?
 
-  \item I have just broken my version of R annoyingly.
-        I want to have a separate set of packages downloaded for my thesis.
-        Done and it is working again nicely now.
+      \item I have just broken my version of R annoyingly. I want to have a
+            separate set of packages downloaded for my thesis. Done and it is
+            working again nicely now.
 
 \end{itemize}
+
+% \subsection{Rough plan} \begin{itemize} \item Intro \begin{enumerate} \item
+% "Latest data from HM Revenue Customs (published in April) showed more than
+% £45bn has been taken from pots since 2015."
+% https://www.ftadviser.com/pensions/2022/02/08/pension-freedoms-were-they-really-a-good-idea/
+% \item what portion of savings are because of retirement savings? \item Also
+% talk about heterogeneity across countries. Some countries want to move towards
+% more annuitisation. This annual review is a good source of info
+% \cite{banks_crawford_ar_2022} \item
+% https://www.moneysavingexpert.com/savings/pension-freedom/
 

### 2023-11-25 R/10__2_summay_stats.R

@@ -211,26 +211,46 @@ writeLines(
     "../Texfiles/tables/cov_balance.tex"
 )
 
-# this seems quite non-sensical
-
-## Now save these variables
-
-
-# use stargazer for now because I know how it works
-# save as tex files in the latex file
 
 
 
+# I want to write my own F-stat function
+# takes multiple models and covariance matrices
+# then calcs F stat for the variable that I care
+
+model_1 = lm(
+    ragender ~  pre_post_ref + rv + rv:pre_post_ref , 
+    data = data_for_regressions
+)
+
+# i.e. constrain that pre_post_ref == 0
+model_2 = lm(
+    ragender ~  rv + rv:pre_post_ref , 
+    data = data_for_regressions
+)
+
+sigma_hat = 1/nrow(data_for_regressions)*(sum(model_1$residuals^2))
+sigma_tilda = 1/nrow(data_for_regressions)*(sum(model_2$residuals^2))
+
+n = nrow(data_for_regressions)
+coefs = length(model_1$coefficients)
+
+F_stat = (sigma_tilda - sigma_hat)/(sigma_hat/(n - coefs))
+
+coeftest(model_1)
+F_stat * 4
+
+library(lmtest)
+
+install.packages("coeftest")
+
+install.packages("lmtest")
+names(model_1)
+
+model_1  %>% 
+lapply(head)
+model_1$qr  %>% names
+
+model_1$qr$rank
 
 
-
-# 0.3 of a year longer
-
-# do this for all the above variables
-# basically want to copy the table from MD
-
-
-
-
-
-# MD has an example of a fuzzy RD that could be good to copy.

### 2024-11-21 Texfiles/5348H.tex
Mostly stlyer changes to R scripts

@@
 \usepackage{amssymb}
-\usepackage[authoryear]{natbib}
+\usepackage[style=authoryear]{biblatex}
+\addbibresource{references.bib} %Import the bibliography file
+
 \usepackage{enumerate}
@@
 \begin{document}
+
 \begin{titlepage}
@@
     \vfill
-        I confirm that this is entirely my own work and has not previously been submitted for assessment, and I have read and understood the University’s and Faculty’s definition of Plagiarism
+    I confirm that this is entirely my own work and has not previously been submitted for assessment, and I have read and understood the University's and Faculty's definition of Plagiarism.
   \end{center}
@@
 guaranteed income stream until the end of life, should be highly valued by
-individuals as a way to insure against late death \citep{yaari_65}. However, in
+individuals as a way to insure against late death (\cite{yaari-65}). However, in
 developed countries, rates of annuitisation are far below the levels that theory
-predicts -- this phenomenon has been termed the ``annuity puzzle". The literature
+predicts --- this phenomenon has been termed the ``annuity puzzle''. The literature
 has suggested several reasons for this, but there is no consensus about which
@@
 tax-advantaged account that is usually supplemented with employer contributions.
-Under the 2010-2015 coalition government in the UK, the law regarding the use of
+Under the 2010--2015 coalition government in the UK, the law regarding the use of
 private DC pensions at retirement changed. Individuals were no longer forced to
@@
 a lump sum withdrawal or income drawdown, involving a steady withdrawal of
-assets from the pension pot -- common advice is to take 4\% a year -- whilst the
+assets from the pension pot---common advice is to take 4\% a year---whilst the
 rest remains invested. Subsequently, the number of annuities sold in the UK
@@
 pessimistic life expectancy. A bequest motive means that individuals want to
-leave an estate for their heirs to inherit -- which they cannot do if they
+leave an estate for their heirs to inherit --- which they cannot do if they
 annuitise their wealth. On the other hand, individuals may believe that they
@@
 age there will be 341 of pensionable age in 2045 compared to 280 in 2020
-\citep{ons_population_predictions_2020}. The increase in absolute and relative
+(\cite{ons-population-predictions-2020}). The increase in absolute and relative
 numbers of elderly people makes retirement policy more consequential. Moreover,
 private DC pensions are becoming increasingly common and are predicted to grow
-as current cohorts age \citep{cribb_karjalainen_ifs_2023}. Therefore, policies
+as current cohorts age (\cite{cribb-karjalainen-ifs-2023}). Therefore, policies
 regarding how private pensions can be accessed will have a larger impact on
@@
 other hand, if retirees over-save and do not spend, perhaps in order to bequest
-wealth, the economy may be dynamically inefficient -- a state in which the
+wealth, the economy may be dynamically inefficient --- a state in which the
 capital stock is too high. In this case consumption, per capita could be
@@
 
-\cite{yaari_65} showed that under standard assumptions with risk averse agents
+\cite{yaari-65} showed that under standard assumptions with risk averse agents
 we would expect individuals to annuitise all of their wealth at retirement, so
 that they are insured against the risk of long life. More recently,
-\cite{davidoff_brown_diamond_aer_2005} show that complete annuitisation is
+\cite{davidoff-brown-diamond-aer-2005} show that complete annuitisation is
 optimal under a broader range of assumptions. Since then, there has been much
 discussion of possible reasons for why people choose not to annuitise.
-\cite{finkelstein_porteba_2002, finkelstein_porteba_2004} find evidence of
-adverse selection into the UK annuity market and \cite{mitchell_et_al_1999} find
+\cite{finkelstein-porteba-2002, finkelstein-porteba-2004} find evidence of
+adverse selection into the UK annuity market and \cite{mitchell-et-al-1999} find
 evidence of the same in the US. This means longer-lived individuals select into
@@
 opposed to the population of annuitants, putting off individuals from buying an
-annuity. However, \cite{mitchell_et_al_1999} also find that theory would still
+annuity. However, \cite{mitchell-et-al-1999} also find that theory would still
 predict annuitisation because the money's worth of annuities is not that far
@@
 
-However, \cite{friedman_warshawsky_qje_1990} show that annuitisation decisions
+However, \cite{friedman-warshawsky-qje-1990} show that annuitisation decisions
 can be fully explained by social security and `actuarially unfair' annuities, at
@@
 they find that individuals would optimally not annuitise much wealth. Similarly,
-to \cite{finkelstein_porteba_2004}, \cite{friedman_warshawsky_chicago_1988} show
+to \cite{finkelstein-porteba-2004}, \cite{friedman-warshawsky-chicago-1988} show
 that there is a significant difference between the life expectancy of annuitants
@@
 
-More recently, \cite{lockwood_red_2012} builds on this by showing that a
+More recently, \cite{lockwood-red-2012} builds on this by showing that a
 realistic bequest motive in lifecycle simulations achieves realistic
@@
 distribution of bequests, such as how much individuals bequeath and the
-financial characteristics of these individuals. \cite{lockwood_aer_2018} also
+financial characteristics of these individuals. \cite{lockwood-aer-2018} also
 finds that a model that treats bequests as luxury goods fits the lack of demand
-for long-term care insurance well. \cite{odea_sturrock_rest_2023} find that
+for long-term care insurance well. \cite{odea-sturrock-rest-2023} find that
 subjective life expectancies can explain a large part of the annuitisation
@@
 medical expenses might cause individuals to prefer holding assets rather than
-an annuity. Moreover, \cite{vidalmelia_lejarragagarcia_munich_2004} suggest
+an annuity. Moreover, \cite{vidalmelia-lejarragagarcia-munich-2004} suggest
 that couples might share the risk of long life by pooling their resources, and
@@
 The main dataset I use is the English Longitudinal Study of Ageing (ELSA)
-\citep{main_elsa_citation}. ELSA picks individuals over the age of 50 to survey
+(\cite{main-elsa-citation}). ELSA picks individuals over the age of 50 to survey
 every two years until death. If individuals leave the survey, ELSA replaces them
@@
 differentiate between individuals who have a DC pension and those who have a DB
-pension. There have been 9 waves of ELSA data collection -- the first was in
+pension. There have been 9 waves of ELSA data collection --- the first was in
 2002/03, and collections happened every two years thereafter.
@@
 adjust them to make death certain at age 110 as is common in the literature
-\citep{odea_sturrock_rest_2023}. I transform these so that I have risk of death
+(\cite{odea-sturrock-rest-2023}). I transform these so that I have risk of death
 conditional on being a given age since this is what I use in the life cycle
@@
 To calculate subjective life expectancies from ELSA data, I follow
-\cite{odea_sturrock_rest_2023}. Individuals were asked “What are the chances
+(\cite{odea-sturrock-rest-2023}). Individuals were asked “What are the chances
 that you will live to be age X or more?” where X changed depending on the age of
@@
 policy reform created a minimum income requirement above which individuals would
-not need to annuitise \citep{finance_act_hmt_2011}, meaning high income
+not need to annuitise (\cite{finance-act-hmt-2011}), meaning high income
 retirees would not need to annuitise their wealth. The income requirement was
@@
 `pensions freedom act', scrapping the minimum income requirement and eliminating
-the compulsory annuities market \citep{pen_freedoms_hmt_2014}. One government
+the compulsory annuities market (\cite{pen-freedoms-hmt-2014}). One government
 minister infamously brought the message home by saying pension pots `can be used
-to buy Lamborghinis' if retirees wanted to \citep{guardian_lambos}. The
+to buy Lamborghinis' if retirees wanted to (\cite{guardian-lambos}). The
 government also announced plans to make switches between DB and DC pension
@@
 The impact of the reforms on annuity demand has been documented by
-\cite{cannon_et_al_nier_2016}. Figure \ref{fig:annovertime} shows
+\cite{cannon-et-al-nier-2016}. Figure \ref{fig:annovertime} shows
 purchases of annuities over time and demonstrates the sharp decrease in purchases
@@
   \label{fig:ann2122}
-
-    \
 \end{figure}
-Figure \ref{fig:ann2122} shows the distribution of methods used to access pension pots
-at retirement in 2021-22. In 2021, 196,736 pots were fully withdrawn at
+
+Figure~\ref{fig:ann2122} shows the distribution of methods used to access pension pots
+at retirement in 2021--22. In 2021, 196,736 pots were fully withdrawn at
 retirement, accounting for over 50\% of pots. Prior to the policy reform this
-was not the case -- most DC pension pots were accessed through
+was not the case --- most DC pension pots were accessed through
 annuitisation.
@@
 until December 2012 to implement the ruling. This change meant that annuity
-products could no longer be priced differently for men and women in the UK --
+products could no longer be priced differently for men and women in the UK ---
 changing the money's worth of an annuity. I run a robustness test in the
@@
 
-Table \ref{tab:sum_stats} shows various summary statistics for each variable of
-interest from ELSA and the ONS.
-
-An individual is defined as treated if they retired between and including 2015
-and 2017, and an individual is in the control group if they retired in 2014 or
-earlier. Since I am interested in consumption in early retirement, an
-individuals interview date needs to be within the first three years of them
-retiring.
-
-The treated retirement group is smaller with just 303 non missing observations
-for gender as opposed to 765 individuals in the control group. The control group
-are more male, retired at a slightly younger age and expected to retire slightly
-earlier. This could be because this period was affected by the increase in the
-state pension age for women from age 60 in 2010 to age 65 in 2018, so early
-cohorts had slightly earlier state pension eligibility ages. But, this change
-happened gradually and ocurred over the whole period, so I do not expect it to
-influence the results. The DifferenceAge row tracks the difference between
-expected and actual retirement age.
+Table~\ref{tab:sum_stats} shows various summary statistics for each variable of interest from ELSA and the ONS.
+
+An individual is defined as treated if they retired between and including 2015 and 2017, and an individual is in the control group if they retired in 2014 or earlier.
+Since I am interested in consumption in early retirement, an individuals interview date needs to be within the first three years of them retiring.
+
+The treated retirement group is smaller with just 303 non missing observations for gender as opposed to 765 individuals in the control group.
+The control group are more male, retired at a slightly younger age and expected to retire slightly earlier.
+This could be because this period was affected by the increase in the
+state pension age for women from age 60 in 2010 to age 65 in 2018, so early cohorts had slightly earlier state pension eligibility ages.
+But, this change happened gradually and ocurred over the whole period, so I do not expect it to influence the results.
+The DifferenceAge row tracks the difference between expected and actual retirement age.
 
@@
 roughly 45\% of individuals across both samples having a DB pension. In general,
-DB pensions are more prevalent than DC pensions in the data -- this reflects the
+DB pensions are more prevalent than DC pensions in the data --- this reflects the
 trend that is observed in the general UK population. Home ownership is roughly
@@
 \end{landscape}
+% \resizebox{\textwidth}{!}{  \input{tables/r_summary_stats.tex}}
 
@@
 \begin{equation*}
-    Y_{i} = \alpha + \beta PostRef_{i}  + \epsilon_{i}
+  Y_{i} = \alpha + \beta_{1} PostRef_{i} + \beta_{2} RetirementYear_{i}  + \beta_{2} RetirementYear_{i}PostRef_{i}  + \epsilon_{i}
 \end{equation*}
@@
 
-Table \ref{tab:cov_balance} shows these results. As expected, retirement year
+Table~\ref{tab:cov_balance} shows these results. As expected, retirement year
 and interview year are greater for the treated group. The difference between
@@
 One threat to validity is manipulation of the running variable, retirement year.
-As observed above in Tables \ref{tab:sum_stats} and \ref{tab:cov_balance}, it
+As observed above in Tables~\ref{tab:sum_stats} and~\ref{tab:cov_balance}, it
 does not appear as though individuals are delaying retirement. However, it is
@@
 surprise the government put forward the reform, with Money Marketing describing
-the change as a ``bombshell" \citep{money_marketing_announcement}.
+the change as a ``bombshell'' (\cite{money-marketing-announcement}).
 
@@
 \begin{equation*}
-    Cons_{it} =  \gamma X_{it} + \beta PostReform_{it} + \epsilon
+  \label{eq:main-rd}
+  Y_{i} = \alpha + \beta_{1} PostRef_{i} + \beta_{2} RetirementYear_{i}  + \beta_{2} RetirementYear_{i} X PostRef_{i}  + \epsilon_{i}
 \end{equation*}
 
-Where $Cons_{it}$ is one of five different consumption variables and $X_{it}$ is
+Where $Y_{i}$ is one of five different consumption variables and $X_{i}$ is
 a set of controls.
 
-Table \ref{tab:DcOnlyRes} shows the results of this specification on the various
+Table~\ref{tab:DcOnlyRes} shows the results of this specification on the various
 consumption indicators. Column 1 has total monthly consumption on the left hand
@@
 further into retirement at the time of interview is associated with higher total
-consumption -- this is surprising since lifecycle models predict consumption
+consumption---this is surprising since lifecycle models predict consumption
 decreases over retirement and this has also been documented empirically
-\citep{hurd_rohwedder_nber_2003}.
+(\cite{hurd-rohwedder-nber-2003}).
 
@@
 to food away from home, to decrease. This wealth effect may be driving the
-results we observe in Table \ref{tab:DcOnlyRes}, since food consumption outside the
-home increases as does expenditure on clothing -- seen in column 5.
+results we observe in Table~\ref{tab:DcOnlyRes}, since food consumption outside the
+home increases as does expenditure on clothing---seen in column 5.
 
@@
 different to those who retired before and perhaps consume more outside the home
-anyway. The financial characteristic data presented in Table \ref{tab:sum_stats}
+anyway. The financial characteristic data presented in Table~\ref{tab:sum_stats}
 would support this hypothesis as it shows all retirees are richer post-reform.
@@
 \begin{equation*}
-    Cons_{it} =  \gamma X_{it} + \beta_{1} PostReform_{it} + \beta_{2} PostReform*DC_{it} + \beta_{3} PostReform*DB_{it}  + \epsilon
+  Cons_{i} =  \gamma X_{i} + \beta_{1} PostReform_{i} + \beta_{2} PostReform*DC_{i} + \beta_{3} PostReform*DB_{i}  + \epsilon
 \end{equation*}
@@
 
-Table \ref{tab:ElsaAllData} shows these results. The interactions at the bottom
+Table~\ref{tab:ElsaAllData} shows these results. The interactions at the bottom
 of the table show us what happened to individuals with a given pension type
@@
 relative to the no pension and DB pension groups. The increase in sample size
-also brings decreases in the standard errors of these variables -- the
+also brings decreases in the standard errors of these variables --- the
 interactions are now all significant a the 5\% level as is the main coefficient
@@
 forced to annuitise a larger amount. So, I interact the policy with DC wealth.
-As in Table \ref{tab:DcOnlyRes}, this regression only uses those individuals who
+As in Table~\ref{tab:DcOnlyRes}, this regression only uses those individuals who
 have a DC pension.
@@
 
-Table \ref{tab:DcOnlyInteract} shows the results. The interaction with DC
+Table~\ref{tab:DcOnlyInteract} shows the results. The interaction with DC
 pension size is positive, meaning that a larger DC pension pot is associated with a
@@
 wave 5 has not been released yet, so for individuals who retire after 2011 I
-have no DC wealth variable. For Table \ref{tab:DcOnlyInteract}, pension wealth
+have no DC wealth variable. For Table~\ref{tab:DcOnlyInteract}, pension wealth
 from the last available year for an individual was given a real return of 3\%
@@
 
-Table \ref{tab:DcOnlyFinWealthInteract} contains these results and shows a
-similar pattern to Table \ref{tab:DcOnlyInteract}. The reform is associated with
+Table~\ref{tab:DcOnlyFinWealthInteract} contains these results and shows a
+similar pattern to Table~\ref{tab:DcOnlyInteract}. The reform is associated with
 an increase in total monthly consumption and more financial wealth increases the
@@
 I carry out two checks for robustness. Firstly, I exclude those individuals who
-retired more than one year away from their expected retirement year -- I define
+retired more than one year away from their expected retirement year --- I define
 an individual's expected retirement year as their expected retirement age in
@@
 individuals either retiring early in order to buy an annuity, since one may be
-cheaper pre-reform because the compulsory market reduces adverse selection
-\cite{finkelstein_porteba_2002}, or retiring later to avoid annuitisation
+cheaper pre-reform because the compulsory market reduces adverse selection~\cite{finkelstein-porteba-2002}, or retiring later to avoid annuitisation
 altogether.
@@
 annuity pricing was made the same for men and women. I drop anyone with a
-retirement year in 2012 or 2011 and repeat the analysis from Table
-\ref{tab:DcOnlyRes}.
-
-Tables \ref{tab:DcOnlyExpOnlyRes} and \ref{tab:DcOnlyNot2012} show these
-results. In Table \ref{tab:DcOnlyExpOnlyRes}, we see that the sign is reversed on
+retirement year in 2012 or 2011 and repeat the analysis from Table~\ref{tab:DcOnlyRes}.
+
+Tables~\ref{tab:DcOnlyExpOnlyRes} and~\ref{tab:DcOnlyNot2012} show these
+results. In Table~\ref{tab:DcOnlyExpOnlyRes}, we see that the sign is reversed on
 PostReform in all consumption categories apart from clothing, however the sample
 size is a third of what it was prior to excluding individuals so there is more
-uncertainty in these estimates. In Table \ref{tab:DcOnlyNot2012}, the sample
+uncertainty in these estimates. In Table~\ref{tab:DcOnlyNot2012}, the sample
 size is again much smaller and the sign has flipped on all consumption
@@
 \begin{equation*}
+  \label{eq:retirement-problem}
   \underset{a_{t+1}, c_{t}}{\max} \{ u(c_{t}) + p_{t}B(a_{t+1}) + \beta(1-p_{t})V(a_{t+1}, y, t+1) \}
@@
 \begin{equation*}
+  \label{eq:budget-constraint}
   c_{t} =a_{t}(1 +r) -  a_{t+1} + y
@@
 rate. I enter $t$ directly into the value function as a state variable to make
-it clear that value is also conditional on age -- through the chance of death
+it clear that value is also conditional on age --- through the chance of death
 and the total number of periods left until terminal age.
@@
 
-As is common in the literature \citep{lockwood_red_2012,de_nardi_et_al_2010}, I
+As is common in the literature (\cite{lockwood-red-2012,de-nardi-et-al-2010}), I
 use a constant relative risk aversion utility function.
 \begin{equation*}
+  \label{eq:utility}
   u(c_{t}) = \frac{c_{t}^{1 - \sigma}}{1 - \sigma}
@@
 In some specifications retirees can leave bequests. I use the bequest function
-from \cite{lockwood_red_2012,lockwood_aer_2018}.
+from~\cite{lockwood-red-2012,lockwood-aer-2018}.
 \begin{equation*}
+  \label{eq:bequest-function}
   B(a_{t+1}) = \bigl( \frac{m}{1 - m} \bigr)^{\sigma}  \frac{(\frac{m}{1 - m}c_{0} + a_{t+1})^{(1 - \sigma)}}{1 - \sigma}
@@
 share of wealth to their heirs. I pick values of $m = 0.95$ and $c_{0} = \pounds
-    20,000$ since these generate low rates of annuitisation. For comparison,
-\cite{lockwood_aer_2018} finds $c_{0}$ is in the range of $\$12,000$ to
+  20,000$ since these generate low rates of annuitisation. For comparison,~\cite{lockwood-aer-2018} finds $c_{0}$ is in the range of $\$12,000$ to
 $\$30,000$ and $m$ is between $0.93$ and $0.96$ in a variety of estimated
@@
 between $0.8$ and $0.9$ for the population as a whole in the UK
-\citep{finkelstein_porteba_2002, finkelstein_porteba_2004, mitchell_et_al_1999}.
+(\cite{finkelstein-porteba-2002, finkelstein-porteba-2004, mitchell-et-al-1999}).
 
@@
 
-I now plot the retirement path of consumption for a median individual in my
-subsample of ELSA. I plot consumption with different proportions of starting wealth
+I now plot the retirement path of consumption for a median individual in my subsample of ELSA.
+I plot consumption with different proportions of starting wealth
 annuitised (either 0\% or 50\%) and the different lifecycle model types.
@@
 
-Figures \ref{fig:AssetPlot} and \ref{fig:ConsumpPlot} show the retirement path
+Figures~\ref{fig:AssetPlot} and~\ref{fig:ConsumpPlot} show the retirement path
 of consumption and assets for a median retiree. This hypothetical female has
@@
 annuity start at exactly the same amount and stay that way until the individual
-reaches her mid seventies -- showing that even quite different lifecycle models
+reaches her mid seventies---showing that even quite different lifecycle models
 can generate similar paths. As expected, the consumption path with the bequest
@@
 matches roughly what is seen in the summary statistics, so, for the untreated
-group -- those who retired pre-reform -- I annuitise half of wealth. I also only
+group---those who retired pre-reform---I annuitise half of wealth. I also only
 annuitise when an individual has financial assets large enough to move up at
@@
 
-Table \ref{tab:simulation_prediction} shows the mean of consumption that each
+Table~\ref{tab:SimulationPrediction} shows the mean of consumption that each
 lifecycle model predicts for the treated and control groups as well as the
@@
 tests on estimated models to see if simulated responses match real world
-responses, for example see \cite{mcgee_2021}. Due to the time constraints of
+responses, for example see~\cite{mcgee-2021}. Due to the time constraints of
 this project, I did not estimate a full structural model but I do check which
@@
   \linespread{1.5}
-
   \input{tables/all_simulated_models.tex }
@@
 
-Table \ref{tab:SubjectiveLifeCycle} shows all these results. Each set of colums shows the
+Table~\ref{tab:SubjectiveLifeCycle} shows all these results. Each set of colums shows the
 results from the empirical models using a different lifecycle model to simulate
@@
 when using the predictions from subjective models, the coefficient on PostReform
-is negative, even though the mean increased for this model as shown in Table
-\ref{tab:simulation_prediction}.
-
-The lifecycle model that is closest to the empirical models shown in Table
-\ref{tab:DcOnlyRes} is the bequest model which gives an increase of 17 in
-monthly consumption. However, this coefficient is just under half of the change
-estimated in Table \ref{tab:DcOnlyRes} and reflects only a 0.6\% rise in
-consumption after the reform.
+is negative, even though the mean increased for this model as shown in Table~\ref{tab:SimulationPrediction}.
+
+The lifecycle model that is closest to the empirical models shown in Table~\ref{tab:DcOnlyRes} is the bequest model which gives an increase of 17 in monthly consumption.
+However, this coefficient is just under half of the change estimated in Table~\ref{tab:DcOnlyRes} and reflects only a 0.6\% rise in consumption after the reform.
 
@@
 with subjective life expectancies has one of -30.9. The bequest and subjective
-coefficients are higher than the coefficient seen in Table \ref{tab:ElsaAllData}
+coefficients are higher than the coefficient seen in Table~\ref{tab:ElsaAllData}
 with the predictions from the standard model matching the empirical model best.
@@
 only needing to know their retirement wealth as we were able to do in
-Figure \ref{fig:AssetPlot}.
+Figure~\ref{fig:AssetPlot}.
 
@@
 
-\bibliographystyle{apa}
-\bibliography{references}
+\printbibliography
+% \bibliography{references}
 \end{document}

### 2024-11-21 R/10__2_summay_stats.R

@@ -5,7 +5,7 @@ library(dplyr)
 library(stringr)
 
 rm(list = ls())
-source("__ELSA_functions.R")
+source("R/__ELSA_functions.R")
 
 
 # Covariate Balance
@@ -17,17 +17,15 @@ source("__ELSA_functions.R")
 # and then focus on the other models on sunday
 
 
-if (!dir.exists("../Texfiles/tables")) {
-    dir.create("../Texfiles/tables")
-}
+if (!dir.exists("Texfiles/tables")) dir.create("/Texfiles/tables")
 
-data_for_regressions <- fread("../../data/ELSA/elsa_to_use/elsa_reg_data.csv") %>%
-    mutate(
-        fin_wealth = fin_wealth / 1000,
-        house_value = house_value / 1000,
-        dc_pot = dc_pot / 1000,
-        public_pension = public_pension / 1000
-    )
+data_for_regressions <- fread("data/ELSA/elsa_to_use/elsa_reg_data.csv") |>
+  _[, ":="(
+    fin_wealth = fin_wealth / 1000,
+    house_value = house_value / 1000,
+    dc_pot = dc_pot / 1000,
+    public_pension = public_pension / 1000
+  )]
 
 search_names(data_for_regressions, "pen")
 search_names(data_for_regressions, "year")
@@ -42,76 +40,76 @@ non_consump_cols <- summary_cols[!grepl("monthly", summary_cols)]
 
 
 ss_stat_out <- data_for_regressions %>%
-    group_by(pre_post_ref) %>%
-    mutate(across(all_of(summary_cols), ~ as.numeric(.x))) %>%
-    summarise(across(
-        all_of(summary_cols),
-        list(
-            mean = \(x) mean(x, na.rm = T),
-            # sd = \(x) sd(x, na.rm = T),
-            min = \(x) min(x, na.rm = T),
-            max = \(x) max(x, na.rm = T),
-            median = \(x) median(x, na.rm = T),
-            non_missing = \(x) sum(!is.na(x))
-        ),
-        .names = "{.col}__{.fn}"
-    )) %>%
-    data.table() %>%
-    melt(id.vars = c("pre_post_ref")) %>%
-    mutate(row_name = str_extract(variable, ".+(?=__)")) %>%
-    # i.e. extract everything before the double underscore
-    mutate(function_name = str_extract(variable, "(?<=__).+")) %>%
-    mutate(function_treatment = paste0(function_name, "_", pre_post_ref)) %>%
[…]

### 2024-11-21 R/10__3_run_regressions.R

@@ -6,20 +6,20 @@ library(data.table)
 source("./__ELSA_functions.R")
 
 data_for_regressions <- fread(
-    "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
+  "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
 ) %>%
-    mutate(
-        fin_wealth = fin_wealth / 1000,
-        house_value = house_value / 1000,
-        dc_pot = dc_pot / 1000,
-        public_pension = public_pension / 1000
-    ) %>%
-    mutate(
-        pre_post_ref_bin = fcase(
-            pre_post_ref == "treat", 1,
-            pre_post_ref == "control", 0
-        )
+  mutate(
+    fin_wealth = fin_wealth / 1000,
+    house_value = house_value / 1000,
+    dc_pot = dc_pot / 1000,
+    public_pension = public_pension / 1000
+  ) %>%
+  mutate(
+    pre_post_ref_bin = fcase(
+      pre_post_ref == "treat", 1,
+      pre_post_ref == "control", 0
     )
+  )
 
 clean_names <- append(clean_names, c("PostReform" = "pre_post_ref_bin"), 0)
 
@@ -53,26 +53,26 @@ names(data_for_regressions)
 
 head(data_for_regressions$idauniq)
 data_for_regressions[, .(count = .N),
-    by = .(idauniq)
+  by = .(idauniq)
 ]
 
 data_for_regressions[, .(count = .N),
-    by = .(retirement_year)
+  by = .(retirement_year)
 ] %>%
-    ggplot(aes(
-        x = retirement_year,
-        y = count
-    )) +
-    geom_col()
+  ggplot(aes(
+    x = retirement_year,
+    y = count
+  )) +
+  geom_col()
 
 data_for_regressions[, .(count = .N),
-    by = .(retirement_year)
+  by = .(retirement_year)
 ] %>%
-    ggplot(aes(
-        x = retirement_year,
-        y = count
-    )) +
-    geom_col()
+  ggplot(aes(
+    x = retirement_year,
+    y = count
+  )) +
+  geom_col()
 data_for_regressions[, rv := retirement_year - 2014]
 
 
@@ -80,7 +80,7 @@ data_for_regressions[, rv := retirement_year - 2014]
 
 # Ok this shold be the spec
 lm(total_monthly_consumption ~ pre_post_ref_bin + rv + rv:pre_post_ref_bin,
-    data = data_for_regressions[retirement_year >= 2012 & retirement_year <= 2016]
[…]

### 2024-11-21 R/11__ELSA_prep_julia.R

@@ -11,22 +11,22 @@ library(patchwork)
 source("__ELSA_functions.R")
 
 reg_data <- fread(
-    "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
+  "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
 )
 
 # reg_data$retirement_year
 # reg_data$retired_age
 
 dt_for_julia <- reg_data[, .(id_wave,
-    gender = ragender, age = age_at_interview,
-    retired_age,
-    retirement_year,
-    public_pension,
-    ever_dc_pen,
-    ever_db_pen,
-    year = date_year,
-    fin_wealth,
-    dc_pot
+  gender = ragender, age = age_at_interview,
+  retired_age,
+  retirement_year,
+  public_pension,
+  ever_dc_pen,
+  ever_db_pen,
+  year = date_year,
+  fin_wealth,
+  dc_pot
 )]
 
 dt_for_julia[, sum(dc_pot > fin_wealth, na.rm = T)]
@@ -38,33 +38,33 @@ dt_for_julia[fin_wealth > dc_pot]
 nrow(dt_for_julia)
 
 dt_for_julia[, gender := fcase(
-    gender == 1, "male",
-    gender == 0, "female"
+  gender == 1, "male",
+  gender == 0, "female"
 )]
 
 
 # , ever_dc_pen, ever_db_pen
 gender_age_year_unique <- dt_for_julia[, .(gender, age, year)] %>%
-    unique()
+  unique()
 
 # check no NAs in any columns
 stopifnot(!any(sapply(gender_age_year_unique, \(x) any(is.na(x)))))
 
 fwrite(
-    gender_age_year_unique,
-    "../../data/ELSA/elsa_to_use/gender_age_year_unique.csv"
+  gender_age_year_unique,
+  "../../data/ELSA/elsa_to_use/gender_age_year_unique.csv"
 )
 # We want to run the life cycle model for each of these gender-pension-age pairs.
 
 
 fwrite(
-    dt_for_julia,
-    "../../data/ELSA/elsa_to_use/for_julia.csv"
+  dt_for_julia,
+  "../../data/ELSA/elsa_to_use/for_julia.csv"
 )
 
 fwrite(
-    select(dt_for_julia, id_wave, age, gender, year),
-    "../../data/ELSA/elsa_to_use/for_julia_if_age_gender_year.csv"
[…]

### 2024-11-21 R/12__cmi_annuity.R

@@ -19,18 +19,18 @@ library(ggplot2)
 
 # annuity life tables in 2008
 ann_lt_08 <- read_excel(
-    "../../data/CMI/raw/92190-cmiwp81-final08series-annuities-rates.xlsx",
-    sheet = "08 annuities qx",
-    skip = 4
+  "../../data/CMI/raw/92190-cmiwp81-final08series-annuities-rates.xlsx",
+  sheet = "08 annuities qx",
+  skip = 4
 ) %>%
-    setnames(str_to_lower) %>%
-    select(age = "age x", "lml08", "lfl08") %>% # get annuitant life tables
-    filter(age >= 60) %>% # only have data for those people who are over 60
-    mutate(year = 2008) %>%
-    rename(
-        male_prob = lml08,
-        female_prob = lfl08
-    )
+  setnames(str_to_lower) %>%
+  select(age = "age x", "lml08", "lfl08") %>% # get annuitant life tables
+  filter(age >= 60) %>% # only have data for those people who are over 60
+  mutate(year = 2008) %>%
+  rename(
+    male_prob = lml08,
+    female_prob = lfl08
+  )
 
 
 
@@ -40,18 +40,18 @@ ann_lt_08 <- read_excel(
 # PFL16	        All products	Female	Lives
 
 ann_lt_16 <- read_excel(
-    "../../data/CMI/raw/Final 16 Series pension annuitant mortality rates v01 2020-07-10.xlsx",
-    sheet = "q",
-    skip = 6
+  "../../data/CMI/raw/Final 16 Series pension annuitant mortality rates v01 2020-07-10.xlsx",
+  sheet = "q",
+  skip = 6
 ) %>%
-    setnames(str_to_lower) %>%
-    select(age, pml16, pfl16) %>%
-    filter(age >= 60) %>%
-    mutate(year = 2016) %>%
-    rename(
-        male_prob = pml16,
-        female_prob = pfl16
-    )
+  setnames(str_to_lower) %>%
+  select(age, pml16, pfl16) %>%
+  filter(age >= 60) %>%
+  mutate(year = 2016) %>%
+  rename(
+    male_prob = pml16,
+    female_prob = pfl16
+  )
 
 # these are just probs of reaching next year
 #   conditional on being a certain age.
@@ -60,25 +60,25 @@ ann_lt_16 <- read_excel(
 
 
 both_years <- rbindlist(list(ann_lt_08, ann_lt_16)) %>%
-    tidyr::pivot_longer(
[…]

### 2024-11-21 R/13__regression_with_lifecycle.R

@@ -11,46 +11,46 @@ clean_names <- append(clean_names, c("PostReform" = "pre_post_ref_bin"), 0)
 # Reading data for regressions
 
 bequest_consump_dt <- fread("../../data/ELSA/lifecycle_outputs/bequest_consumption_simulation.csv") %>%
-    rename(
-        consump_beq_ann = annuity_consump,
-        consump_beq_no_ann = no_annuity_consump
-    )
+  rename(
+    consump_beq_ann = annuity_consump,
+    consump_beq_no_ann = no_annuity_consump
+  )
 
 standard_consump_dt <-
-    fread("../../data/ELSA/lifecycle_outputs/standard_model_consumption_simulation.csv") %>%
-    rename(
-        consump_stan_ann = annuity_consump,
-        consump_stan_no_ann = no_annuity_consump
-    ) %>%
-    select(id_wave, consump_stan_ann, consump_stan_no_ann)
+  fread("../../data/ELSA/lifecycle_outputs/standard_model_consumption_simulation.csv") %>%
+  rename(
+    consump_stan_ann = annuity_consump,
+    consump_stan_no_ann = no_annuity_consump
+  ) %>%
+  select(id_wave, consump_stan_ann, consump_stan_no_ann)
 
 subjective_consump_dt <-
-    fread("../../data/ELSA/lifecycle_outputs/subjective_model_consumption_simulation.csv") %>%
-    rename(
-        consump_sub_ann = annuity_consump,
-        consump_sub_no_ann = no_annuity_consump
-    ) %>%
-    select(id_wave, consump_sub_ann, consump_sub_no_ann)
+  fread("../../data/ELSA/lifecycle_outputs/subjective_model_consumption_simulation.csv") %>%
+  rename(
+    consump_sub_ann = annuity_consump,
+    consump_sub_no_ann = no_annuity_consump
+  ) %>%
+  select(id_wave, consump_sub_ann, consump_sub_no_ann)
 
 
 # standard_consump_dt[, .(annuity_cost, annuity_payment)]
 # bequest_consump_dt[, .(annuity_cost, annuity_payment)]
 
 reg_dt <- fread(
-    "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
+  "../../data/ELSA/elsa_to_use/elsa_reg_data.csv"
 ) %>%
-    mutate(
-        fin_wealth = fin_wealth / 1000,
-        house_value = house_value / 1000,
[…]

### 2024-11-21 R/14__create_lifecycle_plot.R

@@ -9,19 +9,19 @@ source("./__ELSA_functions.R")
 # want to read in all of those strings
 
 vec_dt <- fread(
-    "../../data/ELSA/lifecycle_outputs/model_ass_cons_paths/100006-6.csv",
-    header = F
+  "../../data/ELSA/lifecycle_outputs/model_ass_cons_paths/100006-6.csv",
+  header = F
 )
 
 min_length <- vec_dt[1, V2] %>%
-    strsplit(., ",", fixed = T) %>%
-    unlist() %>%
-    length()
+  strsplit(., ",", fixed = T) %>%
+  unlist() %>%
+  length()
 
 lapply(vec_dt, strsplit, ", ") %>%
-    unlist(recursive = F) %>%
-    lapply(function(x) x[seq(1, min_length)]) %>%
-    lapply(function(x) stringi::stri_replace(x, "", "[", regex = T, mode = "first"))
+  unlist(recursive = F) %>%
+  lapply(function(x) x[seq(1, min_length)]) %>%
+  lapply(function(x) stringi::stri_replace(x, "", "[", regex = T, mode = "first"))
 
 
 
@@ -29,10 +29,10 @@ lapply(vec_dt, strsplit, ", ") %>%
 # until they are same length as other ones
 
 lapply(vec_dt, strsplit, ",") %>%
-    unlist(recursive = F) %>%
-    lapply(length) %>%
-    unlist() %>%
-    min()
+  unlist(recursive = F) %>%
+  lapply(length) %>%
+  unlist() %>%
+  min()
 # take last of this string
 
 

### 2024-11-21 R/1__ELSA_unzip_data.R

@@ -2,9 +2,9 @@
 # All the other scripts should work after it.
 
 if (!dir.exists("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/")) {
-    readLines("Do you want to unzip the ELSA files")
-    unzip(
-        zipfile = "../../data/ELSA/5050tab_BDDEA8F799D95FB6A9F6FB5ADBE43D55F7B50718D8615C88A14F94298AADF043_V1.zip",
-        exdir = "../../data/ELSA/elsa_unziped"
-    )
+  readLines("Do you want to unzip the ELSA files")
+  unzip(
+    zipfile = "../../data/ELSA/5050tab_BDDEA8F799D95FB6A9F6FB5ADBE43D55F7B50718D8615C88A14F94298AADF043_V1.zip",
+    exdir = "../../data/ELSA/elsa_unziped"
+  )
 }

### 2024-11-21 R/3__ELSA_financial.R

@@ -8,12 +8,12 @@ source("__ELSA_functions.R")
 # Compile finance files
 # What is in the finance files?
 all_finance <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-    pattern = "wave",
-    full.names = TRUE
+  pattern = "wave",
+  full.names = TRUE
 ) %>%
-    grep("financial_", ., value = TRUE) %>%
-    lapply(ReadAndSetWave) %>%
-    rbindlist(fill = TRUE, use.names = TRUE)
+  grep("financial_", ., value = TRUE) %>%
+  lapply(ReadAndSetWave) %>%
+  rbindlist(fill = TRUE, use.names = TRUE)
 
 # state pension mean to check weekly or monthly
 all_finance[spen_r_i > 0, mean(spen_r_i, na.rm = T)] # cool yes must be weekly
@@ -48,12 +48,12 @@ all_finance[, mean(ppinc_bu_s > 0, na.rm = T)]
 # doesn't seem to have income variables etc though.
 
 all_ifs <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-    pattern = "wave",
-    full.names = TRUE
+  pattern = "wave",
+  full.names = TRUE
 ) %>%
-    grep("ifs", ., value = TRUE) %>%
-    lapply(ReadAndSetWave) %>%
-    rbindlist(fill = TRUE, use.names = TRUE)
+  grep("ifs", ., value = TRUE) %>%
+  lapply(ReadAndSetWave) %>%
+  rbindlist(fill = TRUE, use.names = TRUE)
 
 search_names(all_ifs, "OUTINDW")
 
@@ -63,7 +63,7 @@ search_names(all_ifs, "OUTINDW")
 all_ifs[, intdate := paste0(intdaty, "-", intdatm)]
 all_ifs[, intdate := lubridate::ym(intdate)]
 hist(all_ifs$intdate,
-    breaks = 50.0
+  breaks = 50.0
 )
 
 
@@ -76,12 +76,12 @@ head(all_ifs$wave)
 # These seem to be more about where the pensions were from in terms of which employer etc
 
 all_pension <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-    pattern = "wave",
-    full.names = TRUE
+  pattern = "wave",
+  full.names = TRUE
 ) %>%
[…]

### 2024-11-21 R/5__ELSA_annuity_play.R

@@ -11,41 +11,41 @@ elsa <- fread("../../data/ELSA/elsa_to_use/elsa_ifs_finance.csv")
 
 # Have a look at attrition in ELSA
 elsa[, .(idauniq, wave)][, .(num_waves = length(unique(wave))), by = .(idauniq)] %>%
-    ggplot(aes(x = num_waves)) +
-    geom_histogram()
+  ggplot(aes(x = num_waves)) +
+  geom_histogram()
 
 # plot percent with annuity income and avg annuity inc
 # by wave and age of respondent.
 annuity_age_wave_means <- elsa[, .(
-    mean_annuity_inc =
-        mean(anin_r_i[anin_r_i > 0], na.rm = T),
-    per_annuity =
-        mean(anin_r_i > 0, na.rm = T)
+  mean_annuity_inc =
+    mean(anin_r_i[anin_r_i > 0], na.rm = T),
+  per_annuity =
+    mean(anin_r_i > 0, na.rm = T)
 ), by = .(age, file_wave)]
 head(annuity_age_wave_means$file_wave)
 
 # Plot percentage who have any annuity income against age shaded by wave.
 ggplot(
-    annuity_age_wave_means[
-        age > 60
-    ][file_wave %in% c("wave_1", "wave_8")],
-    aes(x = age, y = per_annuity, colour = file_wave)
+  annuity_age_wave_means[
+    age > 60
+  ][file_wave %in% c("wave_1", "wave_8")],
+  aes(x = age, y = per_annuity, colour = file_wave)
 ) +
-    geom_point() +
-    geom_line()
+  geom_point() +
+  geom_line()
 # ok nice it looks like fewer people do have annuities
 # but still a small portion of the sample have any annuity income.
 # can we do a cross tab
 
 
 ggplot(
-    annuity_age_wave_means[
-        age > 60
-    ][file_wave %in% c("wave_1", "wave_8")],
-    aes(x = age, y = mean_annuity_inc, colour = file_wave)
+  annuity_age_wave_means[
+    age > 60
+  ][file_wave %in% c("wave_1", "wave_8")],
+  aes(x = age, y = mean_annuity_inc, colour = file_wave)
 ) +
-    geom_point() +
-    geom_line()
+  geom_point() +
+  geom_line()
 
 
 
@@ -64,11 +64,11 @@ length(names(elsa))
 
 
[…]

### 2024-11-21 R/7__ELSA_harmonised.R

@@ -15,10 +15,10 @@ source("__ELSA_functions.R")
 # This is harmonised with HRS.
 
 names_harm <- read_dta(
-    "../../data/ELSA/elsa_unziped/UKDA-5050-tab/code/stata_output_files/H_ELSA_g2_stata12.dta",
-    n_max = 1
+  "../../data/ELSA/elsa_unziped/UKDA-5050-tab/code/stata_output_files/H_ELSA_g2_stata12.dta",
+  n_max = 1
 ) %>%
-    names()
+  names()
 writeLines(names_harm, "name_check.txt")
 ###
 
@@ -33,7 +33,7 @@ writeLines(names_harm, "name_check.txt")
 # )
 
 harm <- fread(
-    "../../data/ELSA/elsa_unziped/UKDA-5050-tab/code/stata_output_files/H_ELSA_g2_stata12.csv"
+  "../../data/ELSA/elsa_unziped/UKDA-5050-tab/code/stata_output_files/H_ELSA_g2_stata12.csv"
 )
 
 
@@ -61,37 +61,37 @@ harm[, sum(!is.na(radyear))]
 
 # Get names of variables that we want to transition from wide to long
 names_patterns_list <- list(
-    "retired" = "r\\d{1}retemp",
-    "pension_type_job1" = "r\\d{1}ptyp1_e",
-    "pension_type_job2" = "r\\d{1}ptyp2_e",
-    "pension_type_job3" = "r\\d{1}ptyp3_e",
-    "retired_age" = "r\\d{1}retage",
-    "expected_retired_age" = "r\\d{1}wretage",
-    "age_at_interview" = "r\\d{1}agey",
-    "in_wave" = "inw\\d{1}$",
-    "date_month" = "r\\d{1}iwindm",
-    "date_year" = "r\\d{1}iwindy",
-    "total_monthly_consumption" = "hh\\d{1}ctot1m",
-    "monthly_food_in" = "hh\\d{1}cfoodi",
-    "monthly_food_out" = "hh\\d{1}cfoodo1m",
-    "monthly_food_total" = "hh\\d{1}cfood1m",
-    "monthly_clothing" = "hh\\d{1}cclo1m",
-    "monthly_leisure" = "hh\\d{1}clei1m",
-    "monthly_rent" = "hh\\d{1}crent",
-    "monthly_utility" = "hh\\d{1}cutil",
[…]

### 2024-11-21 R/8__1_ELSA_subjective_life.R

@@ -18,27 +18,27 @@ source("__ELSA_functions.R")
 
 
 check_files_for_variable_name <- function(var_name) {
-    # returns the files which contain the variable name we are looking for
-    # Just helpful for seeing where variables originally came from
-    files_with_var_name <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-        pattern = "",
-        full.names = TRUE
-    ) %>%
-        lapply(\(filename) names(fread(filename, nrows = 1))) %>%
-        lapply(function(col_names) {
-            grep(var_name,
-                col_names,
-                value = T,
-                ignore.case = T
-            )
-        }) %>%
-        sapply(length) %>%
-        sapply(\(x) x != 0)
+  # returns the files which contain the variable name we are looking for
+  # Just helpful for seeing where variables originally came from
+  files_with_var_name <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "",
+    full.names = TRUE
+  ) %>%
+    lapply(\(filename) names(fread(filename, nrows = 1))) %>%
+    lapply(function(col_names) {
+      grep(var_name,
+        col_names,
+        value = T,
+        ignore.case = T
+      )
+    }) %>%
+    sapply(length) %>%
+    sapply(\(x) x != 0)
 
-    dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-        pattern = "",
-        full.names = TRUE
-    )[files_with_var_name]
+  dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "",
+    full.names = TRUE
+  )[files_with_var_name]
 }
 
 # find out which file the data is in
@@ -50,19 +50,19 @@ check_files_for_variable_name("exlo")
 
 # These are file names
 expectation_cols <- c(
-    "idauniq",
-    "ExLo80",
-    "ExLo90"
+  "idauniq",
+  "ExLo80",
+  "ExLo90"
 )
 
 ## get expectation data
 files_with_exp <- check_files_for_variable_name("exlo")
 
 data_with_exp <- lapply(files_with_exp, ReadAndSetWave) %>%
-    lapply(\(data) setnames(data, str_to_lower)) %>%
-    rbindlist(use.names = T, fill = T) %>%
[…]

### 2024-11-21 R/8__2_ELSA_sub_lifeexps.R

@@ -11,20 +11,20 @@ sub_dt[, number_alive := base_pop * prob]
 setorder(sub_dt, -expected_age)
 
 sub_dt[, life_years_left := cumsum(number_alive),
-    by = .(id_wave)
+  by = .(id_wave)
 ]
 
 sub_dt[, life_exp := life_years_left / number_alive]
 
 sub_life_exp <- sub_dt %>%
-    filter(age == expected_age - 1) %>% # just want life expectancy at interview date
-    mutate(
-        age = expected_age - 1,
-        subjective_life_exp = life_exp
-    ) %>%
-    select(id_wave, age, subjective_life_exp)
+  filter(age == expected_age - 1) %>% # just want life expectancy at interview date
+  mutate(
+    age = expected_age - 1,
+    subjective_life_exp = life_exp
+  ) %>%
+  select(id_wave, age, subjective_life_exp)
 
 fwrite(
-    sub_life_exp,
-    "../../data/ELSA/subjective_tables/subjective_life_expectancies.csv"
+  sub_life_exp,
+  "../../data/ELSA/subjective_tables/subjective_life_expectancies.csv"
 )

### 2024-11-21 R/__ELSA_functions.R

@@ -1,93 +1,93 @@
 ReadAndSetWave <- function(
     filename,
     select_cols = "all") {
-    # Read a tab file and set a wave number.
+  # Read a tab file and set a wave number.
 
-    # Extract wave number from filename
-    WAVE_num <- str_extract(filename, "wave_\\d{1}")
+  # Extract wave number from filename
+  WAVE_num <- str_extract(filename, "wave_\\d{1}")
 
-    # get data, if we know the cols we want then we just select those
+  # get data, if we know the cols we want then we just select those
 
-    if (select_cols[1] == "all") { # bit sloppy
-        data <- fread(filename)
-    } else {
-        data <- fread(filename,
-            select = select_cols
-        )
-    }
-    # Set wave in data
-    data[, file_wave := WAVE_num]
+  if (select_cols[1] == "all") { # bit sloppy
+    data <- fread(filename)
+  } else {
+    data <- fread(filename,
+      select = select_cols
+    )
+  }
+  # Set wave in data
+  data[, file_wave := WAVE_num]
 
-    return(data)
+  return(data)
 }
 
 ELSA_dir <- function(string = "*") {
-    elsa_dir <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
-        pattern = "wave",
-        full.names = TRUE
-    ) %>%
-        grep(string, ., value = TRUE)
+  elsa_dir <- dir("../../data/ELSA/elsa_unziped/UKDA-5050-tab/tab/",
+    pattern = "wave",
+    full.names = TRUE
+  ) %>%
+    grep(string, ., value = TRUE)
 
-    return(elsa_dir)
+  return(elsa_dir)
 }
 
 ReadRTF <- function(filename) {
-    Lines <- readLines(filename)
-    pat <- "^\\\\f0.*\\\\cf0 "
-    g <- grep(pat, Lines, value = TRUE)
-    noq <- gsub("\\\\'", "'", g)
+  Lines <- readLines(filename)
+  pat <- "^\\\\f0.*\\\\cf0 "
+  g <- grep(pat, Lines, value = TRUE)
+  noq <- gsub("\\\\'", "'", g)
 
-    clean <- sub("\\\\.*", "", sub(pat, "", noq))
-    return(clean)
+  clean <- sub("\\\\.*", "", sub(pat, "", noq))
+  return(clean)
 }
 
 search_names <- function(data, string, ...) {
[…]

### 2024-11-21 R/__general_setup.R

@@ -1,7 +1,6 @@
-# Other set up bits 
+# Other set up bits
 
-data.table::update_dev_pkg() # I program on datatable which is only available in the devpkg. 
+data.table::update_dev_pkg() # I program on datatable which is only available in the devpkg.
 
 packages <- utils::installed.packages() # Write out current packages
 utils::write.csv(packages, "current_packages.csv")
-

### 2024-11-23 Texfiles/5348H.tex
Update regression equations

@@
 An individual is defined as treated if they retired between and including 2015 and 2017, and an individual is in the control group if they retired in 2014 or earlier.
-Since I am interested in consumption in early retirement, an individuals interview date needs to be within the first three years of them retiring.
+I use retirement year relative to 2014 as the running variable, and restrict the bandwidth to 3 years.
 
@@
 
-I use consumption data of individuals up to 3 years into their retirement so
-that the sample size is larger. So, if someone retired in 2015 and had
-consumption data in 2015 and 2017, I include both values.
-
 Because of the differences in financial characteristics between the groups, I

### 2024-12-01 Texfiles/5348H.tex
Latest edits to abstract and introduction

@@
 
+\usepackage{setspace}
 \usepackage{enumerate}
@@
 \begin{abstract}
-  I identify the consumption responses of retirees to the UK's `pensions
-  freedom act', a landmark law change that meant defined contribution pension
-  holders were no longer forced to annuitise their funds and could instead
-  access their pensions in a variety of ways. I use a discontinuity in year of
-  retirement and a difference-in-difference approach, harnessing those who
-  were unaffected by the policy reform as a natural control group. I then
-  solve three sets of lifecycle models, tracking assets from retirement until an
-  agent reaches age 110. I simulate their consumption with and without
-  annuities, mimicking the policy reform, and match their consumption
-  responses to what was seen in the empirical models. I find that a lifecycle
-  model with a bequest motive best fits the identified consumption response —
-  a result that could inform policymakers in designing optimal pension
-  policies for an ageing population.
+  \begin{singlespace}
+    Countries around the world are reforming the way individuals contribute to and withdraw from retirement savings accounts.
+    I evaluate the effect that the United Kingdom's `Pensions Freedom Act'---a landmark law change that meant defined contribution pension
+    holders were no longer forced to annuitise their funds---had on consumption in retirement.
+    I use a regression discontinuity specification using the year an individual turns 65 as the running variable.
+    I complement this with a difference-in-difference setup, harnessing those who were unaffected by the policy reform as a natural control group.
+    I find that the rates of annuitisation drop to almost zero after the reform---this is consistent with the literature on the annuity puzzle.
+    I find some evidence that individuals increase consumption in early retirement as a result of the policy change---in particular spending on food outside the house increases.
+    Two features of lifecycle models can explain the annuity puzzle.
+    A lifecycle model with a bequest motive would predict a rise in consumption when individuals no longer annuitise their pensions, whereas a lifecycle model with pessimistic life expectancy would predict a fall in consumption.
+    I show this by solving the consumption paths in retirement with and without annuities, mimicking the policy reform.
+    I then apply the empirical models to this simulated data and measure the consumption change in early retirement that results from annuitisation.
+    I find that a lifecycle model with a bequest motive best fits the identified consumption response---a result that could inform policymakers in designing optimal pension policies for an ageing population.
+  \end{singlespace}
 \end{abstract}
-\newpage
-\tableofcontents
 \newpage
@@
 \section{Introduction}
-Classical economic theory suggests that retirement annuities, which provide a
-guaranteed income stream until the end of life, should be highly valued by
-individuals as a way to insure against late death (\cite{yaari-65}). However, in
-developed countries, rates of annuitisation are far below the levels that theory
-predicts --- this phenomenon has been termed the ``annuity puzzle''. The literature
-has suggested several reasons for this, but there is no consensus about which
-mechanism is dominant. I exploit early retirees' consumption responses to a
-reform to annuity policy in the UK to add to this literature.
 
-In the UK, adults have three methods of funding retirement: the state pension;
-defined benefit (DB) employee pension schemes, which provide an income in
-retirement based on years of service and wages; and defined
-contribution (DC) pensions, under which individuals save and invest in a
-tax-advantaged account that is usually supplemented with employer contributions.
-Under the 2010--2015 coalition government in the UK, the law regarding the use of
-private DC pensions at retirement changed. Individuals were no longer forced to
-annuitise their pension pots and could access them in a variety of ways, such as
-a lump sum withdrawal or income drawdown, involving a steady withdrawal of
-assets from the pension pot---common advice is to take 4\% a year---whilst the
-rest remains invested. Subsequently, the number of annuities sold in the UK
-dropped precipitously.
+The importance of retirement policy to the UK, and across the developed world, is growing.
+The number of individuals of pensionable age is expected to increase from 11.9 million in 2020 to 15.2 million in 2045 according to the latest projections from the Office for National Statistics (ONS), and for every 1,000 people of working age there will be 341 of pensionable age in 2045 compared to 280 in 2020 (\cite{ons-population-predictions-2020}).
+The increase in absolute and relative numbers of elderly people makes retirement policy more consequential.
+Moreover, private DC pensions are becoming increasingly common and are predicted to grow
+as current cohorts age (\cite{cribb-karjalainen-ifs-2023}).
+Therefore, policies regarding how private pensions can be accessed will have a larger impact on the overall welfare of retirees.
 
-In this paper, I first use the policy reform as a discontinuity to measure the
-impact of forced annuitisation on the consumption levels of individuals in the
-first few years of retirement. Given that the reform was implemented suddenly
-and without advanced notice before the Spring 2014 budget, I claim that
-individuals who retired in 2015, 2016 and 2017 are otherwise similar to those
-who retired in 2011, 2012 and 2013 except for the fact that the later retirees
-do not need to annuitise their DC pension pots. Therefore, I can use a
-regression discontinuity to identify the effect that forced annuitisation has on
-the consumption of retirees.
+Moreover, understanding how retirees spend their money over retirement is an important micro and macroeconomic policy question of its own.
+If retirees spend too much in early retirement, the state may need to provide for them towards the end of their lives.
+This has implications for government budgets, especially since population aging means more individuals may require expensive end of life care.
+On the other hand, if retirees over-save and do not spend, perhaps in order to bequest wealth or because of expectations around life expectancy and medical expenditures, the economy may be dynamically inefficient---a state in which the capital stock is too high.
+In this case consumption, per capita could be increased if savings were decreased.
 
-I then test two competing hypotheses for the annuity puzzle: bequests and
-pessimistic life expectancy. A bequest motive means that individuals want to
-leave an estate for their heirs to inherit --- which they cannot do if they
-annuitise their wealth. On the other hand, individuals may believe that they
-will not live as long as annuity providers think they will, and therefore
-annuities that are on the market do not appear to be good value. Depending on
-the reason for the lack of annuitisation in the UK, the consumption response of
-retirees to the pension reform will differ. If individuals do not annuitise
-because of pessimistic life expectancy, their consumption should increase after
-the reform. If, on the other hand, individuals do not annuitise because of a
-bequest motive, consumption should not change as much as a result of the reform.
+In the UK, adults have three methods of funding retirement: the state pension; defined benefit (DB) employee pension schemes, which provide an income in retirement based on some function of tenure and wages; and defined contribution (DC) pensions, under which individuals save and invest in a tax-advantaged account that is usually supplemented with employer contributions.
+Under the 2010--2015 coalition government in the UK, the law regarding the use of private DC pensions at retirement changed.
+Individuals were no longer forced to buy annuities, which provide a guaranteed income stream until the end of life, with their pension pots and could access them in a variety of ways, such as a lump sum withdrawal or income drawdown, involving a steady withdrawal of assets from the pension pot---common advice is to take 4\% a year---whilst the rest remains invested.
+Subsequently, the number of annuities sold in the UK dropped precipitously.
 
-I will solve lifecycle models for both of these cases and simulate consumption
-decisions with and without annuitisation. I will then apply the empirical models
-to this simulated data and measure the consumption change in early retirement
-that results from annuitisation. The sign and magnitude of these coefficients
-relative to the coefficients from the real data will be indicative of the
-mechanism causing the annuitisation puzzle.
+In this paper, I first use the policy reform as a discontinuity to measure the impact of forced annuitisation on the consumption levels of individuals in the first few years of retirement.
+And then I simulate consumption responses to annuities using three different lifecycle models, and compare the predicted effect of forced annuitisation on consumption to the effect observed in the data.
 
-The importance of retirement policy to the UK is growing. The
-number of individuals of pensionable age is expected to increase from 11.9
-million in 2020 to 15.2 million in 2045 according to the latest projections from
-the Office for National Statistics (ONS), and for every 1,000 people of working
-age there will be 341 of pensionable age in 2045 compared to 280 in 2020
-(\cite{ons-population-predictions-2020}). The increase in absolute and relative
-numbers of elderly people makes retirement policy more consequential. Moreover,
-private DC pensions are becoming increasingly common and are predicted to grow
-as current cohorts age (\cite{cribb-karjalainen-ifs-2023}). Therefore, policies
-regarding how private pensions can be accessed will have a larger impact on
-overall welfare for retirees.
+Given that the reform was implemented suddenly and without advanced notice before the Spring 2014 budget, my primary regression discontinuity specification uses the year an individual turns 65, the state pension age in the UK in 2014, as the running variable---if individuals turn 65 after 2014 they are considered treated.
+This rests on the assumption that individuals who retired in 2015, 2016 and 2017 are otherwise similar to those who retired in 2011, 2012 and 2013 except for the fact that the later retirees do not need to annuitise their DC pension pots.
+I also run a difference in difference estimator, comparing the change in consumption at retirement for individuals who retired before the reform to those who retired after the reform.
+I find some evidence that individuals increase consumption in early retirement as a result of the policy change, in particular I find a slight increase in consumption of food away from the home.
 
-Moreover, understanding how retirees spend their money over retirement is an
-important policy question of its own. If retirees spend too much in early
-retirement, the state may need to provide for them towards the end of their
-lives. This has implications for government budgets, especially since population
-aging means more individuals may require expensive end of life care. On the
-other hand, if retirees over-save and do not spend, perhaps in order to bequest
-wealth, the economy may be dynamically inefficient --- a state in which the
-capital stock is too high. In this case consumption, per capita could be
-increased if savings were decreased.
+Classical economic theory suggests that these retirement annuities should be highly valued by individuals as a way to insure against late death (\cite{yaari-65}).
+However, in developed countries, rates of annuitisation are far below the levels that theory predicts---this phenomenon has been termed the ``annuity puzzle''.
+The literature has suggested several reasons for this, but there is no consensus about which mechanism is dominant.
+I exploit early retirees' consumption responses to the change in policy in the UK to add to this literature.
+
+I test two competing hypotheses for the annuity puzzle: bequests and pessimistic life expectancy.
+A bequest motive means that individuals want to leave an estate for their heirs to inherit---which they cannot do if they annuitise their wealth.
+On the other hand, individuals may believe that they will not live as long as annuity providers think they will, and therefore annuities that are on the market do not appear to be good value.
+Depending on the reason for the lack of annuitisation in the UK, the consumption response of retirees to the pension reform will differ.
+If individuals do not annuitise because of pessimistic life expectancy, their consumption should increase after the reform.
+If, on the other hand, individuals do not annuitise because of a bequest motive, consumption should not change as much as a result of the reform.
+
+I solve lifecycle models for both of these cases and simulate consumption decisions with and without annuitisation.
+I then apply the empirical models to this simulated data and measure the consumption change in early retirement that results from annuitisation.
+The sign and magnitude of these coefficients relative to the coefficients from the real data will be indicative of the mechanism causing the annuitisation puzzle.
+I find that the lifecycle model with bequests best matches the coefficients from the regression discontinuity models, and interpret this as evidence that bequests are the dominant reason for the annuity puzzle.
 
@@
 The main dataset I use is the English Longitudinal Study of Ageing (ELSA)
-(\cite{main-elsa-citation}). ELSA picks individuals over the age of 50 to survey
-every two years until death. If individuals leave the survey, ELSA replaces them
-so that it is representative of the over 50 population in the UK. Individuals
-are asked a range of questions relating to their income and wealth as well as
+(\cite{main-elsa-citation}).
+ELSA follows a sample of individuals over the age of 50, interviewing them every two years until death.
+If individuals leave the survey, ELSA replaces them so that it is representative of the over 50 population in the UK.
+Individuals are asked a range of questions relating to their income and wealth as well as
 expectations about the future. One benefit of using ELSA is that it includes
@@
 
-The impact of the reforms on annuity demand has been documented by
-\cite{cannon-et-al-nier-2016}. Figure \ref{fig:annovertime} shows
-purchases of annuities over time and demonstrates the sharp decrease in purchases
-that ocurred from 2014 to 2015. There was an increase in the number of pensions
-that were being accessed using an income drawdown product, but these only partly
-account for the drop in annuitisation as is shown by the graph.
+The impact of the reforms on annuity demand has been documented by~\cite{cannon-et-al-nier-2016}.
+Figure~\ref{fig:annovertime} shows purchases of annuities over time and demonstrates the sharp decrease in purchases
+that occurred from 2014 to 2015.
+There was an increase in the number of pensions that were being accessed using an income drawdown product, but these only partly account for the drop in annuitisation as is shown by the graph.
 
@@
 state pension age for women from age 60 in 2010 to age 65 in 2018, so early cohorts had slightly earlier state pension eligibility ages.
-But, this change happened gradually and ocurred over the whole period, so I do not expect it to influence the results.
+But, this change happened gradually and occurred over the whole period, so I do not expect it to influence the results.
 The DifferenceAge row tracks the difference between expected and actual retirement age.
